\documentclass{article}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{xcolor}
\usepackage[most]{tcolorbox}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, calc, fit, backgrounds, shadows.blur}
\geometry{a4paper, margin=1in}

\newcommand{\code}[1]{\texttt{\small #1}}
\newcommand{\textftol}{\tau_{\mathrm{ftol}}}
\newcommand{\textstag}{\tau_{\mathrm{stag}}}
\newcommand{\textxtol}{\tau_{\mathrm{xtol}}}
\newcommand{\ftolinner}[1]{\tau^{\,\mathrm{inner}}_{\mathrm{ftol},#1}}

% --- Extension highlighting ---------------------------------------------
\definecolor{extcolor}{rgb}{0.00,0.35,0.70}
\newcommand{\ext}[1]{{\color{extcolor}\textbf{[ext]}\ #1}}
\newcommand{\extline}[1]{{\color{extcolor}#1}}

% --- Module-interface card ---------------------------------------------
% A small grey box rendered immediately above each algorithm declaration.
% Arguments: 1=called by, 2=calls, 3=inputs, 4=outputs.
\newtcolorbox{interfacecard}{
    colback=gray!5, colframe=gray!55, boxrule=0.4pt, arc=2pt,
    left=8pt, right=8pt, top=4pt, bottom=4pt,
    width=\textwidth, before skip=4pt, after skip=2pt
}
\newcommand{\moduleinterface}[4]{%
\begin{interfacecard}
\footnotesize
\begin{tabular}{@{}r@{\quad}p{0.82\textwidth}@{}}
\textit{Called by:} & #1 \\
\textit{Calls:}     & #2 \\
\textit{Inputs:}    & #3 \\
\textit{Outputs:}   & #4 \\
\end{tabular}
\end{interfacecard}
}

% Diagram styling
\tikzset{
    modnode/.style    = {draw, rounded corners=3pt, align=center, font=\small,
                         inner sep=5pt, fill=blue!8, drop shadow={opacity=0.2,
                         shadow xshift=0.5pt, shadow yshift=-0.5pt}},
    leafnode/.style   = {modnode, fill=blue!5},
    extnode/.style    = {modnode, fill=extcolor!10, draw=extcolor!70},
    edge/.style       = {-{Stealth[scale=0.9]}, thick, gray!70},
    edgelbl/.style    = {font=\scriptsize\itshape, fill=white, inner sep=1pt},
}

\hypersetup{
    colorlinks=true,
    linkcolor=black,
    urlcolor=black,
    citecolor=black
}

\title{Algorithmic Structure of the Stabilized Porous Navier--Stokes Solver}
\author{Antigravity AI \and Project Authors}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
\sloppy
This document specifies the nonlinear solver pipeline that drives a
single invocation of \code{run\_simulation} in the stabilized porous
Navier--Stokes solver. The pipeline is presented as a strict module
hierarchy: at the top sits the orchestrator (Algorithm O), which calls
two mid-level routines (Algorithms B and C, the Stage I and Stage II
solvers), each of which in turn invokes a shared Newton kernel
(Algorithm A), which is itself composed of three leaf primitives
(Algorithms A.1, A.2, A.3). A graphical map of this hierarchy is given
in Section~\ref{sec:tree}. Each algorithm is documented in two parts:
a brief \emph{interface card} listing its callers, callees, inputs and
outputs; and a procedural pseudocode box.

The presentation is organised in two strict tiers. Part~\ref{part:core}
describes the \emph{core algorithm}. Part~\ref{part:ext} documents two
\emph{optional extensions} that augment the core algorithm in specific
settings: a plateau verifier based on the Method of Manufactured
Solutions, which fires only when a candidate exact solution is supplied;
and a parameter-sweep harness that re-tunes a small number of
tolerances on a per-case basis. The two tiers are kept strictly
separate: passages and algorithm-line additions belonging to the
extensions are typeset in \extline{this colour} throughout.
\end{abstract}

\tableofcontents

% ============================================================
\part{The Core Algorithm}
\label{part:core}
% ============================================================

\section{Overview and Algorithmic Challenges}
\label{sec:overview}

The continuous problem is the stationary, incompressible, generalized
porous Navier--Stokes system on a domain $\Omega \subset \mathbb{R}^d$,
\begin{equation}
    \label{eq:strong}
    \begin{aligned}
    \boldsymbol{u} \cdot \nabla \boldsymbol{u}
        - \nabla \cdot (2 \nu \nabla^{\mathrm{s}} \boldsymbol{u})
        + \sigma(\alpha, \boldsymbol{u}) \boldsymbol{u}
        + \nabla p
        &= \boldsymbol{f}, \\
    \nabla \cdot \boldsymbol{u} &= 0,
    \end{aligned}
\end{equation}
discretised by an equal-order Lagrangian pair $(\boldsymbol{V}_h, Q_h)$
of degree $(k_v, k_p) = (k, k)$ and stabilised by a Variational
Multiscale (VMS) formulation. The drag tensor
$\sigma(\alpha, \boldsymbol{u}) = a(\alpha) + b(\alpha) |\boldsymbol{u}|$
couples the Darcy--Forchheimer regime to the convective momentum
balance; the porosity field $\alpha$ varies in space, and the
high-$\mathrm{Re}$, high-$\mathrm{Da}$ regimes the solver must handle
are exactly those in which the system becomes simultaneously
convection-dominated, reaction-dominated, and stiff.

\paragraph{Nonlinearity from convection and Forchheimer drag.}  Newton's
method converges quadratically once the iterate enters its quadratic
basin, but locating that basin from an arbitrary initial guess is
non-trivial when both $\boldsymbol{u}\cdot\nabla\boldsymbol{u}$ and the
$|\boldsymbol{u}|$ factor in $\sigma$ amplify any error in
$\boldsymbol{u}$. The solver must detect divergence and respond---either
by shortening the step (line search), by switching to an iteration map
with a wider basin (Picard), or by aborting cleanly.

\paragraph{Implicit OSGS projection.}  In the Orthogonal Sub-Grid Scale
formulation the sub-grid component lives in the orthogonal complement
$\mathcal{X}_{h0}^{\perp}$ of the finite element space; in practice the
projection $\boldsymbol{\pi}_h \in \mathcal{X}_h$ of the strong residual
is computed, and the sub-grid scale is identified with
$(I-\Pi_h)\mathcal{R}$. The projection itself depends on
$U_h = (\boldsymbol{u}_h, p_h)$, so a fully monolithic Jacobian would
embed the Fr\'echet derivative of an $L^2$ projection operator. Rather
than carrying this dense block, the solver freezes $\boldsymbol{\pi}_h$,
solves the nonlinear primary system, then re-projects, in a staggered
(block-coordinate) fixed point.

\paragraph{Architecture of the response.}  The architecture that emerges
is a two-stage cascade. \emph{Stage~I} drives the algebraic system to a
converged ASGS state, regardless of the final stabilization method
requested. \emph{Stage~II} either reports the ASGS state directly or,
when the user has selected OSGS, runs a staggered outer loop on
$(U_h, \boldsymbol{\pi}_h)$ initialised at the Stage~I solution. The
outermost orchestration (Algorithm~\ref{alg:O}) wires all of these
together.

\begin{quote}
\itshape
ASGS is always run first; OSGS, when requested, refines the ASGS state
through a staggered projection loop.
\end{quote}

% ============================================================
\section{Module Hierarchy}
\label{sec:tree}

The algorithms below are organised in a strict call hierarchy. The
orchestrator (Algorithm~\ref{alg:O}) sits at the top and dispatches to
two stage solvers (Algorithms~\ref{alg:B} and~\ref{alg:C}), both of
which build on a single shared Newton kernel
(Algorithm~\ref{alg:A}). The Newton kernel itself is built from three
leaf primitives (Algorithms~\ref{alg:A1}, \ref{alg:A2}, \ref{alg:A3}).
Figure~\ref{fig:hierarchy} shows the full tree.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=14mm and 10mm]

\node[modnode] (O)
    {\textbf{Algorithm~\ref{alg:O}}\\$\mathrm{SimulationOrchestration}$\\
     \footnotesize entry point};

\node[modnode, below left=18mm and 22mm of O] (B)
    {\textbf{Algorithm~\ref{alg:B}}\\$\mathrm{RobustNonlinearCascade}$\\
     \footnotesize Newton$\to$Picard$\to$Newton cascade};

\node[modnode, below right=18mm and 22mm of O] (C)
    {\textbf{Algorithm~\ref{alg:C}}\\$\mathrm{OSGSFractionalRelaxation}$\\
     \footnotesize Stage~II (OSGS mode)};

\node[modnode, below=30mm of O] (A)
    {\textbf{Algorithm~\ref{alg:A}}\\$\mathrm{ExactNewtonPipeline}$\\
     \footnotesize shared Newton kernel};

\node[leafnode, below left=14mm and 8mm of A] (A1)
    {\textbf{Algorithm~\ref{alg:A1}}\\$\mathrm{SolveLinearSystem}$};
\node[leafnode, below=14mm of A] (A2)
    {\textbf{Algorithm~\ref{alg:A2}}\\$\mathrm{ArmijoLinesearch}$};
\node[leafnode, below right=14mm and 8mm of A] (A3)
    {\textbf{Algorithm~\ref{alg:A3}}\\$\mathrm{EvaluateSafeguards}$};

\draw[edge] (O.south) to[bend right=12]
    node[edgelbl, midway] {Stage~I} (B.north);
\draw[edge] (O.south) to[bend left=12]
    node[edgelbl, midway] {Stage~II (OSGS only)} (C.north);
\draw[edge] (B.south) to[bend right=5]
    node[edgelbl, midway, fill=white] {Newton / Picard mode} (A.north west);
\draw[edge] (C.south) to[bend left=5]
    node[edgelbl, midway, fill=white] {inner cascade (per outer iter.)} (A.north east);
\draw[edge, dashed] (C.west) to[bend right=20]
    node[edgelbl, midway, fill=white] {reuses (Alg.~\ref{alg:B})} (B.east);
\draw[edge] (A.south) -- (A1.north);
\draw[edge] (A.south) -- (A2.north);
\draw[edge] (A.south) -- (A3.north);

\end{tikzpicture}
\caption{Module hierarchy of the core algorithm. Arrows point from
caller to callee. Each edge is annotated with the calling context.
Algorithm~\ref{alg:A} is the shared workhorse: Algorithm~\ref{alg:B}
invokes it three times in sequence (Newton, then Picard, then Newton)
with different Jacobians, and is itself reused by Algorithm~\ref{alg:C}
once per staggered outer iteration with the projection
$\boldsymbol{\pi}_h$ frozen (the dashed edge). Leaf primitives
A.1--A.3 are pure procedures with no further dependencies.}
\label{fig:hierarchy}
\end{figure}

\paragraph{Reading order.}  This document presents the modules
top-down: first the orchestrator (§\ref{sec:orchestration}), then the
two stage solvers (§\ref{sec:picard}, §\ref{sec:osgs}), then the shared
Newton kernel and its leaves (§\ref{sec:newton}). Each algorithm is
introduced by a short \emph{interface card} that summarises its
position in the tree---callers, callees, inputs, outputs---before the
procedural box. Mathematical preliminaries shared by several algorithms
(the diagonally equilibrated merit function and its directional
derivative identity) are gathered in §\ref{sec:math-prelim}.

% ============================================================
\section{Parameter Taxonomy and Code Equivalence}
\label{sec:params}

\subsection{Numerical parameters of the core algorithm}

Table~\ref{tab:params} enumerates every numerical parameter consumed by
the algorithms of Part~\ref{part:core}, grouped by role. Parameters
that belong to the optional extensions of Part~\ref{part:ext} are
tabulated separately in the corresponding sections.

\begin{table}[h]
\centering
\caption{Numerical parameters of the core nonlinear-solver pipeline.
Throughout the table, $\code{sol} = \code{config.numerical\_method.solver}$
(a flat \code{SolverConfig} \code{@kwdef} struct) and
$\code{stab} = \code{config.numerical\_method.stabilization}$. The
\code{SafeNewtonSolver} struct that runs the inner cascade is built from
these fields by \code{SafeNewtonSolver(::LinearSolver, ::SolverConfig)};
its post-construction field names (e.g.\ \code{c1}, \code{max\_iters})
are shorter aliases of the config field names. Code-side references in
this table are written against the config struct, not the built solver.}
\label{tab:params}
\footnotesize
\begin{tabularx}{\textwidth}{@{} l l X @{}}
\toprule
\textbf{Symbol} & \textbf{Field in code} & \textbf{Role / effect} \\
\midrule
\multicolumn{3}{l}{\emph{Inner Newton iteration controls}} \\
$N_{\mathrm{N}}$         & \code{sol.newton\_iterations}           & Cap on inner Newton iterations. \\
$N_{\mathrm{P}}$         & \code{sol.picard\_iterations}           & Cap on Picard fallback iterations. \\
$N_{\mathrm{inc}}$       & \code{sol.max\_increases}               & Consecutive merit-function expansions tolerated before declaring divergence. \\
$\textftol$              & \code{sol.ftol}                          & Terminal residual tolerance in $\ell^\infty$; the dominant convergence criterion. \\
$\textxtol$              & \code{sol.xtol}                          & Lower bound on the accepted step norm $\|\alpha\delta\boldsymbol{x}\|$ before declaring coordinate stagnation. \\
$\textstag$              & \code{sol.stagnation\_noise\_floor}      & Noise-floor residual; a non-converged exit below this may still be classified as success (Sec.~\ref{sec:falseEquiv}). \\
$\alpha_{\min}$          & \code{sol.linesearch\_alpha\_min}        & Smallest line-search step before giving up. \\
$c_1$                    & \code{sol.armijo\_c1}                    & Armijo / monotone-residual descent constant. \\
$\gamma$                 & \code{sol.linesearch\_contraction\_factor} & Backtracking factor: $\alpha \leftarrow \gamma\alpha$. \\
$N_{\mathrm{ls}}$        & \code{sol.max\_linesearch\_iterations}   & Bound on line-search inner iterations. \\
$F_{\mathrm{div}}$       & \code{sol.divergence\_merit\_factor}     & Threshold by which the merit must expand to count as one ``increase''. \\
$\mathfrak{m}$           & (built into the \code{SafeNewtonSolver} as \code{.mode})  & Cascade mode: \code{:newton} (merit-Armijo) or \code{:picard} (monotone residual reduction); cf.\ §\ref{sec:dir-deriv} and Alg.~\ref{alg:A2}. \\
\code{freeze\_cusp} & \code{sol.freeze\_jacobian\_cusp}   & Boolean: skip Jacobian recomputation at Forchheimer cusps to avoid singular updates. \\
\midrule
\multicolumn{3}{l}{\emph{OSGS staggered outer loop}} \\
$N_{\mathrm{OSGS}}$      & \code{stab.osgs\_iterations}             & Cap on outer (staggered) iterations. \\
$N_{\mathrm{inner}}$     & \code{stab.osgs\_inner\_newton\_iters}   & Cap on inner Newton iterations per outer step. \\
$\tau_{\mathrm{OSGS}}$   & \code{stab.osgs\_tolerance}              & Tolerance on the state drift $d_{\mathrm{state}}^m$. \\
$\tau_{\mathrm{proj}}$   & \code{stab.osgs\_projection\_tolerance}  & Tolerance on the mass-weighted projection drift $d_{\pi}^m$. \\
$m_{\mathrm{w}}$         & \code{stab.osgs\_warmup\_iterations}     & Outer iterations during which the inner Newton uses a relaxed tolerance. \\
$\tau_{\mathrm{w}}$      & \code{stab.osgs\_warmup\_tolerance}      & Relaxed inner tolerance during warmup. \\
$\mathsf{Mode}_{\mathrm{drift}}$ & \code{stab.osgs\_state\_drift\_scale} & Drift metric: \code{Linf} or \code{L2\_mass}. See Sec.~\ref{sec:drift-modes}. \\
$\mathsf{Mode}_{\mathrm{stop}}$  & \code{stab.osgs\_stopping\_mode}      & Convergence predicate: \code{state\_drift}, \code{projection\_drift}, or \code{both}. \\
$m$, $\beta$             & \code{sol.accelerator.\{m,relaxation\_factor\}} & Anderson history depth and damping. \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Algorithm--to--code correspondence}

\begin{table}[h]
\centering
\caption{Algorithm $\leftrightarrow$ code signature mapping (core algorithm).}
\label{tab:functions}
\small
\begin{tabularx}{\textwidth}{@{} l X @{}}
\toprule
\textbf{Algorithm} & \textbf{Code entry point} \\
\midrule
$\mathrm{SimulationOrchestration}$ (Alg.~\ref{alg:O})   & \code{run\_simulation(path)} $\to$ \code{solve\_system(setup, form, solvers, cfg, x0)}. \\
$\mathrm{RobustNonlinearCascade}$ (Alg.~\ref{alg:B})    & The Newton$\to$Picard$\to$Newton cascade inside \code{solve\_system}; appears twice---once for Stage~I (ASGS operators, $\boldsymbol{x}^0_{\mathrm{backup}}$) and once inside the OSGS outer-loop body of Algorithm~\ref{alg:C} (OSGS operators, $\boldsymbol{x}_{\mathrm{prev}}$ at the start of each outer iteration). \\
$\mathrm{OSGSFractionalRelaxation}$ (Alg.~\ref{alg:C})  & The OSGS staggered loop inside \code{solve\_system}. \\
$\mathrm{ExactNewtonPipeline}$ (Alg.~\ref{alg:A})       & \code{\_safe\_solve\_inner!} (\code{src/solvers/nonlinear.jl}), reused with either an exact-Newton or a Picard Jacobian. \\
$\mathrm{SolveLinearSystem}$ (Alg.~\ref{alg:A1})        & \code{eval\_linear\_system\_resolution!(dx, A, b, solver, cache)}. \\
$\mathrm{ArmijoLinesearch}$ (Alg.~\ref{alg:A2})         & \code{eval\_armijo\_linesearch\_pass!}. \\
$\mathrm{EvaluateSafeguards}$ (Alg.~\ref{alg:A3})       & \code{eval\_safeguard\_termination\_bounds!}. \\
\bottomrule
\end{tabularx}
\end{table}

% ============================================================
\section{Mathematical Preliminaries}
\label{sec:math-prelim}

Two mathematical objects are referenced repeatedly by several
algorithm boxes below. We define them once here so that the boxes
themselves remain procedural.

\subsection{The diagonally equilibrated merit function}
\label{sec:merit}

Let $\boldsymbol{b}(\boldsymbol{x}) \in \mathbb{R}^N$ be the assembled
residual of the (possibly stabilised) discrete weak form, and
$\mathcal{J}(\boldsymbol{x}) = \partial \boldsymbol{b}/\partial \boldsymbol{x}$
its Jacobian. The line-search merit function used throughout the
cascade is the diagonally equilibrated quadratic
\begin{equation}
    \label{eq:merit}
    \Phi_{\boldsymbol{w}}(\boldsymbol{x})
        = \tfrac{1}{2} \sum_{k=1}^{N}
            \left( \frac{b_k(\boldsymbol{x})}{w_k} \right)^{2}.
\end{equation}
The weights $w_k$ are recomputed at every Newton iteration from the
diagonal of the current Jacobian. Their definition is mode-dependent:
when the discrete trial space is a multi-field FE space with a
$\mathrm{ConsecutiveMultiFieldStyle}$ DOF ordering (the only case used
in this codebase: velocity $\oplus$ pressure), the equilibration is
performed \emph{per block}. Let $\{\mathcal{B}_q\}_{q}$ denote the
disjoint per-field DOF index ranges (so $\bigcup_q \mathcal{B}_q
= \{1,\dots,N\}$). Then
\begin{equation}
    \label{eq:merit-weights}
    w_k =
    \begin{cases}
        \max\!\bigl(|\mathcal{J}_{kk}|,\ \varepsilon_{\mathrm{mach}}\,
            \max_{j \in \mathcal{B}_{q(k)}} |\mathcal{J}_{jj}|,\
            \varepsilon_{\mathrm{mach}} \bigr)
            & \text{multi-field, $k \in \mathcal{B}_{q(k)}$,} \\[1mm]
        \max\!\bigl(|\mathcal{J}_{kk}|,\ \varepsilon_{\mathrm{mach}}\,
            \|\mathrm{diag}(\mathcal{J})\|_\infty,\
            \varepsilon_{\mathrm{mach}} \bigr)
            & \text{single-field fallback.}
    \end{cases}
\end{equation}
The first form is the production rule for the saddle-point system at
issue; the second is the legacy global rule, retained only for the
unrealistic single-field case in which $\mathcal{B}_q$ cannot be
detected.

\paragraph{Why equilibrate?}  The discrete residual mixes momentum-balance
rows---measured in physical units of force per unit volume, and possibly
amplified by a large Reynolds or Damk\"ohler number---with mass-balance
rows, whose natural magnitude is dimensionally different. A raw unweighted
merit $\tfrac{1}{2}\|\boldsymbol{b}\|^2$ would let the momentum rows
dominate entirely, so the line search would be quietly indifferent to
mass-conservation errors. Equilibrating by $w_k$ rescales every row to
its own natural unit. The lower-bound floor
$\varepsilon_{\mathrm{mach}}\,\max_{j \in \mathcal{B}_{q(k)}} |\mathcal{J}_{jj}|$
prevents a genuinely small diagonal entry from triggering division by
zero.

\paragraph{Why per block?}  A global $\|\mathrm{diag}(\mathcal{J})\|_\infty$
in the floor would be dominated by the momentum-block diagonal whenever
the Reynolds or Darcy regime is severe; the pressure-block floor
would then be so close to $\varepsilon_{\mathrm{mach}}\,$(momentum scale)
that mass-residual rows would still be effectively un-weighted relative
to momentum-residual rows, defeating the equilibration entirely. The
per-block floor scales each field's residual against \emph{its own}
diagonal magnitudes, which is the only choice that gives the line search
a faithful view of mass-residual descent.

\subsection{The Gauss--Newton directional-derivative identity}
\label{sec:dir-deriv}

For a Newton step $\delta\boldsymbol{x}$ satisfying
$\mathcal{J}\delta\boldsymbol{x} = \boldsymbol{b}$, the directional
derivative of $\Phi_{\boldsymbol{w}}$ along $-\delta\boldsymbol{x}$ at
$\boldsymbol{x}$ is
\begin{equation}
    \label{eq:dir-deriv}
    D \;=\; -\sum_k w_k^{-2}\, b_k\,(\mathcal{J}\delta\boldsymbol{x})_k
        \;=\; -\sum_k w_k^{-2}\, b_k^2
        \;=\; -2\,\Phi_{\boldsymbol{w}}(\boldsymbol{x}).
\end{equation}
The Newton-mode Armijo predicate of Algorithm~\ref{alg:A2} therefore
reduces to $\hat\Phi \le \Phi(1 - 2 c_1 \alpha)$, which requires no extra
gradient evaluation.

\paragraph{Picard mode invalidates the identity.}  The identity~\eqref{eq:dir-deriv}
holds \emph{only} because the search direction is the exact Newton step.
The Picard Jacobian drops the linearised convective cross-term and the
linearised Forchheimer term (cf.\ §\ref{sec:picard}), so the resulting
$\delta\boldsymbol{x}$ does not satisfy $\mathcal{J}\delta\boldsymbol{x}
= \boldsymbol{b}$ for the true $\mathcal{J}$; the cancellation in the
second equality above breaks and $D$ is in general not $-2\Phi$. Forcing
the merit-based Armijo predicate on a Picard step would either reject
descent directions for the wrong reason or accept ascent directions
spuriously. The implementation therefore branches the acceptance test by
solver mode: Newton uses the merit-based Armijo above; Picard uses the
monotone residual reduction
$\|\hat{\boldsymbol{b}}\|_\infty \le (1 - c_1 \alpha)\|\boldsymbol{b}\|_\infty$,
which is mode-agnostic because it depends only on the residuals at the
two iterates. This branching is shown explicitly inside
Algorithm~\ref{alg:A2}.

The same mode-awareness extends to the \emph{divergence} safeguard in
Algorithm~\ref{alg:A3}: in Newton mode we compare $\hat\Phi$ against
$\Phi$, because the Armijo line search is itself driving $\Phi$ down
between successive iterates and a sustained $\Phi$-growth signals a
structural problem. In Picard mode we cannot use $\Phi$ for this purpose,
because the merit weights $\boldsymbol{w} = \mathrm{diag}(\mathbf{J})$ are
refreshed every iteration from a Picard Jacobian whose diagonal differs
from Newton's, so $\Phi$ is not even monotone-comparable across iterations
even when $\|\boldsymbol{b}\|_\infty$ is shrinking monotonically. Picard
mode therefore compares $\|\hat{\boldsymbol{b}}\|_\infty$ against
$\|\boldsymbol{b}\|_\infty$ directly. Empirically, applying the
$\Phi$-based check unconditionally produced spurious
$\mathsf{MeritDivergence}$ exits at high Reynolds whenever Picard
fallbacks fired; the corrected mode-aware rule unblocked the
$(\mathrm{Re} = 10^6,\, \mathrm{Da} = 1,\, \alpha_0 = 0.05)$ stiff cell
without any continuation driver (see \code{probe\_stiff\_findings.md},
2026-05-19).

% ============================================================
\section{The Top: Algorithm O (Orchestrator)}
\label{sec:orchestration}

The orchestration drives one simulation end-to-end. Its central
organising principle is a two-stage cascade: an unconditional ASGS
algebraic initialisation followed by a method-dependent dispatch.
Stage~I runs even when the user has requested OSGS, because the
staggered projection loop is much better conditioned when initialised
at a converged ASGS state than at a zero initial guess.

\moduleinterface
    {(top-level entry point; not called from elsewhere)}
    {Algorithm~\ref{alg:B} (Stage~I); Algorithm~\ref{alg:C} (Stage~II, OSGS mode only)}
    {Configuration $\Theta$ (validated JSON)}
    {Final fields $(u_h, p_h, \boldsymbol{\pi}_h)$; success flag $S$}

\begin{algorithm}[H]
\caption{$\mathrm{SimulationOrchestration}$: end-to-end driver.}
\label{alg:O}
\SetAlgoLined
\KwIn{Configuration $\Theta$.}
\KwOut{$(u_h, p_h, \boldsymbol{\pi}_h)$, success flag $S$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}

build mesh $M$, FE pair $(X, Y)$, unconstrained mirrors $(V_{\mathrm{free}}, Q_{\mathrm{free}})$, formulation $F$, topology $T$, initial guess $\boldsymbol{x}^{0}$\;

\vspace{0.15cm}
\tcc{\textbf{Stage I.} Algebraic initialisation in ASGS mode ($\boldsymbol{\pi}_h \equiv 0$).}
$(S, U_h^{\mathrm{ASGS}}) \gets \mathrm{RobustNonlinearCascade}(\boldsymbol{x}^0,\ \boldsymbol{x}^0_{\mathrm{backup}},\ \mathrm{op}_{\mathrm{N}}^{\mathrm{ASGS}},\ \mathrm{op}_{\mathrm{P}}^{\mathrm{ASGS}})$ \tcp*{Alg.~\ref{alg:B}}
\If{\textnormal{\textbf{not }} $S$}{
    \Return failure (no fields written)\;
}

\vspace{0.15cm}
\tcc{\textbf{Stage II.} Method dispatch.}
\If{$\Theta.\mathrm{method} = \mathrm{ASGS}$}{
    \extline{\If{$\code{mms\_cfg.enabled}$}{
        $I_{\mathrm{pass}} \gets 0$;\quad initialise plateau-history $\mathcal{H}$ at $U_h^{\mathrm{ASGS}}$\;
        \For{$\mathrm{cycle} = 1, \dots, K_{\mathrm{extra}}$}{
            run one Newton iteration on top of $U_h^{\mathrm{ASGS}}$\;
            $(S_{\mathrm{plat}}, I_{\mathrm{pass}}) \gets \mathrm{VerifyMMSPlateau}(\dots)$ \tcp*{Alg.~\ref{alg:D}}
            \If{$S_{\mathrm{plat}}$}{ \textbf{break}\; }
        }
    }}
    \Return $(U_h^{\mathrm{ASGS}}, \boldsymbol{\pi}_h \equiv 0)$\;
}
\If{$\Theta.\mathrm{method} = \mathrm{OSGS}$}{
    pre-assemble and LU-factor $M_u = (V_{\mathrm{free}}, U_{\mathrm{free}})_{L^2}$, $M_p = (Q_{\mathrm{free}}, P_{\mathrm{free}})_{L^2}$ once\;
    $\boldsymbol{\pi}_h^{0} \gets 0$ \tcp*{deliberate zero bootstrap; see §\ref{sec:osgs-bootstrap}}
    \extline{\If{$\code{mms\_cfg.enabled}$}{
        $N_{\mathrm{OSGS}}^{\mathrm{eff}} \gets N_{\mathrm{OSGS}} + K_{\mathrm{extra}}$ \tcp*{inflate outer-loop budget for plateau verification}
    }}
    $(S, \boldsymbol{\pi}_h) \gets \mathrm{OSGSFractionalRelaxation}(U_h^{\mathrm{ASGS}}, \boldsymbol{\pi}_h^{0})$ \tcp*{Alg.~\ref{alg:C}; MMS hook in Alg.~\ref{alg:C} fires when \code{mms\_cfg.enabled}}
}

\vspace{0.15cm}
$\mathrm{export\_results}(\mathrm{VTK} \mid \mathrm{HDF5})$\;
\end{algorithm}

\paragraph{Step by step.}  The construction lines build the mesh, the
finite-element pair, and the unconstrained projection mirrors. The
unconstrained spaces $V_{\mathrm{free}}, Q_{\mathrm{free}}$ are required
only by the OSGS projection step (§\ref{sec:osgs}); they carry no
Dirichlet data, because projecting onto a Dirichlet-constrained space
introduces an $O(1)$ boundary residual that destroys the expected
$O(h^{k+1})$ FE convergence rate.

\paragraph{Stage I.}  A single call to $\mathrm{RobustNonlinearCascade}$
(Algorithm~\ref{alg:B}) composes an exact-Newton attempt, a Picard
globalisation pass, and a final Newton polish, all with
$\boldsymbol{\pi}_h \equiv 0$. The output $U_h^{\mathrm{ASGS}}$ is the
algebraic root of the ASGS system; when Stage~II is OSGS, the
staggered loop descends from this state with an already-converged
primary field. The same cascade module is reused inside Stage~II
(Algorithm~\ref{alg:C}) with the OSGS operators and a per-outer-iteration
restore vector.

\paragraph{Stage II.}  Either accept the ASGS state, or hand off to the
OSGS outer loop. The mass matrices $M_u, M_p$ are pre-assembled and
LU-factored exactly once before the OSGS loop begins. Because the
projection target spaces never change during the loop, these
factorisations are reused at every outer iteration, which converts
every $L^2$ projection into a constant-cost forward/back substitution.

\paragraph{Failure propagation.}  If Stage~I returns $S = \mathrm{False}$,
the orchestrator returns immediately without writing output. The
diagnostics cache carries the algebraic exit reason so the calling
script can distinguish a divergent run from a converged one.

% ============================================================
\section{The shared cascade: Algorithm B (RobustNonlinearCascade)}
\label{sec:picard}

When the safeguarded Newton loop terminates without converging, the
orchestration does not give up: it transitions to a Picard iteration
map and then re-engages Newton. Algorithm~\ref{alg:B} packages this
Newton $\to$ Picard $\to$ Newton cascade as a single \emph{generic}
module that does not depend on which stabilisation mode it is driving.
It is therefore reused as-is by both Stage~I (called from
Algorithm~\ref{alg:O} with $\boldsymbol{\pi}_h \equiv 0$ and exact-Newton
and Picard ASGS operators) and the inner solve of Stage~II (called from
Algorithm~\ref{alg:C} once per staggered outer iteration with
$\boldsymbol{\pi}_h$ frozen at the previous projection and OSGS-mode
operators).

The only behavioural axis that varies between the two callers is the
\emph{restore vector} used when Picard's first attempt has to be reset.
Stage~I restores to a copy of the original initial guess
$\boldsymbol{x}^0_{\mathrm{backup}}$ taken at the very start of the
simulation; Stage~II restores to the iterate at the start of the
current outer staggered iteration $\boldsymbol{x}_{\mathrm{prev}}$. The
algorithm box below takes the restore vector as an explicit parameter,
so that both callers reuse a single module without
caller-specific branching.

\moduleinterface
    {Algorithm~\ref{alg:O} (Stage~I); Algorithm~\ref{alg:C} (OSGS inner solve, per outer staggered iteration)}
    {Algorithm~\ref{alg:A} (invoked up to three times in sequence, with different Jacobians)}
    {Initial guess $\boldsymbol{x}^0$; restore vector $\boldsymbol{x}_{\mathrm{restore}}$; exact-Newton and Picard nonlinear operators $\mathrm{op}_{\mathrm{N}}, \mathrm{op}_{\mathrm{P}}$}
    {Success flag $S$ and the converged or best-seen iterate $\boldsymbol{x}^\star$}

\begin{algorithm}[H]
\caption{$\mathrm{RobustNonlinearCascade}$: the Newton $\to$ Picard $\to$ Newton cascade. Stage I instantiates $\boldsymbol{x}_{\mathrm{restore}}$ with a backup copy of the initial guess; Algorithm~\ref{alg:C} instantiates it with the iterate at the start of the current outer staggered iteration.}
\label{alg:B}
\SetAlgoLined
\textbf{Parameters:} $N_{\mathrm{N}}, N_{\mathrm{P}}, \textftol, \textstag$ and all Newton/line-search parameters from Table~\ref{tab:params}.\\
\KwIn{Initial guess $\boldsymbol{x}^0$; restore vector $\boldsymbol{x}_{\mathrm{restore}}$; exact-Newton operator $\mathrm{op}_{\mathrm{N}}$; Picard operator $\mathrm{op}_{\mathrm{P}}$.}
\KwOut{$(S, \boldsymbol{x}^\star)$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}
$(\boldsymbol{x}^\star, S_{\mathrm{N}}, \dots) \gets \mathrm{ExactNewtonPipeline}(\boldsymbol{x}^0,\ \mathrm{op}_{\mathrm{N}})$\;
\If{$S_{\mathrm{N}} = \mathrm{True}$}{
    \Return $(\mathrm{True}, \boldsymbol{x}^\star)$ \tcp*{Newton sufficed; no fallback}
}

restore $\boldsymbol{x} \gets \boldsymbol{x}_{\mathrm{restore}}$\;
$(\boldsymbol{x}^\star, S_{\mathrm{P}}, \dots) \gets \mathrm{ExactNewtonPipeline}(\boldsymbol{x},\ \mathrm{op}_{\mathrm{P}})$\;
\If{any component of $\boldsymbol{x}^\star$ is non-finite}{
    \Return $(\mathrm{False}, \boldsymbol{x}_{\mathrm{restore}})$ \tcp*{catastrophic state}
}
\If{$S_{\mathrm{P}} = \mathrm{True}$}{
    \Return $(\mathrm{True}, \boldsymbol{x}^\star)$\;
}

\tcc{Re-engage Newton from wherever Picard left the iterate.}
$(\boldsymbol{x}^\star, S_{\mathrm{N2}}, \dots) \gets \mathrm{ExactNewtonPipeline}(\boldsymbol{x}^\star,\ \mathrm{op}_{\mathrm{N}})$\;
\Return $(S_{\mathrm{N2}} \vee S_{\mathrm{P}},\ \boldsymbol{x}^\star)$\;
\end{algorithm}

\paragraph{Why Picard works where Newton fails.}  The Newton Jacobian
contains the linearised convective term
$\boldsymbol{u}^k \cdot \nabla \delta\boldsymbol{u} + \delta\boldsymbol{u} \cdot \nabla \boldsymbol{u}^k$
and the linearised Forchheimer term
$\partial\sigma/\partial\boldsymbol{u}$. Both become large in the
difficult regimes, and the Jacobian can develop a near-null direction
where a small change in $\boldsymbol{u}$ produces an outsize change in
the residual. Newton near such a cusp produces a step that diverges.
The Picard Jacobian drops the dependence of
$\boldsymbol{u}\cdot\nabla\boldsymbol{u}$ and of $\sigma$ on the
increment $\delta\boldsymbol{u}$, retaining only the ``frozen''
coefficient form. The resulting iteration map is a contraction in a
wider region than Newton's quadratic basin, at the cost of linear
convergence. Picard does not converge to within $\textftol$ in
general---but it does push the iterate into a region where the
subsequent Newton retry has a fighting chance.

\paragraph{Triggers.}  Any non-converged Newton exit triggers the
fallback, including $\mathsf{Xtol}$ stagnation, $\mathsf{IterationCap}$,
the linear-solver exception path, and any uncaught exception during
the Newton call. All of these states leave $\boldsymbol{x}^0$ outside
the basin, and a Picard pass on top costs the same regardless of which
symptom triggered the entry.

\paragraph{The catastrophic-state guard.}  Picard is not bounded in
general. After the Picard call, the orchestration checks
$\mathrm{any}(!\mathrm{isfinite}, \boldsymbol{x}^\star)$. A NaN or Inf
in the iterate is a hard signal that even Picard could not keep the
model evaluable, and the fallback restores the supplied
$\boldsymbol{x}_{\mathrm{restore}}$ and returns failure. Without this
guard the subsequent Newton call would crash on its first residual
evaluation.

% ============================================================
\section{The Newton Kernel: Algorithm A and Its Leaves}
\label{sec:newton}

Both Algorithms~\ref{alg:B} and~\ref{alg:C} call into a single
safeguarded Newton routine. Beneath this kernel sit three leaf
primitives: a linear-solve helper (A.1), an Armijo backtracking line
search (A.2), and a termination-safeguard arbiter (A.3). This section
documents the kernel first, then its leaves.

% --------------------------------------------------------------
\subsection{Algorithm A: ExactNewtonPipeline}

\moduleinterface
    {Algorithm~\ref{alg:B} (three times per Stage~I); Algorithm~\ref{alg:C} (once per outer staggered iteration)}
    {Algorithm~\ref{alg:A1} (linear solve); Algorithm~\ref{alg:A2} (Armijo line search); Algorithm~\ref{alg:A3} (safeguards)}
    {Initial fields $\boldsymbol{x}^0 \in \mathcal{X}_h$; nonlinear operator $\mathrm{op}$ (encoding the chosen Jacobian variant)}
    {Final iterate $\boldsymbol{x}^\star$; success flag $S$; iteration count $I$; residual norm $\|\boldsymbol{b}^\star\|_\infty$; exit reason $r$ (Table~\ref{tab:reasons})}

\begin{algorithm}[H]
\caption{$\mathrm{ExactNewtonPipeline}$: safeguarded Newton iteration with merit-based line search.}
\label{alg:A}
\SetAlgoLined
\textbf{Parameters:} $N_{\mathrm{N}}$, $\textftol$, $\textxtol$, $\textstag$, $c_1$, $\gamma$, $\alpha_{\min}$, $N_{\mathrm{ls}}$, $F_{\mathrm{div}}$, $N_{\mathrm{inc}}$.\\
\KwIn{$\boldsymbol{x}^0 \in \mathcal{X}_h$, nonlinear operator $\mathrm{op}$.}
\KwOut{$(\boldsymbol{x}^{\star}, S, I, \|\boldsymbol{b}^{\star}\|_\infty, r)$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}

$\boldsymbol{b} \gets \boldsymbol{b}(\boldsymbol{x}^0)$;\quad
\If{$\|\boldsymbol{b}\|_\infty \le \textftol$}{
    \Return $(\boldsymbol{x}^0,\ \mathrm{True},\ 0,\ \|\boldsymbol{b}\|_\infty,\ \mathsf{InitialFtol})$\;
}
$\boldsymbol{x}_{\mathrm{best}} \gets \boldsymbol{x}^0$;\quad
$b_{\mathrm{best}} \gets \|\boldsymbol{b}\|_\infty$;\quad
$I_{\mathrm{up}} \gets 0$\;
\For{$i = 1, \dots, N_{\mathrm{N}}$}{
    $\mathcal{J} \gets \mathcal{J}(\boldsymbol{x})$;\quad
    recompute weights $\boldsymbol{w}$ from $\mathrm{diag}(\mathcal{J})$ as in~\eqref{eq:merit};\quad
    $\Phi \gets \Phi_{\boldsymbol{w}}(\boldsymbol{x})$\;
    $(S_{\mathrm{LS}}^{\mathrm{lin}}, \delta\boldsymbol{x}) \gets \mathrm{SolveLinearSystem}(\mathcal{J}, \boldsymbol{b})$ \tcp*{Alg.~\ref{alg:A1}}
    \If{$S_{\mathrm{LS}}^{\mathrm{lin}} = \mathrm{Failed}$}{
        $\boldsymbol{x} \gets \boldsymbol{x}_{\mathrm{best}}$;\quad
        \Return $(\boldsymbol{x},\ \mathrm{False},\ i,\ b_{\mathrm{best}},\ \mathsf{LinearSolveNaN})$\;
    }
    set $D \gets -2\Phi$ via the identity~\eqref{eq:dir-deriv} (used by the Newton branch of A.2 only)\;
    $(S_{\mathrm{LS}}, \alpha, L, \hat b, \hat \Phi)
        \gets \mathrm{ArmijoLinesearch}(\boldsymbol{x},\boldsymbol{x}_{\mathrm{old}},
              \boldsymbol{b},\delta\boldsymbol{x}, D, \boldsymbol{w}, \Phi, \mathfrak{m})$ \tcp*{Alg.~\ref{alg:A2}, $\mathfrak{m}$ from \code{op}}
    \If{$\hat{b} < b_{\mathrm{best}}$}{ $\boldsymbol{x}_{\mathrm{best}} \gets \boldsymbol{x}$;\quad $b_{\mathrm{best}} \gets \hat{b}$\; }
    $(r, S_{\mathrm{conv}}, S_{\mathrm{brk}}, I_{\mathrm{up}})
        \gets \mathrm{EvaluateSafeguards}(i, S_{\mathrm{LS}}, L, \hat b, \|\boldsymbol{b}\|_\infty, \hat \Phi, \Phi, I_{\mathrm{up}}, \mathfrak{m})$ \tcp*{Alg.~\ref{alg:A3}}
    \If{$S_{\mathrm{brk}}$}{
        \If{$\neg S_{\mathrm{conv}}$ \textnormal{\textbf{and}} $r \notin \{\mathsf{Ftol}, \mathsf{NoiseFloor}, \mathsf{Xtol}\}$}{
            $\boldsymbol{x} \gets \boldsymbol{x}_{\mathrm{best}}$\;
        }
        \Return $(\boldsymbol{x},\ S_{\mathrm{conv}},\ i,\ \|\boldsymbol{b}^{\star}\|_\infty,\ r)$\;
    }
}
\Return $(\boldsymbol{x},\ \mathrm{False},\ N_{\mathrm{N}},\ \|\boldsymbol{b}\|_\infty,\ \mathsf{IterationCap})$\;
\end{algorithm}

\paragraph{Reading the box.}  Each outer iteration computes the Jacobian,
delegates the linear solve to Algorithm~\ref{alg:A1}, performs an
Armijo backtracking line search via Algorithm~\ref{alg:A2}, and asks
Algorithm~\ref{alg:A3} whether to break, continue, or declare
convergence.

\paragraph{Best-state retention.}  Whenever the residual strictly
improves, $(\boldsymbol{x}_{\mathrm{best}}, b_{\mathrm{best}})$ is
updated. On most non-converged exits the iterate is restored to this
best-seen point. Three exits intentionally do not restore:
$\mathsf{Ftol}$ and $\mathsf{NoiseFloor}$ are already at or below the
target, and $\mathsf{Xtol}$ stagnation reflects a state the iterate
could not move beyond---restoring would discard the most current
information.

\paragraph{Mode independence.}  The same loop is reused with either an
exact-Newton or a Picard Jacobian. The merit function, line search,
and safeguards apply unchanged; only the operator \code{op} differs.
This is what allows Algorithm~\ref{alg:B} to compose three calls to
this single kernel.

% --------------------------------------------------------------
\subsection{Algorithm A.1: SolveLinearSystem}

\moduleinterface
    {Algorithm~\ref{alg:A} (once per Newton iteration)}
    {Configured linear solver (LU by default)}
    {Jacobian $\mathcal{J}$; residual $\boldsymbol{b}$}
    {Failure flag $S_{\mathrm{Failed}}$; step $\delta\boldsymbol{x}$}

\begin{algorithm}[H]
\caption{$\mathrm{SolveLinearSystem}$: factor and solve, with NaN guard.}
\label{alg:A1}
\SetAlgoLined
\textbf{Parameters:} (none).\\
\KwIn{$\mathcal{J}$, $\boldsymbol{b}$.}
\KwOut{$(S_{\mathrm{Failed}}, \delta\boldsymbol{x})$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}
\textbf{try} \{ Solve $\mathcal{J} \delta\boldsymbol{x} = \boldsymbol{b}$ via the configured linear solver \} \textbf{catch} \{ \Return $(\mathrm{True}, \delta\boldsymbol{x})$ \tcp*{linear-solver exception} \} \\
\If{any element of $\delta\boldsymbol{x}$ is NaN}{
    \Return $(\mathrm{True}, \delta\boldsymbol{x})$ \tcp*{catastrophic algebraic failure}
}
\Return $(\mathrm{False}, \delta\boldsymbol{x})$\;
\end{algorithm}

The linear solve is the cheapest place to detect a near-singular
Jacobian: either the factorisation aborts with an exception or it
returns a vector contaminated by NaNs. Either way the caller restores
$\boldsymbol{x}_{\mathrm{best}}$ and exits.

% --------------------------------------------------------------
\subsection{Algorithm A.2: ArmijoLinesearch}

\moduleinterface
    {Algorithm~\ref{alg:A} (once per Newton iteration)}
    {(none; pure procedure)}
    {Current iterate $\boldsymbol{x}$; reference iterate $\boldsymbol{x}_{\mathrm{old}}$; residual $\boldsymbol{b}$; step $\delta\boldsymbol{x}$; directional derivative $D = -2\Phi$ (cf.\ §\ref{sec:dir-deriv}); weights $\boldsymbol{w}$; base merit $\Phi$; mode $\mathfrak{m} \in \{\textsc{Newton}, \textsc{Picard}\}$}
    {Acceptance flag $S_{\mathrm{LS}}$; step size $\alpha$; step norm $L$; post-step residual $\|\hat{\boldsymbol{b}}\|_\infty$ and merit $\hat\Phi$}

\begin{algorithm}[H]
\caption{$\mathrm{ArmijoLinesearch}$.}
\label{alg:A2}
\SetAlgoLined
\textbf{Parameters:} $\alpha_{\min}$, $c_1$, $\gamma$, $N_{\mathrm{ls}}$.\\
\KwIn{$\boldsymbol{x}, \boldsymbol{x}_{\mathrm{old}}, \boldsymbol{b}, \delta\boldsymbol{x}, D, \boldsymbol{w}, \Phi, \mathfrak{m}$.}
\KwOut{$(S_{\mathrm{LS}}, \alpha, L, \|\hat{\boldsymbol{b}}\|_\infty, \hat\Phi)$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}
$\alpha \gets 1$;\quad $S_{\mathrm{LS}} \gets \mathrm{False}$\;
\For{$j = 1, \dots, N_{\mathrm{ls}}$}{
    $\boldsymbol{x} \gets \boldsymbol{x}_{\mathrm{old}} - \alpha\,\delta\boldsymbol{x}$;\quad
    $\hat{\boldsymbol{b}} \gets \boldsymbol{b}(\boldsymbol{x})$;\quad
    $\hat{\Phi} \gets \Phi_{\boldsymbol{w}}(\hat{\boldsymbol{b}})$\;
    \If{$\hat\Phi$ \textnormal{\textbf{or}} $\|\hat{\boldsymbol{b}}\|_\infty$ is NaN}{
        $\alpha \gets \gamma \alpha$;\quad
        \textbf{continue} \tcp*{shrink and retry without testing acceptance}
    }
    \If{$\mathfrak{m} = \textsc{Newton}$}{
        \If{$\hat\Phi \le \Phi + c_1 \alpha D$}{
            $S_{\mathrm{LS}} \gets \mathrm{True}$;\quad \textbf{break} \tcp*{merit-Armijo accept}
        }
    }\Else(\tcp*[h]{$\mathfrak{m} = \textsc{Picard}$}){
        \If{$\|\hat{\boldsymbol{b}}\|_\infty \le (1 - c_1\alpha) \|\boldsymbol{b}\|_\infty$}{
            $S_{\mathrm{LS}} \gets \mathrm{True}$;\quad \textbf{break} \tcp*{monotone-residual accept}
        }
    }
    \If{$\alpha \le \alpha_{\min}$}{
        \textbf{break} \tcp*{return $S_{\mathrm{LS}} = \mathrm{False}$}
    }
    $\alpha \gets \gamma \alpha$\;
}
$L \gets \alpha \|\delta\boldsymbol{x}\|_\infty$\;
\Return $(S_{\mathrm{LS}}, \alpha, L, \|\hat{\boldsymbol{b}}\|_\infty, \hat\Phi)$\;
\end{algorithm}

\paragraph{Mode branching.}  The Newton branch implements the standard
sufficient-decrease test on the equilibrated merit (§\ref{sec:dir-deriv});
the Picard branch uses a monotone residual reduction in
$\ell^\infty$. The acceptance criterion is mode-dependent because the
$D = -2\Phi$ identity (eq.~\eqref{eq:dir-deriv}) relies on the search
direction being the exact Newton step. Picard steps drop the linearised
convective cross-term and the Forchheimer derivative from
$\mathcal{J}$, so the identity is generically violated; the merit-based
predicate would then be miscalibrated. Replacing it with a contraction
test on the $\ell^\infty$ residual decouples Picard's accept/reject
loop from $\Phi$ entirely, recovering a sound globalisation step for
the fixed-point map. The condition
$\|\hat{\boldsymbol{b}}\|_\infty \le (1 - c_1\alpha)\|\boldsymbol{b}\|_\infty$
is the standard Bank--Rose monotone-residual test for damped Picard.

\paragraph{NaN handling.}  When the residual evaluation at the trial
point produces a NaN merit \emph{or} a NaN $\ell^\infty$ residual norm
(typically because the Forchheimer factor $|\boldsymbol{u}|$ blew up
under a too-aggressive step), the line search shrinks $\alpha$ without
testing acceptance. The check on $\|\hat{\boldsymbol{b}}\|_\infty$ is
required because the Picard branch never reads $\hat\Phi$, and a
non-finite residual norm could otherwise reach the contraction test.
A Newton step pushing $\boldsymbol{u}$ into a region where the model
momentarily becomes non-evaluable should not be recorded as a failed
acceptance step; it should simply be retried at a smaller size.

% --------------------------------------------------------------
\subsection{Algorithm A.3: EvaluateSafeguards}

\moduleinterface
    {Algorithm~\ref{alg:A} (once per Newton iteration, after the line search)}
    {(none; pure decision procedure)}
    {Outer iteration index $i$; line-search success $S_{\mathrm{LS}}$; step norm $L$; post-step residual $\|\hat{\boldsymbol{b}}\|_\infty$ and merit $\hat\Phi$; base merit $\Phi$; increase counter $I_{\mathrm{up}}$}
    {Exit reason $r$; convergence flag $S_{\mathrm{conv}}$; break flag $S_{\mathrm{brk}}$; updated counter $I_{\mathrm{up}}'$}

\begin{algorithm}[H]
\caption{$\mathrm{EvaluateSafeguards}$: termination arbiter.}
\label{alg:A3}
\SetAlgoLined
\textbf{Parameters:} $\textftol$, $\textstag$, $\textxtol$, $F_{\mathrm{div}}$, $N_{\mathrm{N}}$, $N_{\mathrm{inc}}$.\\
\KwIn{$i, S_{\mathrm{LS}}, L, \|\hat{\boldsymbol{b}}\|_\infty, \|\boldsymbol{b}\|_\infty, \hat\Phi, \Phi, I_{\mathrm{up}}, \mathfrak{m} \in \{\textsc{Newton}, \textsc{Picard}\}$.}
\KwOut{$(r, S_{\mathrm{conv}}, S_{\mathrm{brk}}, I_{\mathrm{up}}')$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}
\If{$\|\hat{\boldsymbol{b}}\|_\infty \le \textftol$}{
    \Return $(\mathsf{Ftol}, \mathrm{True}, \mathrm{True}, I_{\mathrm{up}})$ \tcp*{converged on residual}
}
\If{$\neg S_{\mathrm{LS}}$}{
    \If{$\|\hat{\boldsymbol{b}}\|_\infty \le \textstag$}{
        \Return $(\mathsf{NoiseFloor}, \mathrm{True}, \mathrm{True}, I_{\mathrm{up}})$\;
    }
    \Return $(\mathsf{LinesearchFailed}, \mathrm{False}, \mathrm{True}, I_{\mathrm{up}})$\;
}
\tcp{Mode-aware divergence check: Newton compares the Armijo merit $\Phi$ (which the Newton line search itself drives down); Picard compares $\|\boldsymbol{b}\|_\infty$ directly, because the merit's weights $\boldsymbol{w} = \mathrm{diag}(\mathbf{J})$ are refreshed each iter from a Jacobian that differs structurally between Newton and Picard. $\Phi$ can grow across Picard iterations even when $\|\boldsymbol{b}\|_\infty$ shrinks monotonically; using $\Phi$ in Picard mode causes spurious $\mathsf{MeritDivergence}$ exits at high $\mathrm{Re}$ (settled in \code{probe\_stiff\_findings.md}, 2026-05-19).}
\If{$\mathfrak{m} = \textsc{Newton}$ \textnormal{\textbf{and}} $\hat\Phi > F_{\mathrm{div}} \Phi$ \textnormal{\textbf{or}} $\mathfrak{m} = \textsc{Picard}$ \textnormal{\textbf{and}} $\|\hat{\boldsymbol{b}}\|_\infty > F_{\mathrm{div}} \|\boldsymbol{b}\|_\infty$}{
    $I_{\mathrm{up}} \gets I_{\mathrm{up}} + 1$;\quad
    \If{$I_{\mathrm{up}} \ge N_{\mathrm{inc}}$}{
        \Return $(\mathsf{MeritDivergence}, \mathrm{False}, \mathrm{True}, I_{\mathrm{up}})$\;
    }
}\Else{ $I_{\mathrm{up}} \gets 0$\; }
\If{$L \le \textxtol$}{
    \If{$\|\hat{\boldsymbol{b}}\|_\infty \le \textstag$}{
        \Return $(\mathsf{NoiseFloor}, \mathrm{True}, \mathrm{True}, I_{\mathrm{up}})$\;
    }
    \Return $(\mathsf{Xtol}, \mathrm{False}, \mathrm{True}, I_{\mathrm{up}})$\;
}
\If{$i = N_{\mathrm{N}}$}{
    \If{$\|\hat{\boldsymbol{b}}\|_\infty \le \textstag$}{
        \Return $(\mathsf{NoiseFloor}, \mathrm{True}, \mathrm{True}, I_{\mathrm{up}})$\;
    }
    \Return $(\mathsf{IterationCap}, \mathrm{False}, \mathrm{True}, I_{\mathrm{up}})$\;
}
\Return $(\mathsf{InProgress}, \mathrm{False}, \mathrm{False}, I_{\mathrm{up}})$\;
\end{algorithm}

\noindent In the implementation $I_{\mathrm{up}}$ is carried in the
\code{inc\_count} field of a mutable \code{SafeIterationState} struct
that the routine updates in place; the box treats it as a returned
value for mathematical clarity.

\paragraph{Why six exit reasons?}  The arbiter distinguishes states
that classical Newton implementations conflate, because the choice of
reason changes what the caller does next.
\begin{itemize}
\item $\mathsf{Ftol}$ and $\mathsf{InitialFtol}$ are unambiguous
successes.
\item $\mathsf{NoiseFloor}$ is a \emph{conditional} success: the
residual sits below $\textstag$, which is far above floating-point
noise but well below physical interest. The caller treats this as
success intentionally (§\ref{sec:falseEquiv}).
\item $\mathsf{LinesearchFailed}$ and $\mathsf{MeritDivergence}$ both
trigger the Picard fallback (Algorithm~\ref{alg:B}).
\item $\mathsf{Xtol}$ flags an iterate that has stopped moving even
though the residual is non-trivial; failure unless the residual has
slipped below $\textstag$ (in which case it is silently relabelled as
$\mathsf{NoiseFloor}$).
\item $\mathsf{IterationCap}$ is the budget-exhausted failure.
\item $\mathsf{LinearSolveNaN}$ aborts the kernel unconditionally.
\end{itemize}

\paragraph{The role of $F_{\mathrm{div}}$ and $N_{\mathrm{inc}}$ together.}
A single merit-expansion event is not a divergence: Newton near a
saddle can take one or two steps that increase the merit before the
quadratic basin captures the iterate. Requiring $N_{\mathrm{inc}}$
\emph{consecutive} expansions, each by at least a factor
$F_{\mathrm{div}}$, separates transient overshooting from genuine
repeated divergence.

% ============================================================
\section{Stage II: Algorithm C (OSGSFractionalRelaxation)}
\label{sec:osgs}

The orthogonal sub-grid scale formulation augments the discrete weak
form with a stabilisation term that depends on the $L^2$ projection
$\boldsymbol{\pi}_h$ of the strong residual. To avoid carrying the
Fr\'echet derivative of $\Pi_h$ in the global Jacobian,
Algorithm~\ref{alg:C} treats the pair $(U_h, \boldsymbol{\pi}_h)$ by
block-coordinate relaxation: freeze $\boldsymbol{\pi}_h$, solve the
inner nonlinear primary system for $U_h$ via the Newton kernel
(Algorithm~\ref{alg:A}), then re-project to get the next
$\boldsymbol{\pi}_h$.

\moduleinterface
    {Algorithm~\ref{alg:O} (Stage~II, OSGS mode only)}
    {Algorithm~\ref{alg:B} (once per outer iteration, with $\boldsymbol{\pi}_h$ frozen; the cascade itself invokes Alg.~\ref{alg:A} up to three times)}
    {Stage~I state $U_h^{\mathrm{ASGS}}$; pre-factored mass solvers $(M_u^{-1}, M_p^{-1})$; optional block-wise Anderson accelerators}
    {Success flag $S$; converged projection pair $(\boldsymbol{\pi}_{h,u}, \pi_{h,p})$}

\begin{algorithm}[H]
\caption{$\mathrm{OSGSFractionalRelaxation}$: staggered fixed point on $(U_h, \boldsymbol{\pi}_h)$.}
\label{alg:C}
\SetAlgoLined
\textbf{Parameters:} $N_{\mathrm{OSGS}}, N_{\mathrm{inner}}, m_{\mathrm{w}}, \tau_{\mathrm{w}}, \textftol, \tau_{\mathrm{OSGS}}, \tau_{\mathrm{proj}}, \textstag, \mathsf{Mode}_{\mathrm{drift}}, \mathsf{Mode}_{\mathrm{stop}}$.\\
\KwIn{$U_h^{\mathrm{ASGS}}$; residual operators $(\mathcal{R}_u, \mathcal{R}_p)$; mass-matrix solvers; optional accelerators.}
\KwOut{$(S, \boldsymbol{\pi}_{h,u}, \pi_{h,p})$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}
$U_h^{0} \gets U_h^{\mathrm{ASGS}}$;\quad
$\boldsymbol{\pi}_{h,u}^{0} \gets 0$;\quad $\pi_{h,p}^{0} \gets 0$ \tcp*{see §\ref{sec:osgs-bootstrap}}
\For{$m = 1, \dots, N_{\mathrm{OSGS}}$}{
    set effective inner tolerance $\ftolinner{m} \gets (m \le m_{\mathrm{w}}) \,?\, \max(\textftol, \tau_{\mathrm{w}}) : \textftol$\;
    fix $\boldsymbol{\pi}_h \gets \boldsymbol{\pi}_h^{m-1}$;\quad
    $\boldsymbol{x}_{\mathrm{prev}} \gets U_h^{m-1}$ \tcp*{restore vector for the inner cascade}
    $(S_m^{\mathrm{inner}}, U_h^{m}) \gets \mathrm{RobustNonlinearCascade}(U_h^{m-1},\ \boldsymbol{x}_{\mathrm{prev}},\ \mathrm{op}_{\mathrm{N}}^{\mathrm{OSGS}}(\boldsymbol{\pi}_h),\ \mathrm{op}_{\mathrm{P}}^{\mathrm{OSGS}}(\boldsymbol{\pi}_h))$ \tcp*{Alg.~\ref{alg:B}; inner cap $N_{\mathrm{inner}}$, tolerance $\ftolinner{m}$}
    \If{$S_m^{\mathrm{inner}} = \mathrm{False}$ \textnormal{\textbf{and}} the cascade aborted with a structural failure}{
        \Return $(\mathrm{False}, \boldsymbol{\pi}_{h,u}^{m-1}, \pi_{h,p}^{m-1})$ \tcp*{see §\ref{sec:osgs-inner-failure}}
    }

    $d_{\mathrm{state}}^m \gets \mathrm{StateDrift}(U_h^m, U_h^{m-1};\, \mathsf{Mode}_{\mathrm{drift}})$ \tcp*{see §\ref{sec:drift-modes}}

    re-project: $\boldsymbol{\pi}_{h,u}^{m} \gets \Pi_{V_{\mathrm{free}}}(\mathcal{R}_u(U_h^m))$;\quad
    $\pi_{h,p}^{m} \gets \Pi_{Q_{\mathrm{free}}}(\mathcal{R}_p(U_h^m))$\;

    \If{accelerator present}{
        $\boldsymbol{\pi}_{h,u}^{m} \gets \mathrm{Anderson}_u(\boldsymbol{\pi}_{h,u}^{m-1}, \boldsymbol{\pi}_{h,u}^{m})$ \tcp*{safety fallback (§\ref{sec:anderson}) handled internally}
        $\pi_{h,p}^{m} \gets \mathrm{Anderson}_p(\pi_{h,p}^{m-1}, \pi_{h,p}^{m})$ \tcp*{idem}
    }

    $d_\pi^m \gets \max\!\bigl(
        \|\Delta\boldsymbol{\pi}_{h,u}^m\|_{M_u},\
        \|\Delta\pi_{h,p}^m\|_{M_p}\bigr)$ \tcp*{see §\ref{sec:proj-drift}}

    $\tau_{\mathrm{OSGS}}^{\mathrm{eff}} \gets \max(\tau_{\mathrm{OSGS}}, \textstag)$;\quad
    $\tau_{\mathrm{proj}}^{\mathrm{eff}} \gets \max(\tau_{\mathrm{proj}}, \textstag)$\;
    decide $S_{\mathrm{conv}}$ from $\mathsf{Mode}_{\mathrm{stop}}$, $d_{\mathrm{state}}^m$, $d_\pi^m$ (see §\ref{sec:stopping-mode})\;
    \If{$\mathsf{Mode}_{\mathrm{stop}} = \code{state\_drift}$ \textnormal{\textbf{and}} $m \ge 2$ \textnormal{\textbf{and}} $\|\boldsymbol{X}_h^m - \boldsymbol{X}_h^{m-1}\|_{\ell^\infty} \le \textstag$}{
        $S_{\mathrm{conv}} \gets \mathrm{True}$ \tcp*{absolute noise-floor short circuit; gated to \code{state\_drift} only (§\ref{sec:stopping-mode})}
    }
    \If{$S_{\mathrm{conv}}$}{
        \extline{\If{$\code{mms\_cfg.enabled}$ \textnormal{\textbf{and}} first $S_{\mathrm{conv}}$ event}{
            initialise plateau-history $\mathcal{H}$ at $U_h^m$;\quad $I_{\mathrm{pass}} \gets 0$\;
            \textbf{continue} \tcp*{do not exit yet; collect a plateau sample}
        }
        \If{$\code{mms\_cfg.enabled}$}{
            $(S_{\mathrm{plat}}, I_{\mathrm{pass}}) \gets \mathrm{VerifyMMSPlateau}(U_h^m, U_{\mathrm{ex}}, \mathcal{H}, I_{\mathrm{pass}})$ \tcp*{Alg.~\ref{alg:D}}
            \If{$\neg S_{\mathrm{plat}}$}{ \textbf{continue}\; }
        }}
        \Return $(\mathrm{True}, \boldsymbol{\pi}_{h,u}^m, \pi_{h,p}^m)$\;
    }
}
\Return $(\mathrm{False}, \boldsymbol{\pi}_{h,u}^{N_{\mathrm{OSGS}}}, \pi_{h,p}^{N_{\mathrm{OSGS}}})$\;
\end{algorithm}

\subsection{State drift: \texorpdfstring{$\ell^\infty$ vs continuous $L^2$}{Linf vs L2}}
\label{sec:drift-modes}

The state drift between consecutive outer iterations
\begin{equation}
    d_{\mathrm{state}}^m
    = \begin{cases}
        \max\!\bigl(
            \|\boldsymbol{u}_h^m - \boldsymbol{u}_h^{m-1}\|_{L^2(\Omega)},\
            \|p_h^m - p_h^{m-1}\|_{L^2(\Omega)} \bigr)
            & \mathsf{Mode}_{\mathrm{drift}} = \code{L2\_mass}, \\[1mm]
        \|\boldsymbol{X}_h^m - \boldsymbol{X}_h^{m-1}\|_{\ell^\infty}
            & \mathsf{Mode}_{\mathrm{drift}} = \code{Linf},
    \end{cases}
\end{equation}
admits two metrics with subtly different meanings. The \code{Linf}
mode is the raw vector-space $\ell^\infty$ norm of the concatenated
free-DOF array; it is cheap but pathological in mixed-physics systems
because it lets the field with larger numerical magnitude dominate.
The \code{L2\_mass} mode computes two independent continuous
$L^2(\Omega)$ norms,
$\sqrt{\int_\Omega (\boldsymbol{u}_h^m - \boldsymbol{u}_h^{m-1})\cdot(\boldsymbol{u}_h^m - \boldsymbol{u}_h^{m-1})\,\mathrm{d}\Omega}$
and analogously for the pressure block, evaluated by Gridap quadrature
on the same measure $\mathrm{d}\Omega$ used by the formulation, and
takes their maximum. The distinction matters: \code{L2\_mass} is the
continuous functional norm of the difference FE function, whereas
\code{Linf} is the discrete maximum coefficient deviation, which is
basis-dependent and mixes physical units across blocks. Only
\code{L2\_mass} measures velocity and pressure in their own natural
functional norms.

\subsection{Projection drift in the mass-weighted inner product}
\label{sec:proj-drift}

The projection drift
\begin{equation}
    d_\pi^m
    = \max\!\Bigl(
        \sqrt{(\Delta\boldsymbol{\pi}_{h,u}^m)^\top M_u (\Delta\boldsymbol{\pi}_{h,u}^m)},\
        \sqrt{(\Delta\pi_{h,p}^m)^\top M_p (\Delta\pi_{h,p}^m)} \Bigr)
\end{equation}
is the natural distance in the inner product that defines the $L^2$
projection itself. The Euclidean DOF norm would conflate this with
inverse-mass scaling and report misleading drift values on graded
meshes.

\subsection{Choosing the stopping mode}
\label{sec:stopping-mode}

The user selects $\mathsf{Mode}_{\mathrm{stop}}$ between three options:
\begin{itemize}
\item \code{state\_drift}: terminate when $d_{\mathrm{state}}^m \le \tau_{\mathrm{OSGS}}^{\mathrm{eff}}$.
\item \code{projection\_drift}: terminate when $d_\pi^m \le \tau_{\mathrm{proj}}^{\mathrm{eff}}$.
\item \code{both} (default): require both predicates.
\end{itemize}
\code{both} is the strictest and corresponds to the mathematical
definition of the staggered fixed point. The single-criterion modes
sacrifice one of the two conditions in exchange for fewer iterations
and are appropriate only when the user knows that the slacker
condition is already satisfied. Only under \code{state\_drift} does
an additional absolute short-circuit fire when the raw $\ell^\infty$
DOF drift drops below $\textstag$ for $m \ge 2$; this protects against
limit-cycle oscillations that never cleanly satisfy
$\tau_{\mathrm{OSGS}}$. The short-circuit is \emph{deliberately
disabled} in \code{projection\_drift} and \code{both} modes: when
projection drift is part of the binding criterion, $\boldsymbol{\pi}_h$
may legitimately continue moving even while the primary state appears
stagnant, and short-circuiting on $\ell^\infty$ alone could return a
non-fixed-point on the pair $(U_h, \boldsymbol{\pi}_h)$.

\subsection{Bootstrap, mass-matrix pre-factoring, and warmup}
\label{sec:osgs-bootstrap}

\paragraph{Zero bootstrap of $\boldsymbol{\pi}_h^0$.}  The first sub-grid
projection is initialised to zero, $\boldsymbol{\pi}_h^0 \equiv 0$,
\emph{not} to the $L^2$ projection of the strong residual at
$U_h^{\mathrm{ASGS}}$. The choice is deliberate. Solving the OSGS
inner problem at $m = 1$ with $\boldsymbol{\pi}_h \equiv 0$ is, by
definition, exactly the ASGS problem, and Stage~I has already converged
that system to within $\textftol$. The very first inner solve therefore
starts and ends at $U_h^{\mathrm{ASGS}}$, and the first non-trivial
projection $\boldsymbol{\pi}_h^1$ is produced naturally by the
$L^2$-projection step at the bottom of the outer-loop body. Pre-seeding
$\boldsymbol{\pi}_h^0$ with $\Pi(\mathcal{R}(U_h^{\mathrm{ASGS}}))$
would instead drive the $m = 1$ inner solve away from $U_h^{\mathrm{ASGS}}$
to satisfy a stabilisation term whose projection has not yet been
relaxed by an outer iteration---an ``off-distribution'' bootstrap that
historically required the warmup phase (see below) to unwind.

\paragraph{Mass-matrix pre-factoring.}  Each projection is the solution
of an SPD linear system with $M_u$ or $M_p$. Because the projection
target spaces never change between outer iterations, the matrices and
their LU factorisations are assembled once (in Algorithm~\ref{alg:O})
and reused: every projection is a constant-cost forward/back
substitution.

\paragraph{Warmup phase.}  The warmup phase ($m \le m_{\mathrm{w}}$)
relaxes the inner-cascade tolerance to $\max(\textftol, \tau_{\mathrm{w}})$.
Even with the zero bootstrap of $\boldsymbol{\pi}_h^0$, the projection
$\boldsymbol{\pi}_h^{m}$ for small $m$ is still moving rapidly between
outer iterations; driving the inner solve to a tight $\textftol$ in this
regime would polish the iterate against a projection that gets replaced
immediately. Relaxing the inner tolerance during the warmup trades a
small amount of accuracy in the early iterates for a meaningful cut in
inner-solver budget.

\subsection{Projection policy and the constant-$\sigma$ trim}
\label{sec:proj-policy}

The projection step
$\boldsymbol{\pi}_h^{m} \gets \Pi(\mathcal{R}(U_h^m))$ does not always
project the full strong residual; the policy is configurable, and the
choice is determined jointly by the discrete formulation and the
reaction model.

The default policy projects the full residual. When the reaction
tensor $\sigma(\alpha, \boldsymbol{u})$ is a constant
($\code{reaction\_model = "Constant\_Sigma"}$ in
\code{stabilization\_config}, paired with the
\code{ProjectResidualWithoutReactionWhenConstantSigma} policy), the
reaction term $\sigma \boldsymbol{u}_h$ is omitted from the residual
before projection. The justification is exactness: when $\sigma$ is a
spatial constant, $\sigma \boldsymbol{u}_h$ lies entirely in the FE
velocity space $\boldsymbol{V}_h$ (which is closed under multiplication
by a scalar), so its $L^2$ projection onto $V_{\mathrm{free}}$ is the
function itself; subtracting it again in the orthogonal complement
$I - \Pi_h$ gives exactly zero. Including the reaction term in the
projection step would amount to projecting and re-subtracting an FE
function---a no-op in exact arithmetic, but one that introduces avoidable
floating-point noise. Removing it from the projected residual eliminates
that noise at no cost. This corresponds to Section~4.4 of the parent
paper.

For non-constant $\sigma$ (e.g.\ the Forchheimer--Ergun law), the
reaction term is \emph{not} an FE function and its projection is
non-trivial; the trim does not apply and the policy reverts to the
default of projecting the full residual.

\subsection{Block-wise Anderson acceleration}
\label{sec:anderson}

For stiff fixed points on $\boldsymbol{\pi}_h$, Anderson acceleration
extrapolates a new iterate from the last $m$ fixed-point residuals by
solving a small mass-weighted least-squares problem
\begin{equation}
    \min_\gamma\ \|(\Delta F)\gamma - f_k\|_M^2.
\end{equation}
The accelerator operates independently on $\boldsymbol{\pi}_{h,u}$ and
$\pi_{h,p}$, each with its own mass matrix; combining the blocks would
force a cross-physical mixing that the OSGS formulation goes out of
its way to avoid. The least-squares system uses the same $L^2$ inner
product that defines the underlying projection. A safety fallback
discards the extrapolation and reverts to a relaxed Picard step when
the extrapolated state lies more than a fixed multiple of the
unextrapolated residual away from the raw fixed-point candidate.

\subsection{Failure of the inner cascade}
\label{sec:osgs-inner-failure}

The inner solve inside Algorithm~\ref{alg:C} is the same
$\mathrm{RobustNonlinearCascade}$ used in Stage~I, supplied with
OSGS operators (so the Jacobian depends on the frozen projection
$\boldsymbol{\pi}_h^{m-1}$) and with the iteration cap reduced to
$N_{\mathrm{inner}}$ on each inner $\mathrm{ExactNewtonPipeline}$ call.
The restore vector is $\boldsymbol{x}_{\mathrm{prev}} = U_h^{m-1}$,
i.e.\ the iterate at the start of the current outer iteration; this
differs from Stage~I, which restores to the global initial guess
$\boldsymbol{x}^0_{\mathrm{backup}}$.

Three failure modes are propagated as structural divergence and abort
the outer loop:
\begin{itemize}
\item The first exact-Newton attempt produces a non-finite iterate
\emph{and} the Picard attempt subsequently produces a non-finite
iterate as well (the cascade then returns
$\boldsymbol{x}_{\mathrm{restore}}$ with $S = \mathrm{False}$).
\item The Picard attempt itself raises an unrecoverable exception
(distinct from the benign ``reached maximum iterations'' exception
that is reinterpreted as a completed Picard smoothing pass).
\item The terminal Newton retry, after a Picard smoothing pass,
returns with one of $\mathsf{LinesearchFailed}$,
$\mathsf{MeritDivergence}$, or $\mathsf{LinearSolveNaN}$.
\end{itemize}
All other exit configurations (notably $\mathsf{IterationCap}$ from
exhausting $N_{\mathrm{inner}}$ inside one of the cascade's three
sub-calls) are treated as progress: the next outer iteration simply
re-engages with the updated projection.

% ============================================================
\section{Convergence-Analysis Notes}
\label{sec:falseEquiv}

\subsection{The false-equivalency failsafe}

The single most important invariant of Algorithm~\ref{alg:A3} is the
asymmetry between two exit reasons that, at first glance, look
interchangeable:
\begin{itemize}
\item $\mathsf{IterationCap}$ with $\|\boldsymbol{b}\|_\infty \gg \textstag$:
the iteration ran out of budget while still far from any noise floor.
\item $\mathsf{NoiseFloor}$ with $\|\boldsymbol{b}\|_\infty \le \textstag$:
the iteration may have stopped for any reason, but the residual is below
the noise floor we set for ``effectively converged''.
\end{itemize}
The orchestration treats only the second as success. Coercing the
first into a converged state---``it ran for many iterations and
stopped; close enough''---is forbidden, because doing so would feed an
unresolved iterate into Stage~II of the cascade. In OSGS in particular,
an unresolved $U_h$ produces a stale $\boldsymbol{\pi}_h$ whose
subsequent fixed-point iterations cannot recover.

\subsection{Why the OSGS stopping mode matters}

The default \code{both} predicate corresponds to the mathematical
definition of the staggered fixed point: $(U_h, \boldsymbol{\pi}_h)$
is the converged pair when neither moves between outer iterations. The
relaxed modes are appropriate only when the user knows that the
slacker condition is already satisfied; using a single-criterion mode
without that knowledge introduces the false-convergence trap that the
noise-floor invariant forbids.

% ============================================================
\appendix
\section{Code--Notation Correspondence}
\label{sec:appendix-codes}

Table~\ref{tab:reasons} maps every exit-reason label used by the
algorithm boxes to the string literal returned by the implementation.

\begin{table}[h]
\centering
\caption{Exit reasons in Algorithm~\ref{alg:A3} and their code-side strings.}
\label{tab:reasons}
\small
\begin{tabularx}{\textwidth}{@{} l l X @{}}
\toprule
\textbf{Box label} & \textbf{Code string} & \textbf{Meaning} \\
\midrule
$\mathsf{Ftol}$               & \code{"ftol\_reached"}                 & $\|\boldsymbol{b}\|_\infty \le \textftol$ after a successful line search. \\
$\mathsf{InitialFtol}$        & \code{"initial\_ftol"}                  & Initial guess already satisfies $\textftol$; no Newton step taken. \\
$\mathsf{NoiseFloor}$         & \code{"stagnation\_noise\_floor\_reached"} & $\|\boldsymbol{b}\|_\infty \le \textstag$ on a non-converged exit. \\
$\mathsf{LinesearchFailed}$   & \code{"linesearch\_failed"}             & Line-search step shrank to $\alpha \le \alpha_{\min}$ without descent. \\
$\mathsf{MeritDivergence}$    & \code{"merit\_divergence\_escaped"}     & Merit function expanded by $\ge F_{\mathrm{div}}$ for $N_{\mathrm{inc}}$ consecutive steps. \\
$\mathsf{Xtol}$               & \code{"xtol\_stagnation"}               & Step norm $L \le \textxtol$ with residual still above $\textstag$. \\
$\mathsf{IterationCap}$       & \code{"max\_iters\_stagnation"}         & Iteration budget exhausted with residual still above $\textstag$. \\
$\mathsf{LinearSolveNaN}$     & \code{"linear\_solve\_nan"}             & Linear-solve exception or NaN in $\delta\boldsymbol{x}$. \\
$\mathsf{InProgress}$         & \code{""}                                & Sentinel for ``no break this iteration''. \\
\bottomrule
\end{tabularx}
\end{table}

% ============================================================
\part{Optional Extensions}
\label{part:ext}
% ============================================================

\begin{tcolorbox}[colback=extcolor!5!white, colframe=extcolor, boxrule=0.5pt, arc=2pt, left=6pt, right=6pt, top=4pt, bottom=4pt]
\small
\extline{The remainder of this document describes \emph{optional extensions}
to the core algorithm of Part~\ref{part:core}. These extensions are not
part of a generic \code{run\_simulation} call: each is activated by a
specific configuration setting or a specific calling context (a
parameter sweep, a verification harness, etc.). Where an extension
augments a core algorithm box, the additional lines are shown in this
colour. Removing every \extline{coloured} line from this part yields
the core algorithm of Part~\ref{part:core}, byte-for-byte.}
\end{tcolorbox}

\section{Extension Hierarchy}

Two extensions are documented. The first
(§\ref{sec:mms}) adds an MMS plateau verifier (Algorithm~\ref{alg:D})
that hooks into the orchestrator (Algorithm~\ref{alg:O}) and the OSGS
loop (Algorithm~\ref{alg:C}) by inserting coloured \extline{\extline{}}
lines directly in those algorithm boxes (§\ref{sec:mms-hooks}). The
second (§\ref{sec:harness}) is a set of parameter substitutions applied
by the sweep harness; it does not introduce new algorithm boxes.
Figure~\ref{fig:ext-hierarchy} overlays the extension onto the core
tree of Figure~\ref{fig:hierarchy}.

\begin{figure}[h]
\centering
\begin{tikzpicture}[node distance=14mm and 10mm]

\node[modnode] (O)
    {\textbf{Algorithm~\ref{alg:O}}\\$\mathrm{Orchestration}$};

\node[modnode, below left=14mm and 14mm of O] (B)
    {\textbf{Algorithm~\ref{alg:B}}\\$\mathrm{PicardFallback}$};
\node[modnode, below right=14mm and 14mm of O] (C)
    {\textbf{Algorithm~\ref{alg:C}}\\$\mathrm{OSGS}$};

\node[modnode, below=18mm of O] (A)
    {\textbf{Algorithm~\ref{alg:A}}\\$\mathrm{NewtonKernel}$};

\node[extnode, right=22mm of O] (D)
    {\extline{\textbf{Algorithm~\ref{alg:D}}}\\\extline{$\mathrm{VerifyMMSPlateau}$}\\
     \extline{\footnotesize plateau gate}};

\draw[edge] (O.south) -- (B.north);
\draw[edge] (O.south) -- (C.north);
\draw[edge] (B.south) -- (A.north west);
\draw[edge] (C.south) -- (A.north east);

\draw[->, extcolor, thick, dashed]
    (O.east) to[bend left=15]
    node[edgelbl, midway, fill=white, text=extcolor]
    {\extline{ASGS extension cycles (inline in Alg.~\ref{alg:O})}} (D.west);
\draw[->, extcolor, thick, dashed]
    (C.east) to[bend right=10]
    node[edgelbl, midway, fill=white, text=extcolor]
    {\extline{post-convergence gate (inline in Alg.~\ref{alg:C})}} (D.south west);

\end{tikzpicture}
\caption{Extension hierarchy. The MMS plateau verifier
(\extline{Algorithm~\ref{alg:D}}, blue) is hooked directly into the
core orchestrator (Algorithm~\ref{alg:O}) and the core OSGS loop
(Algorithm~\ref{alg:C}) by inline \extline{\textbackslash extline\{\}}
blue lines, displayed within the respective core algorithm boxes
(§\ref{sec:mms-hooks}). Solid arrows reproduce the core hierarchy of
Figure~\ref{fig:hierarchy}; \extline{dashed coloured arrows} mark the
extension hooks.}
\label{fig:ext-hierarchy}
\end{figure}

\section{Method of Manufactured Solutions Plateau Verifier}
\label{sec:mms}

\subsection{When this extension fires}

This extension is active whenever the solver is supplied with a
configuration \code{mms\_cfg} whose \code{enabled} flag is true. The
\code{mms\_cfg} object carries an analytical reference solution
$U_{\mathrm{exact}} = (\boldsymbol{u}_{\mathrm{ex}}, p_{\mathrm{ex}})$
(via an \code{oracle} callback) and a small set of plateau-test
parameters. When \code{mms\_cfg} is absent or disabled, the core
orchestration runs unchanged.

\subsection{Numerical versus discretisation convergence}

The inner Newton stops when $\|\boldsymbol{b}\|_\infty \le \textftol$.
That is convergence in the \emph{algebraic} sense: the iterate is
close to the root of the discrete system. It says nothing about
whether the discrete system is close to the true PDE solution. The
Method of Manufactured Solutions (MMS) provides exactly that second
test: it injects an exact analytical
$(\boldsymbol{u}_{\mathrm{ex}}, p_{\mathrm{ex}})$, recomputes the
appropriate source terms, and measures $\|U_h - U_{\mathrm{ex}}\|$ in
physical norms. The expected discretisation behaviour is $O(h^{k+1})$
in the $L^2$ norms (one order weaker in the $H^1$-semi norm).

For a single mesh, plateau verification asks a different question:
once the iterate is close enough to the root that further iterations
do not move the FE error, the FE error has \emph{plateaued} at its
discretisation floor.

\subsection{Algorithm D and its interface}

\extline{Algorithm~\ref{alg:D} is itself a leaf---a single side-by-side
comparison of consecutive FE errors against the exact solution. It
does not advance the iterate.}

\begin{tcolorbox}[colback=extcolor!5!white, colframe=extcolor, boxrule=0.4pt, arc=2pt, left=8pt, right=8pt, top=4pt, bottom=4pt, width=\textwidth, before skip=4pt, after skip=2pt]
\footnotesize
\extline{%
\begin{tabular}{@{}r@{\quad}p{0.82\textwidth}@{}}
\textit{Called by:} & Inline blue hook in Algorithm~\ref{alg:O} (per extension cycle in the ASGS path); inline blue hook in Algorithm~\ref{alg:C} (per OSGS staggered iteration once $S_{\mathrm{conv}}$ has fired). See §\ref{sec:mms-hooks}. \\
\textit{Calls:}     & (none; pure procedure) \\
\textit{Inputs:}    & Current iterate $U_h^m$; exact reference $U_{\mathrm{ex}}$; running error history $\mathcal{H}$; pass counter $I_{\mathrm{pass}}$ \\
\textit{Outputs:}   & Plateau flag $S_{\mathrm{plat}}$; updated counter $I_{\mathrm{pass}}'$ \\
\end{tabular}}
\end{tcolorbox}

\begin{algorithm}[H]
\caption{\extline{$\mathrm{VerifyMMSPlateau}$: relative-change plateau test (optional extension).}}
\label{alg:D}
\SetAlgoLined
\textcolor{extcolor}{\textbf{Parameters:} $N_{\mathrm{plat}}, \tau_{\mathrm{err}}, \varepsilon_{u,L^2}, \varepsilon_{u,H^1}, \varepsilon_{p,L^2}$.\\
\KwIn{$U_h^m$, $U_{\mathrm{ex}}$, $\mathcal{H}$, $I_{\mathrm{pass}}$.}
\KwOut{$(S_{\mathrm{plat}}, I_{\mathrm{pass}}')$.}
\vspace{0.2cm}\hrule\vspace{0.2cm}
compute $E_{u,L^2}^m$, $E_{u,H^1}^m$, $E_{p,L^2}^m$\;
append to $\mathcal{H}$; let $E^{m-1}_\bullet$ be the previous entry\;
$r_\bullet \gets |E_\bullet^m - E_\bullet^{m-1}| / \max(E_\bullet^m, E_\bullet^{m-1}, \varepsilon_\bullet)$ for each norm\;
$r_{\max} \gets \max_\bullet r_\bullet$\;
\If{$r_{\max} \le \tau_{\mathrm{err}}$}{
    $I_{\mathrm{pass}} \gets I_{\mathrm{pass}} + 1$\;
}\Else{
    $I_{\mathrm{pass}} \gets 0$\;
}
\If{$I_{\mathrm{pass}} \ge N_{\mathrm{plat}}$}{
    \Return $(\mathrm{True}, I_{\mathrm{pass}})$ \tcp*{plateau satisfied}
}
\Return $(\mathrm{False}, I_{\mathrm{pass}})$ \tcp*{continue extending}}
\end{algorithm}

\extline{The verifier tracks three error quantities, not two: $L^2$
of velocity, $L^2$ of pressure, and the $H^1$-semi-norm of velocity.
The third matters because the $L^2$ errors can plateau while the
velocity gradient---which is what the convective and viscous terms
actually feel---still drifts. The plateau is declared only after
$N_{\mathrm{plat}}$ \emph{consecutive} passes, and the counter resets
to zero on the first failure.}

\subsection{How Algorithm D hooks into the core}
\label{sec:mms-hooks}

\extline{The MMS plateau verifier is wired into the core algorithms by
two short blocks of conditional logic, displayed inline in
Algorithm~\ref{alg:O} (Stage~II dispatch) and Algorithm~\ref{alg:C}
(post-$S_{\mathrm{conv}}$ branch) and typeset in this colour. The
core path is recovered byte-for-byte by deleting every coloured line.}

\extline{In the ASGS branch of Algorithm~\ref{alg:O}, when
\code{mms\_cfg.enabled} is true, the orchestrator enters a bounded
plateau loop: up to $K_{\mathrm{extra}}$ extra single-Newton-step cycles
are run on top of the converged $U_h^{\mathrm{ASGS}}$, each followed
by a call to Algorithm~\ref{alg:D}. The loop exits as soon as the
plateau condition is satisfied or the budget is exhausted.}

\extline{In the OSGS branch, the augmentation has two parts.
The orchestrator inflates the outer-loop budget
($N_{\mathrm{OSGS}}^{\mathrm{eff}} \gets N_{\mathrm{OSGS}} + K_{\mathrm{extra}}$)
to leave room for the plateau samples. Inside Algorithm~\ref{alg:C},
the $S_{\mathrm{conv}}$ branch is no longer terminal: the first
$S_{\mathrm{conv}}$ event is intercepted to seed the plateau history
and is followed by a \texttt{continue}, so the outer loop runs one
more iteration to collect a sample; on subsequent $S_{\mathrm{conv}}$
events, Algorithm~\ref{alg:D} is called and the loop either exits
(plateau satisfied) or continues (still extending).}

\subsection{Parameters of this extension}

\begin{table}[h]
\centering
\caption{\extline{Numerical parameters of the MMS plateau verifier (extension-only).}}
\label{tab:mms-params}
\footnotesize
\textcolor{extcolor}{%
\begin{tabularx}{\textwidth}{@{} l l X @{}}
\toprule
\textbf{Symbol} & \textbf{Field in code} & \textbf{Role / effect} \\
\midrule
$N_{\mathrm{plat}}$ & \code{mms.require\_consecutive\_passes} & Consecutive passes of $r_{\max} \le \tau_{\mathrm{err}}$ required before declaring a plateau. \\
$\tau_{\mathrm{err}}$ & \code{mms.tau\_err} & Relative-change tolerance against which $r_{\max}$ is tested. \\
$\varepsilon_{u,L^2}, \varepsilon_{u,H^1}, \varepsilon_{p,L^2}$ & \code{mms.eps\_u\_l2,\_u\_h1,\_p\_l2} & Per-norm floors guarding the denominator in $r_{\max}$. \\
$K_{\mathrm{extra}}$ & \code{mms.max\_extra\_cycles} & Cap on extension cycles (ASGS path) or staggered iterations added to the OSGS budget (OSGS path) for plateau verification. \\
\bottomrule
\end{tabularx}}
\end{table}

\section{Parameter-Sweep Harness Adaptations}
\label{sec:harness}

\subsection{When this extension fires}

\extline{These adaptations are active only when the solver is driven by
the project's parameter-sweep harness, which wraps the core
\code{run\_simulation} pipeline in a triple loop over $\mathrm{Re}$,
$\mathrm{Da}$, and the mesh-resolution parameter. In an ad-hoc
\code{run\_simulation} call, none of these adaptations fires; the core
algorithm of Part~\ref{part:core} uses the static parameters of
Table~\ref{tab:params} verbatim. The harness changes parameter
\emph{values} only; the algorithm boxes themselves are unchanged, so
no $\mathrm{O}'$ or $\mathrm{B}'$ variants are needed.}

\subsection{Adaptive \texorpdfstring{$\textftol$}{ftol}}
\label{sec:harness-ftol}

\extline{Two error scales matter in an FE method. The \emph{algebraic
error} is the distance between the iterate and the discrete root,
governed by $\textftol$. The \emph{discretisation error} is the
distance between the discrete root and the true PDE solution, which
behaves as $O(h^{k_v+1})$. The harness therefore re-derives $\textftol$
per case:}
\begin{equation}
    \label{eq:dynamic-ftol}
    \textcolor{extcolor}{\textftol^{\,\mathrm{dyn}}
    = \max\!\Bigl(\textftol^{\,\mathrm{static}},\
        \min(C_{\mathrm{ceil}},\ C_{\mathrm{sf}} \cdot h^{k_v+1}) \Bigr).}
\end{equation}
\extline{The right-most term ties the tolerance to the discretisation
scale via a spatial safety factor $C_{\mathrm{sf}}$; the ceiling
$C_{\mathrm{ceil}}$ prevents trivially loose tolerances on coarse
meshes; the outer $\max$ prevents the formula from descending below
the user-supplied $\textftol^{\,\mathrm{static}}$.}

\subsection{Adaptive noise floor}
\label{sec:harness-noise}

\extline{Linear-system conditioning grows as $\mathcal{O}(h^{-2})$ for
elliptic operators and is amplified by convective and reactive terms.
A static $\textstag$ appropriate on a coarse mesh becomes far above
the genuine round-off floor on a fine mesh. The harness scales}
\begin{equation}
    \textcolor{extcolor}{\textstag^{\,\mathrm{dyn}}
    = \min\!\Bigl(\textstag,\ \max(n^{\min},\ n^{\mathrm{base}} \cdot N^2 \cdot \max(1, \mathrm{Re}))\Bigr),}
\end{equation}
\extline{with the additional clamp
$\textstag^{\,\mathrm{dyn}} \gets \max(\textstag^{\,\mathrm{dyn}}, n^{\mathrm{sf}}\textftol^{\,\mathrm{dyn}})$
preserving the strict ordering
$\textstag^{\,\mathrm{dyn}} > \textftol^{\,\mathrm{dyn}}$.}

\subsection{Per-case Picard widening}
\label{sec:harness-picard}

\extline{In a $(\mathrm{Re}, \mathrm{Da}, h)$ sweep, the harness widens
the Picard iteration budget when a stiffness threshold is exceeded:}
\begin{equation}
    \textcolor{extcolor}{\begin{aligned}
    N_{\mathrm{P}}^{\,\mathrm{local}} &\gets N_{\mathrm{P}}, \\
    \text{if } \mathrm{Re} \ge \mathrm{Re}^{\mathrm{th}}: \quad
    N_{\mathrm{P}}^{\,\mathrm{local}} &\gets \max(N_{\mathrm{P}}^{\,\mathrm{local}}, N_{\mathrm{P}}^{\mathrm{Re}}), \\
    \text{if } \mathrm{Da} \ge \mathrm{Da}^{\mathrm{th}}: \quad
    N_{\mathrm{P}}^{\,\mathrm{local}} &\gets \max(N_{\mathrm{P}}^{\,\mathrm{local}}, N_{\mathrm{P}}^{\mathrm{Da}}).
    \end{aligned}}
\end{equation}

\subsection{Parameters of this extension}

\begin{table}[h]
\centering
\caption{\extline{Numerical parameters of the parameter-sweep harness adaptations (extension-only).}}
\label{tab:harness-params}
\footnotesize
\textcolor{extcolor}{%
\begin{tabularx}{\textwidth}{@{} l l X @{}}
\toprule
\textbf{Symbol} & \textbf{Field in code} & \textbf{Role / effect} \\
\midrule
$C_{\mathrm{ceil}}$, $C_{\mathrm{sf}}$ & \code{sol.dynamic\_ftol\_ceiling}, \code{\_spatial\_safety\_factor} & Clamps for the adaptive $\textftol$ rule, \S\ref{sec:harness-ftol}. \\
$n^{\mathrm{base}}, n^{\min}, n^{\mathrm{sf}}$ & \code{sol.condition\_noise\_floor\_*} & Adaptive noise-floor controls, \S\ref{sec:harness-noise}. \\
$\mathrm{Re}^{\mathrm{th}}, N_{\mathrm{P}}^{\mathrm{Re}}$ & \code{sol.dynamic\_picard\_re\_*} & Per-case Picard widening trigger and budget for high Reynolds. \\
$\mathrm{Da}^{\mathrm{th}}, N_{\mathrm{P}}^{\mathrm{Da}}$ & \code{sol.dynamic\_picard\_da\_*} & Per-case Picard widening trigger and budget for high Damk\"ohler. \\
$\mathrm{Re}^{\mathrm{th,N}}, N_{\mathrm{N}}^{\mathrm{Re}}$ & \code{sol.dynamic\_newton\_re\_*} & Per-case Newton widening trigger and budget for high Reynolds. Mirrors the Picard pair; landed 2026-05-19 (\code{fa8aaec}) to absorb the transient smoothing phase the corrected mode-aware divergence safeguard exposes when Newton has to backtrack through residual-landscape spikes at $\mathrm{Re} \ge \mathrm{Re}^{\mathrm{th,N}}$. \\
\bottomrule
\end{tabularx}}
\end{table}

\end{document}
