\documentclass[11pt]{article}

\usepackage{amsmath, amssymb}
\usepackage{booktabs}
\usepackage{geometry}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{siunitx}
\usepackage{hyperref}
\geometry{margin=2.5cm}

\title{Centered-encoding parametrization for the MMS sweep}
\author{Working note (not yet merged into \texttt{article.tex})}
\date{2026-05-28}

% Match the notation conventions of article.tex (lines 140-141):
\newcommand{\Reyn}{\operatorname{\mathit{Re}}}
\newcommand{\Damk}{\operatorname{\mathit{Da}}}

% Convenient shorthand for inline code identifiers (file paths, function names,
% Julia / config keys). Renders the same as \texttt; lets later edits switch the
% style globally without churn.
\newcommand{\code}[1]{\texttt{#1}}

\begin{document}
\maketitle

\noindent\emph{Status note.} The OSGS staggered loop these diagnostics
reference was deleted in the 2026-06-07 coupled-only leaning; the encoding
conclusions are unaffected.

\section{Motivation}

The stabilized FE solver in this codebase is dimensional. Its inputs are
$(\rho, \mu, \sigma, U, L, \boldsymbol{f})$ in a chosen system of units.
The Method of Manufactured Solutions (MMS) test harness, however, organises
runs by the \emph{dimensionless} regime $(\Reyn, \Damk, \alpha_0)$, and is
free to choose any consistent dimensional encoding of each dimensionless
cell. Here $\alpha_0$ is the infimum of the porosity field $\alpha(x)$, and
$\alpha_\infty = 1$ is its supremum (a convention shared with
\texttt{article.tex}).

Until recently, the harness fixed
\begin{equation}
  U = 1, \qquad L = 1,
\end{equation}
and derived $\nu, \sigma$ from the dimensionless definitions used in
\texttt{article.tex}:
\begin{equation}
  \nu \;=\; \frac{U\,L}{\Reyn} \;=\; \frac{1}{\Reyn},
  \qquad
  \sigma \;=\; \frac{\Damk\,\alpha_\infty\,\nu}{L^2}
                          \;=\; \frac{\Damk\,\alpha_\infty}{\Reyn}.
\end{equation}
For the sweep grid $\Reyn \in \{10^{-6}, 1, 10^{6}\}$,
$\Damk \in \{10^{-6}, 1, 10^{6}\}$, $\alpha_0 \in \{1, 0.5, 0.05\}$
(with $\alpha_\infty = 1$), this encoding produces values such as
$\nu = 10^{6}$ and $\sigma = 10^{12}$ for the cell
$(\Reyn=10^{-6}, \Damk=10^{6}, \alpha_0=1)$. Twelve orders of magnitude
separate the smallest from the largest dimensional coefficient that the
solver sees, \emph{all on the same side of unity}.

While the dimensionless physics is unchanged, this one-sided spread breaks
numerical components of the stabilized solver that are not scale-invariant
once round-off enters:
\begin{itemize}
  \item the OSGS strong-residual $L^2$ projection
        $\pi_h = \Pi(R(\boldsymbol{u}))$ evaluates
        $\sigma\,\boldsymbol{u}_h$ and $\nu \Delta \boldsymbol{u}_h$ at
        near-precision-floor $\boldsymbol{u}_h$ when
        $\boldsymbol{u}_h \sim 1/\sigma$, amplifying floating-point noise
        in $\pi_h$;
  \item the assembled Jacobian's row magnitudes span 12 orders, which the
        direct factorisation absorbs but the Newton merit function and
        line-search tolerances do not;
  \item the MMS forcing oracle, computed by symbolic differentiation of the
        analytic exact solution, mixes $\sigma\,\boldsymbol{u}_{\mathrm{ex}}$
        at magnitude $10^{12}$ with sub-dominant terms at $10^{6}$, losing
        six digits of relative precision in the sub-dominant contributions.
\end{itemize}

\section{The centered encoding}

For a fixed dimensionless cell $(\Reyn, \Damk, \alpha_\infty)$, the
harness has \emph{two} free dimensional parameters $(U, L)$. The freedom
follows from the constraint pair (matching \texttt{article.tex} eq.~3.4):
\begin{align}
  \Reyn  &= \frac{U\,L}{\nu}, &
  \Damk  &= \frac{\sigma\,L^2}{\alpha_\infty\,\nu}.
\end{align}
Eliminating $\nu, \sigma$:
\begin{equation}
  \nu\,\sigma
   \;=\; \frac{\alpha_\infty\,\Damk\,U^{2}}{\Reyn^{2}},
  \qquad
  \frac{\nu}{\sigma}
   \;=\; \frac{L^{2}}{\alpha_\infty\,\Damk}.
\end{equation}
So $U$ controls the geometric mean and $L$ controls the spread.

\paragraph{Choice.}
The \emph{centered encoding} chooses $U$ so that
$\sqrt{\nu\,\sigma} = 1$, leaving $L = 1$ because the manufactured solution
shape $\sin(\pi x)$ is hard-coded for a unit domain (see
\texttt{PaperMMS} in \texttt{src/problems/mms\_paper.jl}):
\begin{equation}
  \boxed{\;\;
    U \;=\; \frac{\Reyn}{\sqrt{\alpha_\infty\,\Damk}},
    \qquad
    L \;=\; 1.
  \;\;}
\end{equation}
The resulting dimensional coefficients are
\begin{equation}
  \nu \;=\; \frac{1}{\sqrt{\alpha_\infty\,\Damk}},
  \qquad
  \sigma \;=\; \sqrt{\alpha_\infty\,\Damk}.
\end{equation}
One checks immediately that
$\nu\,\sigma = 1$ (centering condition) and
$\nu/\sigma = 1/(\alpha_\infty\,\Damk)$ (intrinsic spread).
Note that $\nu$ and $\sigma$ no longer depend on $\Reyn$ — only $U$ absorbs
the $\Reyn$ variation. The spread
$\sigma / \nu = \alpha_\infty\,\Damk$ is \emph{intrinsic} to the cell and
cannot be compressed by any choice of $(U, L)$; centering only shifts it
to be symmetric about unity.

\paragraph{What this does \emph{not} change.}
The dimensionless equation (and therefore the dimensionless solution
$\boldsymbol{u}^*, p^*$) is identical between the default and centered
encodings; this is a strict reparametrization. The discrete velocity
\emph{degrees of freedom} are
$\boldsymbol{u}_h^{\mathrm{node}} \approx U\,\boldsymbol{u}^{*,\mathrm{node}}$,
so they take on different absolute magnitudes between encodings, but their
dimensionless counterparts $\boldsymbol{u}^{*,\mathrm{node}}$ are unchanged.
The $H^1$-seminorm error
$|\boldsymbol{u}_h - \boldsymbol{u}_{\mathrm{ex}}|_{H^1(\Omega)}\,/\,(U/L)$
is regime-invariant (verified empirically: agreement to $<\!1\%$ across
encodings on the easy regime C2).

\section{Per-cell parameter table}

The harness sweeps $\Reyn \in \{10^{-6}, 1, 10^{6}\}$,
$\Damk \in \{10^{-6}, 1, 10^{6}\}$,
$\alpha_0 \in \{1, 0.5, 0.05\}$, with $\alpha_\infty = 1$ in all cells.
Three cells are skipped by config
($(\Reyn, \Damk, \alpha_0) = (10^{6}, \star, 0.05)$, three triples),
leaving 24 that are exercised by the sweep. Table~\ref{tab:per-cell}
reports the dimensional values for both encodings. Because the dimensional
$\sigma$ in the code path
(\texttt{sigma\_c = Da * alpha\_infty * nu\_calc / L\^{}2}) uses
$\alpha_\infty = 1$ throughout, the $\sigma$ columns are independent of
$\alpha_0$ — they are determined purely by $(\Reyn, \Damk)$.

\begin{table}[!htbp]
  \centering
  \footnotesize
  \begin{tabular}{l rrr rrr rrr}
    \toprule
    & \multicolumn{3}{c}{Dimensionless inputs} &
      \multicolumn{3}{c}{Default ($U=L=1$)} &
      \multicolumn{3}{c}{Centered} \\
    \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
    \# & $\Reyn$ & $\Damk$ & $\alpha_0$ & $U$ & $\nu$ & $\sigma$ & $U$ & $\nu$ & $\sigma$ \\
    \midrule
    C1  & $10^{-6}$ & $10^{-6}$ & 1.00 & 1 & $10^{6}$    & $1$         & $10^{-3}$ & $10^{3}$    & $10^{-3}$ \\
    C2  & 1         & $10^{-6}$ & 1.00 & 1 & 1           & $10^{-6}$   & $10^{3}$  & $10^{3}$    & $10^{-3}$ \\
    C3  & $10^{6}$  & $10^{-6}$ & 1.00 & 1 & $10^{-6}$   & $10^{-12}$  & $10^{9}$  & $10^{3}$    & $10^{-3}$ \\
    C4  & $10^{-6}$ & 1         & 1.00 & 1 & $10^{6}$    & $10^{6}$    & $10^{-6}$ & 1           & 1         \\
    C5  & 1         & 1         & 1.00 & 1 & 1           & 1           & 1         & 1           & 1         \\
    C6  & $10^{6}$  & 1         & 1.00 & 1 & $10^{-6}$   & $10^{-6}$   & $10^{6}$  & 1           & 1         \\
    C7  & $10^{-6}$ & $10^{6}$  & 1.00 & 1 & $10^{6}$    & $10^{12}$   & $10^{-9}$ & $10^{-3}$   & $10^{3}$  \\
    C8  & 1         & $10^{6}$  & 1.00 & 1 & 1           & $10^{6}$    & $10^{-3}$ & $10^{-3}$   & $10^{3}$  \\
    C9  & $10^{6}$  & $10^{6}$  & 1.00 & 1 & $10^{-6}$   & 1           & $10^{3}$  & $10^{-3}$   & $10^{3}$  \\
    \midrule
    C10 & $10^{-6}$ & $10^{-6}$ & 0.50 & 1 & $10^{6}$    & $1$         & $10^{-3}$ & $10^{3}$    & $10^{-3}$ \\
    C11 & 1         & $10^{-6}$ & 0.50 & 1 & 1           & $10^{-6}$   & $10^{3}$  & $10^{3}$    & $10^{-3}$ \\
    C12 & $10^{6}$  & $10^{-6}$ & 0.50 & 1 & $10^{-6}$   & $10^{-12}$  & $10^{9}$  & $10^{3}$    & $10^{-3}$ \\
    C13 & $10^{-6}$ & 1         & 0.50 & 1 & $10^{6}$    & $10^{6}$    & $10^{-6}$ & 1           & 1         \\
    C14 & 1         & 1         & 0.50 & 1 & 1           & 1           & 1         & 1           & 1         \\
    C15 & $10^{6}$  & 1         & 0.50 & 1 & $10^{-6}$   & $10^{-6}$   & $10^{6}$  & 1           & 1         \\
    C16 & $10^{-6}$ & $10^{6}$  & 0.50 & 1 & $10^{6}$    & $10^{12}$   & $10^{-9}$ & $10^{-3}$   & $10^{3}$  \\
    C17 & 1         & $10^{6}$  & 0.50 & 1 & 1           & $10^{6}$    & $10^{-3}$ & $10^{-3}$   & $10^{3}$  \\
    C18 & $10^{6}$  & $10^{6}$  & 0.50 & 1 & $10^{-6}$   & 1           & $10^{3}$  & $10^{-3}$   & $10^{3}$  \\
    \midrule
    C19 & $10^{-6}$ & $10^{-6}$ & 0.05 & 1 & $10^{6}$    & $1$         & $10^{-3}$ & $10^{3}$    & $10^{-3}$ \\
    C20 & 1         & $10^{-6}$ & 0.05 & 1 & 1           & $10^{-6}$   & $10^{3}$  & $10^{3}$    & $10^{-3}$ \\
    \multicolumn{10}{l}{\emph{(skipped: $\Reyn = 10^{6}, \Damk = 10^{-6}, \alpha_0 = 0.05$)}} \\
    C21 & $10^{-6}$ & 1         & 0.05 & 1 & $10^{6}$    & $10^{6}$    & $10^{-6}$ & 1           & 1         \\
    C22 & 1         & 1         & 0.05 & 1 & 1           & 1           & 1         & 1           & 1         \\
    \multicolumn{10}{l}{\emph{(skipped: $\Reyn = 10^{6}, \Damk = 1, \alpha_0 = 0.05$)}} \\
    C23 & $10^{-6}$ & $10^{6}$  & 0.05 & 1 & $10^{6}$    & $10^{12}$   & $10^{-9}$ & $10^{-3}$   & $10^{3}$  \\
    C24 & 1         & $10^{6}$  & 0.05 & 1 & 1           & $10^{6}$    & $10^{-3}$ & $10^{-3}$   & $10^{3}$  \\
    \multicolumn{10}{l}{\emph{(skipped: $\Reyn = 10^{6}, \Damk = 10^{6}, \alpha_0 = 0.05$)}} \\
    \bottomrule
  \end{tabular}
  \caption{Dimensional parameter values for each dimensionless cell in the
    \texttt{phase1\_quad\_k1} sweep, under the default encoding
    ($U=L=1$) and the centered encoding
    ($U = \Reyn/\sqrt{\alpha_\infty \Damk}$, $L=1$), with
    $\alpha_\infty = 1$ as run by the test harness
    (\texttt{run\_test.jl:76}). The dimensionless physics is identical
    between the two columns; only the dimensional representation differs.
    Under the centered encoding, $\nu$ and $\sigma$ depend only on
    $\Damk$, never on $\Reyn$ or $\alpha_0$.}
  \label{tab:per-cell}
\end{table}

\paragraph{Maximum dimensional coefficient.}
The worst cells under the default encoding (C7, C16, C23) hit
$\sigma = 10^{12}$. Under the centered encoding the worst is
$\sigma = 10^{3}$ (any cell with $\Damk = 10^{6}$). The maximum dimensional
coefficient drops by nine orders of magnitude across the sweep.

\section{Verification evidence}

Two single-cell probes
(\texttt{test/extended/ManufacturedSolutions/diagnostics/jacobian\_equilibration\_osgs\_probe.jl})
validated the reparametrization before the full rerun. Note that the
probe used a slightly different centering formula
$U = \Reyn/\sqrt{\alpha_0 \Damk}$ (with $\alpha_0 = 0.5$ for C16), so
$U = 1.41\cdot10^{-9}$ and $\sigma = 1.41\cdot10^{3}$ in the probe;
the harness rerun uses $U = \Reyn/\sqrt{\alpha_\infty \Damk}$ with
$\alpha_\infty = 1$, giving $U = 10^{-9}$ and $\sigma = 10^{3}$ exactly.

\paragraph{C16 OSGS (high-$\sigma$, the canonical fold case).}
\begin{center}
\begin{tabular}{lrr}
  \toprule
  & Default & Centered (probe) \\
  \midrule
  $\nu$                           & $10^{6}$            & $1.41\cdot10^{-3}$ \\
  $\sigma$                        & $10^{12}$           & $1.41\cdot10^{3}$ \\
  OSGS staggered iters$^{\dagger}$            & 25 (budget)         & 0 (immediate) \\
  $\pi_u$-drift$^{\dagger}$                   & $7.6\cdot10^{8}$    & $1.03\cdot10^{-6}$ \\
  \texttt{overall\_converged}     & \texttt{false}      & \texttt{true} \\
  Initial residual                & $1.27\cdot10^{-3}$  & $3.08\cdot10^{-10}$ \\
  $L^{2}(\boldsymbol{u})$         & $1.07\cdot10^{-3}$  & $4.72\cdot10^{-4}$ \\
  $H^{1}(\boldsymbol{u})$         & $2.45\cdot10^{-1}$  & $1.04\cdot10^{-1}$ \\
  \bottomrule
\end{tabular}
\end{center}
\noindent ($^{\dagger}$ The ``OSGS staggered iters'' and ``$\pi_u$-drift''
rows are pre-leaning (2026-05-28) staggered-loop diagnostics, retained here
because they quantify the strong-residual-projection conditioning that the
encoding argument turns on; the staggered loop itself was deleted in the
2026-06-07 coupled-only leaning. The conditioning conclusion is unchanged.)
The strong-residual projection drift drops by
\emph{fourteen orders of magnitude}. Same physics; the difference is the
floating-point conditioning of the strong-residual projection.

\paragraph{C2 (easy regime; consistency check).}
With $\Reyn=1$, $\Damk=10^{-6}$, $\alpha_0=1$, both encodings should give
the same answer up to round-off.
\begin{center}
\begin{tabular}{lrr}
  \toprule
  & Default ($U=1$) & Centered ($U=10^{3}$) \\
  \midrule
  $H^{1}(\boldsymbol{u})$ & $3.612\cdot10^{-2}$ & $3.564\cdot10^{-2}$ \quad ($0.7\%$ diff) \\
  $L^{2}(\boldsymbol{u})$ & $5.92\cdot10^{-4}$  & $1.18\cdot10^{-4}$  \quad (different noise floor) \\
  \bottomrule
\end{tabular}
\end{center}
\noindent The $H^{1}$ semi-norm error agrees to under $1\%$ between
encodings, confirming that the centered encoding represents the same
discrete physics. The $L^{2}$ values differ by a constant noise-floor
factor: in both runs the absolute residual at convergence is $O(10^{-9})$,
and that absolute noise is divided by $U_c = U$ when forming the
normalized error — a $1000\times$ different denominator produces the
apparent $5\times$ discrepancy without any discretisation disagreement.

\section{Implementation of centered encoding}
\label{sec:impl-centered}

The centered encoding alone is a single-cell-loop change in the MMS
harness; no modification to the dimensional solver. The original
\code{U\_amp = 1.0}, \code{L = 1.0} line is replaced by:
\begin{center}
\begin{minipage}{0.9\linewidth}
\begin{lstlisting}[basicstyle=\ttfamily\small, frame=single]
# test/extended/ManufacturedSolutions/run_test.jl

(L_cell, U_amp) = compute_L_and_U(encoding_strategy,
                                  Re, Da, alpha_infty)
\end{lstlisting}
\end{minipage}
\end{center}
\noindent where \code{compute\_L\_and\_U} returns $(L=1, U=\Reyn/\sqrt{\alpha_\infty\Damk})$
for the centered strategy (Section~2), $(L=1, U=1)$ for the legacy
default, and the variable-$L$ values for balanced and minmax
(Section~\ref{sec:varL}). The downstream pipeline
(\code{build\_mms\_formulation}, \code{PaperMMS},
\code{evaluate\_exactness\_diagnostics}, \code{get\_characteristic\_scales})
threads $U$ and $L$ through correctly: the MMS exact velocity is multiplied
by $U$, the manufactured forcing is computed analytically from that
scaled solution, and error norms divide out the appropriate
dimensionless combination (Section~\ref{sec:varL-norm} corrects the
naive divisor to give genuinely $L$-invariant quantities). The
variable-$L$ generalisation — polynomial frequency $\pi \to \pi/L$,
domain bounding box $\Omega \to L \cdot \Omega$, porosity radii
$r_i \to L \cdot r_i$, per-cell mesh rebuild — is the additional refactor
documented in §\ref{sec:varL-impl}.

\section{Variable $L$ generalisation}
\label{sec:varL}

The centered encoding of Section~2 spends the $U$ degree of freedom and
hardcodes $L = 1$ because the manufactured shape
$\sin(\pi x)$ was wired to a unit period. Generalising the polynomial
to $\sin(\pi x / L)$ and coupling the domain to $L$ (so the dimensionless
mesh ratio $h_{\text{dim}} = h / L = 1/N$ and the similarity analysis
remain intact under any choice of $L$) unlocks the second degree of
freedom.

\paragraph{Status (2026-06-01; default updated 2026-06-09).} The
generalisation is implemented. The harness default is now
\texttt{encoding\_strategy = "minmax"} (\texttt{run\_test.jl}:
\texttt{get(test\_dict, "encoding\_strategy", "minmax")}); the
\texttt{"centered"} strategy ($L=1$) preserves the pre-generalisation
behaviour bit-identically and remains available, alongside
\texttt{"balanced"} and \texttt{"minmax"}, which expose the second-DOF
machinery for use in MMS sweeps. The remainder of this section presents
the math (§\ref{sec:varL-identities}--\ref{sec:varL-comparison}), the
implementation (§\ref{sec:varL-impl}), the empirical validation
(§\ref{sec:varL-empirical}), the Float64-precision interpretation
(§\ref{sec:varL-fp}), and the dimensionless-error normalisation
correction (§\ref{sec:varL-norm}).

\subsection{Identities and the design space}
\label{sec:varL-identities}

Recall the two identities from Section~2:
\begin{equation}
  \nu \, \sigma \;=\; \frac{\alpha_\infty\,\Damk\,U^{2}}{\Reyn^{2}},
  \qquad
  \frac{\nu}{\sigma} \;=\; \frac{L^{2}}{\alpha_\infty\,\Damk}.
  \label{eq:vL-identities}
\end{equation}
The first identity says $U$ alone controls the geometric mean
$\sqrt{\nu\sigma}$ — $L$ has no effect on it. The second says $L$ alone
controls the ratio $\nu/\sigma$ — $U$ has no effect on it. The two
degrees of freedom are therefore \emph{orthogonal} in log-space: one
slides the pair $(\nu, \sigma)$ along the line $\log\nu + \log\sigma =
\mathrm{const}$, the other rotates the pair around its geometric mean.

\subsection{Balanced-coefficient encoding ($\nu = \sigma$)}

The natural use of the $L$ degree of freedom is to set $\nu/\sigma = 1$,
making the diffusion and reaction coefficients dimensionally identical
on every cell. From the second identity in~\eqref{eq:vL-identities}:
\begin{equation}
  L^{\star} \;=\; \sqrt{\alpha_\infty\,\Damk}.
  \label{eq:vL-L}
\end{equation}
Note that, like $\nu$ and $\sigma$ in the centered encoding, $L^\star$
depends only on $\Damk$ (and $\alpha_\infty$), not on $\Reyn$ or
$\alpha_0$.

The remaining choice is $U$. The centered selection
$U = \Reyn / \sqrt{\alpha_\infty \Damk}$ collapses the geometric mean to
unity ($\sqrt{\nu\sigma} = 1$), so combined with~\eqref{eq:vL-L} it
yields $\nu = \sigma = 1$ on every cell. This is the most aggressive
flattening of the operator coefficients, but it leaves $U$ unchanged
from the centered encoding — the same factor of $10^9$ excursion to
$U = 10^{-9}$ at the C7 corner.

\subsection{Minmax encoding: trading $U$ headroom against $\nu, \sigma$}

A different use of the $L$ DOF is to minimise the worst-case excursion
across all three of $U, \nu, \sigma$ simultaneously. Concretely, we
solve
\begin{equation}
  \min_{U, L}\;\; \max\!\bigl(\,|\log_{10} U|,\,|\log_{10} \nu|,\,|\log_{10} \sigma|\,\bigr).
\end{equation}
Keeping $\nu = \sigma$ (which sets $L = L^\star$ as above) and choosing
$U$ to equalise $|\log U|$ with $|\log \nu|$ gives a closed form. Define
\begin{equation}
  A \;\equiv\; \tfrac{1}{2}\log_{10}(\alpha_\infty \Damk) \;-\; \log_{10}\Reyn.
\end{equation}
$A$ is the excursion of $\log\nu$ from $\log U$ at the centered choice.
The minmax optimum is achieved at
\begin{equation}
  \boxed{\;\;
    \log_{10} U^{\star\star} \;=\; -\,\frac{A}{2},
    \qquad
    \log_{10} L^{\star\star} \;=\; \tfrac{1}{2}\log_{10}(\alpha_\infty \Damk),
    \qquad
    \log_{10}\nu^{\star\star} \;=\; \log_{10}\sigma^{\star\star} \;=\; +\,\frac{A}{2}.
  \;\;}
  \label{eq:vL-minmax}
\end{equation}
At this point, $|\log U| = |\log \nu| = |\log \sigma| = |A|/2$ — all
three dimensional quantities sit at the same distance from unity, with
$U$ on the opposite side of $\sigma$ in log-space. Closed forms:
\begin{equation}
  U^{\star\star} \;=\; \left(\frac{\Reyn^{2}}{\alpha_\infty\,\Damk}\right)^{1/4},
  \qquad
  L^{\star\star} \;=\; \sqrt{\alpha_\infty\,\Damk},
  \qquad
  \nu^{\star\star} \;=\; \sigma^{\star\star} \;=\; \left(\frac{\alpha_\infty\,\Damk}{\Reyn^{2}}\right)^{1/4}.
\end{equation}
Notice the satisfying symmetry $U^{\star\star} \cdot \nu^{\star\star} =
1$ on every cell.

\paragraph{Spread reduction.}
Under the centered encoding the worst-case excursion is $|A|$:
$\log\nu = -A/2 \cdot 2 = $ (signed) $A$ and $\log U =$ opposite-sign
$A/2$, so $\max(|\log U|,|\log\nu|,|\log\sigma|) = |A|$. Under minmax
that worst case drops to $|A|/2$ — \textbf{a factor-of-two reduction in
log-magnitude}, i.e.\ the worst dimensional quantity halves its number
of digits away from unity, on every cell.

\subsection{Per-cell numerical values}

Table~\ref{tab:per-cell-varL} reports the dimensional values under
both new strategies, alongside the centered baseline for reference.
All values shown for $\alpha_\infty = 1$ as run by the harness.

\begin{table}[!htbp]
  \centering
  \footnotesize
  \setlength{\tabcolsep}{3pt}
  \begin{tabular}{l rrr rrr rrrr rrrr}
    \toprule
    & \multicolumn{3}{c}{Dim'less} &
      \multicolumn{3}{c}{Centered ($L=1$)} &
      \multicolumn{4}{c}{Balanced ($\nu = \sigma$, $U$ centered)} &
      \multicolumn{4}{c}{Minmax ($\nu = \sigma = 1/U$)} \\
    \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-11} \cmidrule(lr){12-15}
    \# & $\Reyn$ & $\Damk$ & $\alpha_0$
       & $U$ & $\nu$ & $\sigma$
       & $U$ & $L$ & $\nu$ & $\sigma$
       & $U$ & $L$ & $\nu$ & $\sigma$ \\
    \midrule
    C1  & $10^{-6}$ & $10^{-6}$ & 1.00
        & $10^{-3}$ & $10^{3}$ & $10^{-3}$
        & $10^{-3}$ & $10^{-3}$ & 1 & 1
        & $10^{-1.5}$ & $10^{-3}$ & $10^{1.5}$ & $10^{1.5}$ \\
    C2  & 1 & $10^{-6}$ & 1.00
        & $10^{3}$ & $10^{3}$ & $10^{-3}$
        & $10^{3}$ & $10^{-3}$ & 1 & 1
        & $10^{1.5}$ & $10^{-3}$ & $10^{-1.5}$ & $10^{-1.5}$ \\
    C3  & $10^{6}$ & $10^{-6}$ & 1.00
        & $10^{9}$ & $10^{3}$ & $10^{-3}$
        & $10^{9}$ & $10^{-3}$ & 1 & 1
        & $10^{4.5}$ & $10^{-3}$ & $10^{-4.5}$ & $10^{-4.5}$ \\
    C4  & $10^{-6}$ & 1 & 1.00
        & $10^{-6}$ & 1 & 1
        & $10^{-6}$ & 1 & 1 & 1
        & $10^{-3}$ & 1 & $10^{3}$ & $10^{3}$ \\
    C5  & 1 & 1 & 1.00
        & 1 & 1 & 1
        & 1 & 1 & 1 & 1
        & 1 & 1 & 1 & 1 \\
    C6  & $10^{6}$ & 1 & 1.00
        & $10^{6}$ & 1 & 1
        & $10^{6}$ & 1 & 1 & 1
        & $10^{3}$ & 1 & $10^{-3}$ & $10^{-3}$ \\
    C7  & $10^{-6}$ & $10^{6}$ & 1.00
        & $10^{-9}$ & $10^{-3}$ & $10^{3}$
        & $10^{-9}$ & $10^{3}$ & 1 & 1
        & $10^{-4.5}$ & $10^{3}$ & $10^{4.5}$ & $10^{4.5}$ \\
    C8  & 1 & $10^{6}$ & 1.00
        & $10^{-3}$ & $10^{-3}$ & $10^{3}$
        & $10^{-3}$ & $10^{3}$ & 1 & 1
        & $10^{-1.5}$ & $10^{3}$ & $10^{1.5}$ & $10^{1.5}$ \\
    C9  & $10^{6}$ & $10^{6}$ & 1.00
        & $10^{3}$ & $10^{-3}$ & $10^{3}$
        & $10^{3}$ & $10^{3}$ & 1 & 1
        & $10^{1.5}$ & $10^{3}$ & $10^{-1.5}$ & $10^{-1.5}$ \\
    \midrule
    \multicolumn{15}{l}{\emph{$\alpha_0 = 0.5$ cells (C10--C18): identical $U, \nu, \sigma, L$ values
                                  as C1--C9 because the harness uses
                                  $\alpha_\infty = 1$.}} \\
    \midrule
    C19 & $10^{-6}$ & $10^{-6}$ & 0.05
        & $10^{-3}$ & $10^{3}$ & $10^{-3}$
        & $10^{-3}$ & $10^{-3}$ & 1 & 1
        & $10^{-1.5}$ & $10^{-3}$ & $10^{1.5}$ & $10^{1.5}$ \\
    C20 & 1 & $10^{-6}$ & 0.05
        & $10^{3}$ & $10^{3}$ & $10^{-3}$
        & $10^{3}$ & $10^{-3}$ & 1 & 1
        & $10^{1.5}$ & $10^{-3}$ & $10^{-1.5}$ & $10^{-1.5}$ \\
    C21 & $10^{-6}$ & 1 & 0.05
        & $10^{-6}$ & 1 & 1
        & $10^{-6}$ & 1 & 1 & 1
        & $10^{-3}$ & 1 & $10^{3}$ & $10^{3}$ \\
    C22 & 1 & 1 & 0.05
        & 1 & 1 & 1
        & 1 & 1 & 1 & 1
        & 1 & 1 & 1 & 1 \\
    C23 & $10^{-6}$ & $10^{6}$ & 0.05
        & $10^{-9}$ & $10^{-3}$ & $10^{3}$
        & $10^{-9}$ & $10^{3}$ & 1 & 1
        & $10^{-4.5}$ & $10^{3}$ & $10^{4.5}$ & $10^{4.5}$ \\
    C24 & 1 & $10^{6}$ & 0.05
        & $10^{-3}$ & $10^{-3}$ & $10^{3}$
        & $10^{-3}$ & $10^{3}$ & 1 & 1
        & $10^{-1.5}$ & $10^{3}$ & $10^{1.5}$ & $10^{1.5}$ \\
    \bottomrule
  \end{tabular}
  \caption{Per-cell dimensional values under three encodings: centered
    ($L=1$ fixed), balanced ($\nu = \sigma$ via $L = \sqrt{\alpha_\infty \Damk}$
    with centered $U$), and minmax (jointly optimal $U$ and $L$).
    Skipped cells $(\Reyn=10^{6}, \alpha_0=0.05)$ omitted. Values shown
    with $\alpha_\infty = 1$ as run by the test harness; under this
    convention all three encodings are independent of $\alpha_0$.}
  \label{tab:per-cell-varL}
\end{table}

\subsection{Comparison summary}
\label{sec:varL-comparison}

For each encoding and each cell, the worst-case dimensional excursion
$M = \max(|\log_{10} U|, |\log_{10}\nu|, |\log_{10}\sigma|)$ measures
how far from unity the most extreme coefficient lies. The sweep maxima
over all 24 cells are:

\begin{center}
\begin{tabular}{lrl}
  \toprule
  Encoding & $\max_{\text{cells}} M$ & Worst-case cells \\
  \midrule
  Default ($U = L = 1$)             & 12  & C3, C7, C12, C16, C23 \\
  Centered ($L = 1$)                & 9   & C3, C7, C12, C16, C23 \\
  Balanced ($\nu = \sigma$, $U$ centered) & 9 & C3, C7, C12, C16, C23 (set by $|\log U|$) \\
  Minmax (jointly optimal $U, L$)   & 4.5 & C3, C7, C12, C16, C23 \\
  \bottomrule
\end{tabular}
\end{center}

\noindent The same five $(\mathrm{Re}, \mathrm{Da}) \in
\{(10^{6}, 10^{-6}), (10^{-6}, 10^{6})\}$ corner cells dominate in
every encoding — they are the diagonally-opposite corners of the
$(\mathrm{Re}, \mathrm{Da})$ plane that the sweep grid samples. The
balanced encoding flattens the operator coefficients to
$\nu = \sigma = 1$ but does not improve the worst-case excursion
because $|\log U|$ is unchanged from centered. The minmax encoding
redistributes the excursion across all three quantities and halves
the worst case from 9 to 4.5 orders of magnitude.

\subsection{Implementation}
\label{sec:varL-impl}

The generalisation was implemented on 2026-06-01. The diff is contained
to two files plus blitz tests:

\paragraph{\texttt{src/problems/mms\_paper\_2d.jl}.}  The hard-coded
spatial frequency $\pi$ is replaced by $k = \pi / L$ throughout the
analytic chain. The struct \code{UExFunc} gains an
\code{L::Float64} field; the operator
$f \mapsto f(x) = U \alpha_0 \alpha(x)^{-1} S(x/L)$ is now genuinely
$L$-aware. The five analytic-derivative functions
\code{grad\_u\_ex}, \code{lap\_u\_ex}, \code{grad\_div\_u\_ex},
\code{get\_p\_ex}, and the pressure-gradient block in
\code{evaluate\_exactness\_diagnostics} each pick up the appropriate
$1/L$ from the chain rule: first derivatives by $1/L$, second
derivatives by $1/L^2$. A backward-compatible three-argument constructor
\code{UExFunc(U, alpha\_0, alpha\_field)} with $L = 1$ default is added
so legacy probes that pre-date the generalisation continue to compile.

\paragraph{\texttt{test/extended/ManufacturedSolutions/run\_test.jl}.}
A new helper
\code{compute\_L\_and\_U(strategy, Re, Da, alpha\_infty)}
returns the pair $(L, U_{\text{amp}})$ for the four named strategies
\code{"centered"} (default — $L = 1$, $U = \mathrm{Re}/\sqrt{\alpha_\infty \mathrm{Da}}$),
\code{"balanced"} ($L = \sqrt{\alpha_\infty \mathrm{Da}}$,
same $U$),
\code{"minmax"} ($L = \sqrt{\alpha_\infty \mathrm{Da}}$,
$U = (\mathrm{Re}^2 / (\alpha_\infty \mathrm{Da}))^{1/4}$), and
\code{"unit"} ($L = U = 1$, kept for diagnostic probes). The strategy
is read once at the top of \code{run\_mms} via
\code{get(test\_dict, "encoding\_strategy", "centered")}.

The harness builds the per-cell $L$ \emph{inside} the
$(\mathrm{Re}, \mathrm{Da}, \alpha_0)$ triple loop because $L$ varies
per cell under non-centered strategies. The mesh and FE spaces are
therefore rebuilt per cell rather than once per~$N$ — a few extra
seconds of wall-clock per sweep, negligible against solve time.
\code{\_build\_local\_mesh} accepts $L$ and tuples
\code{L .* domain\_cfg.bounding\_box}; \code{build\_porosity\_field}
accepts $L$ and constructs \code{SmoothRadialPorosity} with
$L \cdot r_1$ and $L \cdot r_2$. The homotopy perturbation
\code{h\_raw\_func} is rescaled with $k = \pi/L$ on the polynomial
frequencies and with $1/L^8$ on the boundary bubble $B_{\text{fn}}$,
so $\|h\|_{L^2(\Omega)}$ stays $O(L^{d/2})$ matching the manufactured
solution.

\paragraph{Similarity invariants preserved by the implementation.} The
key check that the dimensionless physics is genuinely the same across
encodings:
\begin{itemize}
\item $\mathrm{Re}_{\text{check}} = U \cdot L / \nu = U L / (UL/\mathrm{Re}) = \mathrm{Re}$
      — preserved algebraically.
\item $\mathrm{Da}_{\text{check}} = \alpha_\infty \sigma L^2 / \nu
      = \alpha_\infty (\alpha_\infty \mathrm{Da} \nu / L^2) L^2 / (\alpha_\infty \nu) = \mathrm{Da}$
      — preserved.
\item $u_{ex}(L\hat{x}) / U_{\text{amp}} = \alpha_0 / \alpha(L\hat{x}) \cdot \hat{S}(\hat{x})$.
      For this ratio to be $L$-invariant, $\alpha(L\hat{x})$ as a
      function of $\hat{x}$ must equal $\alpha(\hat{x})$ at $L=1$.
      Writing $\tilde{r}_i = L \cdot r_i$ for the L-scaled radii stored
      in the field (set by \code{build\_porosity\_field}), the
      smoothing parameter at physical $r$ is
      $\eta(r; L) = (r^2 - \tilde{r}_1^2)/(\tilde{r}_2^2 - \tilde{r}_1^2)$.
      At $r = L\hat{r}$:
      \[
        \eta(L\hat{r}; L) \;=\; \frac{L^2 \hat{r}^2 - L^2 r_1^2}{L^2 r_2^2 - L^2 r_1^2}
        \;=\; \frac{\hat{r}^2 - r_1^2}{r_2^2 - r_1^2}
        \;=\; \eta(\hat{r}; 1).
      \]
      The $L^2$ factor cancels between numerator and denominator, so
      $\alpha$ at the corresponding dimensionless position is identical
      to the $L=1$ baseline.
\item Physical $\nabla_x \alpha$ scales as $1/L$ — exactly as a physical
      gradient should under coordinate dilation. From the same formula,
      $d\eta/dr = 2r/(\tilde{r}_2^2 - \tilde{r}_1^2) = 2r/(L^2 (r_2^2 - r_1^2))$.
      At $r = L \hat{r}$ this gives $(2\hat{r}/(r_2^2 - r_1^2)) \cdot 1/L$,
      so the dimensionless $L \cdot \nabla \alpha = 2 \hat{r}/(r_2^2 - r_1^2)$
      is encoding-invariant.
\item The grid Reynolds number $\mathrm{Re}_h = U h / \nu
      = (U L / \nu) \cdot (h/L) = \mathrm{Re} / N$ — independent of $L$.
      Stabilisation $\tau_{\text{NS}}$ at the same $N$ should therefore have
      the same dimensionless value across encodings.
\end{itemize}

\subsection{Empirical validation}
\label{sec:varL-empirical}

Two end-to-end checks were performed.

\paragraph{$L=1$ bit-identical regression.} Cell C5 (Re$=1$, Da$=1$,
$\alpha_0=1$, OSGS) run at $N \in \{10, 20, 40, 80\}$ through the new
harness with \code{encoding\_strategy = "centered"} reproduces the
master HDF5 values to all seven reported digits:

\begin{center}
\begin{tabular}{lrrr|rrr}
  \toprule
  & \multicolumn{3}{c|}{Master} & \multicolumn{3}{c}{New harness (centered)} \\
  $N$ & $L^2_u$ & $L^2_p$ & $H^1_u$ & $L^2_u$ & $L^2_p$ & $H^1_u$ \\
  \midrule
  10 & $9.530767\!\times\!10^{-3}$ & $2.298378\!\times\!10^{-2}$ & $2.858131\!\times\!10^{-1}$
     & $9.530767\!\times\!10^{-3}$ & $2.298378\!\times\!10^{-2}$ & $2.858131\!\times\!10^{-1}$ \\
  20 & $1.841275\!\times\!10^{-3}$ & $1.928104\!\times\!10^{-2}$ & $1.438814\!\times\!10^{-1}$
     & $1.841275\!\times\!10^{-3}$ & $1.928104\!\times\!10^{-2}$ & $1.438814\!\times\!10^{-1}$ \\
  40 & $4.633695\!\times\!10^{-4}$ & $5.695171\!\times\!10^{-3}$ & $7.150274\!\times\!10^{-2}$
     & $4.633695\!\times\!10^{-4}$ & $5.695171\!\times\!10^{-3}$ & $7.150274\!\times\!10^{-2}$ \\
  80 & $1.313346\!\times\!10^{-4}$ & $1.976955\!\times\!10^{-3}$ & $3.568200\!\times\!10^{-2}$
     & $1.313346\!\times\!10^{-4}$ & $1.976955\!\times\!10^{-3}$ & $3.568200\!\times\!10^{-2}$ \\
  \bottomrule
\end{tabular}
\end{center}

\noindent The agreement confirms that the per-cell mesh rebuild, the
$k = \pi/L$ substitution, the porosity-radius scaling, and the
normalisation correction (§\ref{sec:varL-norm}) collapse exactly to the
pre-generalisation formula at $L = 1$.

\paragraph{C7 minmax probe.} The corner cell C7 (Re$=10^{-6}$,
Da$=10^{6}$, $\alpha_0=1$, OSGS) was the empirical motivation: under the
centered encoding, the $L^2_u$ slope at the finest mesh pair drops from
$2.04$ at $N=80 \to 160$ to $0.84$ at $N=160 \to 320$ — a stall
attributed in Section~4 to the solution-scale precision floor at
$U_{\text{amp}} = 10^{-9}$. Running the same cell at
\code{encoding\_strategy = "minmax"} (which gives $U_{\text{amp}} =
10^{-4.5}$, $L = 10^3$, $\nu = \sigma = 10^{4.5}$) yields the
complementary trajectory:

\begin{center}
\begin{tabular}{lrr}
  \toprule
  $N_{\text{coarse}} \to N_{\text{fine}}$ & Centered slope & Minmax slope \\
  \midrule
  $10 \to 20$   & $1.96$ & $0.94$ \\
  $20 \to 40$   & $1.98$ & $1.28$ \\
  $40 \to 80$   & $2.01$ & $1.47$ \\
  $80 \to 160$  & $2.04$ & $1.63$ \\
  $160 \to 320$ & $\mathbf{0.84}$ \emph{(stalled)} & $\mathbf{1.93}$ \emph{(recovered)} \\
  \bottomrule
\end{tabular}
\end{center}

\noindent Centered is clean at coarse meshes and stalls at the finest
pair; minmax is pre-asymptotic at coarse meshes and recovers at the
finest pair. The trajectories cross. The continuous problems and the
dimensionless meshes are identical between the two; the difference is
the Float64 realisation of the discrete operator, discussed next.

\subsection{Floating-point realisation}
\label{sec:varL-fp}

The continuous dimensionless problem
\begin{equation}
  -\tfrac{1}{\Reyn} \widehat{\Delta} \hat{u}
  + \tfrac{\alpha_\infty \Damk}{\Reyn} \hat{u}
  + (\hat{u} \cdot \widehat{\nabla}) \hat{u}
  + \widehat{\nabla} \hat{p}
  = \hat{f}
  \label{eq:dimless-ns}
\end{equation}
is invariant under the choice of $(U_{\text{amp}}, L)$ — both encodings
discretise the same equation, and the algebraic similarity invariants
of §\ref{sec:varL-impl} (Re, Da, $\alpha$ at corresponding positions,
$\nabla\alpha$ scaling, grid Reynolds $\mathrm{Re}_h = \mathrm{Re}/N$)
all hold exactly. The discrete realisation is invariant only to the
precision of the floating-point system. Float64's $\sim 16$ digits of
relative precision are preserved at each elementary operation, but the
\emph{same} dimensionless quantity computed via \emph{different
sequences of dimensional operations} at different orders of magnitude
accumulates round-off differently. The empirical L=1 bit-identity
(§\ref{sec:varL-empirical}) and the dimensional-similarity algebra
together pin the source to discrete operations whose absolute operand
magnitudes change with the encoding; three plausible sites:

\begin{enumerate}
\item \textbf{OSGS strong-residual projection.} The discrete projection
      $\boldsymbol{\pi}_h = M^{-1} R(u_h)$ inverts the dimensional mass
      matrix $M$ against the dimensional strong residual $R$. For C7 at
      $N=10$, where the cell measure is $|\Omega|/N^2$:

      \begin{itemize}
      \item Centered: $M_{ij} \sim |\Omega|/N^2 \sim 10^{-2}$;
            the residual entries $R_i$ sum contributions
            $\sigma u_h \sim 10^{-6}$, $-\nu \Delta u_h \sim 10^{-11}$,
            $\nabla p_h \sim 3 \cdot 10^{-6}$, all dimensional. The
            $R_i$ extracted by quadrature is on order
            $10^{-8}$--$10^{-7}$; the projected $\pi$ on order $10^{-5}$.
      \item Minmax: $M_{ij} \sim L^2 |\Omega|/N^2 \sim 10^{4}$;
            $R_i$ summands are $\sigma u_h \sim 1$,
            $-\nu \Delta u_h \sim 10^{-5.5}$,
            $\nabla p_h \sim 0.3$ — five orders of magnitude apart in
            the $\sigma u$ vs $\nu \Delta u$ summands rather than
            centered's six.
      \end{itemize}

      Both encodings produce the same dimensionless $\hat{\boldsymbol{\pi}}_h$
      \emph{exactly} (the dimensional rescaling cancels term by term),
      but the round-off in computing the dimensional $R$ as a sum of
      four contributions at different magnitudes accumulates
      differently. The relative position of summands in log-space
      changes between encodings, and Float64 cancels the smaller
      ones differently when summed with the larger ones.
\item \textbf{Momentum stabilisation inverse-scale $\tau_1^{-1}$.} The
      dimensional formula
      $\tau_1^{-1} = \alpha\,c_1 \nu / h^2 + \alpha\,c_2 |u|/h + \sigma$
      (with $c_1 = 4k^4$, $c_2 = 2k^2$; note the reaction $\sigma$ enters
      $\tau_1$ directly, \emph{not} the Navier--Stokes parameter
      $\tau_{\text{NS}} = (c_1 \nu / h^2 + c_2 |u|/h)^{-1}$, which carries no
      reaction term) contains summands that scale unequally per encoding.
      For C7 at $N=10$ the diffusion summand and the reaction summand
      $\sigma$ sit four orders apart:
      \begin{itemize}
      \item Centered: $\alpha\,c_1 \nu / h^2 \approx 0.4\,c_1$, $\sigma \approx 2000$ — summands span four orders.
      \item Minmax:   $\alpha\,c_1 \nu / h^2 \approx 12.6\,c_1$, $\sigma \approx 6.3 \times 10^4$ — summands span four orders again, but the diffusion summand sits at a different relative weight and is more vulnerable to floating-point obliteration when added to the reaction summand.
      \end{itemize}
      $\hat{\tau} = \tau \cdot U_c / L$ is encoding-invariant
      algebraically; the dimensional summation is not invariant in
      finite precision.
\item \textbf{Newton residual.} $\boldsymbol{r} = \boldsymbol{b} -
      \mathcal{J}\, \delta\boldsymbol{x}$ in centered has
      $|\boldsymbol{b}|, |\mathcal{J} \delta\boldsymbol{x}| \sim 10^{-7}$
      cancelling to extract $|\boldsymbol{r}|$ at machine precision; in
      minmax the same cancellation happens between operands at $\sim
      10^{4.5}$. Same relative residual reduction, different absolute
      round-off in the extraction.
\end{enumerate}

\noindent At fine $N$ the FE truncation grows back above these
floating-point floors and the dimensionless slopes converge to the
$h^{k+1}$ optimum in \emph{both} encodings. At coarse $N$ and at the
mesh where the corresponding floor is binding, encoding-dependent
deviations show: centered fails first at solution-scale (C7 at $N=320$
where $U_{\text{amp}} h^{k+1} \sim \varepsilon_{\text{mach}}$); minmax
fails first at operator-scale (the same cell at coarse $N$ where
$\sigma h^2 \sim$ summand-cancellation noise). The full sweep of the
24~cells shows the bulk of cells lying away from either floor for both
encodings; the slope trajectory is essentially identical for all cells
where neither floor is binding (most of the
$|\log\nu|, |\log\sigma|, |\log U| < 6$ interior).

\subsection{Dimensionless error normalisation}
\label{sec:varL-norm}

The harness's error norms were historically computed as
\begin{equation}
  \widetilde{e}_{L^2, u} \;=\; \frac{\|\boldsymbol{e}_u\|_{L^2(\Omega)}}{U_c},
  \qquad
  \widetilde{e}_{H^1, u} \;=\; \frac{\|\nabla \boldsymbol{e}_u\|_{L^2(\Omega)}}{U_c / L}.
\end{equation}
At $L = 1$ with a unit-area baseline bounding box these are numerically
the dimensionless quantities $\|\hat{e}_u\|_{L^2(\hat\Omega)}$ and
$\|\widehat{\nabla} \hat{e}_u\|_{L^2(\hat\Omega)}$, but at $L \neq 1$
they carry a residual factor of $L$ — they are \emph{not} fully
dimensionless. Reading the $L^2$ similarity in detail
($\boldsymbol{e}_u(x) = U_c \cdot \hat{\boldsymbol{e}}_u(x/L)$,
$\mathrm{d}x = L^d \mathrm{d}\hat{x}$):
\begin{align}
  \|\boldsymbol{e}_u\|_{L^2(\Omega)}^2
  \;&=\; U_c^2 \cdot L^d \cdot \|\hat{\boldsymbol{e}}_u\|_{L^2(\hat\Omega)}^2 \notag \\
  \Rightarrow \quad
  \|\boldsymbol{e}_u\|_{L^2(\Omega)}
  \;&=\; U_c \cdot \sqrt{|\Omega|} \cdot \|\hat{\boldsymbol{e}}_u\|_{L^2(\hat\Omega)}.
  \label{eq:l2-similarity}
\end{align}
For the gradient, $\nabla_x \boldsymbol{e}_u = (U_c / L) \nabla_{\hat{x}} \hat{\boldsymbol{e}}_u$,
and the $L^2$ similarity inherits a $(1/L)^2 \cdot L^d$ factor that in
2D reduces to unity:
\begin{equation}
  \|\nabla \boldsymbol{e}_u\|_{L^2(\Omega)}
  \;=\; U_c \cdot \|\widehat{\nabla} \hat{\boldsymbol{e}}_u\|_{L^2(\hat\Omega)}
  \qquad (\text{2D}).
  \label{eq:h1-similarity}
\end{equation}
The dimensionless quantities are therefore
\begin{equation}
  \boxed{\quad
    e_{L^2, u} \;=\; \frac{\|\boldsymbol{e}_u\|_{L^2(\Omega)}}{U_c \cdot \sqrt{|\Omega|}},
    \qquad
    e_{H^1, u} \;=\; \frac{\|\nabla \boldsymbol{e}_u\|_{L^2(\Omega)}}{U_c}
    \quad
  }
  \label{eq:norm-fix}
\end{equation}
(and analogously $P_c \cdot \sqrt{|\Omega|}$ and $P_c$ for the pressure).
The corrected formulas are implemented in
\code{calculate\_normalized\_errors} in
\code{test/extended/ManufacturedSolutions/run\_test.jl}
(line~150 of the 2026-06-01 commit). For
the standard baseline domain $[-0.5, 0.5]^2$ at $L = 1$, $|\Omega| = 1$
and the corrected formulas collapse to the legacy ones bit-identically —
which is exactly what the regression of §\ref{sec:varL-empirical}
confirms.

\section{What is deliberately \emph{not} addressed here}

\begin{itemize}
  \item \textbf{Higher-precision arithmetic} as a Float64-floor remedy.
        The encoding-dependent slope drift documented in
        §\ref{sec:varL-fp} is a property of Float64's 16-digit relative
        precision applied to dimensional operands at heterogeneous
        magnitudes. Switching the LU back-substitution (and the
        $\tau$ and OSGS-projection arithmetic) to extended precision
        — \code{DoubleFloats.jl} ($\sim 32$ digits) or quad precision —
        would eliminate the floor at the cost of a $\sim 5$--$10\times$
        runtime overhead. This is a single-cell diagnostic worth
        running if a future paper requires the C7-corner slope cleanly
        at the finest mesh under \emph{either} encoding; the variable-$L$
        machinery already moves the floor by 4.5 orders of magnitude,
        which is sufficient for the present sweep.
  \item \textbf{The pressure penalty $\eta = \mathrm{eps\_val}$.}
        The dimensional pressure amplitude $P_c$ varies with the encoding,
        which means $\eta \cdot p^2$ has a different effective weight per
        cell. Empirically the discrete LBB stabilization is the primary
        gauge fix and $\eta$ acts only as a safety net; the rerun will
        validate this in practice. If the sweep reveals an $\eta$-related
        regression, a regime-aware scaling
        $\eta_{\mathrm{cell}} \propto P_c^{-2}$ is a focused follow-up.
  \item \textbf{Solver tolerances.}
        \texttt{ftol}, \texttt{xtol}, \texttt{stagnation\_noise\_floor}
        are absolute. Under the centered encoding the residual magnitudes
        shrink for low-$\Reyn$ cells; the existing
        \texttt{dynamic\_ftol\_spatial\_safety\_factor} machinery scales
        \texttt{ftol} with $h^{k+1}$, which empirically reaches sensible
        values in both encodings (the probes converged to high precision).
        If a regression appears, the dynamic-tolerance formula is the
        first place to look.
\end{itemize}

\section*{Status (2026-06-01)}

Centered encoding (Section~2) is the harness default and the
\code{phase1\_quad\_k1} master HDF5 was produced under it. Variable-$L$
generalisation (Section~\ref{sec:varL}) is implemented in the harness
with $L = 1$ centered preserved bit-identically; \code{"balanced"} and
\code{"minmax"} are available as opt-in strategies. The corner case
C7 (Re$=10^{-6}$, Da$=10^{6}$, $\alpha_0=1$) was the empirical
diagnosis target: its Float64 noise-floor stall under centered at
$N=320$ is recovered by minmax at the same $N$, confirming that the
mechanism is solution-scale precision (not algorithm-side). The
remainder of the (Re, Da, $\alpha$) sweep behaves equivalently across
encodings for cells where neither floor is binding.

This document is kept in \texttt{theory/} as a standalone working
note. Once a final paper-grade $k=2$ sweep is performed (deciding
between centered and minmax based on which side's Float64 floor is
binding for that polynomial order), the contents here should be
condensed into \code{article.tex} as a verification-section remark.

\end{document}
