% ===========================================================================
% osgs_appendix_commented.tex --- osgs_appendix.tex + pedagogical commentary
% ---------------------------------------------------------------------------
% The mathematical content below is VERBATIM osgs_appendix.tex: no statement,
% proof, equation or label has been altered, and the numbering is unchanged
% (the commentary adds no numbered objects). To use, replace
%       \input{osgs_appendix.tex}
% by    \input{osgs_appendix_commented.tex}
% in article_v2.tex. To hide the commentary without switching files, change
% \pedagogyshowtrue to \pedagogyshowfalse below (the boxes are then typeset
% into a discarded box, so pagination-independent content is unaffected).
%
% The commentary layer is addressed to a reader fluent in the ASGS analysis
% (app:Continuity) who is meeting the OSGS proof for the first time. Boxes
% marked "Audit note" flag wording the commentator believes could be
% tightened; they are suggestions about the prose, not applied corrections.
% ===========================================================================
\makeatletter
\newif\ifpedagogyshow
\pedagogyshowtrue % <-- change to \pedagogyshowfalse to hide every box
\@ifundefined{definecolor}{%
  \providecommand{\pedagogy@color}{}%
}{%
  \definecolor{PedagogyColor}{rgb}{0.05,0.32,0.60}%
  \providecommand{\pedagogy@color}{\color{PedagogyColor}}%
}
\newsavebox{\pedagogy@bin}
\ifpedagogyshow
\newenvironment{pedagogy}[1][Commentary]{%
  \par\medskip
  \begingroup\pedagogy@color\small
  \list{}{\leftmargin 1.4em \rightmargin 0.4em \topsep 0pt
          \parsep 0.4\baselineskip \listparindent 0pt \itemindent 0pt}%
  \item[]\textbf{\textsf{#1.}}\enspace\ignorespaces
}{%
  \endlist\endgroup\par\medskip
}
\else
\newenvironment{pedagogy}[1][Commentary]{%
  \begingroup\setbox\pedagogy@bin\vbox\bgroup\small
}{%
  \egroup\endgroup
}
\fi
\makeatother

\section{Convergence analysis of the OSGS method}\label{app:osgs}

This appendix establishes, for the linearized OSGS variant, the stability
(\cref{oa:th:stability}) and convergence (\cref{oa:th:convergence}) results stated as
\cref{th:StabilityOSGS,th:ConvergenceOSGS} in \cref{sec:StabilityOSGS}. It follows the orthogonal-subscale template
of~\cite{Codina2008FiniteEA}, transported to the porosity-weighted operator, the Darcy
reaction term, and the artificial-compressibility perturbation. The proof is developed in
full; only the material shared with the ASGS analysis is not re-derived. Specifically, the
continuous problem, the Galerkin bilinear form and its coercivity identity, the stabilization
parameters, the mesh and interpolation-error setting, and the common standing assumptions are
stated once in the main text (\cref{sec:CommonSetting,sec:StabilizationDesign}) and, for the
ASGS route, in \cref{app:Continuity}; here they are only recalled and referenced, so that this
appendix develops the ingredients specific to the OSGS route---the orthogonal projection, the
weighted best-approximation and smoothing lemmas, the inf--sup argument, and the
Damk\"ohler-weighted error functional. Differences from the ASGS analysis of
\cref{app:Continuity} are flagged where they arise.

\begin{pedagogy}[How to read this appendix]
These colored boxes form a commentary layer addressed to a reader who knows the
ASGS analysis of \cref{app:Continuity} in detail and is meeting the OSGS analysis
for the first time; they can be removed without altering any numbered content.
Both analyses follow the three-act structure common to all quasi-optimal finite
element error analyses: \emph{stability} (the discrete bilinear form does not
shrink any discrete direction), \emph{consistency} (inserting the exact solution
into the discrete equations leaves only a small defect), and \emph{continuity
with respect to the interpolation error} (the form applied to the interpolation
error is bounded by interpolation quantities times the test norm). Given the
three, convergence follows by a two-line argument (\cref{oa:th:convergence}),
identical in shape to \cref{thm:convergence}. The routes differ in how each act
is performed. For ASGS: stability is \emph{coercivity} (\cref{lem:coercivity}),
obtained by testing with the solution itself; consistency is \emph{exact}
(\cref{eq:consistency}), because the stabilization is built from the full
residual, which vanishes at the exact solution; continuity is the long
\cref{lem:continterp}, with the telescopings, the elementwise integrations by
parts and the jump machinery. For OSGS all three change, and for a single
structural reason established first (\cref{oa:lem:annihilation}): the
orthogonal projection deletes the reactive (and viscous) pieces from residual
and test slots. Consequences: (a)~the $\sigma \to \sigmatilde$ renormalization
of \cref{eq:sigmatilde}/\cref{eq:reactrenorm} never happens, so the working norm
keeps the \emph{full} reaction $\norm{\sigma^{1/2}\bv}$---a stronger norm;
(b)~testing with $U_h$ no longer produces the full residual square, so
coercivity in that stronger norm fails, and stability must be rebuilt as an
\emph{inf--sup} bound with a hand-built test function (\cref{oa:th:stability});
(c)~exact consistency fails, and is replaced by the principle that
\emph{fluctuations of smooth functions are small}
(\cref{oa:lem:bestapprox,oa:lem:consistency}). A suggested reading order:
\cref{oa:lem:annihilation} first, then the norm (\cref{oa:sec:norm}), then
stability, then convergence, consulting \cref{oa:sec:technical} on demand.
\end{pedagogy}

\subsection{The linearized problem and the OSGS method}
\label{oa:sec:setting}

\subsubsection{Continuous problem and Galerkin form}
\label{oa:sec:continuous}

The analyzed linearized problem, the Galerkin bilinear form $B(\ba;U,V)$
(\cref{eq:BilinearForm}) and the composite first-order quantity
$X(V) \coloneqq \ba\cdot\nabla\bv + \nabla q$ are those of the common setting
(\cref{sec:CommonSetting}) and of the ASGS analysis (\cref{sec:setting} of
\cref{app:Continuity}, where $X$ is denoted as in \cref{eq:XG}); we use them without
restatement. We recall only that, in the porous setting, the momentum residual carries the
weight $\alpha$, so the natural stabilized quantity is $\alpha X(V)$, and that
$(\nabla\bv, \alpha\nu\ViscProj\nabla\bu) = (\ViscProj\nabla\bv, \alpha\nu\ViscProj\nabla\bu)$
since $\ViscProj$ is a pointwise orthogonal projection and $\alpha\nu$ is scalar. We write
$(\cdot,\cdot)$ and $\norm{\cdot}$ for the $L^2(\Omega)$ inner product and norm (of scalars,
vectors or tensors), $\normK{\cdot} \coloneqq \norm{\cdot}_{L^2(K)}$,
$\norm{\cdot}_{m,K} \coloneqq \norm{\cdot}_{H^m(K)}$; the velocity and pressure spaces are
$V_0 \coloneqq H_0^1(\Omega)^d$ and $Q_0$ (the zero-mean subspace of $L^2(\Omega)$; for
$\varepsilon > 0$ the zero mean is automatic, since integrating the mass equation
$\varepsilon p + \nabla\cdot(\alpha\bu) = 0$ over $\Omega$ and using $\bu \in V_0$ gives
$\varepsilon\int_\Omega p = 0$), and $\mathcal{X}_0 \coloneqq V_0\times Q_0$.

Testing $B$ with $V = U$ and using $\nabla\cdot(\alpha\ba) = 0$ together with
$\bu|_{\partial\Omega} = \boldsymbol{0}$---so that the pressure--divergence pair cancels by
integration by parts and the convective term vanishes---gives the Galerkin coercivity
identity of the main text: for every $U \in \mathcal{X}_0$ with $\bu \in H^1_0(\Omega)^d$,
\begin{equation}
  \label{oa:eq:CoercivityIdentity}
  B(\ba; U, U)
  = 2\nu\,\bigl\lVert \alpha^{1/2}\ViscProj\nabla\bu \bigr\rVert^2
  + \norm{\sigma^{1/2}\bu}^2
  + \varepsilon\norm{p}^2 .
\end{equation}
This is \cref{eq:GalerkinCoercivityIdentity} of the main text, recalled here in the notation
of this appendix; the derivation is given there.

\begin{pedagogy}[The energy identity is the shared floor]
Identity \cref{oa:eq:CoercivityIdentity} is the same Galerkin energy budget both
methods start from (\cref{eq:GalerkinCoercivityIdentity}). Its mechanism, worth
keeping in mind: with $V = U$, the convective term
$(\bu, \alpha\ba\cdot\nabla\bu) = \tfrac12\int_\Omega \alpha\ba\cdot\nabla|\bu|^2$
integrates by parts to zero using $\nabla\cdot(\alpha\ba) = 0$ and the no-slip
data, and the pressure--divergence pair cancels by the same integration by
parts; what survives are the genuinely dissipative or reactive squares. Two
details matter downstream. First, only the \emph{deviatoric} part
$\ViscProj\nabla\bu$ of the gradient is dissipated---this is why proving that
\cref{oa:eq:TripleNorm} is a norm needs a Korn-type argument
(\cref{oa:lem:definiteness}) rather than being immediate. Second, everything a
stabilized method adds sits \emph{on top} of this identity: the ASGS addition is
indefinite and must be dominated (\cref{eq:stabdiag1}--\cref{eq:viscabsorb}); the
OSGS addition is a sum of squares for free (\cref{oa:eq:Diagonal})---but of the
\emph{wrong} squares, as discussed at \cref{oa:th:stability} below.
\end{pedagogy}

\subsubsection{Meshes, finite element spaces and interpolation}
\label{oa:sec:meshes}

The mesh nondegeneracy, the inverse estimate \cref{eq:InverseEstimateFiniteOrderNorm} and the
finite element spaces are the common ones of the analysis (\cref{sec:CommonSetting}), shared
with the ASGS setting of \cref{app:Continuity}; the OSGS route additionally uses the patch
structure and element{\hyp}size surrogates of \cref{H:patch} and the interpolation operators
recalled below.

We write $\mathcal{F} = \{\mathcal{T}_h\}_{h>0}$ for the (nondegenerate, shape-regular) mesh
family, $h_K \coloneqq \operatorname{diam}(K)$, $h \coloneqq \max_K h_K$; the continuous
degree-$k_u,k_p$ spaces $V_h \subset H^1(\Omega)^d$ and
$Q_h \subset H^1(\Omega)\cap C^0(\overline\Omega)$ (constants in $Q_h$, $k_p \le k_u+1$); and
the constrained subspaces $V_{h0} \coloneqq V_h \cap H^1_0(\Omega)^d$,
$Q_{h0} \coloneqq Q_h \cap Q_0$, $\Xhz \coloneqq V_{h0}\times Q_{h0}$.

\emph{Patch structure and element{\hyp}size surrogates} (\cref{H:patch}). We use the patch
quasi{\hyp}uniformity $\underline{C}\, h_{K} \le h_{K'} \le \overline{C}\, h_K$ for all
$K' \subset S_K$~\cite{AinsworthOden} and the continuous surrogates $\chi_1(\cdot,h),
\chi_2(\cdot,h)$ of \cref{H:patch}, with the element length taken as
$h_{K} \coloneqq \chi_1(\boldsymbol{x}_K, h)$ (as in~\cite[Lemma~1]{Codina2008FiniteEA}); this
last choice alters each stabilization parameter by a bounded factor ($\le \chi_0^2$) relative
to the raw diameter and is immaterial to every estimate, so we do not distinguish the two
notationally.

\emph{Interpolation operators.} For $v \in H^{r}(\Omega)$ with $r \ge 2$ (hence
$v \in C^0(\overline\Omega)$ for $d \le 3$) the nodal interpolant $\hat v_h$ satisfies
\begin{equation}
  \label{oa:eq:NodalInterp}
  \norm{v - \hat v_h}_{m,K} \le C_I\, h_K^{n-m}\norm{v}_{n,K},
  \qquad 0 \le m \le n \le \min(r,\, k+1),
\end{equation}
with $k$ the polynomial degree of the target space, and it preserves homogeneous Dirichlet
data (this is the estimate of the ASGS analysis, \cref{eq:interp} of \cref{app:Continuity},
in the norm form used below). For $v \in H^{r}(\Omega)$ with $r \ge 1$ the Scott--Zhang
quasi-interpolant~\cite{ScottZhang} satisfies the same estimate with $\norm{v}_{n,K}$
replaced by $\norm{v}_{H^n(S_K)}$ and also preserves homogeneous boundary values.

\begin{pedagogy}[Two interpolants, and why both are needed]
Background. The \emph{nodal} interpolant evaluates the target function at the
finite element nodes; this requires point values, hence continuity, which is why
$\bu \in H^{k_u+1}(\Omega)^d$ is paired with the Sobolev embedding
$H^{s} \hookrightarrow C^0$ for $s > d/2$ (here $k_u+1 \ge 2 > 3/2$ suffices for
$d \le 3$). Its virtues are locality and exact preservation of Dirichlet data
(boundary nodes receive the exact boundary values), which is what licenses the
boundary-term-free global integrations by parts in (I2) and (I5) of
\cref{oa:lem:interpolation}. The \emph{Scott--Zhang} quasi-interpolant replaces
point evaluation by local averages, so it is defined for any $H^1$ function---in
particular for the products $\alpha z$ fed to it in \cref{oa:lem:bestapprox}---at
the price that its error over $K$ is measured on the \emph{patch} $S_K$. That
price is the reason patches appear in this analysis at all: patch
quasi-uniformity (\cref{H:patch}) and patch equivalence of the parameters
(\cref{oa:lem:patch}) are exactly what converts patch-indexed sums back into
element-indexed sums with bounded overlap. Finally, the continuous surrogate
$\chi_1(\cdot,h)$ exists because the smoothing operation of \cref{oa:eq:SmulDef}
must evaluate ``the element size at a node'', and a node belongs to several
elements; \cite[Lemma~1]{Codina2008FiniteEA} builds a continuous field uniformly
equivalent to the local mesh size, and adopting it changes each $\tau$ by a
bounded factor only. None of this machinery has an ASGS counterpart: the
coercivity route never builds test functions from $\tau$, and its interpolation
lemma \cref{lem:continterp} uses only the nodal interpolant.
\end{pedagogy}

\subsubsection{Stabilization parameters}
\label{oa:sec:parameters}

The stabilization parameters $\tns$, $\tau_{1,K}$ and $\tau_{2,K}$ are those of the main text
(\cref{eq:TauNavierStokes,eq:Tau1Final,eq:Tau2Final}), shared with \cref{app:Continuity}, with
$\alpha_K \coloneqq \sup_K \alpha$, $|\ba|_{\infty,K} \coloneqq \norm{\ba}_{L^\infty(K)}$ and
design constants $c_1, c_2 > 0$ subject to \cref{H:design} of \cref{sec:CommonSetting}. Specific to the OSGS analysis are the
two dimensionless numbers and the $\boldsymbol{\tau}$-weighted broken products introduced now.
Expanding $\tau_2$,
\begin{equation}
  \label{oa:eq:Tau2Expanded}
  \tau_{2,K}
  = \frac{\nu}{\alpha_K}\Bigl( 1 + \frac{c_2}{c_1}\,\Reh \Bigr),
  \qquad
  \Reh \coloneqq \frac{|\ba|_{\infty,K}\, h_K}{\nu},
\end{equation}
so that $\tau_{2,K} \ge \nu/\alpha_K$ always. We will also use the elementwise mesh
Damk\"ohler number $\Dah \coloneqq \sigma h_K^2/(\alpha_K\nu)$. Weighted broken inner
products and norms are defined, for $i = 1,2$, by
\begin{equation}
  \label{oa:eq:TauInner}
  (f, g)_{\tau_i} \coloneqq \sum_K \tau_{i,K}\,(f, g)_{K},
  \qquad
  \norm{f}_{\tau_i} \coloneqq (f,f)_{\tau_i}^{1/2},
\end{equation}
for functions that need only be square integrable elementwise; note
$\norm{f}_{\tau_i} = \lVert \tau_i^{1/2} f\rVert_h$ in the broken-norm notation
of~the main text.

\begin{pedagogy}[Fixing the units the whole analysis is measured in]
This short subsection does one job: it fixes the currency in which every later
bound is paid. \cref{oa:eq:Tau2Expanded} rewrites the pressure parameter as
$\tau_{2,K} = (\nu/\alpha_K)\bigl(1 + (c_2/c_1)\Reh\bigr)$. Since the design
constants satisfy $c_1,c_2>0$ (\cref{H:design}) and $\Reh\ge 0$, the bracket is at
least $1$; hence $\tau_{2,K}\ge \nu/\alpha_K$ always---a floor the parameter
inequalities lean on repeatedly. The two dimensionless numbers each compare two
competing effects \emph{on one element, at that element's own length} $h_K$: the
mesh Reynolds number $\Reh = |\ba|_{\infty,K}h_K/\nu$ weighs advection against
diffusion, and the mesh Damk\"ohler number $\Dah = \sigma h_K^2/(\alpha_K\nu)$
weighs reaction against diffusion. Watch $\Dah$: it is the Damk\"ohler factor that
will price almost every convergence constant to come, all the way to the main
estimate \cref{oa:eq:MainResult}. Finally, ``broken'' in \cref{oa:eq:TauInner}
means summed element by element with no inter-element continuity demanded (the
inputs need only be square integrable on each $K$), each term weighted by the
local $\tau_{i,K}$; the identity $\norm{f}_{\tau_i}=\lVert\tau_i^{1/2}f\rVert_h$
just records that this weight can be folded into the function. These
$\boldsymbol{\tau}$-weighted norms are the natural currency here because the
stabilization terms are themselves $\tau$-weighted.
\end{pedagogy}

\subsubsection{Projections and the analyzed OSGS method}
\label{oa:sec:method}

Let $\Pone$ denote the $\tau_1$-weighted $L^2$ projection onto the
\emph{unconstrained} velocity space $V_h$ (acting componentwise), $\Ptwo$ the
$\tau_2$-weighted projection onto the unconstrained pressure space $Q_h$, and
$\Poneo \coloneqq I - \Pone$, $\Ptwoo \coloneqq I - \Ptwo$ the complementary
(fluctuation) operators. We will also need $\Ponez$, the $\tau_1$-weighted projection
onto the \emph{constrained} space $V_{h0}$, which enters only the stability proof
through the special test function, exactly as in~\cite{Codina2008FiniteEA}. Projecting onto the
unconstrained spaces in the method matches both the implementation described
in~the main text (footnote to the OSGS system there) and the discussion
in~\cite[Remark~1]{Codina2008FiniteEA}: projecting onto the constrained spaces would simplify
the analysis (assumption \cref{H:projection} would not be needed) but produces spurious numerical
boundary layers.

The \emph{analyzed} OSGS method reads: find $U_h \in \Xhz$ such that
\begin{equation}
  \label{oa:eq:Method}
  B_{\mathrm{osgs}}(\ba; U_h, V_h) = L(V_h) \coloneqq (\bff, \bv_h)
  \qquad \forall\, V_h \in \Xhz,
\end{equation}
with
\begin{equation}
  \label{oa:eq:Bosgs}
  B_{\mathrm{osgs}}(\ba; U, V)
  \coloneqq
  B(\ba; U, V) + S_1(U, V) + S_2(U, V),
\end{equation}
\begin{equation}
  \label{oa:eq:Sdefs}
  \begin{aligned}
    S_1(U, V) &\coloneqq \bigl( \Poneo[\alpha X(U)],\, \alpha X(V) \bigr)_{\tau_1}, \\
    S_2(U, V) &\coloneqq \bigl( \Ptwoo[\nabla\cdot(\alpha\bu)],\, \nabla\cdot(\alpha\bv) \bigr)_{\tau_2}.
  \end{aligned}
\end{equation}
Since $(\Poneo z, \Pone w)_{\tau_1} = 0$ for all $z, w$, both stabilization terms are
symmetric and positive semidefinite:
\begin{equation}
  \label{oa:eq:Ssym}
\begin{aligned}
  S_1(U, V) &= \bigl( \Poneo[\alpha X(U)],\, \Poneo[\alpha X(V)] \bigr)_{\tau_1},
  \qquad \\
  S_2(U, V) &= \bigl( \Ptwoo[\nabla\cdot(\alpha\bu)],\, \Ptwoo[\nabla\cdot(\alpha\bv)] \bigr)_{\tau_2}.
\end{aligned}
\end{equation}
In particular, with \cref{oa:eq:CoercivityIdentity},
\begin{equation}
  \label{oa:eq:Diagonal}
  B_{\mathrm{osgs}}(\ba; U_h, U_h)
  = 2\nu\,\bigl\lVert \alpha^{1/2}\ViscProj\nabla\bu_h \bigr\rVert^2
  + \norm{\sigma^{1/2}\bu_h}^2
  + \varepsilon\norm{p_h}^2
  + \norm{\Poneo \xi_h}_{\tau_1}^2
  + \norm{\Ptwoo \delta_h}_{\tau_2}^2,
\end{equation}
where here and below we abbreviate
\begin{equation}
  \label{oa:eq:XiDelta}
  \xi_h \coloneqq \alpha X(U_h) = \alpha\bigl(\ba\cdot\nabla\bu_h + \nabla p_h\bigr),
  \qquad
  \delta_h \coloneqq \nabla\cdot(\alpha\bu_h).
\end{equation}

\begin{pedagogy}[Reading the fluctuation stabilization]
$\Pone$ is the $\tau_1$-weighted $L^2$ projection onto the velocity space $V_h$: it returns the finite element field closest to its argument in the weighted product $(\cdot,\cdot)_{\tau_1}$ of \cref{oa:eq:TauInner}; $\Ptwo$ does the same in pressure. The complements $\Poneo = I - \Pone$ and $\Ptwoo = I - \Ptwo$ therefore extract the \emph{fluctuation}---the part of a field the mesh cannot represent. The method \cref{oa:eq:Method} adds the terms $S_1, S_2$ of \cref{oa:eq:Sdefs}, which weigh the fluctuation of the first-order residual ($\xi_h$ in momentum, $\delta_h$ in mass; these are just shorthand, \cref{oa:eq:XiDelta}) against the test slot. Since a fluctuation is $\tau_i$-orthogonal to every projected field, $(\Poneo z, \Pone w)_{\tau_1} = 0$, only the fluctuation of the test slot survives; this is \cref{oa:eq:Ssym}, exhibiting $S_1, S_2$ as symmetric and positive semidefinite. Taking $V = U_h$ then gives \cref{oa:eq:Diagonal}: the Galerkin energy \cref{oa:eq:CoercivityIdentity} plus two fluctuation squares, gained for free---though the ``wrong'' squares, as \cref{oa:th:stability} shows.
\end{pedagogy}

The next elementary observation locates the analyzed method precisely relative to the
implemented one, in which the orthogonal projection is applied to the \emph{full}
residual and the test slot is the full adjoint.

\begin{lemma}[Exact annihilations]
\label{oa:lem:annihilation}
Let $\sigma$ and $\varepsilon$ be constant. Then, for all $U_h, V_h \in \Xhz$:
\begin{enumerate}[label=\textup{(\roman*)}, itemsep=0.2ex]
  \item $\Poneo(\sigma\bu_h) = \boldsymbol{0}$ and $\Ptwoo(\varepsilon p_h) = 0$
        (reactive and compressibility parts of the residual have no fluctuation);
  \item $\bigl( \Poneo z,\, \sigma\bv_h \bigr)_{\tau_1} = 0$ and
        $\bigl( \Ptwoo w,\, \varepsilon q_h \bigr)_{\tau_2} = 0$ for any $z, w$
        (reactive and compressibility parts of the adjoint test slot are invisible to
        the fluctuations).
\end{enumerate}
Consequently, the formulation obtained from the implemented OSGS method by projecting
the full residual and testing with the full adjoint differs from
\cref{oa:eq:Method}--\cref{oa:eq:Sdefs} only through (a) the fluctuations of the viscous
residual term $2\nabla\cdot(\alpha\nu\ViscProj\nabla\bu_h)$ and of the force $\bff$,
and (b) the viscous part of the adjoint test slot---exactly the Method~I simplification
of~\cite{Codina2008FiniteEA}.
\end{lemma}

\begin{proof}
For (i), $\sigma\bu_h \in V_h$ and $\varepsilon p_h \in Q_h$, and the fluctuation of
any element of the projection space vanishes. For (ii), $\sigma\bv_h \in V_h$ and
$\varepsilon q_h \in Q_h$, and fluctuations are $\tau_i$-orthogonal to the projection
spaces. The final statement follows by expanding the implemented stabilization
term by linearity.
\end{proof}

\begin{pedagogy}[Why the projection erases the reactive term---in both slots]
This short lemma is the hinge the whole appendix turns on, so read it slowly. One
fact does all the work: the projection $\Pone$ reproduces anything already in $V_h$,
so its fluctuation $\Poneo = I - \Pone$ kills every element of $V_h$. Now if $\sigma$
is \emph{constant}, then $\sigma\bu_h$ is just a fixed multiple of $\bu_h \in V_h$ and
hence lies in $V_h$ itself; therefore $\Poneo(\sigma\bu_h) = \boldsymbol{0}$, and
identically $\Ptwoo(\varepsilon p_h) = 0$---that is part~(i). Part~(ii) is the same
fact seen from the \emph{test} slot: fluctuations are $\tau_i$-orthogonal to the
projection space by construction, and $\sigma\bv_h$, $\varepsilon q_h$ live in that
space, so both pairings vanish. Constancy is essential---a variable $\sigma$ would
carry $\sigma\bu_h$ out of $V_h$. Significance: the reactive and compressibility
pieces drop from residual \emph{and} test slots, so---unlike ASGS---the OSGS route
never renormalizes $\sigma$ to $\sigmatilde$. That is exactly why the working norm
keeps the full $\norm{\sigma^{1/2}\bv}$ and why coercivity must give way to the
inf--sup argument of \cref{oa:th:stability} (see \cref{oa:rem:mechanism}).
\end{pedagogy}

\begin{remark}[Analyzed versus implemented formulation]
\label{oa:rem:analyzed}
For later reference we list the exact distance between \cref{oa:eq:Method} and the OSGS
method as implemented in~the main text. (i)~\emph{Weighted versus plain projection}:
the analysis uses the $\boldsymbol{\tau}$-weighted projections of
\cref{oa:eq:TauInner}; the implementation uses the plain $L^2$ projection. The two
coincide when $\tau_i$ is globally constant; otherwise the weighted product is the one
for which the best-approximation property of \cref{oa:lem:bestapprox} holds, which is
essential to the consistency estimate (cf.~\cite[Remark~4]{Codina2008FiniteEA}).
(ii)~\emph{First-order truncation}: residual and test slots are truncated to their
first-order parts; by \cref{oa:lem:annihilation} the only genuinely neglected pieces are
the viscous fluctuation, the force fluctuation, and the viscous test term. For $k_u = 1$
and elementwise-constant $\alpha\nu$ the viscous residual vanishes on element interiors;
with variable $\alpha$ it contributes $2\nu(\ViscProj\nabla\bu_h)\nabla\alpha$, which is
small under the resolution condition of \cref{H:porosity}. (iii)~\emph{Converged versus lagged
projection}: the analysis takes the projection equation solved exactly (the coupled
system), whereas the implementation lags the projection along the nonlinear iteration.
(iv)~\emph{Unconstrained projection space}: analyzed and implemented formulations
agree (both project onto $V_h\times Q_h$ without Dirichlet constraints); note that the
display defining the OSGS projection in~the main text shows the projection onto
$\Xhz$, while the accompanying footnote states the implemented choice, which is the one
analyzed here and the one recommended by~\cite[Remark~1]{Codina2008FiniteEA}.
(v)~\emph{Elementwise-constant versus pointwise parameters}: as for the ASGS analysis
of~the main text, $\tau_{i}$ is elementwise constant in the analysis and pointwise
variable in the computations.
\end{remark}

\subsection{Standing assumptions and notation}
\label{oa:sec:hypotheses}

Throughout, $C$ denotes a generic positive constant, independent of $h$, $\nu$,
$\sigma$, $\varepsilon$ and of $\alpha$ and $\ba$ except through the constants named
below; its value may change between appearances. The analysis is carried out under the
standing assumptions collected in \cref{sec:CommonSetting}: the assumptions
\cref{H:mesh}--\cref{H:regularity} common to both methods, and the OSGS{\hyp}specific
assumptions \cref{H:design}--\cref{H:advectionsmooth}---namely the design constants
\cref{H:design} (with the compressibility bound $\varepsilon\tau_{2,K}\le C_2$), the projection
stability \cref{H:projection} (\cref{eq:ProjectionStability}), and the technical strengthenings
\cref{H:patch} (patch structure and element{\hyp}size surrogates), \cref{H:porositysmooth}
(porosity smoothness) and \cref{H:advectionsmooth} (advection smoothness) of the common mesh,
porosity and advection assumptions. The strengthenings \cref{H:porositysmooth,H:advectionsmooth}
enter only the consistency estimate. The ASGS analysis of \cref{app:Continuity} instead requires
the interior{\hyp}face jump condition \cref{H:jump}; the two hypothesis sets are therefore not
nested (see \cref{oa:rem:dispensed}). We use the abbreviations $\Ponez z_h$ and $\Poneo z_h$ of
\cref{oa:sec:method} for the constrained projection and the fluctuation appearing in
\cref{H:projection}, and, for \cref{H:porositysmooth}, $C_{\alpha,m}$ ($m\ge2$) as defined there,
setting $C_{\alpha,1}\coloneqq 1+\norm{\nabla\alpha}_{L^\infty(\Omega)}/\alpha_0$ for the
first{\hyp}order case.

\begin{pedagogy}[A map of the hypotheses]
Before the technical work, a one-glance grouping of the assumptions this
subsection collects. \Cref{H:mesh}--\cref{H:regularity} are the \emph{common}
ones, shared with the ASGS analysis: mesh nondegeneracy, coefficient and
solution regularity, and the finite element spaces. The rest are
\emph{OSGS-specific}, \cref{H:design}--\cref{H:advectionsmooth}. Among these,
note how little some are asked to do: \cref{H:porositysmooth} and
\cref{H:advectionsmooth}---extra derivatives of $\alpha$ and $\ba$---enter
\emph{only} the consistency estimate, where the Leibniz rule differentiates the
products handed to the projection, and play no part in stability. The single
assumption that is \emph{not} an automatic consequence of the common
mesh-and-data setting is the projection stability \cref{H:projection}, a genuine
property of the finite element pair; the stability proof draws on it twice
(Steps~2 and~5 of \cref{oa:th:stability}). The ASGS route pays instead with the
interior-face jump condition \cref{H:jump}, so the two hypothesis sets are not
nested---the precise anatomy of that non-nesting, and why \cref{H:jump} is
nonetheless dispensable here, is deferred to \cref{oa:rem:dispensed}.
\end{pedagogy}

\subsection{Technical results}
\label{oa:sec:technical}

\subsubsection{Parameter inequalities}

Items (P1)--(P3) below are elementary parameter bounds shared with the ASGS analysis: they are
recalled from \cref{lem:parameters} of \cref{app:Continuity}---in the notation of this appendix,
through $\varphi_1=\alpha_K\tns^{-1}$---and are not re-derived. Items (P4)--(P6) and the
inverse{\hyp}estimate inequalities (K1)--(K3) are specific to the OSGS route (they feed the
special test function of \cref{oa:th:stability} and the Damk\"ohler-weighted conversions of the
convergence proof).

\begin{lemma}[Parameter inequalities]
\label{oa:lem:parameters}
Under \cref{H:design}, the parameters \cref{eq:Tau1Final,eq:Tau2Final} satisfy, on every element $K$
(element subscripts omitted inside the formulas):
\begin{align}
  &\sigma\tau_1 \le 1,
  \qquad
  \alpha_K \tau_1 \le \tns,
  \qquad
  \varepsilon \le C_2\, \tau_2^{-1},
  \tag{P1}\label{oa:eq:P1}\\
  &\tau_1\,\tau_2 \le \frac{h_K^2}{c_1\,\alpha_K^2},
  \tag{P2}\label{oa:eq:P2}\\
  &h_K^2\,\tns^{-1} = c_1\,\alpha_K\,\tau_2,
  \quad\implies\quad
  \bigl(\alpha_K\,\tns^{-1}\bigr)^{1/2} = \sqrt{c_1}\,\frac{\alpha_K}{h_K}\,\tau_2^{1/2},
  \tag{P3}\label{oa:eq:P3}\\
  &\nu \le \alpha_K\,\tau_2,
  \qquad
  \tau_1^{1/2}\,|\ba|_{\infty,K} \le \frac{\sqrt{c_1}}{c_2}\,\tau_2^{1/2},
  \tag{P4}\label{oa:eq:P4}\\
  &\alpha_K^2\,\tau_1\,\tau_2 \;\ge\; \frac{h_K^2}{c_1 + \Dah},
  \quad\implies\quad
  \tau_2^{-1/2} \le \bigl(c_1 + \Dah\bigr)^{1/2}\,\frac{\alpha_K}{h_K}\,\tau_1^{1/2},
  \tag{P5}\label{oa:eq:P5}\\
  &\sigma^{1/2} \;\le\; \Dah^{1/2}\,\frac{\alpha_K}{h_K}\,\tau_2^{1/2}.
  \tag{P6}\label{oa:eq:P6}
\end{align}
Moreover, with $\gamma$ as in \cref{H:design},
\begin{align}
  &C_{\mathrm{inv}}\,\frac{(\alpha_K\nu)^{1/2}}{h_K} \le \frac{1}{\gamma}\,\tau_1^{-1/2},
  \tag{K1}\label{oa:eq:K1}\\
  &C_{\mathrm{inv}}\,\frac{\alpha_K\,|\ba|_{\infty,K}}{h_K} \le \frac{1}{\gamma}\,\tau_1^{-1},
  \tag{K2}\label{oa:eq:K2}\\
  &C_{\mathrm{inv}}\,\frac{\alpha_K\,(\tau_1\tau_2)^{1/2}}{h_K} \le \frac{1}{\gamma}.
  \tag{K3}\label{oa:eq:K3}
\end{align}
\end{lemma}

\begin{pedagogy}[Deriving (P5) and (P6): substitute the definitions, then bound once]
These are the only two conversions that carry a Damk\"ohler premium, and each is
just a substitution followed by a single inequality. For \cref{oa:eq:P5}, insert the
relation $\tau_1^{-1}=\alpha_K\tns^{-1}+\sigma$ and, from \cref{oa:eq:P3}, the
rewriting $\tau_2=h_K^2/(c_1\alpha_K\tns)$:
\[
  \alpha_K^2\tau_1\tau_2
  = \frac{\alpha_K^2}{\alpha_K\tns^{-1}+\sigma}\cdot\frac{h_K^2}{c_1\alpha_K\tns}
  = \frac{h_K^2}{c_1\bigl(1+\sigma\tns/\alpha_K\bigr)}.
\]
So far this is an identity. Only now does an inequality enter, the bound
$\tns\le h_K^2/(c_1\nu)$ that the proof invokes: it makes
$\sigma\tns/\alpha_K\le\sigma h_K^2/(c_1\alpha_K\nu)=\Dah/c_1$, so the denominator
is at most $c_1+\Dah$, and---because it sits in a denominator---the whole quantity
is at least $h_K^2/(c_1+\Dah)$. For \cref{oa:eq:P6}, the equality
$\sigma^{1/2}=\Dah^{1/2}(\alpha_K\nu)^{1/2}/h_K$ is nothing but the definition
$\Dah=\sigma h_K^2/(\alpha_K\nu)$ solved for $\sigma^{1/2}$; then the first
inequality of \cref{oa:eq:P4}, $\nu\le\alpha_K\tau_2$, replaces $(\alpha_K\nu)^{1/2}$
by $\alpha_K\tau_2^{1/2}$.
\end{pedagogy}

\begin{proof}
The first two inequalities of \cref{oa:eq:P1} ($\sigma\tau_1\le1$ and $\alpha_K\tau_1\le\tns$)
and the identity \cref{oa:eq:P3} are established in \cref{lem:parameters} of
\cref{app:Continuity} (with $\varphi_1=\alpha_K\tns^{-1}$) and are not re-derived here; the
third inequality of \cref{oa:eq:P1}, $\varepsilon\le C_2\tau_2^{-1}$, is \cref{H:design}, and
\cref{oa:eq:P2} follows on combining $\alpha_K\tau_1\le\tns$ with the definition of $\tau_2$.

We prove the OSGS{\hyp}specific items:
\begin{itemize}[leftmargin=*]
  \item \cref{oa:eq:P4}: The first inequality is \cref{oa:eq:Tau2Expanded}; for the second, $\tau_1 \le \tns/\alpha_K \le h_K/(c_2\alpha_K|\ba|_{\infty,K})$ gives $\tau_1|\ba|_{\infty,K}^2 \le |\ba|_{\infty,K}h_K/(c_2\alpha_K)$, while \cref{oa:eq:Tau2Expanded} gives $\tau_2 \ge (c_2/c_1)\,|\ba|_{\infty,K}h_K/\alpha_K$; divide.
  \item \cref{oa:eq:P5}: Using $\tns \le h_K^2/(c_1\nu)$,
\[
\begin{aligned}
  \alpha_K^2\tau_1\tau_2
  &= \frac{\alpha_K^2}{\alpha_K\tns^{-1}+\sigma}\cdot\frac{h_K^2}{c_1\alpha_K\tns}
  = \frac{h_K^2}{c_1\bigl(1 + \sigma\tns/\alpha_K\bigr)} \\
  &\ \ge\ \frac{h_K^2}{c_1\bigl(1 + \sigma h_K^2/(c_1\alpha_K\nu)\bigr)}
  \ =\ \frac{h_K^2}{c_1 + \Dah}.
\end{aligned}
\]
  \item \cref{oa:eq:P6}: $\sigma^{1/2} = \Dah^{1/2}(\alpha_K\nu)^{1/2}/h_K \le \Dah^{1/2}\alpha_K\tau_2^{1/2}/h_K$ by the first inequality of \cref{oa:eq:P4}.
  \item \cref{oa:eq:K1}: $\tau_1^{-1} \ge \alpha_K\tns^{-1} \ge c_1\alpha_K\nu/h_K^2$, so $C_{\mathrm{inv}}(\alpha_K\nu)^{1/2}/h_K \le (C_{\mathrm{inv}}/\sqrt{c_1})\tau_1^{-1/2} \le \gamma^{-1}\tau_1^{-1/2}$ by \cref{H:design}.
  \item \cref{oa:eq:K2}: $\tau_1^{-1} \ge \alpha_K\tns^{-1} \ge c_2\alpha_K|\ba|_{\infty,K}/h_K$ and $c_2 \ge \gamma C_{\mathrm{inv}}$.
  \item \cref{oa:eq:K3}: Combine \cref{oa:eq:P2} with $c_1 \ge \gamma^2C_{\mathrm{inv}}^2$.
\end{itemize}
\end{proof}

\begin{pedagogy}[Reading (K1)--(K3): a two-step lower bound in $\tau$-currency]
Each $(\mathrm{K})$-inequality is proved by the same elementary move: bound
$\tau_1^{-1}$ from below, then let a design constant of \cref{H:design} absorb the
rest. Take \cref{oa:eq:K1}. First drop the nonnegative reaction,
$\tau_1^{-1}=\alpha_K\tns^{-1}+\sigma\ge\alpha_K\tns^{-1}$; then use
$\tns^{-1}\ge c_1\nu/h_K^2$ (the same bound $\tns\le h_K^2/(c_1\nu)$ as above) to get
$\tau_1^{-1}\ge c_1\alpha_K\nu/h_K^2$. Taking square roots and dividing,
$C_{\mathrm{inv}}(\alpha_K\nu)^{1/2}/h_K\le(C_{\mathrm{inv}}/\sqrt{c_1})\,\tau_1^{-1/2}$,
and the choice $c_1\ge\gamma^2 C_{\mathrm{inv}}^2$ makes $C_{\mathrm{inv}}/\sqrt{c_1}\le1/\gamma$.
\Cref{oa:eq:K2} is identical with the convective floor
$\tns^{-1}\ge c_2|\ba|_{\infty,K}/h_K$ and $c_2\ge\gamma C_{\mathrm{inv}}$;
\cref{oa:eq:K3} instead feeds \cref{oa:eq:P2} into
$\alpha_K(\tau_1\tau_2)^{1/2}/h_K\le1/\sqrt{c_1}$, whence the same $c_1\ge\gamma^2 C_{\mathrm{inv}}^2$
closes it. In every case it is the design constant, not the geometry, that buys the
factor $1/\gamma$---which is exactly why \cref{H:design} is stated the way it is.
\end{pedagogy}

\begin{pedagogy}[Reading (P1)--(P6) and (K1)--(K3) as exchange rates]
A useful way to hold this lemma in mind: the convergence proof is a sequence of
currency conversions. Interpolation errors arrive in raw sizes ($h^{k+1}$ in
$L^2$, $h^{k}$ in $H^1$, per element); the test norm only pays in the five
weighted currencies of \cref{oa:eq:TripleNorm}; each estimate of
\cref{oa:lem:consistency,oa:lem:interpolation} converts the former into the
latter at a posted rate. (P3) is an \emph{exact} rate: the weight
$(\alpha_K\tns^{-1})^{1/2}$, ubiquitous because
$\tau_1^{-1} = \alpha_K\tns^{-1}+\sigma$, equals
$\sqrt{c_1}\,(\alpha_K/h_K)\tau_2^{1/2}$ identically. (P4) converts convective
weights into $\tau_2$-currency at rate $\sqrt{c_1}/c_2$. The only two rates that
cost a Damk\"ohler premium are (P5), which prices the pressure weight
$\tau_2^{-1/2}$, and (P6), which prices the raw reaction weight $\sigma^{1/2}$;
\emph{every} occurrence of $(c_1+\Dah)^{1/2}$ in this appendix is one
application of (P5) or (P6) and nothing else, so the final constant can be
audited by counting their uses ((I2)--(I5) of \cref{oa:lem:interpolation} and
the $\varepsilon$-part of \cref{oa:lem:consistency}). For the ASGS-fluent
reader: \cref{lem:parameters} plays the same role there, with
\cref{lem:parameters}\ref{P:sigmatilde} (the $\sigmatilde$ identities) as the
star; (P4)--(P6) have no ASGS counterpart because the ASGS norm never pays in
the currencies $\sigma^{1/2}$ or $\tau_2^{-1/2}$---its reactive currency is the
damped $\sigmatilde^{1/2}$, which \cref{eq:sigmatilde} keeps commensurate with
$(\alpha_K/h_K)\tau_2^{1/2}$. The (K)-inequalities are different in kind: they
are \emph{calibrations} of the inverse estimate against the design constants,
each stating that a specific inverse-estimate factor costs at most $1/\gamma$
in $\tau$-currency; they exist so that in the stability proof every
perturbation term appears with a coefficient that \cref{H:design} can make as
small as needed. Background, for completeness: inverse estimates such as
\cref{eq:InverseEstimateFiniteOrderNorm} hold because on the fixed reference
element all norms on the finite-dimensional polynomial space are equivalent,
and mapping to a physical element of size $h_K$ converts each derivative into a
factor $h_K^{-1}$; they are properties of \emph{discrete} functions only, which
is why they may be applied to $\bv_h$ or $\bw_h$ but never to $\bu$ or
$\hat{\boldsymbol e}_u$.
\end{pedagogy}

\subsubsection{Patch equivalence and the smoothing operation}

\begin{lemma}[Patch equivalence]
\label{oa:lem:patch}
Under \cref{H:mesh}, \cref{H:porosity}, \cref{H:advection} and \cref{H:patch} there are constants
$0 < \underline{c} \le \overline{c}$, independent of $h$, such that for $i = 1,2$,
\begin{equation}
  \label{oa:eq:PatchEquivalence}
  \underline{c}\;\tau_{i,K} \le \tau_{i,K'} \le \overline{c}\;\tau_{i,K},
  \qquad
  \underline{c}\;\alpha_{K} \le \alpha_{K'} \le \overline{c}\;\alpha_{K},
  \qquad \forall\, K' \subset S_K,\ \forall\, K.
\end{equation}
\end{lemma}

\begin{proof}
Patch quasi-uniformity gives $h_{K'} \eqsim h_K$. Since $\alpha$ is Lipschitz and
bounded below by $\alpha_0 > 0$, and $\operatorname{diam}(S_K) \le C h_K$, one has
$|\alpha_{K'} - \alpha_K| \le C h_K\norm{\nabla\alpha}_{L^\infty(S_K)} \le
C\,C_{\nabla\alpha}\,\max_{K''\subset S_K}\alpha_{K''}$, from which the uniform equivalence of
the $\alpha_{K}$ over the patch follows for $C\,C_{\nabla\alpha}$ bounded (and, in general, from
$\alpha_{K'}/\alpha_K \le \alpha_\infty/\alpha_0$; the sharper route through $C_{\nabla\alpha}$
keeps the constant independent of $\alpha_\infty/\alpha_0$). Likewise
$\ba \in C^0(\overline\Omega)$ implies $|\ba|_{\infty,K'} \le |\ba|_{\infty,K} +
\omega_{\ba}(Ch_K)$, where $\omega_{\ba}$ is the modulus of continuity, and the
resulting perturbations of $\tns^{-1}$, $\tau_1^{-1}$ and $\tau_2$ are controlled as
in the proof of \cref{oa:lem:smoothing} below. The claim follows from the explicit
formulas \cref{eq:Tau1Final,eq:Tau2Final}.
\end{proof}

\begin{pedagogy}[Why patches; a bridge to the ASGS jump condition]
This lemma says the parameters are \emph{locally flat}: across a patch,
$\tau_i$ and $\alpha_K$ vary by bounded factors, uniformly in $h$. It is
consumed twice: to re-sum Scott--Zhang patch errors into element sums
(\cref{oa:lem:bestapprox}), and, in nodal form, in the smoothing lemma below.
For the reader coming from \cref{app:Continuity}, one identification is worth
making explicit. Restricted to two elements sharing a face, the comparability
proved here is precisely the content of the jump condition \cref{H:jump} used
in the ASGS continuity proof (there through \cref{lem:jump}): both assert
facewise comparability of $\varphi_1 = \alpha_K\tns^{-1}$, and $\sigma$ cancels
in the jump of $\tau_1^{-1} = \varphi_1 + \sigma$, so neither constrains the
reaction. The proof here derives that comparability from the standing
hypotheses---the porosity-resolution argument for $\alpha_K$, mesh
nondegeneracy for $h_K$, continuity of $\ba$ with the viscous term as a floor
for the advective part. The same regularity substrate is therefore available to
both analyses; the ASGS text states it as a hypothesis where needed, the OSGS
text consumes it as a lemma (see also the commentary at
\cref{oa:rem:dispensed}). Note the quiet appearance of the floor $\nu > 0$: the
advective oscillation is absorbed \emph{relative to} the viscous part of
$\tns^{-1}$; the same floor returns, more visibly, as the $\nu$-dependence of
$\psi(h)$ in \cref{oa:lem:smoothing} and of the threshold $h_0$ in
\cref{oa:th:stability}.
\end{pedagogy}

Following~\cite{Codina2008FiniteEA}, for $\tau \in \{\tau_1, \tau_2\}$ and a finite element
function $v_h$ (scalar or vector) we define the discontinuous and the smoothed
products
\begin{equation}
  \label{oa:eq:SmulDef}
  (\tau \pmul v_h)\big|_K \coloneqq \tau_{K}\, v_h\big|_K,
  \qquad
  (\tau \smul v_h)\big|_K \coloneqq \sum_{a \in N_K} N_a(\boldsymbol{x})\big|_K\,
  \bar\tau_{a}\, v_a,
\end{equation}
where $N_a$ are the nodal shape functions, $v_a$ the nodal values of $v_h$, and the
nodal parameters $\bar\tau_{i,a}$ are obtained by evaluating the formulas
\cref{eq:Tau1Final,eq:Tau2Final} with $h_K$ replaced by $\chi_1(\boldsymbol{x}_a, h)$,
$\alpha_K$ by $\alpha(\boldsymbol{x}_a)$ and $|\ba|_{\infty,K}$ by
$|\ba(\boldsymbol{x}_a)|$. Thus $\tau \smul v_h$ is a continuous finite element
function, whereas $\tau \pmul v_h$ is not.

\begin{pedagogy}[Why a smoothed product must exist at all]
The natural test velocity for harvesting the momentum residual would be
$\tau_1 \pmul \Ponez\xi_h$---multiply the projected residual by the parameter,
elementwise. But $\tau_1$ is elementwise constant, so $\tau_1 \pmul v_h$ jumps
across interior faces and is \emph{not} an element of $V_h$: a conforming
method requires test functions in $H^1$, and for piecewise polynomials
$H^1$-conformity is interelement continuity (a jump would put a surface Dirac
into the distributional gradient). The smoothed product $\smul$ repairs this in
the only way available: rebuild $\tau v_h$ as a genuine finite element function
by sampling parameter and function at the \emph{nodes} and expanding in the
shape functions. This is legal (the result lies in $V_h$ by construction and
preserves nodal zeros, hence homogeneous boundary values) but no longer exact:
$\smul \ne \pmul$, and the entire purpose of \cref{oa:lem:smoothing} is to
certify that the discrepancy is a uniform relative factor $\psi(h) \to 0$. The
ASGS analysis has no counterpart to any of this---coercivity tests with $U_h$
itself, and no $\tau$-weighted test function is ever built. The comparison
worth internalizing: the discontinuity of the elementwise $\tau$ taxes
\emph{both} analyses, each at the point where its own manoeuvre touches
faces---the ASGS proof pays through the $[\tau_1]$ face terms of its elementwise
integrations by parts (\cref{eq:jumptotal}, \cref{eq:IIIface}), the OSGS proof
pays through the smoothing discrepancy $\psi(h)$ here.
\end{pedagogy}

\begin{lemma}[Smoothing; adaptation of~{\cite[Lemma~2]{Codina2008FiniteEA}}]
\label{oa:lem:smoothing}
Under \cref{H:mesh}, \cref{H:porosity}, \cref{H:advection} and \cref{H:patch} there is a function $\psi(h) \to 0$
as $h \to 0$ such that, for $\tau \in \{\tau_1, \tau_2\}$ and any finite element
function $v_h$,
\begin{equation}
  \label{oa:eq:Smoothing}
  \norm{\tau\pmul v_h - \tau\smul v_h}_{K} \le \tau_K\,\psi(h)\,\normK{v_h},
  \quad\implies\quad
  \norm{\tau\smul v_h}_{K} \le (1+\psi(h))\,\tau_K\,\normK{v_h}.
\end{equation}
If, in addition, $\ba$ is Lipschitz on $\overline\Omega$, then $\psi(h) = O(h)$.
\end{lemma}

\begin{proof}
As in~\cite[Lemma~2]{Codina2008FiniteEA}, the equivalence
$C_1 h_K^{d/2}\norm{w_h}_{L^\infty(K)} \le \normK{w_h} \le
C_2 h_K^{d/2}\norm{w_h}_{L^\infty(K)}$ for finite element functions on nondegenerate
meshes reduces \cref{oa:eq:Smoothing} to the statement
$\max_{a\in N_K} |\bar\tau_a/\tau_K - 1| \le C\,\psi(h)$ with $\psi(h)\to 0$
uniformly in $K$. Write $\Delta(\cdot)$ for the difference between the nodal and the
elemental value of a quantity. For $\tau_1$,
$\tau_{1}|\Delta(\tau_1^{-1})| \le
\tau_1\bigl[\,|\Delta\alpha|\,\tns^{-1}
+ \alpha_\infty\, c_2\,|\Delta|\ba||/h_K
+ \alpha_K\, c_1\nu\,|\Delta(h^{-2})| + \alpha_K c_2 |\ba|_{\infty,K}|\Delta(h^{-1})|\,\bigr]$.
Using $\tau_1 \le \tns/\alpha_0$ and $\tau_1 \le h_K^2/(c_1\alpha_0\nu)$,
the first term is bounded by $\omega_\alpha(Ch)/\alpha_0$, the second by
$\alpha_\infty\,(c_2/(c_1\alpha_0\nu))\,h\,\omega_{\ba}(Ch)$, and the last two by
$C\,\omega_{\chi}(h)$, where $\omega_\alpha, \omega_{\ba}$ are the moduli of
continuity of $\alpha$ and $\ba$ and $\omega_\chi(h)\to 0$ quantifies the relative
oscillation of $\chi_1(\cdot,h)$ over an element, which vanishes as $h \to 0$ by the
construction of~\cite[Lemma~1]{Codina2008FiniteEA}. All right-hand sides tend to zero
uniformly; here the strict positivity $\alpha_0 > 0$ and $\nu > 0$ are used, and for
Lipschitz data all moduli are $O(h)$. Then
$|\bar\tau_{1,a}/\tau_{1,K} - 1| = \bar\tau_{1,a}\,|\Delta(\tau_1^{-1})|
\le (1+o(1))\,\tau_{1,K}|\Delta(\tau_1^{-1})|$, which proves the claim for $\tau_1$;
the same computation applied to the explicit form \cref{oa:eq:Tau2Expanded} of $\tau_2$
(a continuous, positive expression in $\alpha$, $|\ba|$ and $h$, bounded away from
zero by $\nu/\alpha_\infty$) proves it for $\tau_2$.
\end{proof}

\begin{pedagogy}[What $\psi(h)$ measures, and what degrades it]
The proof has two moves. First, for finite element functions the $L^2$ and
$L^\infty$ norms on one element are equivalent up to the volume factor
$h_K^{d/2}$ (finite dimensionality again---the same mechanism as the inverse
estimates), which reduces a norm statement to a statement about \emph{nodal
versus elemental parameter values}. Second, the relative deviation of
$\tau^{-1}$ is expanded source by source: porosity contributes its modulus of
continuity over the floor $\alpha_0$; advection contributes $h\,\omega_{\ba}(Ch)$
over the viscous floor $c_1\nu$; the surrogate mesh-size field contributes its
own vanishing relative oscillation. Hence $\psi(h) \to 0$ always, and
$\psi(h) = O(h)$ for Lipschitz data---but \emph{not uniformly} in $\nu$ or
$\alpha_0$: at fixed $h$, shrinking $\nu$ inflates $\psi$. This is the honest
content of the caveat in \cref{oa:th:stability} that $h_0$ depends on the
coefficient data, and it is the same viscous floor already met in
\cref{oa:lem:patch}. Note also what is absent: $\sigma$ appears in no bound
here---large reaction does \emph{not} degrade $\psi$, because $\sigma$ enters
$\tau_1^{-1}$ as a global constant and cancels from every nodal-versus-elemental
difference.
\end{pedagogy}

\subsubsection{Best approximation by the weighted projections}

\begin{lemma}[Best approximation and weighted residual approximation]
\label{oa:lem:bestapprox}
Let $i \in \{1,2\}$ and let $\Pi_i$ be as in \cref{oa:sec:method}. Then, for any
elementwise square-integrable $z$ and any finite element function $z_h$ in the
corresponding projection space,
\begin{equation}
  \label{oa:eq:BestApprox}
  \norm{\Pi_i^{\perp} z}_{\tau_i} \le \norm{z - z_h}_{\tau_i}.
\end{equation}
Moreover, let $z \in H^{r}(\Omega)^s$ with $1 \le r \le k_u+1$ (for $i = 1$) or
$1 \le r \le k_p + 1$ (for $i = 2$), and let $\alpha$ satisfy \cref{H:porositysmooth} with
$m \ge r$ wherever $r \ge 2$. Then
\begin{equation}
  \label{oa:eq:WeightedApprox}
  \norm{\Pi_i^{\perp}(\alpha z)}_{\tau_i}
  \;\le\;
  C\, C_{\alpha,r}
  \Bigl( \sum_K \tau_{i,K}\,\alpha_K^2\, h_K^{2r}\,\norm{z}_{H^{r}(K)}^2 \Bigr)^{1/2},
\end{equation}
with $C$ depending only on the interpolation constant, the patch-overlap constant and
the constants of \cref{oa:lem:patch}, and $C_{\alpha,r}$ as in \cref{H:porositysmooth}
(with $C_{\alpha,1} \coloneqq 1 + \norm{\nabla\alpha}_{L^\infty(\Omega)}/\alpha_0$).
\end{lemma}

\begin{proof}
\cref{oa:eq:BestApprox} is the minimizing property of orthogonal projections in their own
inner product, cf.~\cite[Remark~4]{Codina2008FiniteEA}. For \cref{oa:eq:WeightedApprox}, take
$z_h$ to be the Scott--Zhang interpolant of $\alpha z$ in the projection space, so that
$\norm{\alpha z - z_h}_{K} \le C h_K^{r} \norm{\alpha z}_{H^{r}(S_K)}$
(here $r \le k+1$ for the relevant degree $k$, by \cref{H:spaces} and the stated restrictions on
$r$; in particular the case $i=1$, $z = \nabla p$, $r = k_p$ uses $k_p \le k_u+1$).
By the Leibniz rule and \cref{H:porositysmooth},
$\norm{\alpha z}_{H^{r}(S_K)}
\le \bigl( \alpha_{\infty,S_K} + \sum_{j=1}^{r}\binom{r}{j}\norm{D^j\alpha}_{L^\infty}\bigr)
\norm{z}_{H^{r}(S_K)}
\le C\,C_{\alpha,r}\,\alpha_{\infty,S_K}\norm{z}_{H^{r}(S_K)}$,
using $\norm{D^j\alpha}_{L^\infty} \le (C_{\alpha,r}-1)\alpha_0 \le
(C_{\alpha,r}-1)\alpha_{\infty,S_K}$. Insert this into \cref{oa:eq:BestApprox}, replace
$\alpha_{\infty,S_K}$ and $\tau_{i,S_K}$-values by their values on $K$ using
\cref{oa:lem:patch}, and re-sum over patches using the bounded overlap implied by
nondegeneracy, absorbing $\norm{z}_{H^r(S_K)}$ into element contributions.
\end{proof}

\begin{pedagogy}[The two halves of the best-approximation lemma]
Both halves are short; neither is decorative. Half one, \cref{oa:eq:BestApprox},
is the Pythagorean minimizing property of orthogonal projections---
$\norm{\Pi^\perp z}_{\tau} \le \norm{z - z_h}_{\tau}$ for any competitor $z_h$
in the projection space, since $z - \Pi z \perp_\tau (\Pi z - z_h)$. Trivial as
it looks, it is the reason the \emph{analysis} uses $\tau$-weighted projections
while the \emph{code} uses the plain $L^2$ projection
(\cref{oa:rem:analyzed}(i)): the minimizing property must hold in the same
weighted product in which the stabilization pays, or the consistency estimate
below collapses; cf.~\cite[Remark~4]{Codina2008FiniteEA}. Half two,
\cref{oa:eq:WeightedApprox}, is the engine of the whole consistency argument:
\emph{fluctuations of smooth functions are as small as interpolation errors}.
The proof pattern is worth absorbing once: to bound $\Poneo(\alpha z)$ or
$\Ptwoo(\alpha z)$, insert the best competitor available in closed form---the
Scott--Zhang interpolant of $\alpha z$---then handle the product by the Leibniz
rule (this is the \emph{only} place \cref{H:porositysmooth} is used, through
$C_{\alpha,r}$), and convert patch sums to element sums by \cref{oa:lem:patch}
plus bounded patch overlap (each element lies in a uniformly bounded number of
patches, by shape regularity). The ASGS analysis needs no analogue: its
stabilization is residual-based and exactly consistent (\cref{eq:consistency}),
so the exact solution never leaves a stabilization footprint to estimate. Here
the projection destroys exact consistency, and this lemma is precisely what is
received in exchange.
\end{pedagogy}

\subsubsection{The OSGS triple norm}
\label{oa:sec:norm}

The stability and convergence analysis is carried out in the mesh-dependent OSGS norm of the
main text (\cref{eq:TripleNormOSGS}), recalled here for convenience,
\begin{equation}
  \label{oa:eq:TripleNorm}
  \tnorm{V}_{\mathrm O}
  \coloneqq
  \Bigl(
  2\nu\,\bigl\lVert \alpha^{1/2}\ViscProj\nabla\bv \bigr\rVert^2
  + \norm{\sigma^{1/2}\bv}^2
  + \varepsilon\norm{q}^2
  + \norm{\alpha X(V)}_{\tau_1}^2
  + \norm{\nabla\cdot(\alpha\bv)}_{\tau_2}^2
  \Bigr)^{1/2}.
\end{equation}
Throughout this appendix we keep the method subscripts of the main text:
$\tnorm{\cdot}_{\mathrm O}$ denotes this OSGS norm (\cref{eq:TripleNormOSGS}), while
$\tnorm{\cdot}_{\mathrm S}$ denotes the ASGS norm of \cref{lemma:Stability} of the main text,
which controls only the damped reaction $\sigmatilde$ (\cref{oa:lem:normcomparison}). The bare
triple bar is not used.

\begin{lemma}[Definiteness]
\label{oa:lem:definiteness}
Under \cref{H:mesh}--\cref{H:porosity}, and with the pressure normalized to zero mean when
$\varepsilon = 0$, \cref{oa:eq:TripleNorm} is a norm on $\Xhz$; in particular $\tnorm{V}_{\mathrm{O}} = 0$
implies $V = 0$.
\end{lemma}

\begin{proof}
Homogeneity, symmetry and the triangle inequality are immediate; only definiteness needs an
argument, and it is here that the full Dirichlet data enter. Let $\tnorm{V}_{\mathrm{O}} = 0$ for
$V = [\bv; q] \in \Xhz$. Then $\ViscProj\nabla\bv \equiv 0$, i.e.\ the symmetric gradient of
$\bv$ is spherical almost everywhere; since $\bv \in H^1_0(\Omega)^d$, the deviatoric
(conformal) Korn inequality on the bounded domain $\Omega$---whose conformal--Killing kernel
is excluded by the homogeneous boundary data---gives $\norm{\nabla\bv} \le
C\,\bigl\lVert\ViscProj\nabla\bv\bigr\rVert = 0$, so $\bv \equiv 0$. With $\bv = 0$, the term
$\norm{\alpha X(V)}_{\tau_1} = 0$ reduces to $\norm{\alpha\nabla q}_{\tau_1} = 0$; as
$\alpha \ge \alpha_0 > 0$ and $\tau_1 > 0$, this forces $\nabla q \equiv 0$, so $q$ is
constant, and the normalization ($\varepsilon\norm{q} = 0$ when $\varepsilon > 0$, or the
zero-mean constraint when $\varepsilon = 0$) forces $q \equiv 0$. Hence $\tnorm{\cdot}_{\mathrm{O}}$ is
definite. (Absent full Dirichlet data or pressure normalization it is only a seminorm.)
\end{proof}

\begin{lemma}[Comparison with the ASGS norm]
\label{oa:lem:normcomparison}
Let $\tnorm{\cdot}_{\mathrm{S}}$ denote the triple norm of the ASGS stability lemma
of~the main text,
\[
\begin{aligned}
  \tnorm{V}_{\mathrm{S}}^2
  &= \nu\,\bigl\lVert \alpha^{1/2}\ViscProj\nabla\bv \bigr\rVert^2
  + \bigl\lVert \widetilde{\sigma}_\alpha^{1/2}\bv \bigr\rVert_h^2
  + \varepsilon\norm{q}^2
  + \norm{\alpha X(V)}_{\tau_1}^2
  + \norm{\nabla\cdot(\alpha\bv)}_{\tau_2}^2,
  \qquad \\
  \widetilde{\sigma}_\alpha
  &= \frac{\tns^{-1}\sigma}{\tns^{-1} + \sigma/\alpha_K}.
\end{aligned}
\]
Then $\tnorm{V}_{\mathrm{S}} \le \tnorm{V}_{\mathrm{O}}$ for all $V$.
\end{lemma}

\begin{proof}
Term by term: $\nu \le 2\nu$; on each element
$\widetilde{\sigma}_\alpha = \sigma\,\tns^{-1}/(\tns^{-1}+\sigma/\alpha_K) \le \sigma$;
the remaining three terms coincide.
\end{proof}

\begin{pedagogy}[Anatomy of the norm; the Korn subtlety; a one-way comparison]
Read \cref{oa:eq:TripleNorm} as a price list: its five components are the only
currencies the analysis can pay estimates with. Four coincide (up to the factor
$2$ on the viscous term) with the ASGS norm \cref{eq:triplenorm}; the single
substantive difference is the reactive component---full $\norm{\sigma^{1/2}\bv}$
here, damped $\lVert\sigmatilde^{1/2}\bv\rVert_h$ there. By
\cref{eq:sigmatilde}, $\sigmatilde = \sigma\varphi_1/(\varphi_1+\sigma) \le
\min(\sigma, \varphi_1)$: in the reaction-dominated regime $\sigmatilde$
saturates at $\varphi_1 = c_1\alpha_K^2\tau_2/h_K^2 \ll \sigma$, so the OSGS
norm is genuinely, unboundedly stronger there, and \cref{oa:lem:normcomparison}
is the precise one-way street ($\tnorm{\cdot}_{\mathrm S} \le
\tnorm{\cdot}_{\mathrm O}$, never the reverse with a uniform constant). Keep
this asymmetry in mind when comparing the two convergence theorems: they bound
different left-hand sides. On definiteness: the point needing proof is that the
viscous term sees only the \emph{deviatoric symmetric} gradient. Velocity
fields with $\ViscProj\nabla\bv \equiv 0$ form the finite-dimensional
conformal-Killing kernel (infinitesimal rigid motions and dilation-like
fields), and the deviatoric Korn inequality states that once the boundary data
exclude that kernel---as homogeneous Dirichlet data do---the full gradient is
controlled by its deviatoric symmetric part. Without full Dirichlet data (or
another kernel-excluding constraint), \cref{oa:eq:TripleNorm} degrades to a
seminorm, as the parenthetical in the proof warns.
\end{pedagogy}

\subsection{Stability}
\label{oa:sec:stability}

\begin{theorem}[Inf--sup stability]
\label{oa:th:stability}
Under \cref{H:mesh}--\cref{H:spaces}, \cref{H:design}--\cref{H:patch} there exist $h_0 > 0$ and $\beta_{\mathrm{osgs}} > 0$,
depending only on $\beta_0$, $C_2$, $C_{\nabla\alpha}/C_{\mathrm{inv}}$ and the constants of
\cref{oa:sec:technical}, such that for all $h \le h_0$,
\begin{equation}
  \label{oa:eq:InfSup}
  \inf_{U_h \in \Xhz\setminus\{0\}}\;
  \sup_{V_h \in \Xhz\setminus\{0\}}\;
  \frac{B_{\mathrm{osgs}}(\ba; U_h, V_h)}{\tnorm{U_h}_{\mathrm{O}}\,\tnorm{V_h}_{\mathrm{O}}}
  \;\ge\; \beta_{\mathrm{osgs}} .
\end{equation}
An admissible choice of the constants in \cref{H:design} is
$\gamma_0 = 10\,(1 + M_2)$ with $M_2 \coloneqq 1 + C_{\nabla\alpha}/C_{\mathrm{inv}}$,
together with $h_0$ small enough that
$\psi(h_0) \le \min\{1/8,\ \beta_0^2/16\}$, where $\psi$ is the function of
\cref{oa:lem:smoothing}. These are \emph{sufficient} conditions, and $\gamma_0$ is not claimed
sharp: with $\gamma_0 \ge 20$ they demand $c_1 \gtrsim 400\,C_{\mathrm{inv}}^2$ and
$c_2 \gtrsim 20\,C_{\mathrm{inv}}$, well above the values $c_1 = 4k^4$ (or $16k^4$),
$c_2 = 2k^2$ used in the experiments of~the main text, which are not shown to satisfy
them. The result is thus a \emph{conditional} inf--sup stability theorem, not a calibration
or validation of the stabilization constants employed in the computations; the
threshold~$h_0$ likewise depends on the coefficient data through $\psi$
(\cref{oa:lem:smoothing}) and is not uniform in $\alpha_0$, $\nu$ or the moduli of continuity
of $\alpha$ and $\ba$.
\end{theorem}

\begin{pedagogy}[What inf--sup means, and why coercivity is unavailable here]
Background, in finite dimensions. For a square discrete system,
\emph{coercivity} $B(U,U) \ge c\norm{U}^2$ gives solvability and stability by
testing each $U$ with itself. The \emph{inf--sup} condition \cref{oa:eq:InfSup}
is the general sufficient condition: for every $U_h$ there exists a
\emph{witness} $V_h$ with $B(U_h,V_h) \ge \beta\,\tnorm{U_h}_{\mathrm
O}\tnorm{V_h}_{\mathrm O}$; equivalently, the system matrix is injective with a
quantitative bound, and a square injective matrix is invertible. Coercivity is
the special case in which the witness is always $U_h$ itself. Why that special
case fails here \emph{in this norm}: the diagonal identity \cref{oa:eq:Diagonal}
is a sum of squares, so $B_{\mathrm{osgs}}$ \emph{is} nonnegative on the
diagonal---but of the wrong quantities. Splitting $\norm{\xi_h}_{\tau_1}^2 =
\norm{\Pone\xi_h}_{\tau_1}^2 + O_1$ (Pythagoras for the unconstrained
projection), the norm \cref{oa:eq:TripleNorm} contains the \emph{full}
$\norm{\xi_h}_{\tau_1}^2$, while the diagonal delivers only the fluctuation
parts $O_1$, $O_2$: the projected components are invisible to
$B_{\mathrm{osgs}}(U_h,U_h)$, and no constant relates them to what the diagonal
controls. Contrast with the ASGS diagonal, familiar from \cref{lem:coercivity}:
there the stabilization at $V=U$ produces the \emph{full} residual square minus
a viscous--reactive penalty (\cref{eq:stabdiag1}--\cref{eq:Gexpand}); the
penalty is absorbed by the design conditions (\cref{eq:viscabsorb}) and, for
the reactive part, absorbed \emph{exactly} at the price of renormalizing
$\sigma$ down to $\sigmatilde$ (\cref{eq:reactrenorm}). Deleting the reactive
and viscous slots (\cref{oa:lem:annihilation}) removes both the penalty and the
renormalization: the full $D_\sigma$ survives on the diagonal---which is why the
stronger norm is attainable at all---but the projected residual components must
now be \emph{harvested} by a purpose-built witness. That is the entire business
of the proof below. On the statement's honesty paragraph: the constants are
sufficient-side only ($\gamma_0 = 10(1+M_2)$ demands $c_1 \gtrsim
400\,C_{\mathrm{inv}}^2$, far above the values used in the computations), and
$h_0$ inherits the data-dependence of $\psi$; the theorem certifies the
mechanism, it does not calibrate the code.
\end{pedagogy}

\begin{proof}
Fix $U_h = [\bu_h; p_h] \in \Xhz$ and recall the abbreviations \cref{oa:eq:XiDelta}.
Write $\psi \coloneqq \psi(h)$ and set
\[
  D_\nu \coloneqq 2\nu\,\bigl\lVert \alpha^{1/2}\ViscProj\nabla\bu_h \bigr\rVert^2,
  \quad
  D_\sigma \coloneqq \norm{\sigma^{1/2}\bu_h}^2,
  \quad
  D_\varepsilon \coloneqq \varepsilon\norm{p_h}^2,
\]
\[
  O_1 \coloneqq \norm{\Poneo\xi_h}_{\tau_1}^2,
  \quad
  O_2 \coloneqq \norm{\Ptwoo\delta_h}_{\tau_2}^2,
  \quad
  P_1 \coloneqq \norm{\Ponez\xi_h}_{\tau_1}^2,
  \quad
  P_2 \coloneqq \norm{\Ptwo\delta_h}_{\tau_2}^2 .
\]

\medskip\noindent\emph{Step 1: the special test function.}
Define
\[
  \bw_h \coloneqq \tau_1 \smul \Ponez(\xi_h) \;\in\; V_{h0},
  \qquad
  r_h \coloneqq \tau_2 \smul \Ptwo(\delta_h) - m_h \;\in\; Q_{h0},
\]
where $m_h$ is the mean of $\tau_2\smul\Ptwo(\delta_h)$ over $\Omega$ whenever
$Q_{h0}$ consists of zero-mean functions, and $m_h \coloneqq 0$ otherwise. Note that
$\bw_h$ is an admissible velocity test function ($\Ponez\xi_h \in V_{h0}$ vanishes on
$\partial\Omega$, and the smoothed product preserves nodal zeros), and $r_h$ an
admissible pressure test function since constants belong to $Q_h$ by \cref{H:spaces}. Set
$V^0 \coloneqq [\bw_h; r_h]$ and take as test function $W \coloneqq U_h + V^0$.
By \cref{oa:eq:Diagonal},
\begin{equation}
  \label{oa:eq:StepDiag}
  B_{\mathrm{osgs}}(\ba; U_h, U_h) = D_\nu + D_\sigma + D_\varepsilon + O_1 + O_2 ,
\end{equation}
and it remains to bound $B_{\mathrm{osgs}}(\ba; U_h, V^0)$ from below. Expanding
\cref{eq:BilinearForm} and \cref{oa:eq:Sdefs},
\begin{align}
  B_{\mathrm{osgs}}(\ba; U_h, V^0)
  ={}& 2\nu\,(\alpha\,\ViscProj\nabla\bu_h, \ViscProj\nabla\bw_h)
  \tag{T1}\label{oa:eq:T1}\\
  &+ (\bw_h, \xi_h)
  \tag{T2}\label{oa:eq:T2}\\
  &+ \sigma(\bu_h, \bw_h)
  \tag{T3}\label{oa:eq:T3}\\
  &+ \varepsilon(p_h, r_h)
  \tag{T4}\label{oa:eq:T4}\\
  &+ (r_h, \delta_h)
  \tag{T5}\label{oa:eq:T5}\\
  &+ \bigl(\Poneo\xi_h,\ \alpha(\ba\cdot\nabla\bw_h + \nabla r_h)\bigr)_{\tau_1}
  \tag{T6}\label{oa:eq:T6}\\
  &+ \bigl(\Ptwoo\delta_h,\ \nabla\cdot(\alpha\bw_h)\bigr)_{\tau_2},
  \tag{T7}\label{oa:eq:T7}
\end{align}
where in \cref{oa:eq:T2} we used
$(\bw_h, \alpha\ba\cdot\nabla\bu_h) + (\bw_h, \alpha\nabla p_h) = (\bw_h, \xi_h)$.
The constant shift $m_h$ is invisible throughout: it does not affect $\nabla r_h$ in
\cref{oa:eq:T6}; in \cref{oa:eq:T5} it contributes $-m_h\int_\Omega \delta_h = 0$ because
$\int_\Omega \nabla\cdot(\alpha\bu_h) = 0$ by the divergence theorem and
$\bu_h \in V_{h0}$; and in \cref{oa:eq:T4} it contributes
$-\varepsilon\, m_h \int_\Omega p_h = 0$, since either $m_h = 0$ or $p_h$ has zero
mean. We now estimate \cref{oa:eq:T1}--\cref{oa:eq:T7} term by term. Throughout,
\cref{oa:eq:Smoothing} is used in the form
$\normK{\bw_h} \le (1+\psi)\,\tau_{1,K}\normK{\Ponez\xi_h}$ and
$\normK{r_h + m_h} \le (1+\psi)\,\tau_{2,K}\normK{\Ptwo\delta_h}$ (recall also that
subtracting the mean does not increase the $L^2$ norm, so the same bound holds
globally for $r_h$).

\begin{pedagogy}[Step 1 read slowly: the witness, component by component]
The design recipe (from~\cite{Codina2008FiniteEA}) is: \emph{to harvest a
square $\norm{z}_w^2$, test with $w\,z$}---the Galerkin pairing of the
corresponding equation against that test function then produces the square.
Here the two quantities missing from the diagonal are the projected momentum
residual and the projected mass residual, so the witness has one component for
each: $\bw_h = \tau_1\smul\Ponez\xi_h$ and $r_h = \tau_2\smul\Ptwo\delta_h -
m_h$. Each ingredient is forced. The \emph{constrained} projection $\Ponez$
(not $\Pone$): test velocities must vanish on the boundary, and
$\Ponez\xi_h \in V_{h0}$ does; the gap between $\Ponez$ and $\Pone$ is exactly
what \cref{H:projection} will price. The smoothing $\smul$ (not $\pmul$):
conformity, as prepared in \cref{oa:lem:smoothing}. The mean shift $m_h$:
membership in $Q_{h0}$; it is then checked to be invisible in every
pairing---note the small pleasure of the argument that $\int_\Omega\delta_h = 0$
by the divergence theorem, so even \cref{oa:eq:T5} ignores it. Why no pressure
analogue of \cref{H:projection} is needed is explained at the end of Step~2:
$\delta_h$ already has zero mean and constants belong to $Q_h$, so projecting
onto the \emph{unconstrained} space and shifting the test function costs
nothing---``$\beta_0 = 1$'' for the mass slot
(\cref{oa:rem:proofstructure}(iv)). Finally, the expansion
\cref{oa:eq:T1}--\cref{oa:eq:T7} is the campaign map: \cref{oa:eq:T2} and
\cref{oa:eq:T5} are the \emph{productive} terms (they will yield $P_1$ and
$P_2$); \cref{oa:eq:T1}, \cref{oa:eq:T3}, \cref{oa:eq:T4} are Galerkin
cross-talk between $U_h$ and the witness; \cref{oa:eq:T6}, \cref{oa:eq:T7} are
stabilization cross-talk. The ledger to be established: every cross term is at
most a small multiple of a quantity the diagonal already controls, plus a
fraction strictly less than one of the freshly harvested $P$'s.
\end{pedagogy}

\medskip\noindent\emph{Step 2: the two productive terms \cref{oa:eq:T2} and \cref{oa:eq:T5}.}
Splitting $\tau_1\smul(\cdot) = \tau_1\pmul(\cdot) + (\tau_1\smul - \tau_1\pmul)(\cdot)$
and using the $\tau_1$-orthogonality
$(\xi_h, \Ponez\xi_h)_{\tau_1} = \norm{\Ponez\xi_h}_{\tau_1}^2$,
\[
\begin{aligned}
  (\bw_h, \xi_h)
  &= P_1 + \bigl((\tau_1\smul - \tau_1\pmul)\Ponez\xi_h,\ \xi_h\bigr) \\
  &\;\ge\; P_1 - \psi\,\norm{\Ponez\xi_h}_{\tau_1}\norm{\xi_h}_{\tau_1}
  \;\ge\; P_1 - \frac{2\psi}{\beta_0^2}\,(P_1 + O_1),
\end{aligned}
\]
where the last step used the nonexpansiveness of $\Ponez$ in $\norm{\cdot}_{\tau_1}$
together with \cref{H:projection}, in the form
$\norm{\xi_h}_{\tau_1}^2 \le (2/\beta_0^2)(P_1 + O_1)$. For \cref{oa:eq:T5}, since
$\delta_h$ has zero mean the shift is invisible as noted, and by the exact
$\tau_2$-orthogonal decomposition of $\delta_h$ with respect to $Q_h$,
$\norm{\delta_h}_{\tau_2}^2 = P_2 + O_2$; hence
\[
  (r_h, \delta_h)
  = P_2 + \bigl((\tau_2\smul - \tau_2\pmul)\Ptwo\delta_h,\ \delta_h\bigr)
  \;\ge\; P_2 - \psi\,(P_2 + O_2).
\]
This is the point where the mass slot enjoys the exact analogue of \cref{H:projection} with
constant one: because $\delta_h$ has zero mean, projecting onto $Q_h$ and correcting
the test function by a constant replaces the constrained projection
$\Pi_{2,0}$ of~\cite{Codina2008FiniteEA} at no cost.

\begin{pedagogy}[Step 2 read slowly: the harvest, and the first bite of \cref{H:projection}]
Splitting $\smul = \pmul + (\smul-\pmul)$: the $\pmul$-part of \cref{oa:eq:T2}
is $(\tau_1\pmul\Ponez\xi_h,\ \xi_h) = (\Ponez\xi_h,\ \xi_h)_{\tau_1} = P_1$,
the promised square, by $\tau_1$-orthogonality
($\xi_h - \Ponez\xi_h \perp_{\tau_1} V_{h0} \ni \Ponez\xi_h$); the
$(\smul-\pmul)$-part is priced by $\psi$ (\cref{oa:lem:smoothing}) times
$\norm{\Ponez\xi_h}_{\tau_1}\norm{\xi_h}_{\tau_1}$. To convert the second
factor into ledger currency, \cref{H:projection} is used in squared form:
\cref{eq:ProjectionStability} plus $(a+b)^2 \le 2(a^2+b^2)$ give
$\norm{\xi_h}_{\tau_1}^2 \le (2/\beta_0^2)(P_1+O_1)$. This is the exact point
of the hypothesis: it quantifies \emph{how much of the momentum residual the
constrained projection can see}. If $\beta_0$ could degenerate with $h$, the
harvest $P_1$ could be an arbitrarily small fraction of
$\norm{\Pone\xi_h}_{\tau_1}$, and Step~5 could not reconstruct the norm. This
Codina--Blasco-type condition is the one hypothesis of this appendix that is a
genuine, method-specific property of the finite element pair rather than a
consequence of the standing mesh and data assumptions (see the commentary at
\cref{oa:rem:dispensed}). The mass computation is the same but cleaner: the
decomposition $\norm{\delta_h}_{\tau_2}^2 = P_2 + O_2$ is \emph{exact}
(unconstrained projection, genuine Pythagoras), so only the smoothing penalty
$\psi\,(P_2+O_2)$ appears.
\end{pedagogy}

\medskip\noindent\emph{Step 3: the perturbation terms.}
For \cref{oa:eq:T1}, by Cauchy--Schwarz, $|\alpha| \le \alpha_K$ elementwise,
$|\ViscProj \boldsymbol{M}| \le |\boldsymbol{M}|$ pointwise, the inverse estimate
\cref{eq:InverseEstimateFiniteOrderNorm}, \cref{oa:eq:Smoothing} and \cref{oa:eq:K1},
\begin{align*}
  \text{\cref{oa:eq:T1}}
  &\ \ge\ -\,2\sum_K \nu^{1/2}\,\bigl\lVert\alpha^{1/2}\ViscProj\nabla\bu_h\bigr\rVert_K
  \cdot \nu^{1/2}\alpha_K^{1/2}\,\frac{C_{\mathrm{inv}}}{h_K}\,(1+\psi)\,\tau_{1,K}
  \norm{\Ponez\xi_h}_K\\
  &\ \ge\ -\,2 M_1 \Bigl(\frac{D_\nu}{2}\Bigr)^{1/2} P_1^{1/2},
\end{align*}
with $M_1 \coloneqq (1+\psi)/\gamma$, whence, by Young's inequality,
\[
  \text{\cref{oa:eq:T1}} \ \ge\ -\frac{D_\nu}{4} - 2M_1^2\, P_1 .
\]
For \cref{oa:eq:T3}, using $\sigma\tau_1 \le 1$ from \cref{oa:eq:P1},
\begin{align*}
  \text{\cref{oa:eq:T3}}
  &\ \ge\ -(1+\psi)\sum_K \sigma\,\tau_{1,K}\normK{\bu_h}\,\normK{\Ponez\xi_h}\\
  &\ \ge\ -(1+\psi)\, D_\sigma^{1/2} P_1^{1/2}
  \ \ge\ -\frac{D_\sigma}{2} - \frac{(1+\psi)^2}{2}\,P_1 .
\end{align*}
For \cref{oa:eq:T4}, using $\varepsilon\tau_2 \le C_2$ from \cref{oa:eq:P1},
\begin{align*}
  \text{\cref{oa:eq:T4}}
  &\ \ge\ -(1+\psi)\sum_K \varepsilon\,\tau_{2,K}\normK{p_h}\normK{\Ptwo\delta_h}\\
  &\ \ge\ -(1+\psi)\,C_2^{1/2}\, D_\varepsilon^{1/2} P_2^{1/2}
  \ \ge\ -\frac{D_\varepsilon}{2} - \frac{(1+\psi)^2 C_2}{2}\,P_2 .
\end{align*}
For \cref{oa:eq:T6} we bound the two contributions to
$\alpha X(V^0) = \alpha\ba\cdot\nabla\bw_h + \alpha\nabla r_h$ elementwise. By the
inverse estimate, \cref{oa:eq:Smoothing} and \cref{oa:eq:K2},
\[
\begin{aligned}
  \tau_{1,K}^{1/2}\normK{\alpha\,\ba\cdot\nabla\bw_h}
  &\le \tau_{1,K}^{1/2}\,\alpha_K |\ba|_{\infty,K}\frac{C_{\mathrm{inv}}}{h_K}
  (1+\psi)\,\tau_{1,K}\normK{\Ponez\xi_h} \\
  &\le \frac{1+\psi}{\gamma}\,\tau_{1,K}^{1/2}\normK{\Ponez\xi_h},
\end{aligned}
\]
and by \cref{oa:eq:K3},
\[
  \tau_{1,K}^{1/2}\normK{\alpha\nabla r_h}
  \le \tau_{1,K}^{1/2}\,\alpha_K\,\frac{C_{\mathrm{inv}}}{h_K}(1+\psi)\,
  \tau_{2,K}\normK{\Ptwo\delta_h}
  \le \frac{1+\psi}{\gamma}\,\tau_{2,K}^{1/2}\normK{\Ptwo\delta_h}.
\]
Hence, by Cauchy--Schwarz over the elements and $ab \le (a^2+b^2)/2$,
\[
  \text{\cref{oa:eq:T6}}
  \ \ge\ -\frac{1+\psi}{\gamma}\, O_1^{1/2}\bigl( P_1^{1/2} + P_2^{1/2} \bigr)
  \ \ge\ -\frac{1+\psi}{\gamma}\Bigl( O_1 + \frac{P_1 + P_2}{2} \Bigr).
\]
For \cref{oa:eq:T7}, expanding
$\nabla\cdot(\alpha\bw_h) = \alpha\,\nabla\cdot\bw_h + \nabla\alpha\cdot\bw_h$ and
using \cref{H:porosity} in the form
$\norm{\nabla\alpha}_{L^\infty(K)} \le C_{\nabla\alpha}\,\alpha_K/h_K$,
\[
\begin{aligned}
  \tau_{2,K}^{1/2}\normK{\nabla\cdot(\alpha\bw_h)}
  &\le \Bigl(1 + \frac{C_{\nabla\alpha}}{C_{\mathrm{inv}}}\Bigr)
  \tau_{2,K}^{1/2}\,\alpha_K\,\frac{C_{\mathrm{inv}}}{h_K}\,
  (1+\psi)\,\tau_{1,K}\normK{\Ponez\xi_h} \\
  &\le \frac{M_2(1+\psi)}{\gamma}\,\tau_{1,K}^{1/2}\normK{\Ponez\xi_h},
\end{aligned}
\]
by \cref{oa:eq:K3}, with $M_2 = 1 + C_{\nabla\alpha}/C_{\mathrm{inv}}$; hence
\[
  \text{\cref{oa:eq:T7}}
  \ \ge\ -\frac{M_2(1+\psi)}{\gamma}\, O_2^{1/2} P_1^{1/2}
  \ \ge\ -\frac{M_2(1+\psi)}{2\gamma}\,\bigl( O_2 + P_1 \bigr).
\]

\begin{pedagogy}[Step 3 read slowly: the perturbation ledger and the Young workhorse]
Every estimate here has the same shape: $|\text{cross}| \le \kappa\,A^{1/2}
P^{1/2}$ with $A$ a quantity the diagonal controls and $\kappa$ a coefficient
made small by design, followed by Young's inequality $ab \le
\tfrac{\lambda}{2}a^2 + \tfrac{1}{2\lambda}b^2$, with $\lambda$ chosen to
sacrifice only a \emph{fraction} of $A$ (for \cref{oa:eq:T1}, a quarter of
$D_\nu$) against a controlled multiple of $P$. Where each $\kappa$'s smallness
comes from is the content of the (K)-inequalities: \cref{oa:eq:K1} for the
viscous cross-talk---note, per \cref{oa:rem:proofstructure}(i), that only the
\emph{plain} inverse estimate is needed, because the analyzed OSGS form
contains no viscous stabilization, so the enlarged constant of
\cref{eq:winv-divvisc} and its calibration burden never appear here;
\cref{oa:eq:K2} and \cref{oa:eq:K3} for the two halves of the witness's own
first-order quantity in \cref{oa:eq:T6}; \cref{oa:eq:K3} again, with the
porosity surcharge $M_2 = 1 + C_{\nabla\alpha}/C_{\mathrm{inv}}$, for
\cref{oa:eq:T7}, where $\nabla\cdot(\alpha\bw_h)$ picks up
$\nabla\alpha\cdot\bw_h$---the only entry point of porosity \emph{variation}
into stability, priced by the resolution condition of \cref{H:porosity}. The
exception to the pattern is deliberately the most important two lines of the
appendix: \cref{oa:eq:T3}, the reactive cross-talk $\sigma(\bu_h,\bw_h)$, is
bounded using nothing but $\sigma\tau_1 \le 1$ and absorbed by the \emph{full}
Galerkin reaction $D_\sigma$ at rate one half. The reader coming from
\cref{app:Continuity} should pause here: this is the pairing whose ASGS
analogue is renormalized by the stabilization partner $T_{14}$ into the damped
$\sigmatilde$-pairing (\cref{eq:reactivestep}); the OSGS projection deleted
that partner, and the full-strength diagonal $D_\sigma$---the very term the
deletion preserved---absorbs it instead. The stronger norm is paid for at
exactly this spot, and the price is only a Young split.
\end{pedagogy}

\medskip\noindent\emph{Step 4: collection.}
Adding \cref{oa:eq:StepDiag} and the bounds of Steps~2--3,
\begin{align*}
  B_{\mathrm{osgs}}(\ba; U_h, W)
  \ \ge\ & \tfrac34\, D_\nu + \tfrac12\, D_\sigma + \tfrac12\, D_\varepsilon\\
  &+ \Bigl( 1 - \frac{2\psi}{\beta_0^2} - \frac{1+\psi}{\gamma} \Bigr)\, O_1
  + \Bigl( 1 - \psi - \frac{M_2(1+\psi)}{2\gamma} \Bigr)\, O_2\\
  &+ \Bigl( 1 - \frac{2\psi}{\beta_0^2} - 2M_1^2 - \frac{(1+\psi)^2}{2}
  - \frac{1+\psi}{2\gamma} - \frac{M_2(1+\psi)}{2\gamma} \Bigr)\, P_1\\
  &+ \Bigl( 1 - \psi - \frac{(1+\psi)^2 C_2}{2} - \frac{1+\psi}{2\gamma} \Bigr)\, P_2 .
\end{align*}
With $\gamma \ge \gamma_0 = 10(1+M_2)$ and
$\psi \le \psi_0 = \min\{1/8, \beta_0^2/16\}$ (so that $2\psi/\beta_0^2 \le 1/8$ and
$1+\psi \le 9/8$), a direct evaluation shows that every coefficient is at least
$1/8$ (for $P_1$: $1 - 0.125 - 0.007 - 0.633 - 0.029 - 0.057 \ge 0.14$; for $P_2$,
using $C_2 \le 1$: $1 - 0.125 - 0.633 - 0.029 \ge 0.21$; the others are larger).
Hence, with $C_0 = 1/8$,
\begin{equation}
  \label{oa:eq:Collected}
  B_{\mathrm{osgs}}(\ba; U_h, W)
  \ \ge\ C_0\,\bigl( D_\nu + D_\sigma + D_\varepsilon + O_1 + O_2 + P_1 + P_2 \bigr).
\end{equation}

\begin{pedagogy}[Step 4: the audit]
Nothing conceptual happens here; it is the audit that the seven balances stay
positive. The arithmetic worth internalizing: with $\gamma \ge 10(1+M_2)$,
every $(1+\psi)/\gamma$-type coefficient is a few percent;
$\psi \le \min\{1/8,\beta_0^2/16\}$ caps both $\psi$ and $2\psi/\beta_0^2$ at
$1/8$; the dominant bite on $P_1$ is the $0.633 = (1+\psi)^2/2$ from the Young
split of \cref{oa:eq:T3}---the cost of absorbing the reactive cross-talk at
rate one half---and even so the $P_1$ balance closes at $0.14$. The uniform
floor is $C_0 = 1/8$. Note what the collection does \emph{not} contain: any
$\sigma$-, $\nu$-, $\varepsilon$- or $\Dah$-dependent coefficient. Stability is
unconditional in the physical parameters; all parameter sensitivity of the
final convergence estimate enters later, through the interpolation
conversions.
\end{pedagogy}

\medskip\noindent\emph{Step 5: recovery of the triple norm and boundedness of the test
function.}
By \cref{H:projection} and the exact decomposition of the mass slot,
\[
  \tnorm{U_h}_{\mathrm{O}}^2
  = D_\nu + D_\sigma + D_\varepsilon + \norm{\xi_h}_{\tau_1}^2 + \norm{\delta_h}_{\tau_2}^2
  \le D_\nu + D_\sigma + D_\varepsilon + \frac{2}{\beta_0^2}(P_1 + O_1) + (P_2 + O_2),
\]
so \cref{oa:eq:Collected} yields
$B_{\mathrm{osgs}}(\ba; U_h, W) \ge C_0 \min\{1, \beta_0^2/2\}\,\tnorm{U_h}_{\mathrm{O}}^2$.
It remains to bound $\tnorm{W}_{\mathrm{O}} \le \tnorm{U_h}_{\mathrm{O}} + \tnorm{V^0}_{\mathrm{O}}$. Using
\cref{oa:eq:Smoothing} elementwise together with \cref{oa:eq:K1} for the viscous term,
$\sigma\tau_1 \le 1$ for the reactive term, $\varepsilon\tau_2 \le C_2$ for the
compressibility term, and the bounds of Step~3 for
$\norm{\alpha X(V^0)}_{\tau_1}$ and $\norm{\nabla\cdot(\alpha\bw_h)}_{\tau_2}$,
\begin{align*}
  \tnorm{V^0}_{\mathrm{O}}^2
  &\le \Bigl[ \frac{2(1+\psi)^2}{\gamma^2}
  + (1+\psi)^2
  + \frac{2(1+\psi)^2}{\gamma^2}
  + \frac{M_2^2(1+\psi)^2}{\gamma^2} \Bigr] P_1 \\
  &+ \Bigl[ (1+\psi)^2 C_2 + \frac{2(1+\psi)^2}{\gamma^2} \Bigr] P_2\\
  &\le C\,(P_1 + P_2),
\end{align*}
and $P_1 \le \norm{\xi_h}_{\tau_1}^2$, $P_2 \le \norm{\delta_h}_{\tau_2}^2$ give
$\tnorm{V^0}_{\mathrm{O}} \le C\,\tnorm{U_h}_{\mathrm{O}}$, hence $\tnorm{W}_{\mathrm{O}} \le C\,\tnorm{U_h}_{\mathrm{O}}$. Combining,
\[
  \sup_{V_h\in\Xhz} \frac{B_{\mathrm{osgs}}(\ba; U_h, V_h)}{\tnorm{V_h}_{\mathrm{O}}}
  \ \ge\ \frac{B_{\mathrm{osgs}}(\ba; U_h, W)}{\tnorm{W}_{\mathrm{O}}}
  \ \ge\ \beta_{\mathrm{osgs}}\, \tnorm{U_h}_{\mathrm{O}}
\]
with $\beta_{\mathrm{osgs}} = C_0\min\{1,\beta_0^2/2\}/C > 0$, which is
\cref{oa:eq:InfSup}.
\end{proof}

\begin{pedagogy}[Step 5: reconstruction and boundedness---the two closing duties]
An inf--sup proof has two closing obligations that a coercivity proof does not.
(a)~\emph{Reconstruction}: the collected lower bound is in the bookkeeping
variables $D$, $O$, $P$, but the norm contains $\norm{\xi_h}_{\tau_1}^2$ and
$\norm{\delta_h}_{\tau_2}^2$; the passage back is a second (and final)
application of \cref{H:projection} for the momentum slot and the exact
Pythagoras for the mass slot. (b)~\emph{Boundedness of the witness}: the
inf--sup quotient divides by $\tnorm{W}_{\mathrm O}$, so a witness of unbounded
size would prove nothing; the estimate $\tnorm{V^0}_{\mathrm O} \le
C\,\tnorm{U_h}_{\mathrm O}$ deliberately reuses the elementwise bounds of
Step~3---an economy worth noticing: the same (K)-conversions price both the
cross-talk and the witness's size. The final constant $\beta_{\mathrm{osgs}} =
C_0\min\{1,\beta_0^2/2\}/C$ assembles the three ingredients: collection floor,
reconstruction rate, witness size. Two structural notes. First, the pattern
``test with $U_h$ plus a weighted correction, harvest, absorb, reconstruct,
bound the witness'' is the standard template for inf--sup proofs of stabilized
and mixed methods; met once here, it will be recognized throughout that
literature. Second, well-posedness of \cref{oa:eq:Method} is a footnote in
finite dimensions: the system is square, and \cref{oa:eq:InfSup} makes its
matrix injective, hence invertible---the statement at the head of
\cref{oa:sec:convergence}.
\end{pedagogy}

\begin{remark}[On the structure of the proof]
\label{oa:rem:proofstructure}
The proof is that of~\cite[Theorem~1]{Codina2008FiniteEA} with four modifications. (i)~All
inverse-estimate steps carry the porosity weight, absorbed through
\cref{oa:eq:K1}--\cref{oa:eq:K3}; only the plain inverse estimate
\cref{eq:InverseEstimateFiniteOrderNorm} is used---in contrast with the ASGS stability proof
of~the main text, no estimate for $\nabla\cdot(\alpha\nu\ViscProj\nabla\bv_h)$ is
needed, because the analyzed OSGS form contains no viscous stabilization term.
(ii)~The reactive pairing \cref{oa:eq:T3} is new; it is absorbed by the \emph{full}
Galerkin reaction term $D_\sigma$ at the price of half of $P_1$, using nothing more
than $\sigma\tau_1 \le 1$. This is where the OSGS analysis retains
$\norm{\sigma^{1/2}\bu_h}$ in the norm, where the ASGS analysis retains only the
damped $\widetilde{\sigma}_\alpha$. (iii)~The compressibility pairing \cref{oa:eq:T4} is
new and is controlled by the same condition $\varepsilon\tau_2 \le C_2$ assumed
in~the main text. (iv)~The constrained pressure projection $\Pi_{2,0}$
of~\cite{Codina2008FiniteEA} is replaced by the unconstrained projection plus a mean
correction of the test function, which is invisible in every term; this makes the
mass-slot stability constant exactly one and is the reason no analogue of
\cref{H:projection} is required for the divergence.
\end{remark}

\begin{pedagogy}[Cross-walk to the ASGS stability proof]
A dictionary, for the reader holding both proofs in mind. \emph{Test function}:
$U_h$ itself (\cref{lem:coercivity}) versus $U_h + V^0$ here. \emph{Diagonal}:
full residual square minus an indefinite penalty (\cref{eq:stabdiag1}) versus
fluctuation squares with no penalty (\cref{oa:eq:Diagonal}).
\emph{Absorptions}: viscous penalty via the enlarged inverse estimate and
$c_1 > 2\xi C_{\mathrm{inv}}^2$ (\cref{eq:viscabsorb}, \cref{H:coercivity})
versus none needed. \emph{Reaction}: renormalized to $\sigmatilde$
(\cref{eq:reactrenorm}) versus retained in full, at the cost of item~(ii)
here. \emph{Extra machinery}: none versus smoothing, patches, constrained
projection. \emph{Extra hypothesis}: none versus \cref{H:projection}.
\emph{Output norm}: \cref{eq:triplenorm} with $\sigmatilde$ versus
\cref{oa:eq:TripleNorm} with $\sigma$. The trade is clean: one page and a
heavier constants condition buy the damped norm; five steps and one genuinely
non-automatic hypothesis buy the full norm.
\end{pedagogy}

\subsection{Convergence}
\label{oa:sec:convergence}

Throughout this section, $U = [\bu; p]$ denotes the solution of
the linearized problem (\cref{sec:CommonSetting}) under the regularity \cref{H:regularity}, and $U_h \in \Xhz$ the solution of
\cref{oa:eq:Method}. That solution exists and is unique: for $h \le h_0$ the inf--sup estimate
\cref{oa:eq:InfSup} of \cref{oa:th:stability}, on the square linear system with equal trial and
test spaces $\Xhz$, makes the system operator injective and hence invertible, so
\cref{oa:eq:Method} is well posed. We use the local interpolation{\hyp}error quantities
$\Eint{K}{\bu}$ and $\Eint{K}{p}$ of the main text (\cref{eq:InterpolationError}), and state
the convergence estimate in the porosity-weighted broken $\ell^2$ functional
\begin{equation}
  \label{oa:eq:ErrorFunctionL2}
  \Psi_{\mathrm O}(h)
  \coloneqq
  \Bigl( \sum_K \bigl( c_1 + \Dah \bigr)\,\frac{\alpha_K^2}{h_K^2}
  \bigl( \tau_{2,K}\,\Eint{K}{\bu}^2 + \tau_{1,K}\,\Eint{K}{p}^2 \bigr) \Bigr)^{1/2},
\end{equation}
whose coarser $\ell^1$ majorant is the error function
\begin{equation}
  \label{oa:eq:ErrorFunction}
  \mathcal{E}(h)
  \coloneqq
  \sum_K \frac{\alpha_K}{h_K}\,\bigl(1 + \Dah\bigr)^{1/2}
  \Bigl( \tau_{2,K}^{1/2}\,\Eint{K}{\bu} + \tau_{1,K}^{1/2}\,\Eint{K}{p} \Bigr),
\end{equation}
i.e., the error function of the ASGS convergence theorem of~the main text with the
local factor $(1+\Dah)^{1/2}$ inserted. The two are related by
$\Psi_{\mathrm O}(h) \le c_1^{1/2}\,\mathcal{E}(h)$ (the discrete $\ell^2\subset\ell^1$
inequality together with $c_1+\Dah \le c_1(1+\Dah)$). We state the theorem in
$\Psi_{\mathrm O}$: the $\ell^1$ passage to $\mathcal{E}$ can lose a mesh-cardinality
factor $\sim h^{-d/2}$ on a quasi-uniform family, so the optimal-order corollaries
must be read off the $\ell^2$ functional.

\begin{pedagogy}[Two error functionals; how to read orders]
$\Psi_{\mathrm O}$ is what the proofs naturally produce: every estimate ends in
Cauchy--Schwarz over elements, i.e.\ a broken $\ell^2$ sum. $\mathcal E$ exists
only to display the result in the same $\ell^1$ format as the ASGS theorem of
the main text. The warning deserves amplification, because convergence orders
are easily misread otherwise: passing from $\ell^2$ to $\ell^1$ over
$N \sim h^{-d}$ elements can inflate by $N^{1/2} \sim h^{-d/2}$
(Cauchy--Schwarz is sharp when elementwise contributions are equidistributed,
the generic case for smooth solutions on quasi-uniform meshes). So optimal
orders---and the corollaries below---must be read from $\Psi_{\mathrm O}$;
$\mathcal E$ is a display convention, not the sharp quantity. The ASGS appendix
makes the same distinction with its $\ell^2$ functional in \cref{eq:psih}
versus the $\ell^1$ display of \cref{thm:convergence}, so
\cref{oa:cor:asgsnorm} below compares like with like. Also note the cosmetic
difference between $(c_1+\Dah)$ here and $(1+\Dah)$ in
\cref{oa:eq:ErrorFunction}: they differ by a bounded factor, absorbed into $C$;
the functional keeps $c_1$ visible because that is what (P5) actually
produces.
\end{pedagogy}

\subsubsection{Consistency}

The analyzed method is not consistent in the classical variational sense
(cf.~\cite[Eq.~(44)]{Codina2008FiniteEA}); its consistency error is the value of the
stabilization terms at the exact solution, and it is of the order of the error
function:

\begin{lemma}[Consistency error]
\label{oa:lem:consistency}
Under \cref{H:mesh}--\cref{H:regularity}, \cref{H:design}--\cref{H:advectionsmooth}, for all $V_h \in \Xhz$,
\begin{equation}
  \label{oa:eq:ConsistencyId}
  B_{\mathrm{osgs}}(\ba;\, U - U_h,\, V_h)
  = S_1(U, V_h) + S_2(U, V_h),
\end{equation}
and
\begin{equation}
  \label{oa:eq:ConsistencyBound}
  \bigl| B_{\mathrm{osgs}}(\ba;\, U - U_h,\, V_h) \bigr|
  \;\le\; C\,\Psi_{\mathrm O}(h)\,\tnorm{V_h}_{\mathrm{O}}.
\end{equation}
\end{lemma}

\begin{proof}
Since $U$ solves the linearized problem of \cref{sec:CommonSetting} with the stated regularity, testing the
strong equations against $V_h \in \Xhz$ and integrating the viscous term by parts
gives $B(\ba; U, V_h) = L(V_h)$; adding $S_1(U,V_h) + S_2(U,V_h)$ to both sides and
subtracting \cref{oa:eq:Method} yields \cref{oa:eq:ConsistencyId}.

For the momentum part, split $\alpha X(U) = \alpha(\ba\cdot\nabla\bu) + \alpha\nabla p$
and use, respectively, Cauchy--Schwarz in $(\cdot,\cdot)_{\tau_1}$, linearity of
$\Poneo$ and \cref{oa:lem:bestapprox} with $z = \ba\cdot\nabla\bu$, $r = k_u$ and with
$z = \nabla p$, $r = k_p$:
\[
  |S_1(U, V_h)|
  \le \Bigl( \norm{\Poneo[\alpha\,\ba\cdot\nabla\bu]}_{\tau_1}
  + \norm{\Poneo[\alpha\nabla p]}_{\tau_1} \Bigr)\,\norm{\alpha X(V_h)}_{\tau_1}.
\]
By \cref{H:advectionsmooth} and the Leibniz rule,
$\norm{\ba\cdot\nabla\bu}_{H^{k_u}(K)} \le C(1+C_D)\,|\ba|_{\infty,K}\,
\norm{\bu}_{H^{k_u+1}(K)}$, so \cref{oa:eq:WeightedApprox} and the second inequality of
\cref{oa:eq:P4} give
\[
\begin{aligned}
  \norm{\Poneo[\alpha\,\ba\cdot\nabla\bu]}_{\tau_1}
  &\le C \Bigl( \sum_K \tau_{1,K}\,\alpha_K^2\,|\ba|_{\infty,K}^2\, h_K^{2k_u}
  \norm{\bu}_{H^{k_u+1}(K)}^2 \Bigr)^{1/2} \\
  &\le C \Bigl( \sum_K \Bigl(\frac{\alpha_K}{h_K}\Bigr)^{2} \tau_{2,K}\,
  \Eint{K}{\bu}^2 \Bigr)^{1/2},
\end{aligned}
\]
while $\norm{\nabla p}_{H^{k_p}(K)} \le \norm{p}_{H^{k_p+1}(K)}$ and
\cref{oa:eq:WeightedApprox} give directly
\[
\begin{aligned}
  \norm{\Poneo[\alpha\nabla p]}_{\tau_1}
  &\le C \Bigl( \sum_K \tau_{1,K}\,\alpha_K^2\, h_K^{2k_p}
  \norm{p}_{H^{k_p+1}(K)}^2 \Bigr)^{1/2} \\
  &= C \Bigl( \sum_K \Bigl(\frac{\alpha_K}{h_K}\Bigr)^{2} \tau_{1,K}\,
  \Eint{K}{p}^2 \Bigr)^{1/2}.
\end{aligned}
\]
Both right-hand sides are bounded by $C\,\Psi_{\mathrm O}(h)$ (with room to spare, since
$c_1+\Dah \ge 1$ and no Damk\"ohler factor arises here), and
$\norm{\alpha X(V_h)}_{\tau_1} \le \tnorm{V_h}_{\mathrm{O}}$.

For the mass part, the exact mass equation \cref{eq:StrongMassEquation} gives
$\nabla\cdot(\alpha\bu) = -\varepsilon p$, hence, by \cref{oa:lem:annihilation}-style
linearity,
\[
  S_2(U, V_h) = -\varepsilon\,\bigl( \Ptwoo p,\ \nabla\cdot(\alpha\bv_h)
  \bigr)_{\tau_2},
\]
which vanishes when $\varepsilon = 0$. For $\varepsilon > 0$, by Cauchy--Schwarz,
\cref{oa:lem:bestapprox} (with $\alpha$ replaced by $1$, $z = p$, $r = k_p+1$),
$\varepsilon\tau_{2} \le C_2 \le 1$ and $\varepsilon \le C_2\tau_2^{-1}$ from
\cref{oa:eq:P1}, and finally \cref{oa:eq:P5},
\begin{align*}
  |S_2(U, V_h)|
  &\le \varepsilon \norm{\Ptwoo p}_{\tau_2}\,\norm{\nabla\cdot(\alpha\bv_h)}_{\tau_2}
  \le C \Bigl( \sum_K \varepsilon^2 \tau_{2,K}\, h_K^{2(k_p+1)}
  \norm{p}_{H^{k_p+1}(K)}^2 \Bigr)^{1/2} \tnorm{V_h}_{\mathrm{O}}\\
  &\le C\,C_2 \Bigl( \sum_K \tau_{2,K}^{-1}\, \Eint{K}{p}^2 \Bigr)^{1/2}
  \tnorm{V_h}_{\mathrm{O}}\\
  &\le C \Bigl( \sum_K \bigl(c_1 + \Dah\bigr)
  \Bigl(\frac{\alpha_K}{h_K}\Bigr)^{2}\tau_{1,K}\,\Eint{K}{p}^2 \Bigr)^{1/2}
  \tnorm{V_h}_{\mathrm{O}}
\end{align*}
(in the last step, $\varepsilon\tau_2 \le C_2 \le 1$ was applied once more, followed
by \cref{oa:eq:P5}),
which is exactly $C\,\Psi_{\mathrm O}(h)\,\tnorm{V_h}_{\mathrm{O}}$. Each of the two broken $\ell^2$
right-hand sides is thus dominated by $\Psi_{\mathrm O}(h)$, so summing the two parts
proves \cref{oa:eq:ConsistencyBound} in the $\ell^2$ functional (with no $\ell^2\subset\ell^1$
enlargement).
\end{proof}

\begin{pedagogy}[What ``consistency error'' means here, and why it is small]
For the ASGS method, consistency is \emph{exact} (\cref{eq:consistency}): the
stabilization is assembled from the full residual, the exact solution's
residual vanishes, hence the exact solution satisfies the discrete equations
and Galerkin orthogonality holds. Here it fails, for a reason visible in
\cref{oa:eq:Sdefs}: the stabilized quantity is the truncated first-order
expression $\alpha X(U)$, which does \emph{not} vanish at the exact solution
(it is not the residual). Identity \cref{oa:eq:ConsistencyId} quantifies the
failure: the defect is exactly the stabilization evaluated at $U$. Salvation
comes from the symmetric form \cref{oa:eq:Ssym}: only the \emph{fluctuation}
$\Poneo[\alpha X(U)]$ enters, and by \cref{oa:lem:bestapprox} fluctuations of
smooth functions are interpolation-small. Three observations on the proof.
(1)~The momentum part needs \cref{H:advectionsmooth} only to
Leibniz-differentiate $\ba\cdot\nabla\bu$ up to order $k_u$---this, and
\cref{H:porositysmooth} inside \cref{oa:lem:bestapprox}, is the entire role of
the two smoothness strengthenings, and it carries no Damk\"ohler cost (``room
to spare''). (2)~The mass part is an $\varepsilon$-effect \emph{exactly}: the
strong mass equation gives $\nabla\cdot(\alpha\bu) = -\varepsilon p$, so
$S_2(U,\cdot)$ vanishes identically for $\varepsilon = 0$; for
$\varepsilon > 0$ its estimate is the one place a Damk\"ohler factor enters
consistency, through (P5) pricing the $\tau_2$-weights. (3)~The estimate lands
directly in the $\ell^2$ functional---no $\ell^1$ enlargement---which is what
permits stating \cref{oa:th:convergence} in $\Psi_{\mathrm O}$. The take-away
principle, stated once: orthogonal-subscale methods trade exact consistency
for consistency-up-to-fluctuations; at the level of convergence rates the
trade is free, but it is paid for in data-smoothness hypotheses.
\end{pedagogy}

\subsubsection{Interpolation estimates}

Let $\hat U_h = [\hat\bu_h; \hat p_h]$ where $\hat\bu_h \in V_{h0}$ is the nodal
interpolant of $\bu$ (well defined since $\bu \in H^{k_u+1} \subset
C^0(\overline\Omega)$ for $d\le 3$, and boundary-value preserving) and
$\hat p_h \in Q_{h0}$ is the nodal interpolant of $p$, shifted by its mean when
$Q_{h0}$ consists of zero-mean functions. Write
$\hat E \coloneqq U - \hat U_h = [\hat{\boldsymbol e}_u; \hat e_p]$.

\begin{remark}[The pressure mean shift]
\label{oa:rem:meanshift}
Since $\int_\Omega p = 0$, the shift constant is bounded by the raw interpolation
error, $|\Omega|^{1/2}|m| = |\int_\Omega (\hat p_h - p)|\,|\Omega|^{-1/2}
\le \norm{p - \hat p_h}$, so all interpolation estimates for $\hat e_p$ hold with at
most doubled constants. Moreover the constant part of $\hat e_p$ is invisible in the
pairings below: against $\nabla\cdot(\alpha\bv_h)$ it integrates to zero by the
divergence theorem, against $\varepsilon q_h$ it vanishes because $q_h$ has zero mean
(or no shift was performed), and it has no gradient. We therefore do not distinguish
the shifted interpolant notationally.
\end{remark}

\begin{pedagogy}[Why the mean shift deserves a remark]
A bookkeeping point that would otherwise nag: the pressure interpolant must
live in $Q_{h0}$, and shifting by the mean could in principle pollute every
estimate. The remark disposes of this once: the shift is itself
interpolation-small (because $p$ has zero mean), and the \emph{constant} part
of $\hat e_p$ is invisible in each pairing it meets---by the divergence
theorem, by the zero mean of $q_h$, or by differentiation. The same three-way
check was performed for $m_h$ inside the stability proof; it is the standard
tax of working in zero-mean pressure spaces.
\end{pedagogy}

\begin{lemma}[Interpolation continuity]
\label{oa:lem:interpolation}
Under \cref{H:mesh}--\cref{H:regularity}, \cref{H:design}--\cref{H:advectionsmooth}, for all $V_h \in \Xhz$,
\begin{equation}
  \label{oa:eq:InterpolationBound}
  \bigl| B_{\mathrm{osgs}}(\ba;\, \hat E,\, V_h) \bigr|
  \;\le\; C\,\Psi_{\mathrm O}(h)\,\tnorm{V_h}_{\mathrm{O}}.
\end{equation}
\end{lemma}

\begin{proof}
Expand $B_{\mathrm{osgs}}(\ba; \hat E, V_h)$ into the Galerkin terms and the
stabilization terms and estimate each. Below, all interpolation estimates are
\cref{oa:eq:NodalInterp} with $n = k_u+1$ or $n = k_p+1$ and $m \in \{0,1\}$.

\smallskip\noindent
\emph{(I1) Viscous term.} Using \cref{oa:eq:P4} ($\nu \le \alpha_K\tau_{2,K}$),
\begin{align*}
  2\nu(\alpha\ViscProj\nabla\hat{\boldsymbol e}_u, \ViscProj\nabla\bv_h)
  &\le \sqrt{2}\,\tnorm{V_h}_{\mathrm{O}}\Bigl( \sum_K 2\nu\alpha_K\,
  \norm{\nabla\hat{\boldsymbol e}_u}_K^2 \Bigr)^{1/2}\\
  &\le C\,\tnorm{V_h}_{\mathrm{O}}\Bigl( \sum_K
  \Bigl(\frac{\alpha_K}{h_K}\Bigr)^2\tau_{2,K}\,\Eint{K}{\bu}^2 \Bigr)^{1/2}.
\end{align*}

\smallskip\noindent
\emph{(I2) Convective and mass--divergence terms, jointly.} Using
$\nabla\cdot(\alpha\ba) = 0$, $\hat{\boldsymbol e}_u \in H^1_0(\Omega)^d$
(nodal interpolation preserves the boundary values of $\bu$) and the continuity of the
pressure space, global integration by parts gives
\[
  (\bv_h, \alpha\ba\cdot\nabla\hat{\boldsymbol e}_u)
  + (q_h, \nabla\cdot(\alpha\hat{\boldsymbol e}_u))
  = -\bigl( \hat{\boldsymbol e}_u,\ \alpha\bigl(\ba\cdot\nabla\bv_h + \nabla q_h\bigr)
  \bigr)
  = -\bigl( \hat{\boldsymbol e}_u,\ \alpha X(V_h) \bigr).
\]
Hence, splitting $\tau_{1,K}^{-1/2} \le (\alpha_K\tns^{-1})^{1/2} + \sigma^{1/2}$ and
using \cref{oa:eq:P3} for the first part and \cref{oa:eq:P6} for the second,
\[
\begin{aligned}
  \bigl|\bigl(\hat{\boldsymbol e}_u, \alpha X(V_h)\bigr)\bigr|
  &\le \tnorm{V_h}_{\mathrm{O}} \Bigl( \sum_K \tau_{1,K}^{-1} \norm{\hat{\boldsymbol e}_u}_K^2
  \Bigr)^{1/2} \\
  &\le C\,\tnorm{V_h}_{\mathrm{O}}\Bigl( \sum_K \bigl(c_1 + \Dah\bigr)
  \Bigl(\frac{\alpha_K}{h_K}\Bigr)^2 \tau_{2,K}\,\Eint{K}{\bu}^2 \Bigr)^{1/2}.
\end{aligned}
\]
This is the first of the two places where the Damk\"ohler factor enters.

\begin{pedagogy}[(I2) read slowly: one integration by parts, two terms at once]
Watch the opening move of (I2) in slow motion. The idea is to shift each
derivative off the interpolation error $\hat{\boldsymbol e}_u$ and onto the
smooth discrete field. For the convective term, $\nabla\cdot(\alpha\ba)=0$ makes
the weighted advection skew-adjoint, so
$(\bv_h,\alpha\ba\cdot\nabla\hat{\boldsymbol e}_u) =
-(\hat{\boldsymbol e}_u,\alpha\ba\cdot\nabla\bv_h)$; the mass term becomes
$(q_h,\nabla\cdot(\alpha\hat{\boldsymbol e}_u)) =
-(\hat{\boldsymbol e}_u,\alpha\nabla q_h)$. Both boundary integrals vanish
because $\hat{\boldsymbol e}_u\in H^1_0(\Omega)^d$: $\bu$ carries the no-slip
data and the nodal interpolant reproduces it exactly at boundary nodes. Adding
the two recombines them into the single pairing
$-(\hat{\boldsymbol e}_u,\alpha X(V_h))$ against the composite first-order
quantity $X(V_h)$---and that is the point, since $\alpha X(V_h)$ is exactly what the
$\tau_1$-part of $\tnorm{V_h}_{\mathrm O}$ controls (\cref{oa:eq:TripleNorm}), leaving the whole
convective-plus-mass block one weighted Cauchy--Schwarz from the test norm.
(Term~(I5) repeats the manoeuvre for the pressure gradient, its boundary term
vanishing instead because $\bv_h\in V_{h0}$.)
\end{pedagogy}

\smallskip\noindent
\emph{(I3) Reactive term.} Using the full reactive control in the norm
\cref{oa:eq:TripleNorm} and \cref{oa:eq:P6},
\[
\begin{aligned}
  \sigma(\hat{\boldsymbol e}_u, \bv_h)
  &\le \norm{\sigma^{1/2}\bv_h}\Bigl( \sum_K \sigma\,
  \norm{\hat{\boldsymbol e}_u}_K^2 \Bigr)^{1/2} \\
  &\le C\,\tnorm{V_h}_{\mathrm{O}}\Bigl( \sum_K \Dah
  \Bigl(\frac{\alpha_K}{h_K}\Bigr)^2\tau_{2,K}\,\Eint{K}{\bu}^2 \Bigr)^{1/2}.
\end{aligned}
\]

\smallskip\noindent
\emph{(I4) Compressibility term.} Using $\varepsilon \le C_2\tau_2^{-1}$ and
\cref{oa:eq:P5},
\[
\begin{aligned}
  \varepsilon(\hat e_p, q_h)
  &\le \varepsilon^{1/2}\norm{\varepsilon^{1/2}q_h}
  \Bigl( \sum_K \norm{\hat e_p}_K^2 \Bigr)^{1/2} \\
  &\le C\,\tnorm{V_h}_{\mathrm{O}}\Bigl( \sum_K \bigl(c_1+\Dah\bigr)
  \Bigl(\frac{\alpha_K}{h_K}\Bigr)^2\tau_{1,K}\,\Eint{K}{p}^2 \Bigr)^{1/2}.
\end{aligned}
\]

\smallskip\noindent
\emph{(I5) Pressure-gradient term.} Global integration by parts
($\bv_h \in V_{h0}$) and \cref{oa:eq:P5} give
\[
\begin{aligned}
  (\bv_h, \alpha\nabla\hat e_p)
  &= -(\hat e_p, \nabla\cdot(\alpha\bv_h)) \\
  &\le \norm{\nabla\cdot(\alpha\bv_h)}_{\tau_2}
  \Bigl( \sum_K \tau_{2,K}^{-1}\norm{\hat e_p}_K^2 \Bigr)^{1/2} \\
  &\le C\,\tnorm{V_h}_{\mathrm{O}}\Bigl( \sum_K \bigl(c_1+\Dah\bigr)
  \Bigl(\frac{\alpha_K}{h_K}\Bigr)^2\tau_{1,K}\,\Eint{K}{p}^2 \Bigr)^{1/2},
\end{aligned}
\]
the second place where the Damk\"ohler factor enters (through
$\tau_2^{-1/2} \le (c_1+\Dah)^{1/2}(\alpha_K/h_K)\tau_1^{1/2}$).

\smallskip\noindent
\emph{(I6) Momentum stabilization.} Since fluctuation operators are nonexpansive,
\[
  |S_1(\hat E, V_h)|
  \le \bigl( \norm{\alpha\,\ba\cdot\nabla\hat{\boldsymbol e}_u}_{\tau_1}
  + \norm{\alpha\nabla\hat e_p}_{\tau_1} \bigr)\,\tnorm{V_h}_{\mathrm{O}}.
\]
Elementwise, by \cref{oa:eq:P4},
\[
\begin{aligned}
  \tau_{1,K}^{1/2}\,\alpha_K\,|\ba|_{\infty,K}\norm{\nabla\hat{\boldsymbol e}_u}_K
  &\le C\,\frac{\alpha_K}{h_K}\,\tau_{2,K}^{1/2}\,\Eint{K}{\bu},
  \qquad \\
  \tau_{1,K}^{1/2}\,\alpha_K\norm{\nabla\hat e_p}_K
  &\le C\,\frac{\alpha_K}{h_K}\,\tau_{1,K}^{1/2}\,\Eint{K}{p},
\end{aligned}
\]
the second bound holding exactly; no Damk\"ohler factor arises.

\smallskip\noindent
\emph{(I7) Mass stabilization.} Likewise,
$|S_2(\hat E, V_h)| \le
\norm{\nabla\cdot(\alpha\hat{\boldsymbol e}_u)}_{\tau_2}\tnorm{V_h}_{\mathrm{O}}$, and elementwise,
using \cref{H:porosity} in the form $\norm{\nabla\alpha}_{L^\infty(K)} \le C_{\nabla\alpha}\alpha_K/h_K$,
\[
  \tau_{2,K}^{1/2}\norm{\nabla\cdot(\alpha\hat{\boldsymbol e}_u)}_K
  \le \tau_{2,K}^{1/2}\bigl( \alpha_K \norm{\nabla\hat{\boldsymbol e}_u}_K
  + C_{\nabla\alpha}\tfrac{\alpha_K}{h_K}\norm{\hat{\boldsymbol e}_u}_K \bigr)
  \le C(1+C_{\nabla\alpha})\,\frac{\alpha_K}{h_K}\,\tau_{2,K}^{1/2}\,\Eint{K}{\bu}.
\]

Collecting (I1)--(I7), each square-summed right-hand side is dominated by
$\Psi_{\mathrm O}(h)$---the momentum, reaction and viscous groups through
$c_1+\Dah \ge 1$, the convective, compressibility and pressure-gradient groups through their explicit
$(c_1+\Dah)$ factor---and, absorbing $c_1$-dependent constants into $C$, we obtain
\cref{oa:eq:InterpolationBound} directly in the $\ell^2$ functional.
\end{proof}

\begin{pedagogy}[The interpolation ledger, and the term-by-term ASGS contrast]
This lemma is the OSGS counterpart of \cref{lem:continterp}, and holding the
two side by side is the single most instructive comparison in the two
appendices. First the ledger here. Damk\"ohler-free: (I1) (viscous, priced by
$\nu \le \alpha_K\tau_2$) and (I6)--(I7) (the stabilization terms, priced by
nonexpansiveness plus (P4); note the pressure half of (I6) is exact, no
conversion at all). Damk\"ohler-priced: (I2), through the split
$\tau_1^{-1/2} \le (\alpha_K\tns^{-1})^{1/2} + \sigma^{1/2}$ (from
$\tau_1^{-1} = \alpha_K\tns^{-1}+\sigma$ and $\sqrt{a+b} \le \sqrt a + \sqrt
b$), whose first piece converts by (P3) at no cost and whose second converts by
(P6) at cost $\Dah^{1/2}$; (I3), the head-on reactive pairing, again by (P6);
(I4) and (I5), both through (P5). \textbf{Audit note.} The prose tags (I2) as
``the first of the two places where the Damk\"ohler factor enters'' and (I5) as
``the second''; as the displayed bounds show, (I3) and (I4) also carry the
factor, so the count ``two'' is accurate only for the two \emph{global
integration-by-parts} terms, not for the factor's entry points. A revision
might tag all four of (I2)--(I5), or reword to ``the first of the places''.
Now the contrast the ASGS-fluent reader should draw, pairing by pairing. The
pairing of (I2) is the Galerkin convective-plus-mass combination that the ASGS
proof writes as [V] and \emph{telescopes} with the stabilization partner
$\sigma(\tau_1\bu, \alpha\X(V))_h$ (its [II]) through
$1-\sigma\tau_1 = \varphi_1\tau_1$, yielding the Damk\"ohler-free bound
\cref{eq:IIVterm}. The pairing of (I3) is the ASGS reactive pair
$T_3 + T_{14}$, renormalized to the $\sigmatilde$-pairing at zero cost
(\cref{eq:reactivestep}). The pairing of (I5) is the Galerkin pressure pairing
that the ASGS proof treats jointly with its stabilization partner [I] as
[I]+[IV], telescoped elementwise at the price of the $[\tau_1]$ face terms and
the jump condition (\cref{eq:IandIV}, \cref{eq:jumptotal}). In all three cases
the ASGS proof escapes the Damk\"ohler factor by pairing the dangerous Galerkin
term with a $\sigma$-weighted stabilization partner; the OSGS projection
annihilates precisely those partners (\cref{oa:lem:annihilation}), so the same
pairings must be bounded head-on, and (P5)/(P6) are the best head-on rates
available. The factor $(c_1+\Dah)^{1/2}$ in \cref{oa:th:convergence} is exactly
the sum of these head-on payments---no more, no less.
\end{pedagogy}

\begin{lemma}[Size of the interpolation error in the triple norm]
\label{oa:lem:interpsize}
Under \cref{H:mesh}--\cref{H:regularity}, \cref{H:design}--\cref{H:advectionsmooth},
$\tnorm{\hat E}_{\mathrm{O}} \le C\,\Psi_{\mathrm O}(h)$.
\end{lemma}

\begin{proof}
Estimate the five terms of \cref{oa:eq:TripleNorm} at $V = \hat E$: the viscous term as
in (I1); $\norm{\sigma^{1/2}\hat{\boldsymbol e}_u}$ as in (I3);
$\varepsilon^{1/2}\norm{\hat e_p}$ as in (I4);
$\norm{\alpha X(\hat E)}_{\tau_1}$ as in (I6); and
$\norm{\nabla\cdot(\alpha\hat{\boldsymbol e}_u)}_{\tau_2}$ as in (I7).
\end{proof}

\begin{pedagogy}[Why this lemma exists, and a hidden moral]
The triangle inequality at the end needs the size of $\hat E$ itself in the
norm, not just how the form sees it; the proof is pure reuse. One consequence
deserves flagging, because it reframes the Damk\"ohler factor. The reactive
component of $\tnorm{\hat E}_{\mathrm O}$ is
$\norm{\sigma^{1/2}\hat{\boldsymbol e}_u} \sim \sigma^{1/2}h^{k_u+1}$, and by
(P6) this is $\Dah^{1/2}$ times the $(\alpha_K/h_K)\tau_2^{1/2}$-weighted
velocity interpolation error---i.e.\ the interpolation error \emph{alone},
measured in this norm, is already of size $\Dah^{1/2}$ times the ASGS-type
functional. Since no approximation can beat the best approximation, any
convergence estimate in a norm containing the full $\norm{\sigma^{1/2}\cdot}$
must carry the factor; it is a property of the \emph{norm}, not of the proof
strategy. This is the cleanest way to see that the $(1+\Dah)^{1/2}$ of
\cref{oa:th:convergence} cannot be removed by sharper estimates---only by
weakening the display (\cref{oa:cor:asgsnorm}) or by dividing the reaction back
out (\cref{oa:cor:Ltwo}).
\end{pedagogy}

\subsubsection{Main theorem and corollaries}

\begin{theorem}[Convergence of the OSGS method]
\label{oa:th:convergence}
Let $U$ solve the linearized problem (\cref{sec:CommonSetting}) and let
$U_h \in \Xhz$ solve \cref{oa:eq:Method}. Under \cref{H:mesh}--\cref{H:regularity}, \cref{H:design}--\cref{H:advectionsmooth} and for
$h \le h_0$ as in \cref{oa:th:stability}, there is a constant $C$, independent of $h$,
$\nu$, $\sigma$, $\varepsilon$ and of $\alpha$, $\ba$ except through the constants of
\cref{oa:sec:hypotheses}, such that
\begin{equation}
  \label{oa:eq:MainResult}
\begin{aligned}
  \tnorm{U - U_h}_{\mathrm{O}}
  &\;\le\;
  C\,\Psi_{\mathrm O}(h) \\
  &= C\Bigl( \sum_K \bigl( c_1 + \Dah \bigr)\,\frac{\alpha_K^2}{h_K^2}
  \bigl( \tau_{2,K}\,\Eint{K}{\bu}^2 + \tau_{1,K}\,\Eint{K}{p}^2 \bigr) \Bigr)^{1/2} \\
  &\;\le\; C\,\mathcal{E}(h),
\end{aligned}
\end{equation}
with $\Psi_{\mathrm O}(h)$ and its $\ell^1$ majorant $\mathcal{E}(h)$ as in
\cref{oa:eq:ErrorFunctionL2,oa:eq:ErrorFunction}.
\end{theorem}

\begin{proof}
Write $U - U_h = \hat E + \eta_h$ with $\eta_h \coloneqq \hat U_h - U_h \in \Xhz$.
By \cref{oa:th:stability} there is $V_h \in \Xhz$, $\tnorm{V_h}_{\mathrm{O}} = 1$, with
\[
  \beta_{\mathrm{osgs}}\,\tnorm{\eta_h}_{\mathrm{O}}
  \le B_{\mathrm{osgs}}(\ba; \eta_h, V_h)
  = -\,B_{\mathrm{osgs}}(\ba; \hat E, V_h)
  + B_{\mathrm{osgs}}(\ba; U - U_h, V_h)
  \le C\,\Psi_{\mathrm O}(h),
\]
by \cref{oa:lem:interpolation,oa:lem:consistency}. The triangle inequality and
\cref{oa:lem:interpsize} conclude, both in the $\ell^2$ functional $\Psi_{\mathrm O}(h)$.
\end{proof}

\begin{pedagogy}[The two-line endgame]
This is the abstract pattern \emph{stability + consistency + approximability
$\Rightarrow$ convergence}, in its inf--sup form: split $U - U_h = \hat E +
\eta_h$ with $\eta_h$ discrete; apply the inf--sup bound to $\eta_h$, which
converts its norm into the form evaluated against a unit witness; expand
$B_{\mathrm{osgs}}(\eta_h, V_h) = B_{\mathrm{osgs}}(U-U_h, V_h) -
B_{\mathrm{osgs}}(\hat E, V_h)$, the first term being the consistency defect
and the second the interpolation continuity; finish by triangle inequality with
\cref{oa:lem:interpsize}. \Cref{thm:convergence} is the same three lines with
the coercivity constant in place of $\beta_{\mathrm{osgs}}$ and Galerkin
orthogonality in place of the consistency lemma. Once this pattern is
internalized, every quasi-optimal error analysis in the finite element
literature reads as an instance of it.
\end{pedagogy}

\begin{corollary}[Estimate in the ASGS norm]
\label{oa:cor:asgsnorm}
Under the hypotheses of \cref{oa:th:convergence}, the OSGS solution also satisfies the
estimate of the ASGS convergence theorem of~the main text up to the local factor
$(c_1+\Dah)^{1/2}$, in the same porosity-weighted $\ell^2$ form:
\[
\begin{aligned}
  \tnorm{U - U_h}_{\mathrm{S}}
  &\;\le\;
  C\,\Psi_{\mathrm O}(h) \\
  &= C\Bigl( \sum_K \bigl( c_1 + \Dah \bigr)\,\frac{\alpha_K^2}{h_K^2}
  \bigl( \tau_{2,K}\,\Eint{K}{\bu}^2 + \tau_{1,K}\,\Eint{K}{p}^2 \bigr) \Bigr)^{1/2} .
\end{aligned}
\]
\end{corollary}

\begin{proof}
Combine \cref{oa:th:convergence} with \cref{oa:lem:normcomparison}.
\end{proof}

\begin{pedagogy}[Reading the ASGS-norm corollary]
Because \cref{oa:lem:normcomparison} is one-way, this corollary is the
legitimate apples-to-apples display: the same left-hand side as the ASGS
theorem, the same $\ell^2$ format, and the difference isolated in the local
factor $(c_1+\Dah)^{1/2}$. Two cautions, both already in the text but worth
repeating: the comparison orders the \emph{bounds}, not the methods
(\cref{oa:rem:open}(iii)); and \cref{oa:th:convergence} actually controls the
OSGS error in the \emph{stronger} norm---this corollary discards information,
it does not add hypotheses.
\end{pedagogy}

\begin{corollary}[$\sigma$-robust $L^2$ velocity estimate]
\label{oa:cor:Ltwo}
Let $\sigma > 0$. Under the hypotheses of \cref{oa:th:convergence},
\begin{equation}
  \label{oa:eq:LtwoVelocity}
  \norm{\bu - \bu_h}
  \;\le\;
  C \Bigl[ \sum_K \Bigl( 1 + \frac{c_1}{\Dah} \Bigr)
  \Bigl(
  \bigl( 1 + \tfrac{c_2}{c_1}\Reh \bigr)\, \Eint{K}{\bu}^2
  + \frac{\alpha_K}{\sigma\nu}\, \Eint{K}{p}^2
  \Bigr) \Bigr]^{1/2}.
\end{equation}
The factor is written $1+c_1/\Dah$, not $1+\Dah^{-1}$: dividing $\Psi_{\mathrm O}$ by
$\sigma^{1/2}$ carries the full $c_1+\Dah$, and absorbing $c_1$ into that factor would hide
its dependence. In particular, wherever $\Dah \gtrsim c_1$ (the reaction-dominated
pre-asymptotic regime), the local factor $1+c_1/\Dah$ is $O(1)$ and, on a quasi-uniform
mesh, the $L^2$ velocity error is of full interpolation order $h^{k_u+1}$ with a constant
uniform in $\sigma$; the pressure contribution moreover \emph{decreases} as
$(\sigma\nu)^{-1/2}$ as the reaction strengthens.
\end{corollary}

\begin{proof}
Since the norm \cref{oa:eq:TripleNorm} controls $\norm{\sigma^{1/2}(\bu-\bu_h)}$,
$\norm{\bu - \bu_h} \le \sigma^{-1/2}\tnorm{U - U_h}_{\mathrm{O}} \le
C\,\sigma^{-1/2}\Psi_{\mathrm O}(h)$ by \cref{oa:th:convergence}. In each summand of
$\sigma^{-1}\Psi_{\mathrm O}(h)^2$, use $\sigma^{-1} = h_K^2/(\Dah\,\alpha_K\nu)$ to write
\[
  \sigma^{-1}\,(c_1+\Dah)\,\frac{\alpha_K^2}{h_K^2}
  = \Bigl(1 + \frac{c_1}{\Dah}\Bigr)\,\frac{\alpha_K}{\nu};
\]
then \cref{oa:eq:Tau2Expanded} gives $(\alpha_K/\nu)\,\tau_{2,K} = 1 + (c_2/c_1)\Reh$ for the
velocity term, while $\tau_{1,K} \le 1/\sigma$ from \cref{oa:eq:P1} gives
$(\alpha_K/\nu)\,\tau_{1,K} \le \alpha_K/(\sigma\nu)$ for the pressure term. Substituting
these into $\sigma^{-1/2}\Psi_{\mathrm O}(h)$ yields \cref{oa:eq:LtwoVelocity}. \qedhere
\end{proof}

\begin{pedagogy}[Dividing by $\sigma^{1/2}$: the payoff of the stronger norm]
Here the full-$\sigma$ component, which forced the Damk\"ohler payments
upstream, pays out: since the norm controls $\norm{\sigma^{1/2}(\bu-\bu_h)}$,
dividing by $\sigma^{1/2}$ yields an $L^2$ velocity estimate whose quality
\emph{improves} as the reaction strengthens---the factor $1 + c_1/\Dah$ tends
to one, and the pressure contribution decays like $(\sigma\nu)^{-1/2}$. Do not
over-read this as an OSGS-only asset: the ASGS route reaches a comparable
reaction-limit conclusion by a different mechanism, since in that regime
$\sigmatilde$ saturates at $\varphi_1$, and dividing the ASGS estimate by
$\sigmatilde^{1/2} \approx \varphi_1^{1/2}$ likewise returns a full-order,
$\sigma$-uniform $L^2$ velocity bound. What the corollary adds is a single
formula valid across regimes, with the crossover made explicit by $c_1/\Dah$
and the pressure decay quantified. This is also the reading that makes
fact~(i) of \cref{oa:rem:numerics}---indistinguishable $L^2$ velocity
errors---unsurprising from both sides of the theory.
\end{pedagogy}

\subsection{Discussion}
\label{oa:sec:discussion}

\begin{remark}[The mechanism: $\sigma$ versus $\widetilde{\sigma}_\alpha$]
\label{oa:rem:mechanism}
The single structural difference between the two analyses is the fate of the reactive
term. In the ASGS quadratic form the stabilization contains the reactive residual, and
the diagonal reactive contribution is
$\sigma\norm{\bu_h}_K^2 - \sigma^2\tau_{1}\norm{\bu_h}_K^2
= \sigma(1-\sigma\tau_1)\norm{\bu_h}_K^2$, with
$\sigma(1-\sigma\tau_1) = \sigma\,\alpha_K\tns^{-1}\tau_1 \eqsim
\widetilde{\sigma}_\alpha$: the subscales \emph{renormalize} the reaction down to
$\widetilde{\sigma}_\alpha \le \alpha_K\tns^{-1} = c_1\alpha_K^2\tau_2/h_K^2$, a
quantity always commensurate with the divergence-stabilization scale, so all reactive
pairings in the ASGS convergence proof can be traded against
$(\alpha_K/h_K)\tau_2^{1/2}$-weighted quantities with no Damk\"ohler penalty. In the
OSGS variant the orthogonal projection deletes the reactive residual identically
(\cref{oa:lem:annihilation}), so no renormalization occurs: the diagonal keeps the full
$\norm{\sigma^{1/2}\bu_h}^2$ (a stronger left-hand side, cf.\
\cref{oa:lem:normcomparison}), but the reactive and pressure interpolation pairings must
be bounded head-on, and the best conversions available to the present proof---\cref{oa:eq:P5}
and \cref{oa:eq:P6}---cost factors $(c_1+\Dah)^{1/2}$ and $\Dah^{1/2}\le(c_1+\Dah)^{1/2}$, respectively. The Damk\"ohler factor in
\cref{oa:eq:MainResult} is thus the factor these conversions introduce when the reactive
subscale is deleted, while the strengthened norm (in particular \cref{oa:cor:Ltwo}) is the
reaction control retained in return. Both are properties of the present (one-sided)
estimate, not established as sharp or necessary.
\end{remark}

\begin{pedagogy}[The identity to memorize]
This remark is the executive summary of the appendix, and the identity chain to
memorize is $\sigma(1-\sigma\tau_1) = \sigma\varphi_1\tau_1 = \sigmatilde$
(\cref{eq:sigmatilde}): the ASGS diagonal renormalizes, the OSGS projection
deletes the renormalizer. Note the closing sentence's discipline
(``one-sided, not established as sharp or necessary'') and read it together
with the commentary at \cref{oa:lem:interpsize}, which shows the direction in
which sharpness \emph{can} be argued: within this norm the factor is forced at
the level of best approximation, so it is not an artifact of the estimates.
\end{pedagogy}

\begin{remark}[Consistency with the numerical campaign of~the main text]
\label{oa:rem:numerics}
The reaction-dominated experiments of~the main text (manufactured
solutions, $\mathrm{Da} = 10^6$) report two facts about the OSGS variant relative to the ASGS
one: (i)~the $L^2$ velocity errors of the two methods are indistinguishable at all
resolutions, and (ii)~the $H^1$-type velocity error of the OSGS method exceeds the
ASGS one by a factor of about $3.5$ on the finest $P_1$ meshes---the largest velocity
discrepancy of the whole campaign---with the gap closing under refinement at an
accelerating relative rate. Both observations are compatible in order of magnitude with
the present, one-sided bounds. Fact (i) is \cref{oa:cor:Ltwo}: in the regime $\Dah \gg 1$ the
OSGS $L^2$ velocity estimate is of full order with a $\sigma$-uniform constant, so no excess
is predicted in $L^2$ (a one-sided bound cannot, of course, forbid one). Fact (ii) is consistent with the
factor $(1+\Dah)^{1/2}$ in \cref{oa:eq:MainResult}: on the finest meshes of those
experiments the global mesh Damk\"ohler number is of order ten (its precise value
depends on the element-length convention entering $\tau$), so the bound permits an
excess of about $(1+\Dah)^{1/2} \approx 3$; on the interior plateau of the porosity
profile the elementwise values are larger and permit factors of $4$--$5$. Moreover,
$\Dah^{1/2} = (\sigma/(\alpha_K\nu))^{1/2}\, h_K$ decays \emph{linearly} in $h$ while
the bound itself decays like $h^{k_u}$, which reproduces the observed accelerating
relative decay of the gap. Two caveats are mandatory. First, upper bounds are
one-sided: the analysis shows that the observed excess is \emph{permitted} and
supplies a mechanism for it (\cref{oa:rem:mechanism}); it does not prove that the excess
must occur, and the absence of the factor from the ASGS bound does not by itself prove
ASGS superiority. Second, the near-coincidence of the permitted and observed factors
at the finest resolution should be read as an order-of-magnitude consistency check,
not a validated prediction, since the constants in \cref{oa:eq:MainResult} are not
tracked sharply.
\end{remark}

\begin{pedagogy}[One-sided-bound epistemics]
The epistemics here are worth imitating: an upper bound can only ever
\emph{permit} an observation, and the remark scrupulously claims no more. The
genuinely falsifiable content is the scaling argument: $\Dah^{1/2} \propto
h_K$, so the permitted excess must close linearly in $h$ relative to the
bound---an accelerating relative gap closure, which is what the campaign
observed. Had the observed gap been $h$-independent, the analysis would have
been in tension with the data; consistency checks of this kind, not the
coincidence of prefactors, carry the evidential weight.
\end{pedagogy}

\begin{remark}[Hypotheses dispensed with]
\label{oa:rem:dispensed}
Two technical hypotheses of the ASGS continuity analysis of~the main text are not
needed here: the interior-face jump condition on $\tau_1$ (required there when
$\sigma > 0$, following Eq.~(39) of~\cite{codina2001stabilized}), and the enlarged
inverse-estimate constant for quantities of the form
$\nabla\cdot(\alpha\nu\ViscProj\nabla\bv_h)$. The reason is structural: the analyzed
OSGS form contains no viscous or reactive stabilization terms, so the proof never
integrates a stabilization term by parts elementwise---the only integrations by parts
are global ones, in (I2) and (I5) of \cref{oa:lem:interpolation}. In exchange, the OSGS
analysis needs the projection stability condition \cref{H:projection}, which has no ASGS
counterpart. The two hypothesis sets are thus not nested; they reflect the different
routes (coercivity for ASGS, inf--sup for OSGS).
\end{remark}

\begin{pedagogy}[The hypothesis ledger, audited]
\textbf{Audit note.} ``Dispensed'' is accurate about the \emph{proof} but is
easily over-read about the \emph{hypotheses}, and the reader should hold both
halves of the ledger. (a)~The jump condition's content---facewise comparability
of $\varphi_1$, with $\sigma$ cancelling in the jump of $\tau_1^{-1}$---is
implied by the standing hypotheses: it is \cref{oa:lem:patch} restricted to
face-adjacent elements, derived there from porosity resolution, mesh
nondegeneracy and continuity of $\ba$ over the viscous floor. So the ASGS
analysis, in assuming \cref{H:jump}, assumes nothing beyond what this appendix
proves and consumes anyway; dispensing with it is presentation economy, not a
weakening. The same holds for the enlarged inverse constant, which the ASGS
appendix itself controls through porosity-resolution factors
(\cref{rem:winvconst}). (b)~Conversely, \cref{H:projection} is \emph{not}
implied by the standing hypotheses: it is a genuine property of the finite
element pair, of Codina--Blasco type, to be verified per element family. The
honest net position: the two hypothesis sets are indeed not nested, but the
only assumption in either analysis that is not automatic under the common
standing assumptions is \cref{H:projection}, on the OSGS side. The structural
point of the remark---no elementwise integration by parts of stabilization
terms ever occurs here, because the projection annihilated the terms that
would require it---is exactly right and is the deeper fact: granting the jump
condition to the OSGS analysis would change nothing, since no $[\tau_1]$ face
term exists anywhere in this appendix for it to price.
\end{pedagogy}

\begin{remark}[A Method II variant]
\label{oa:rem:methodII}
As in~\cite[\S 2.2]{Codina2008FiniteEA}, one may control the convective and pressure-gradient
fluctuations separately, replacing $S_1$ by
\[
  S_1^{\mathrm{II}}(U,V) \coloneqq
  \bigl( \Poneo[\alpha\,\ba\cdot\nabla\bu],\ \alpha\,\ba\cdot\nabla\bv \bigr)_{\tau_1}
  + \bigl( \Poneo[\alpha\nabla p],\ \alpha\nabla q \bigr)_{\tau_1}.
\]
The diagonal then controls the two fluctuations individually, and the same arguments
(with the elementary inequality $a^2 + b^2 \ge \tfrac13(a^2+b^2) + \tfrac13(a+b)^2$
used in~\cite[Theorem~3]{Codina2008FiniteEA}) are expected to give stability in the correspondingly
finer norm and a convergence estimate of the same form as \cref{oa:eq:MainResult}. We state
this only as a prospective observation---the modified proof is not written out here---having
analyzed Method~I because it is the variant implemented in~the main text (a single
projected quantity per equation).
\end{remark}

\begin{pedagogy}[What the Method II variant would buy]
With separate projections of the convective and pressure-gradient parts, the
diagonal controls each fluctuation individually, hence a finer norm. The quoted
inequality $a^2+b^2 \ge \tfrac13(a^2+b^2)+\tfrac13(a+b)^2$ is the device for
recovering control of the \emph{combined} quantity (whose square is of
$(a+b)^2$ type) from the separate squares---the reverse of the triangle
inequality, at the cost of a factor. Analyzing Method~I instead is a fidelity
choice: it is what the code implements.
\end{pedagogy}

\begin{remark}[Robustness regimes and the screening length]
\label{oa:rem:robustness}
Repeating the regime normalizations of the robustness section of~the main text
with \cref{oa:eq:MainResult} in place of the ASGS estimate simply multiplies each regime
estimate by $(1+\Dah)^{1/2}$. In the viscous- and convection-dominated regimes
$\Dah \le 1$ on any reasonable mesh and the two variants obey identical bounds up to
constants; asymptotically, $\Dah \propto h_K^2$, so the OSGS bound exceeds the ASGS
one by $1 + O(h_K^2)$ only. The factor matters solely in the pre-asymptotic
reaction-dominated regime, and it has a clean physical reading: introducing the
porous screening length $\ell_\sigma \coloneqq (\alpha_K\nu/\sigma)^{1/2}$ (the decay
length of velocity perturbations in a Brinkman medium), one has
$\Dah = (h_K/\ell_\sigma)^2$, so the penalty persists exactly while the mesh
under-resolves $\ell_\sigma$ and disappears once $h_K \lesssim \ell_\sigma$. In the
$\mathrm{Da} = 10^6$ experiments of~the main text, $\ell_\sigma \sim 10^{-3}L$,
below the finest resolutions employed, which is why the excess is still visible there.
\end{remark}

\begin{pedagogy}[The physical dictionary]
The screening length gives the physical reading of the entire Damk\"ohler
discussion: $\Dah = (h_K/\ell_\sigma)^2$, so the penalty is precisely a
statement about \emph{resolving the Brinkman layer width} $\ell_\sigma =
(\alpha_K\nu/\sigma)^{1/2}$. Under-resolve it and the OSGS bound concedes the
factor $h_K/\ell_\sigma$; resolve it and the two analyses coincide up to
constants. This also explains why the factor is invisible in the viscous- and
convection-dominated regimes and why, asymptotically in $h$, it is a
second-order effect ($1 + O(h_K^2)$).
\end{pedagogy}

\begin{remark}[What this analysis does not show]
\label{oa:rem:open}
(i)~The numerical campaign of~the main text observes, for $k \ge 2$, an $L^2$
pressure convergence order for the OSGS variant one unit \emph{higher} than for the
ASGS variant. Nothing in the present analysis (nor in~\cite{Codina2008FiniteEA}, whose norms
contain no $L^2$ pressure term except in the viscous-dominated variant with
Reynolds-degrading constants) explains this; a duality argument exploiting the
orthogonality of the consistency error to the finite element space would be the
natural route, and remains open. (ii)~The analysis takes the projection equation
solved exactly; the lagged projection of the implemented staggered iteration, and the
nonlinear problem, are not covered. (iii)~The estimate \cref{oa:eq:MainResult} is an
upper bound; no lower bound is claimed, so the comparison with the ASGS estimate in
\cref{oa:cor:asgsnorm} orders the \emph{bounds}, not the methods. (iv)~For $\alpha \equiv 1$
and $\sigma = \varepsilon = 0$ the present theorem reduces to
\cite[Theorem~2]{Codina2008FiniteEA} with the parameter dictionary $c_3 = 1$,
$c_4 = c_2/c_1$, which serves as a consistency check of the constants.
\end{remark}

\begin{pedagogy}[Reading the open items]
Item~(i) names the one observation the theory genuinely does not reach: the
extra order of $L^2$ \emph{pressure} convergence for OSGS at $k \ge 2$. The
suggested route is the right one to attempt: Aubin--Nitsche duality arguments
convert norm-level estimates into lower-norm gains precisely when the
consistency defect carries orthogonality against the discrete space---and the
OSGS defect \cref{oa:eq:ConsistencyId} is a pairing against fluctuations,
which are by construction $\tau_1$-orthogonal to the discrete velocity space.
That orthogonality is an OSGS-specific asset with no ASGS analogue, which is
at least consistent with the phenomenon being OSGS-specific. Items
(ii)--(iii) are scope statements to keep in view when quoting the results;
item~(iv)'s reduction to \cite[Theorem~2]{Codina2008FiniteEA} with $c_3 = 1$,
$c_4 = c_2/c_1$ is the cheap but valuable sanity check that no porous-weight
bookkeeping error crept in.
\end{pedagogy}


