% =====================================================================
%  coq_coverage.tex
%
%  A reader's map from the paper's theory (theory/paper/article.tex and
%  its appendices) to the machine-checked Coq development in
%  proof_verification/coq-formal/.
%
%  Standalone document: compile with
%      pdflatex coq_coverage.tex   (run twice for cross-references)
%  or  latexmk -pdf coq_coverage.tex
%  Uses only standard TeX Live packages (amsmath, amssymb, booktabs,
%  longtable, listings, hyperref, geometry, xcolor).
% =====================================================================
\documentclass[11pt]{article}

\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{xcolor}
\usepackage{listings}
\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}

% ---- Coq identifier / label helpers -------------------------------------
% \detokenize renders underscores literally, so Coq names need no escaping.
\newcommand{\coq}[1]{\texttt{\detokenize{#1}}}
\newcommand{\lbl}[1]{\texttt{\detokenize{#1}}}
\newcommand{\file}[1]{\texttt{\detokenize{#1}}}

% Long typewriter identifiers/paths are unbreakable; let TeX break the line
% before them rather than run them into the margin.
\emergencystretch=3em
\sloppy

% ---- a lightweight Coq listing style ------------------------------------
\definecolor{coqkw}{rgb}{0.30,0.12,0.55}
\definecolor{coqcmt}{rgb}{0.30,0.45,0.30}
\lstdefinelanguage{Coq}{
  morekeywords={Theorem,Lemma,Corollary,Definition,Record,Proof,Qed,forall,exists,
    Section,Variable,Variables,Hypothesis,Hypotheses,Require,Import,From,fun,
    intros,unfold,rewrite,apply,field,split,nra,lra,psatz,pose,assert},
  sensitive=true,
  morecomment=[n]{(*}{*)},
}
\lstset{
  language=Coq,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{coqkw}\bfseries,
  commentstyle=\color{coqcmt}\itshape,
  columns=fullflexible,
  keepspaces=true,
  breaklines=true,
  frame=single,
  framesep=4pt,
  xleftmargin=2pt,
}

% ---- math shorthands matching the paper ---------------------------------
\newcommand{\sigt}{\widetilde{\sigma}}
\newcommand{\Cinv}{C_{\mathrm{inv}}}
\newcommand{\Cbar}{\bar C_{\mathrm{inv}}}
\newcommand{\aK}{\alpha_K}
\newcommand{\Reh}{\mathrm{Re}_h}
\newcommand{\Dah}{\mathrm{Da}_h}
\newcommand{\Cstab}{C_{\mathrm{stab}}}
\newcommand{\Th}{\mathcal{T}_h}          % the triangulation, as in the paper
\renewcommand{\arraystretch}{1.25}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}

\title{\bfseries Machine-checked coverage map:\\
       from the paper's theory to the Coq formalisation}
\author{Coverage note accompanying \file{proof_verification/coq-formal/}}
\date{Revised 2026-07-17 \quad (18 files, $\approx$789 theorems; all compiled
      and \coq{coqchk}-verified on Coq~8.18.0)}

\begin{document}
\maketitle

\begin{abstract}
\noindent
This note is a reader's map between two artefacts in this repository: the
mathematical theory of the paper \emph{``A stabilized finite element method
for incompressible, inertial flows in inhomogeneous porous media''}
(Casas, Gonz\'alez-Us\'ua, Codina, de-Pouplana), whose \LaTeX{} sources live
under \file{theory/paper/}, and the Coq formalisation under
\file{proof_verification/coq-formal/}, which re-proves the paper's algebraic
and analytic \emph{claims} from the axioms of the real numbers and has them
re-checked by Coq's trusted kernel. \textbf{The headline of the current
revision: the paper's entire a priori chain --- \lbl{lemma:Stability},
\lbl{lemma:Continuity}, \lbl{lem:continterp} and \lbl{thm:convergence} ---
now consists of complete machine-checked theorems}
(\coq{abstract_stability}, \coq{abstract_continuity},
\coq{abstract_continterp}, \coq{abstract_convergence}), proved over an
abstract inner-product / mesh interface from a short, \emph{named,
quantitative} trusted base of standard analytic facts --- under exactly the
manuscript's own hypotheses, $\sigma\ge0$ and $\varepsilon\ge0$ included, so
that the reaction-free case $\sigma=0$ is a genuine instance; the convergence
theorem is a genuine kernel-level \emph{composition} of the closed
stability and interpolation-continuity lemmas, with the triple-norm
triangle inequality proved rather than assumed. Section~\ref{sec:coq}
is a primer on how Coq (now \emph{Rocq}) works and what ``machine-checked''
buys. Section~\ref{sec:map} explains the translation convention --- including
how a lemma with functional-analytic hypotheses is honestly encoded.
Section~\ref{sec:crosswalk} is the crosswalk itself, one table per Coq file.
Section~\ref{sec:assumptions} is the heart of the transparency claim: it reads
out \emph{every} assumption the machine-checked theorems make --- the explicit
hypotheses, and equally the implicit ones that live in definitions rather than
in \coq{Hypothesis} lines --- each in its exact Coq form, its mathematical form,
and with a discussion of its role, its justification, and whether it could be
tightened; it also settles the question of whether the hypothesis bundle is
consistent at all. Section~\ref{sec:left} then states what is \emph{not} inside
the kernel and by what other means (a hand audit, the SymPy suite, a Lean~4
roadmap, or the convergence tests) each residual item is addressed.
Section~\ref{sec:amend} records the theory-side amendments applied to keep the
manuscript $100\%$ compatible with what was verified.
Section~\ref{sec:conclusion} states, in plain terms, how strong the resulting
guarantee is and where residual trust remains.
\end{abstract}

\tableofcontents

% =====================================================================
\section{Purpose and scope}\label{sec:purpose}
% =====================================================================

The Coq development mirrors, file by file, the SymPy ``verification scripts''
suite in \file{proof_verification/sympy/}: where SymPy checks an identity
by symbolic computation, the Coq files \emph{prove} it. The suite is
\textbf{18 files, $\approx 789$ theorems, all proved} --- there is no
\coq{Admitted}, no \coq{Axiom} and no \coq{admit} anywhere in the sources
(\coq{grep} confirms this, and so, more strongly, does \coq{Print Assumptions}:
\S\ref{sec:stratum0}). It uses \emph{only} the standard library, and the whole
suite has been compiled and independently kernel-checked on Coq~8.18.0: eighteen
\coq{[PASS]} lines followed by \emph{``coqchk: modules successfully checked by
the trusted kernel''} (\S\ref{sec:repro}).

What changed since the previous revision is a step up in \emph{kind}, not just
in count. Previously the Coq layer certified the \emph{algebraic skeleton} of
the two central lemmas (parameter identities, coefficient expansions, the jump
algebra) while the surrounding \emph{functional-analytic} argument (testing
with a discrete field, integration by parts, inverse estimates, summation over
the mesh) was only hand-audited. Four new files
(\file{AbstractSums.v}, \file{InnerSpace.v}, \file{AbstractStability.v},
\file{AbstractContinuity.v}) close that gap: they prove \lbl{lemma:Stability}
and \lbl{lemma:Continuity} \emph{as complete theorems} over an abstract
inner-product / mesh interface, reducing each to a short list of
explicitly-named standard analytic facts (Green-type identities and
inverse-type estimates) that appear as \emph{hypotheses} of the abstract
theorem. The elemental and discrete Cauchy--Schwarz inequalities, the Young
step, the difference-of-squares cancellation, all eighteen term bounds, the
jump treatment, and the assembly over elements and faces are now inside the
kernel.

The present revision closes the last remaining corner. Two new files ---
\file{AbstractInterpolation.v} and \file{AbstractConvergence.v} --- prove
\lbl{lem:continterp} and \lbl{thm:convergence} as complete theorems. The
first reruns the continuity machinery with the first argument equal to the
interpolation error, exactly as the appendix's proof prescribes: the
discrete inverse estimates on the first slot are replaced by the
interpolation estimates \lbl{eq:interpdivvisc}--\lbl{eq:interpinftyE}
(seven named hypotheses \coq{HI_*}, one constant), while the discrete
second argument keeps its hypotheses verbatim. The second composes the two
closed lemmas \emph{inside the kernel}: \coq{abstract_stability} is
applied to the discrete error, \coq{abstract_continterp} to the
(interpolation error, discrete error) pair, the two are glued by Galerkin
orthogonality (\coq{Horth}) --- the tested identity for the
discrete-error pair, formerly the hypothesis \coq{HBS_W}, is no longer needed at
all, since \coq{abstract_stability} is applied at those atoms and carries it ---
and the triangle inequality for the mesh-dependent triple norm is \emph{proved}
from the pre-Hilbert axioms.
Even the agreement of the $\tau_1,\tau_2,\sigt$ formulas across the two
scalar-algebra modules is machine-checked (\coq{tau1_agree},
\coq{tau2_agree}, \coq{sigt_agree}).

This revision also \emph{shrinks} the trusted base rather than merely
describing it: from $53$ named hypotheses to $50$. Exactly \emph{three} rows
leave, \emph{no} row is added, and one hypothesis is strengthened to pay for two
of the departures. (i) The tested identity \coq{HBS} is \emph{proved} --- from
the two \emph{diagonal} Green identities it was silently bundling --- so the
class-G bundle leaves the base and the two encodings of $B_S$ are reconciled
inside the kernel. The previous revision described \coq{HBS} here as the single
largest assumption in the development; that label was wrong and is
\emph{withdrawn} --- the row was a free real pinned by its own defining equation,
hence logically eliminable --- and \S\ref{sec:stratumG}~G1 is the honest
accounting of what its removal did and did not buy. The two diagonal identities
that discharge it are \emph{not} new to the inventory: \file{AbstractConvergence.v}
already carried them (items 28--29), so the stability file now \emph{shares} them
rather than adding them, at the price of the one \emph{increase} in the
development --- its own count goes $16\to17$. (ii) The nonnegativity of the two
interpolation-error functionals, previously assumed
($E_{\mathrm{int},K}(\mathbf u)\ge0$, $E_{\mathrm{int},K}(p)\ge0$), is
\emph{derived} from the interpolation estimates and the pre-Hilbert axioms --- at
the price of strengthening $C_I\ge0$ to $C_I>0$, a trade we flag rather than
bank (\S\ref{sec:stratumD}, D1).

Earlier tightenings still stand, and are recorded here because the development
rests on them --- not because this revision performed them. \coq{HBS}'s
convergence twin \coq{HBS_W} left the trusted base one revision earlier; what
\emph{this} revision adds is that even the theorem which replaced it is gone, the
convergence file now \emph{applying} \coq{abstract_stability} at the
discrete-error atoms, so one proof serves both diagonals. The two elemental
integration-by-parts hypotheses are \emph{proved}: their face terms were free
variables, hence constrained nothing, and are now defined as the
integration-by-parts defect. The strictness $\sigma>0$, $\varepsilon>0$,
$C_2>0$ is gone: the development assumes exactly \lbl{H:data}, so the
reaction-free case $\sigma=0$ is a genuine instance. The numerical-parameter
condition is assumed in its \emph{sharp} form $c_1>\xi\Cbar^2$, with a separate
theorem (\coq{C_stab_margin}) recording what the manuscript's two-fold stronger
condition buys. And the analysed-versus-implemented $\tau_2$ bridge is wired
end to end: \coq{abstract_convergence_implemented} states \lbl{thm:convergence}
for the parameter the solver actually forms, with the constant inflated by at
most $\sqrt{1+C_2}<\sqrt2$. Section~\ref{sec:tighten} is the ledger.

Three further files --- \file{NonVacuity.v}, \file{NonVacuityInterp.v} and
\file{NonVacuityConv.v} --- answer the one question a hypothesis-heavy
formalisation always invites: \emph{could the hypotheses be contradictory,
making the theorems vacuously true?} Each exhibits an explicit instance --- the
carrier $\mathbb R$ with $\langle x,y\rangle=xy$, a small mesh, and rational data
--- and lets the kernel check that it satisfies every hypothesis of an abstract
theorem, yielding a \emph{non-degenerate} conclusion with both sides strictly
positive (for \coq{abstract_stability}, $B_S=\tfrac78\ge\Cstab|||U_h|||^2=\tfrac7{16}$).
Three of the four abstract theorems are covered: \coq{abstract_stability} ---
with a \emph{second} witness carrying a strictly positive advection field
$|\mathbf a|_\infty=1$ --- \coq{abstract_continterp} and
\coq{abstract_convergence}, the last two being the first witnesses for the
strengthened hypothesis \coq{CI_pos} ($C_I>0$). \coq{abstract_continuity} is
the one theorem still resting on a hand argument. Consistency of these bundles
is thus itself machine-checked, not argued (\S\ref{sec:vacuity}).

Two boundaries still frame everything.
\begin{itemize}[nosep]
  \item \textbf{Coverage boundary.} Coq certifies the mathematics down to a
    named trusted base. That base --- the divergence-theorem / integration-by-parts
    identities and the finite-element inverse estimates --- is textbook material
    (Brenner--Scott, Ern--Guermond), hand-audited against the manuscript and
    slated for a Lean~4 formalisation
    (Section~\ref{sec:left}). It is \emph{stated quantitatively} as hypotheses,
    so the abstract theorems are genuine implications, not sketches. The
    base now also contains the interpolation estimates of
    \lbl{eq:interp}/\lbl{eq:interpinfty} (Bramble--Hilbert technology),
    which enter \lbl{lem:continterp} as the named hypotheses \coq{HI_*},
    and the Galerkin-orthogonality item \coq{Horth} of the convergence
    file.
  \item \textbf{Meaning boundary.} Coq verifies the \emph{mathematics as
    stated}: that each identity, inequality or limit follows from the
    real-number axioms and the named base. It does \emph{not} certify that the
    statements model the physics, nor that the Julia/Kratos solver implements
    them --- that is the job of the method-of-manufactured-solutions
    convergence tests elsewhere in the codebase.
\end{itemize}

% =====================================================================
\section{A short primer: how Coq works and what ``verified'' means}\label{sec:coq}
% =====================================================================

\subsection{Proof assistant, kernel, and the de Bruijn criterion}

Coq (renamed \emph{Rocq} from version~9) is an interactive proof assistant
built on a small, fixed logic --- the Calculus of Inductive Constructions. A
Coq \emph{proof} is not prose: it is a formal term whose \emph{type} is the
statement being proved (the ``propositions-as-types'' correspondence). The
user constructs that term interactively with \emph{tactics}, but tactics are
untrusted conveniences; what matters is the object they build.

The trust rests on the \textbf{de Bruijn criterion}: a large, human-friendly
front end may do arbitrary heuristics, but the final proof term is re-examined
by a \emph{small, fixed kernel} that only knows the typing rules of the logic.
If the kernel accepts the term, the theorem holds. In this development the
kernel is exercised twice:
\begin{enumerate}[nosep]
  \item \coq{coqc} type-checks each file as it compiles it;
  \item \coq{coqchk} --- a \emph{separate}, independent re-implementation of
    the kernel --- re-verifies the compiled objects (\coq{.vo} files) from
    scratch, with no trust in \coq{coqc}. The build script runs it and reports
    \emph{``coqchk: modules successfully checked by the trusted kernel''}.
\end{enumerate}

\subsection{The trust base --- and why ``no \texttt{Admitted}, no \texttt{Axiom}'' matters}

Believing a Coq theorem means trusting exactly three things: the logic itself
(the Calculus of Inductive Constructions, studied for decades), the kernel
implementation, and the \emph{statement} --- one must still read what was
proved, including its hypotheses. Everything else (tactics, notations, the
elaborator) cannot make a false theorem be accepted.

Two escape hatches would silently widen that base, and both are audited out
here:
\begin{itemize}[nosep]
  \item \coq{Admitted} / \coq{admit} closes a goal \emph{without} a proof; a
    single one would let anything ``follow''.
  \item \coq{Axiom} postulates a statement as true without proof.
\end{itemize}
Neither appears anywhere in the eighteen files. Consequently every result below
is reduced to the standard library's real numbers --- constructed, not
postulated, but resting on a small fixed set of classical principles
(\S\ref{sec:stratum0}) --- plus, for the four abstract theorems, the
explicitly-listed analytic hypotheses of Section~\ref{sec:left}. The suite uses \emph{no external libraries} (no
Mathcomp, no Coquelicot): the analysis is done on \coq{Coq.Reals} with the
arithmetic tactics \coq{lra}/\coq{nra} (linear and nonlinear real arithmetic)
and \coq{field} (rational-function identities). Those tactics are
decision/normalisation procedures whose \emph{output} is a kernel-checked
term --- using them does not weaken the guarantee.

\subsection{The abstract-interface method (new in this revision)}

The two main lemmas are proved not against a concrete function space but
against a \emph{minimal algebraic interface} that the concrete objects satisfy.
Two devices carry this.

\emph{A pre-Hilbert record.} \file{InnerSpace.v} models the elemental
$L^2(K)$ pairings by a four-field record: a carrier with addition and scalar
multiplication and a form \coq{ip} that is symmetric, additive and homogeneous
in its first argument, and positive-semidefinite.

\begin{lstlisting}
Record PreHilbert := mkPreHilbert {
  carrier  :> Type;
  vadd     : carrier -> carrier -> carrier;
  vscal    : R -> carrier -> carrier;
  ip       : carrier -> carrier -> R;
  ip_sym   : forall x y, ip x y = ip y x;
  ip_add_l : forall x y z, ip (vadd x y) z = ip x z + ip y z;
  ip_scal_l: forall a x y, ip (vscal a x) y = a * ip x y;
  ip_pos   : forall x, 0 <= ip x x }.
\end{lstlisting}

\noindent
From these four axioms the file \emph{derives} (does not assume) full
bilinearity, the Cauchy--Schwarz inequality (by the discriminant argument,
with the degenerate direction handled), the norm expansions of sums and
differences, the triangle inequality, the parametrized Young inequality at
vector level, and the difference-of-squares identity
$\langle -a+b,\,a+b\rangle = \lVert b\rVert^2 - \lVert a\rVert^2$ that drives
the ASGS cancellation. Any concrete weighted $L^2(K)$ is an instance, so the
proofs transfer for free.

\emph{A named trusted base.} A lemma such as \lbl{lemma:Stability} needs
genuinely analytic inputs --- a Green-type testing identity and a weighted
inverse estimate --- that stdlib Coq cannot prove without the variational setup
and finite-element scaling theory. Rather than pretend otherwise, each is
written as an explicit, \emph{quantitative} \coq{Hypothesis} of the theorem
(for \coq{abstract_stability} there are exactly four: the two diagonal Green
identities \coq{H_skew_diag} and \coq{H_ibp_diag}, the weighted inverse estimate
\coq{S3}, and the $\varepsilon$-condition \coq{Heps}; see
\S\ref{sec:assumptions}). The theorem is then a real implication:
\emph{if} the inverse estimate holds with the stated constant, \emph{then} the
stability bound holds with the stated constant. This is the honest encoding of
a ``cited-standard'' fact, and it makes the residual trusted base a finite,
readable list (Section~\ref{sec:left}) rather than a gap.

\subsection{Reading Coq syntax}\label{sec:syntax}

Table~\ref{tab:syntax} is enough to read every entry in the crosswalk.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}L{0.34\textwidth}L{0.58\textwidth}@{}}
\toprule
\textbf{Coq} & \textbf{Meaning} \\
\midrule
\coq{Definition n : R := e.}       & Names a real-valued quantity $n := e$. \\
\coq{Record R := mk { f : T; ax : P }.} & A structure bundling fields \coq{f} with axioms \coq{ax} they satisfy. \\
\coq{Theorem}/\coq{Lemma}/\coq{Corollary n : P.} & Asserts $P$; the following \coq{Proof. ... Qed.} is its (kernel-checked) proof. \\
\coq{Section S. ... End S.}        & A scope sharing common parameters and hypotheses. \\
\coq{Variables (x y : R).}         & Introduces symbolic reals $x,y$ for the section. \\
\coq{Hypothesis h : 0 < x.}        & A standing assumption; every theorem in the section is conditioned on it. \\
\coq{forall x, ...} / \coq{exists x, ...} & Universal / existential quantifier. \\
\coq{<< x , y >>}                  & The pre-Hilbert inner product \coq{ip x y}; \coq{nrm x} $= \sqrt{\langle x,x\rangle}$. \\
\coq{Rsum Th f}                    & The finite sum $\sum_{k\in\mathtt{Th}} f(k)$ over the mesh element (or face) list. \\
\coq{a <= b}, \coq{a < b}, \coq{a = b} & Order and equality on $\mathbb{R}$. \\
\coq{sqrt}, \coq{Rabs}, \coq{exp}, \coq{sin} & Standard-library real functions. \\
\coq{derivable_pt_lim f x d}       & ``$f$ has derivative $d$ at $x$'' (the stdlib derivative predicate). \\
\bottomrule
\end{tabular}
\caption{The Coq vocabulary used in the statements.}
\label{tab:syntax}
\end{table}

A note on total division. In \coq{Coq.Reals}, $x/0$ is defined and equals $0$;
the development never relies on this --- every division either is harmless by
construction (its statement carries no hypothesis and the denominator is a
declared positive quantity) or is guarded by an explicit non-vanishing
hypothesis.

% =====================================================================
\section{How paper statements map to Coq}\label{sec:map}
% =====================================================================

\subsection{Standing hypotheses become section variables}

The paper carries global positivity conventions --- mesh size $h>0$, viscosity
$\nu>0$, elemental porosity $\aK>0$ (with $\aK\le 1$ in the appendix),
$\sigma\ge 0$, non-dimensional numbers $\Reh,\Dah\ge 0$, constants
$c_1,c_2,\Cinv,\xi>0$, profile $0<\alpha_0<1$. In Coq these are the
\coq{Hypothesis} lines of each \coq{Section}; every theorem there is read as
``under these positivity assumptions, \dots''.

\subsection{Naming dictionary}

The Coq identifiers transliterate the paper's symbols. The recurring ones:

\begin{center}
\small
\begin{tabular}{@{}llll@{}}
\toprule
Paper & Coq & Paper & Coq \\
\midrule
$\tau_1$ (\lbl{eq:Tau1Final})       & \coq{tau1}       & $\varphi_1=\aK\,\tau_{1,\mathrm{NS}}^{-1}$ & \coq{phi1} \\
$\tau_2$ (\lbl{eq:Tau2Final})       & \coq{tau2}       & $\tau_{1,\mathrm{NS}}$ (\lbl{eq:TauNavierStokes}) & \coq{tau1NS} \\
$\sigt$ (\lbl{eq:SigmaAlpha})       & \coq{sigt}/\coq{sg} & $\bar C_{\mathrm{inv}}$ (\lbl{lem:winv}) & \coq{Cb} \\
$\aK$                               & \coq{alphaK}/\coq{aK}     & $|\mathbf a|_{\infty,K}$   & \coq{am} \\
$\Cinv$ (\lbl{eq:inverse})          & \coq{Cinv}       & $\xi$ (Young parameter)   & \coq{xi} \\
$\Reh,\Dah$                         & \coq{Re},\coq{Da} & $\varepsilon$ (mass penalty) & \coq{eps} \\
$B_S(U_h,U_h)$                      & \coq{BS}         & $|||U_h|||^2$ (\lbl{eq:triplenorm}) & \coq{NormSq} \\
\bottomrule
\end{tabular}
\end{center}

\subsection{A worked example, read side by side}

The sharpest scalar statement of the stability section is the \emph{velocity
slack identity}: the gap between the velocity coefficient of the final
stability estimate (\lbl{eq:VelocityCoefficientBound}) and $C_u\,\sigt$ is not
merely non-negative --- it equals a specific product \emph{exactly}. In Coq:

\begin{lstlisting}
Definition C_u : R := 1 - xi * Cinv^2 / c1.

Theorem velocity_slack_identity :
  u_855 - C_u * sigt
    = alphaK * tau1 * sigma * (xi * Cinv^2 / c1) * (c2 * amag / h).

Corollary velocity_coefficient_lower_bound : u_855 >= C_u * sigt.
\end{lstlisting}

\noindent
Read: \emph{under the section hypotheses}, the difference of the two
real-valued expressions equals $\aK\tau_1\sigma\,(\xi\Cinv^2/c_1)(c_2|\mathbf
a|/h)\ge 0$; hence the lower bound. The proof rewrites with the identity
$\sigt=\sigma\varphi_1\tau_1$ and discharges the rational-function equality
with \coq{field}.

\subsection{How a lemma with analytic hypotheses maps}\label{sec:absexample}

The abstract theorems read the same way, with the trusted-base facts as named
hypotheses. \lbl{lemma:Stability} is \coq{abstract_stability}: under the
section's positivity plus $c_1>\xi\Cbar^{2}$ (the \emph{sharp} form of
\lbl{eq:conditions_on_num_param}; the manuscript's $c_1>2\xi\Cbar^2$ implies it
--- \S\ref{sec:stratumD}, D3), $\xi>2$, $C_2<1$, the three analytic hypotheses
\coq{H_skew_diag}, \coq{H_ibp_diag} and \coq{S3} shown below, and the
$\varepsilon$-condition \coq{Heps} (class~D, item~21 of
Table~\ref{tab:inventory})

\begin{lstlisting}
(* the two DIAGONAL Green identities: eq:skew and eq:globalibp with both
   slots carrying the same field.  Together they are what used to be
   bundled inside the monolithic Hypothesis HBS, which is now a Theorem
   (see sec:stratumG, G1). *)
Hypothesis H_skew_diag :
  Rsum Th (fun k => << uu k , cxu k >>) = 0.

Hypothesis H_ibp_diag :
  Rsum Th (fun k => << uu k , gpu k >>)
  = - Rsum Th (fun k => << pp k , divu k >>).

(* S3: the weighted inverse estimate eq:winv-divvisc (lem:winv) *)
Hypothesis S3 :
  forall k, nrm (du k) <= (2*nu*Cb / hK k) * sqrt (aK k) * nrm (gu k).
\end{lstlisting}

\noindent
the theorem concludes
\begin{lstlisting}
Theorem abstract_stability :
  0 < C_stab c1 Cb xi C2  /\  BS >= C_stab c1 Cb xi C2 * NormSq.
\end{lstlisting}

\noindent
which is exactly $B_S(U_h,U_h)\ge \Cstab\,|||U_h|||^2$ with the explicit
$\Cstab=\min\{C_{\mathrm{visc}},C_u,1-C_2,1\}>0$. Everything between the
hypotheses and the conclusion --- the difference-of-squares cancellation, the
expansion of $\lVert G(u)\rVert^2$, the elemental Cauchy--Schwarz, the Young
step with parameter $\xi\bar C_{\mathrm{inv}}^2\nu\aK/h^2$, the exact
cancellation of the inverse-estimate constant, the coefficient collection, and
the summation over the mesh --- is machine-checked.

\subsection{What a green check does and does not assert}

An entry marked as covered means: \emph{the displayed relation is a theorem of
real arithmetic under the stated hypotheses}. For the algebraic files the
hypotheses are only positivity, so the theorem is unconditional in the
mathematics. For the four abstract theorems the hypotheses additionally include
the named analytic base of Section~\ref{sec:left}; the green check asserts the
\emph{implication}, and the honesty of the result rests on that base being
exactly the standard, hand-audited facts it is labelled as.

% =====================================================================
\section{The crosswalk: theory \texorpdfstring{$\to$}{->} Coq}\label{sec:crosswalk}
% =====================================================================

Dependency and topic map of the eighteen files. \file{Limits.v}/\file{DerivKit.v}
serve the concrete analytic files; \file{AbstractSums.v}/\file{InnerSpace.v}
serve the four abstract-theorem files, which also import the closed algebraic
theorems of \file{StabilityAlgebra.v}/\file{ContinuityAlgebra.v} and apply them
per element.

\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
Coq file & Paper location & Role \\
\midrule
\file{Limits.v}      & --- & infrastructure: $\varepsilon$--$\delta$ limit theory \\
\file{DerivKit.v}    & --- & infrastructure: derivative toolkit \\
\file{AbstractSums.v}  & --- & infrastructure: finite mesh/face sums, discrete C--S \\
\file{InnerSpace.v}    & --- & infrastructure: abstract pre-Hilbert space \\
\file{InverseEstimates.v} & \lbl{lem:winv} & infrastructure: inverse-estimate schema \& composition \\
\file{StabilityAlgebra.v}  & \S\ref{sec:crossStab} (\S5)     & stability estimate, scalar algebra \\
\file{ContinuityAlgebra.v} & \S\ref{sec:crossCont} (App.\ B) & continuity proof, scalar algebra \\
\file{AbstractStability.v} & \S\ref{sec:crossAbsStab} (\S5) & \lbl{lemma:Stability}, complete theorem \\
\file{AbstractContinuity.v}& \S\ref{sec:crossAbsCont} (App.\ B) & \lbl{lemma:Continuity}, complete theorem \\
\file{AbstractInterpolation.v}& \S\ref{sec:crossAbsInterp} (App.\ B) & \lbl{lem:continterp}, complete theorem \\
\file{AbstractConvergence.v}& \S\ref{sec:crossAbsConv} (App.\ B) & \lbl{thm:convergence}, complete theorem \\
\file{TauDesign.v}   & \S\ref{sec:crossTau} (\S4 + App.\ C) & Fourier design of $\tau_1,\tau_2$ \\
\file{Asymptotics.v} & \S\ref{sec:crossAsym} (\S6)     & robustness in $(\Reh,\Dah)$ \\
\file{ManufacturedSolution.v} & \S\ref{sec:crossMMS} (\S7) & manufactured solution \\
\file{PlateauBump.v} & \S\ref{sec:crossBump} (\S7)     & plateau bump, through $C^1$ \\
\file{NonVacuity.v}  & \S\ref{sec:vacuity}             & consistency witness for \lbl{lemma:Stability} (two instances) \\
\file{NonVacuityInterp.v} & \S\ref{sec:vacuity}        & consistency witness for \lbl{lem:continterp} \\
\file{NonVacuityConv.v}  & \S\ref{sec:vacuity}         & consistency witness for \lbl{thm:convergence} \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Infrastructure: \texttt{InnerSpace.v}, \texttt{InverseEstimates.v} and \texttt{AbstractSums.v}}\label{sec:crossInfra}

\begin{longtable}{@{}L{0.42\textwidth}L{0.53\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Content (all \emph{derived}, not assumed)} \\
\midrule \endhead
\multicolumn{2}{@{}l}{\emph{\file{InnerSpace.v} --- the elemental $L^2(K)$ pairing as a pre-Hilbert space}}\\
\coq{PreHilbert} (\coq{ip_sym}, \coq{ip_add_l}, \coq{ip_scal_l}, \coq{ip_pos}) & the four axioms modelling a symmetric positive-semidefinite $L^2(K)$ form. \\
\coq{ip_add_r}, \coq{ip_scal_r}, \coq{ip_sub_l/r}, \coq{ip_opp_l/r} & full bilinearity, from left-linearity $+$ symmetry. \\
\coq{CS_sq}, \coq{CS}, \coq{CS_le}, \coq{CS_ge} & Cauchy--Schwarz $|\langle x,y\rangle|\le\lVert x\rVert\lVert y\rVert$ (discriminant argument). \\
\coq{nrm_sq}, \coq{ip_expand_add}, \coq{ip_expand_sub} & Gram expansions $\lVert x\pm y\rVert^2$. \\
\coq{nrm_triangle} & the triangle inequality. \\
\coq{young_scalar}, \coq{young_vector} & the parametrized Young inequality at vector level. \\
\coq{diff_of_squares} & $\langle -a+b,\,a+b\rangle=\lVert b\rVert^2-\lVert a\rVert^2$: the ASGS cancellation. \\
\midrule
\multicolumn{2}{@{}l}{\emph{\file{InverseEstimates.v} --- the \lbl{lem:winv} weighted inverse-estimate shape, named once (full account: \S\ref{sec:stratumI}, I1)}}\\
\coq{winv_est}\,$C\,W\,A\,B$ & the common shape $\lVert A_K\rVert_K\le(C/h_K)\,W_K\,\lVert B_K\rVert_K$ that the nine \coq{Hw_*} hypotheses each instantiate; a \emph{notational} predicate, not a hypothesis --- it replaces none and adds none, so Table~\ref{tab:inventory} keeps its row count. \\
\coq{winv_compose} & composes two \coq{winv_est} instances sharing a middle atom into the double estimate $\lVert A_K\rVert_K\le C_1C_2(W_1W_2)_K\,h_K^{-2}\,\lVert B_K\rVert_K$; the single proved lemma that retires the byte-identical \coq{double_inv_u}/\coq{double_inv_v} triplication. \\
\midrule
\multicolumn{2}{@{}l}{\emph{\file{AbstractSums.v} --- sums over the element / face list}}\\
\coq{Rsum}, \coq{Rsum_plus/minus/opp/scal}, \coq{Rsum_le}, \coq{Rsum_nonneg} & linearity and monotonicity of $\sum_{k\in\mathrm{Th}}$. \\
\coq{Rsum_abs_le} & the triangle inequality for sums. \\
\coq{Rsum_CS_sq}, \coq{Rsum_CS} & the discrete Cauchy--Schwarz over the mesh, by induction (the paper's ``C--S for the sums over elements''). \\
\coq{Rsum_sq_mono}, \coq{sqrt_plus_le}, \coq{sqrt_mono} & squared-sum monotonicity; subadditivity/monotonicity of $\sqrt{\cdot}$. \\
\bottomrule
\end{longtable}

\subsection{\texttt{StabilityAlgebra.v} --- \S5: the stability estimate (scalar algebra)}\label{sec:crossStab}

Section hypotheses: $\nu,h,\aK>0$, $\sigma\ge0$, $|\mathbf a|\ge 0$,
$c_1,c_2,\Cinv,\xi>0$; and, in its second section, $\varepsilon\ge0$,
$C_2\ge0$ --- exactly \lbl{H:data} (\S\ref{sec:stratumD}, D1).
Mirrors \file{stability_estimate_verification.py}.

\emph{A naming caution.} \file{StabilityAlgebra.v}'s section variable is spelled
\coq{Cinv}, but it is a \emph{slot}: every use in the abstract layer fills it
with \coq{Cb} $=\Cbar$ (every instantiation in \file{AbstractStability.v} does
so). The constants below are therefore written with $\Cbar$ --- which is exactly the
substitution \lbl{rem:winvconst} prescribes, and what \lbl{prop:stability}
states. (Renaming the Coq variable to \coq{Cb} would remove the trap.)

\begin{longtable}{@{}L{0.42\textwidth}L{0.53\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Paper content} \\
\midrule \endhead
\coq{sigt_form_phi}, \coq{sigt_form_tau1}, \coq{sigt_form_key} & The three forms of $\sigt$ (\lbl{eq:SigmaAlpha}). \\
\coq{young_perfect_square}, \coq{young_inequality} & Young's step $-2xy\ge -x^2/\xi-\xi y^2$. \\
\coq{viscous_coefficient_expansion} & Viscous coefficient of \lbl{eq:StabilityEstimateFinal} $=$ the expanded \lbl{eq:ViscousCoefficientBound}. \\
\coq{velocity_coefficient_expansion}, \coq{velocity_slack_identity}, \coq{velocity_coefficient_lower_bound}, \coq{u_final_lower_bound} & \lbl{eq:VelocityCoefficientBound} and its \emph{exact} slack over $C_u\sigt$, $C_u=1-\xi\Cbar^2/c_1$. \\
\coq{eps_max_tau2_identity}, \coq{eps_tau2_le_C2}, \coq{pressure_term_coercive}, \coq{epsilon_smallness_chain} & The $\varepsilon$-smallness chain (\lbl{eq:UpperBoundOnEpsilon}): $\varepsilon\tau_2\le C_2$, pressure coeff $\ge(1-C_2)\varepsilon$. \\
\coq{viscous_coefficient_lower_bound}, \coq{visc_final_lower_bound} & Viscous bracket $\ge C_{\mathrm{visc}}\nu$, $C_{\mathrm{visc}}=\min\{2-4\Cbar^2/c_1,2(1-2/\xi)\}$. \\
\coq{stability_constants_positive_sharp} & $c_1>\xi\Cbar^2$, $\xi>2$ $\Rightarrow$ coefficients positive: the \emph{sharp} form of \lbl{eq:conditions_on_num_param} (a factor two weaker than the manuscript's). \\
\coq{stability_constants_positive} & the manuscript's $c_1>2\xi\Cbar^2$, $\xi>2$; a corollary of the sharp form. \\
\coq{C_stab_margin} & what the manuscript's extra factor of two buys: $C_u>\tfrac12$, $2-4\Cbar^2/c_1>1$, hence the $\Cbar$-free floor $\Cstab\ge\min\{2(1-2/\xi),\tfrac12,1-C_2\}$. \\
\coq{elemental_coercivity} & Elemental assembly $\ge\Cstab\times$ the triple-norm contributions, $\Cstab=\min\{C_{\mathrm{visc}},C_u,1-C_2,1\}$. \\
\bottomrule
\end{longtable}

\subsection{\texttt{ContinuityAlgebra.v} --- Appendix B: the continuity proof (scalar algebra)}\label{sec:crossCont}

Section hypotheses: $\nu,h,\aK,c_1,c_2>0$, $\aK\le 1$, $\sigma\ge 0$,
$|\mathbf a|\ge 0$. Explicit constants wherever the appendix writes a generic
$C$.

\begin{longtable}{@{}L{0.44\textwidth}L{0.51\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Appendix content} \\
\midrule \endhead
\coq{P1_*}, \coq{P2_*}, \coq{P3_*}, \coq{P4_*}, \coq{P5_*} & \lbl{lem:parameters} (P1)--(P5) and the square-root forms. \\
\coq{sigt_id_tau}, \coq{sigt_id_sub}, \coq{one_minus_sigma_tau1} & \lbl{eq:sigmatilde} and the workhorse $1-\sigma\tau_1=\varphi_1\tau_1$. \\
\coq{keyvisc_sq}, \coq{keyvisc_sqrt}, \coq{T8_chain}, \coq{T13_chain} & \lbl{eq:keyvisc} (Step~4); the $\sigma\varphi_1\tau_1=\sigt$ absorptions (\lbl{eq:T8bound}, \lbl{eq:T13conv}). \\
\coq{volpart_scalar}, \coq{sqrtphi_tau_le}, \coq{phi_sqrttau_le}, \coq{sqrtphi_sqrttau_le_1} & \lbl{eq:volpart} (Step~6b) and Steps~6c--6d. \\
\coq{absorb1_*}, \coq{absorb2_*}, \coq{absorb4_*} & the \lbl{eq:absorb1}--\lbl{eq:absorb4} coefficient chains (Step~9). \\
\coq{jump_formula}, \coq{phi_diff_bound}, \coq{tau_comp_AB/_BA}, \coq{sg_comp_A_le_B}, \coq{jump_bound_A}, \coq{jumpsplit_AB}, \coq{III_absorb_AB/_BA} & \lbl{lem:jump}, \lbl{eq:jumpest}, \lbl{eq:jumpsplit}, Step~6d --- constants explicit ($C_J/c_J$). \\
\coq{delta_alpha_bound}, \coq{winv_ratio} & \lbl{H:porosity}'s derived bound and the \lbl{lem:winv} coefficient step $\alpha_\infty/\alpha_0^{1/2}\le\delta^{1/2}\alpha_\infty^{1/2}$. \\
\coq{lagrange2/3}, \coq{CS2}, \coq{CS3}, \coq{step1_bound}, \coq{norm5_absorption} & the finite Cauchy--Schwarz of Steps~1,~8 and the Step~9 squaring-and-summing. \\
\coq{tau2impl_le_tau2}, \coq{tau2_le_scaled_impl} & analysis-$\tau_2$ (\lbl{eq:Tau2Final}) vs implemented $\tau_2$ (\lbl{eq:Tau2}): equivalent up to $1+C_2$ (amendment~F6, \S\ref{sec:amend}). \\
\bottomrule
\end{longtable}

\subsection{\texttt{AbstractStability.v} --- \lbl{lemma:Stability} as a complete theorem}\label{sec:crossAbsStab}

Models each field's elemental restriction as a \coq{PreHilbert} vector and the
parameters via the closed \file{StabilityAlgebra.v} formulas. Conclusion:
\coq{abstract_stability} (see \S\ref{sec:absexample}). Trusted base:
\coq{H_skew_diag}, \coq{H_ibp_diag}, \coq{S3}, \coq{Heps} (the four of
Section~\ref{sec:left}). $B_S(U_h,U_h)$ is \emph{not} among them: it is a
\coq{Definition} --- \file{AbstractInterpolation.v}'s eighteen-term expression
at the diagonal atoms --- and the tested identity is the \emph{theorem}
\coq{HBS}, proved from the first two (\S\ref{sec:stratumG}, G1). This is why the
file imports \file{AbstractInterpolation.v}.

\begin{longtable}{@{}L{0.40\textwidth}L{0.55\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Paper content (all machine-checked)} \\
\midrule \endhead
\coq{momentum_pairing}, \coq{mass_pairing} & the adjoint / difference-of-squares expansions of the residual pairings (\lbl{eq:StabilityEstimate}). \\
\coq{Bv_expand}, \coq{du_sq_bound} & expansion of $\lVert G(u)\rVert^2$; the Young step with the exact cancellation of the inverse-estimate constant $\bar C_{\mathrm{inv}}$ (\lbl{eq:BoundOfCrossedTerm}). \\
\coq{element_lower_bound}, \coq{element_coercive} & the per-element coercivity, importing \coq{elemental_coercivity}. \\
\coq{diag_zero}, \coq{momentum_group}, \coq{mass_group}, \coq{HBS} & the tested identity $B_S(U_h,U_h)$ of \lbl{eq:StabilityEstimate}, \textbf{proved} (was the hypothesis \coq{HBS}): the eighteen-term $B_S$ at the diagonal is re-assembled into the stability form, the $T_2+T_4$ remainder vanishing by the two diagonal Green identities. \\
\coq{BS_as_sum} & re-summation of the tested form over the mesh. \\
\coq{abstract_stability} & \textbf{\lbl{lemma:Stability}}: $0<\Cstab$ and $B_S(U_h,U_h)\ge\Cstab\,|||U_h|||^2$. \\
\bottomrule
\end{longtable}

\subsection{\texttt{AbstractContinuity.v} --- \lbl{lemma:Continuity} as a complete theorem}\label{sec:crossAbsCont}

Adds a face list with the two neighbouring elements $e_1,e_2$ per interior
face. Conclusion: \coq{abstract_continuity}. Trusted base: the IBP / facewise
/ face-estimate / jump / mesh / inverse-estimate / $\varepsilon$ hypotheses of
Section~\ref{sec:left}.

\begin{longtable}{@{}L{0.40\textwidth}L{0.55\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Appendix content (all machine-checked)} \\
\midrule \endhead
\coq{BS} $= $ \coq{T1} $+\cdots+$ \coq{T18} & the eighteen-term decomposition of $B_S$ (\lbl{eq:Bstab}, Step~0). \\
\coq{bound_T1}, \coq{bound_T3_T14}, \coq{bound_T5}, \coq{bound_T10}, \coq{bound_T15}, \coq{bound_T16}, \coq{bound_T17}, \coq{bound_T18}, \coq{keyvisc_coef} & the per-term bounds of Steps~1--5 and~7. \\
\coq{step6a} & the exact five-term rewriting of Step~6a (\lbl{eq:fiveterms}), consuming the IBP identities \lbl{eq:elemibp} etc. \\
\coq{bound_VolP}, \coq{bound_JmpP} (\coq{Jb11}--\coq{Jb22}) & Step~6b: volumetric part via $\varphi_1\tau_1\le\sqrt{\varphi_1}\sqrt{\tau_1}$ and (P3); jump part via \lbl{lem:jump} through four face-resolved bounds, facewise C--S, bounded multiplicity. \\
\coq{bound_II_V} & Step~6c. \\
\coq{bound_III} (\coq{JD11}--\coq{JD22}) & Step~6d, with the four face-difference bounds. \\
\coq{face_CS_pair} & facewise Cauchy--Schwarz with the multiplicity hypotheses. \\
\coq{abstract_continuity_sharp} & Step~8: the sharp porosity-weighted assembly (\lbl{eq:assembly}). \\
\coq{absorb_elem}, \coq{step9} & Step~9: the \lbl{eq:absorb1}--normconv absorption chain. \\
\coq{abstract_continuity} & \textbf{\lbl{lemma:Continuity}}: $B_S(U,V)\le C_{\mathrm{tot}}(\mathrm{Br}_U+\mathrm{Br}_P)\,|||V|||$, $C_{\mathrm{tot}}$ fully explicit. \\
\bottomrule
\end{longtable}

\subsection{\texttt{AbstractInterpolation.v} --- \lbl{lem:continterp} as a complete theorem}\label{sec:crossAbsInterp}

The continuity machinery rerun with the first argument equal to the
interpolation error $E=U-\widehat U_h$, exactly as the appendix's proof
prescribes. The first-slot vector family now models $E$; two nonnegative
elemental size functionals $\mathtt{IU}\,k\sim E_{\mathrm{int},K}(\mathbf u)$
and $\mathtt{IP}\,k\sim E_{\mathrm{int},K}(p)$ enter through the seven
interpolation hypotheses \coq{HI_*} (the replacement table
\lbl{eq:interpdivvisc}--\lbl{eq:interpzero}, one constant \coq{CI}) and
through the two face-integral hypotheses, whose first-argument $L^\infty$
content is now the interpolation estimate \lbl{eq:interpinftyE}. The
discrete second argument keeps its hypotheses verbatim (the appendix's
observation~(1): the exact identities remain valid for $E$).

\begin{longtable}{@{}L{0.40\textwidth}L{0.55\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Appendix content (all machine-checked)} \\
\midrule \endhead
\coq{bound_T6}, \coq{bound_T7}, \coq{bound_T8} & the Step-4 chains with
\lbl{eq:interpdivvisc} in place of the double inverse estimate; explicit
constants $4\bar C_{\mathrm{inv}}C_I/c_1$, $2C_I/\sqrt{c_1}$. \\
\coq{bound_VolP}, \coq{bound_II_V} & Steps 6b--6c with the $m'{=}0$
estimates \lbl{eq:interpzero} on the first slot. \\
\coq{bound_JmpP}, \coq{bound_III} & the jump terms with
\lbl{eq:interpinftyE} replacing the $L^\infty$ inverse estimate on the
first argument (V-side kept), through the same facewise C--S and
multiplicity machinery. \\
\coq{absorb_elem}, \coq{interp_triple_norm} & the Step-9 absorption with
the interpolation table; \lbl{eq:Enorm} in its sharp $\ell^2$ form,
$|||E|||\le\sqrt2\,K_{\mathrm{agg}}^{I}\,(\Psi_U+\Psi_P)$. \\
\coq{abstract_continterp_sharp} & the three-group assembly, in
absolute-value form. \\
\coq{abstract_continterp} & \textbf{\lbl{lem:continterp}}:
$|B_S(E,V_h)|\le C_{\mathrm{tot}}^{I}\,(\Psi_U+\Psi_P)\,|||V_h|||$, the
$\ell^2$ (porosity-weighted) form of $\psi(h)$; constants fully explicit. \\
\coq{abstract_continterp_l1} & the appendix's $\ell^1$ form of
\lbl{eq:psih}, via the discrete $\ell^2\subset\ell^1$ inequality
(\coq{sqrt_sum_sq_le_sum}, proved in \file{AbstractSums.v}). \\
\bottomrule
\end{longtable}

\subsection{\texttt{AbstractConvergence.v} --- \lbl{thm:convergence} as a kernel-level composition}\label{sec:crossAbsConv}

The convergence theorem proved by literally \emph{applying} the two closed
theorems inside Coq --- not by re-deriving them and not by assuming their
conclusions. \coq{abstract_stability} is instantiated at the discrete
error $E_h=\widehat U_h-U_h$ (its inverse-estimate input is literally the
V-side hypothesis \coq{Hw_dv}; its $\varepsilon$-condition is
definitionally the same \lbl{eq:epscond}), \coq{abstract_continterp} at
the pair $(E,E_h)$. The only trusted additions over the union of the two
prior files are the two \emph{diagonal} Green identities \coq{H_skew_diag} and
\coq{H_ibp_diag} --- and those are not new to the development either, being
literally the pair that \file{AbstractStability.v} already asks for, here stated
at the $W$ atoms; the tested identity for the discrete-error pair, formerly the
hypothesis \coq{HBS_W}, is gone, since \coq{abstract_stability} applied at those
atoms carries it --- together with \coq{Horth} (Galerkin orthogonality
\lbl{eq:consistency} plus bilinearity of $B_S$ in its first argument), which is
the one genuinely new trusted item of this file.

\begin{longtable}{@{}L{0.40\textwidth}L{0.55\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Content (all machine-checked)} \\
\midrule \endhead
\coq{tau1_agree}, \coq{tau2_agree}, \coq{sigt_agree} & the
$\tau_1,\tau_2,\sigt$ formulas of \file{StabilityAlgebra.v} and
\file{ContinuityAlgebra.v} agree (the first two definitionally, the third
as a field identity): the two developments provably speak about the same
parameters. \\
\coq{BSWW_is_ASBS}, \coq{stab_W} & the tested identity for the discrete-error
pair, formerly the hypothesis \coq{HBS_W}: \emph{eliminated}, not merely proved.
\coq{BSWW} is definitionally \file{AbstractStability.v}'s \coq{BS} at the $W$
atoms (\coq{reflexivity}), so \coq{abstract_stability} is simply \emph{applied}
there and carries its own \coq{HBS} with it. \\
\coq{t2i}, \coq{NErrI}, \coq{PsiI}, \coq{abstract_convergence_implemented} & \lbl{thm:convergence} for the \emph{implemented} $\tau_2$ (\lbl{eq:Tau2}): same estimate, constant inflated by at most $\sqrt{1+C_2}<\sqrt2$. \\
\coq{NW2_bridge} & the stability file's $|||E_h|||^2$ equals the
continuity file's, term by term over the mesh. \\
\coq{stab_W}, \coq{cont_EW} & the two closed theorems, applied. \\
\coq{NW_bound} & $|||E_h|||\le(C_{\mathrm{tot}}^{I}/\Cstab)\,\psi$: the
appendix's chain
$C|||E_h|||^2\le B_S(E_h,E_h)=-B_S(E,E_h)\le C\psi\,|||E_h|||$. \\
\coq{Qk_CS}, \coq{sumQ_CS}, \coq{NErr_triangle} & the triangle inequality
for the mesh-dependent triple norm, \emph{proved} from the pre-Hilbert
axioms: per-element five-fold Cauchy--Schwarz (the discrete C--S applied
over an explicit five-element index list), then the discrete C--S over the
mesh. \\
\coq{abstract_convergence} & \textbf{\lbl{thm:convergence}}:
$|||U-U_h|||\le(K_{\mathrm{abs}}^{I}+C_{\mathrm{tot}}^{I}/\Cstab)
(\Psi_U+\Psi_P)$ --- the porosity-weighted sharp form of
\lbl{rem:sharperconv}. \\
\coq{abstract_convergence_l1} & \lbl{eq:convergence}'s $\ell^1$ line. \\
\bottomrule
\end{longtable}

\subsection{\texttt{TauDesign.v} --- \S4 / Appendix C: Fourier design of \texorpdfstring{$\tau$}{tau}}\label{sec:crossTau}

Mirrors \file{fourier_tau_verification.py}. Section hypotheses: $\nu,\alpha,h>0$.

\begin{longtable}{@{}L{0.40\textwidth}L{0.55\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Paper content} \\
\midrule \endhead
\coq{viscous_symbol_closed_form} & $\sum_{ij}k_ik_j\mathbf K_{ij}=\alpha\nu(|k|^2 I+\tfrac13 kk^\top)$, $d{=}3$ (\lbl{eq:matrices_stationary_strong_problem}). \\
\coq{viscous_parallel_eigenpair}, \coq{viscous_transverse_eigenpair}, \coq{parallel_factor_d3/d2} & viscous spectrum $\tfrac43\alpha\nu|k|^2$, $\alpha\nu|k|^2$; the $d{=}3,2$ prefactors. \\
\coq{convective_symbol_scalar}, \coq{convective_spectral_radius} & convective symbol and radius $(\alpha/h)|w\cdot k|$. \\
\coq{coupling_*}, \coq{coupling_spectrum}, \coq{coupling_spectral_radius} & pressure--velocity coupling $\mathbf C$: spectrum $\{0,\pm|k|\}$. \\
\coq{taub_pair_solves}, \coq{taub_pair_unique} & $XY=c^2,\;X/Y=\lambda$ $\Rightarrow$ unique positive $(c\sqrt\lambda,c/\sqrt\lambda)$. \\
\coq{sqrt_lam_choice}, \coq{tau1_inv_assembled}, \coq{tau2_assembled}, \coq{visc_plus_conv_momentum} & assembly of \lbl{eq:Tau1}, \lbl{eq:Tau2}, \lbl{eq:CAlpha}, \lbl{eq:TauNavierStokes}. \\
\bottomrule
\end{longtable}

\subsection{\texttt{Asymptotics.v} --- \S6: robustness in \texorpdfstring{$(\Reh,\Dah)$}{(Re,Da)}}\label{sec:crossAsym}

Mirrors \file{robustness_asymptotics_verification.py}. Section hypotheses:
$\nu,h,\aK>0$; $\Reh,\Dah\ge0$.

\begin{longtable}{@{}L{0.42\textwidth}L{0.53\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Paper content} \\
\midrule \endhead
\coq{tau1_general_form}, \coq{tau2_general_form}, \coq{sigt_general_form} & closed forms in $(\Reh,\Dah)$ (\lbl{eq:GeneralAsymptoticBehaviourOfParameters}). \\
\coq{dom_viscosity_*} & dominant viscous diffusion ($\Reh,\Dah\to0$) (\S6.1). \\
\coq{dom_convection_*} & dominant convection ($\Reh\to\infty$): genuine \coq{tendsto_at_top} limits (\S6.2). \\
\coq{dom_reaction_*} & dominant reaction ($\Dah\to\infty$): $\tau_1\sigma\to1$, $\Dah$-independence of $\tau_2$ (paper \S6.3). \\
\coq{convection_coefficient}, \coq{viscous_coefficient}, \coq{pressure_coefficient}, \coq{reaction_coefficient}, \coq{forcing_coefficient} & nondimensionalisation coefficients $\Reh,1,(1{+}\Reh{+}\Dah),\Dah$ (\lbl{eq:DimensionlessMomentumEquation}, \lbl{eq:PressureScale}). \\
\bottomrule
\end{longtable}

\subsection{\texttt{ManufacturedSolution.v} --- \S7: the manufactured solution}\label{sec:crossMMS}

Mirrors \file{manufactured_solution_verification.py}; genuine
\coq{derivable_pt_lim} derivatives.

\begin{longtable}{@{}L{0.42\textwidth}L{0.53\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Paper content} \\
\midrule \endhead
\coq{manufactured_mass_conservation_2d/3d} & $\nabla\!\cdot(\alpha\mathbf u)=0$ for an arbitrary \emph{positive} $\alpha$ --- differentiability not assumed, since $\alpha$ cancels in $\alpha\mathbf u$ (\coq{alpha_u1}, \coq{alpha_u2}) (\lbl{eq:ManufacturedProblem}). \\
\coq{boundary_trace_nonzero} & $u_2(0,0)>0$: the trace does not vanish (amendment A11). \\
\coq{rhs_closed_form}, \coq{denominator_gap}, \coq{mid_le_rhs}, \coq{eps_choice_margin} & the \lbl{eq:EpsilonRef} chain at $C_2=10^{-2}$, and $\varepsilon:=10^{-4}\varepsilon_{\mathrm{ref}}$. \\
\coq{dbf_a_nonneg}, \coq{dbf_b_nonneg} & DBF coefficients admissible on $0<\alpha\le1$ (\lbl{eq:DBFResistanceTerm}). \\
\coq{alpha_lit_at_1}, \coq{alpha_lit_deriv(_pos)}, \coq{alpha_lit_at_0_range} & literature profile $\alpha(y)=0.45+0.55e^{y-1}$ (paper \S7.3). \\
\bottomrule
\end{longtable}

\subsection{\texttt{PlateauBump.v} --- \S7: the plateau bump, through \texorpdfstring{$C^1$}{C1}}\label{sec:crossBump}

$\gamma(\eta)=(2\eta-1)/(\eta(1-\eta))$,
$\alpha_{\mathrm{mid}}=1-(1-\alpha_0)/(1+e^\gamma)$
(\lbl{eq:PlateauBumpFunction}, \lbl{eq:Gamma}). Explicit $\varepsilon$--$\delta$
on $t<e^t$ and $t^3/27\le e^t$.

\begin{longtable}{@{}L{0.40\textwidth}L{0.55\textwidth}@{}}
\toprule
\textbf{Coq statement} & \textbf{Paper content} \\
\midrule \endhead
\coq{gam_deriv}, \coq{gam'_pos}, \coq{q_deriv}, \coq{am_deriv}, \coq{am'_pos} & genuine derivatives of $\gamma$ and $\alpha_{\mathrm{mid}}$; strict monotonicity. \\
\coq{am_range} & $\alpha_0<\alpha_{\mathrm{mid}}<1$ unconditionally. \\
\coq{exp_gam_small}, \coq{exp_gam_cubic_small}, \coq{gam'_upper_small/_big} (and mirrors) & quantitative decay bounds near $0$ and $1$. \\
\coq{am_limit_0}, \coq{am_limit_1} & continuity of the joins ($C^0$). \\
\coq{am'_limit_0}, \coq{am'_limit_1} & vanishing derivative at the joins ($C^1$). \\
\bottomrule
\end{longtable}

The paper's $C^\infty$ claim rests on every derivative vanishing at the joins;
the \emph{first-order} case (which the SymPy script also checks) is what is
formalised.

\subsection{Infrastructure: \texttt{Limits.v} and \texttt{DerivKit.v}}

\file{Limits.v} defines \coq{tendsto_at_top}, \coq{tendsto_at_0plus},
\coq{tendsto_at_1minus} (explicit $\varepsilon$--$\delta$, no filters), transfer
lemmas, and the generic rational limits consumed by \file{Asymptotics.v} and
\file{PlateauBump.v}. \file{DerivKit.v} provides \coq{derivable_pt_lim}
helpers and derivatives of $a\cdot t$, $\sin(at)$, $\cos(at)$, $c\cdot f(t)$,
$e^{t-1}$. Neither is paper-specific.

% =====================================================================
\section{The assumptions, read in full}\label{sec:assumptions}
% =====================================================================

A machine-checked theorem is an \emph{implication}. The kernel guarantees, with
a certainty no human referee can match, that the conclusion follows from the
hypotheses; it says nothing whatever about whether the hypotheses are true, or
even about whether they mean what their names suggest. All of the residual risk
therefore migrates into two places: reading the conclusion correctly, and
auditing the hypotheses. The crosswalk of Section~\ref{sec:crosswalk} discharges
the first. This section discharges the second, exhaustively: every assumption
the four abstract theorems make --- explicit or implicit --- is reproduced here
in the exact form the kernel sees it, translated into ordinary mathematics,
matched to the manuscript, and assessed.

\subsection{Why the list below is complete}\label{sec:complete}

Before the list, the structural fact that gives it its value.

The four abstract theorems are not stated about Sobolev spaces. They are stated
about an \emph{abstract interface}: the fields are opaque vectors in a
pre-Hilbert space, the mesh is an abstract list, and the elemental quantities
$h_K,\aK,|\mathbf a|_{\infty,K}$ are abstract real-valued functions on it.
\textbf{Coq does not know what a gradient is.} It has never heard of
$\alpha$, of a simplex, or of a polynomial.

The consequence is decisive. In an ordinary paper proof, a step may quietly
invoke a property of $\nabla$, of the mesh, or of the finite element space
without the author noticing that a hypothesis was needed --- this is precisely
how gaps survive refereeing. Here that is impossible. If a step of the argument
requires a fact about the objects, the fact is not available unless it has been
written down as a hypothesis, and if it has not been written down the file
simply does not compile. \emph{The hypothesis list is therefore complete by
construction, not by diligence.} It is not a list of the assumptions the authors
remembered to declare; it is a list of every assumption the proof makes
\emph{about the objects} --- every relation between the atoms that the kernel is
permitted to use --- enforced by a machine that cannot be persuaded to overlook
one.

One thing this guarantee does \emph{not} cover, and it is important: the
\emph{encoding} itself. Whether the abstract atoms and the abstract $B_S$ really
stand for the objects the paper is about is a question the kernel cannot even
ask. Assumptions of that kind hide in \coq{Definition}s and \coq{Variable}s
rather than in \coq{Hypothesis} lines, and they appear in no trusted-base table.
They are the subject of \S\ref{sec:stratum1}, and they are the reason this
section has four strata rather than two.

That guarantee is worth more than any individual entry below, and it is the
reason the abstract-interface method was chosen over a direct formalisation.
Its price is that the reader must now do the one thing the machine cannot: check
that each hypothesis, read as a statement about the real finite element objects,
is true. The rest of this section is written to make that check as easy as
possible.

We organise the assumptions into four strata, in increasing order of
mathematical content:

\begin{enumerate}[label=\textbf{Stratum \arabic*.},leftmargin=6.2em,nosep]
  \setcounter{enumi}{-1}
  \item The \emph{logic} --- what Coq itself rests on (\S\ref{sec:stratum0}).
  \item The \emph{interface} --- the modelling identifications made when the
    finite element objects were encoded as abstract vectors. These are the
    \emph{implicit} assumptions: they live in \coq{Definition}s and
    \coq{Variable}s rather than in \coq{Hypothesis} lines, so they appear in no
    trusted-base table, and they are exactly what a reader is most likely to
    over-read (\S\ref{sec:stratum1}).
  \item The \emph{data, mesh and parameter conditions} --- Class~D. Hypotheses
    of the paper's theorems in any formal system, so there is nothing to prove
    --- with two disclosed exceptions where Coq asks for more
    (\S\ref{sec:stratumD}).
  \item The \emph{analytic base} --- Classes~G and~I: the Green-type identities
    and the inverse/interpolation estimates. Textbook facts, cited rather than
    proved, and the only place where genuine mathematics is taken on trust
    (\S\ref{sec:stratumG}--\S\ref{sec:stratumI}).
\end{enumerate}

% ---------------------------------------------------------------------
\subsection{Stratum 0: the logical base}\label{sec:stratum0}
% ---------------------------------------------------------------------

Three things, and no more, sit under every theorem in the development.

\emph{The logic.} The Calculus of Inductive Constructions, as implemented by the
Coq/Rocq kernel and independently re-checked by \coq{coqchk}
(\S\ref{sec:coq}). No \coq{Axiom}, no \coq{Admitted}, no \coq{admit} occurs in
the eighteen files, so nothing has been postulated behind the reader's back.

\emph{The real numbers.} The development imports \coq{Coq.Reals} (with
\coq{Lra}, \coq{Lia}, \coq{Psatz}, \coq{List}) and nothing else --- no Mathcomp,
no Coquelicot. Coq's $\mathbb R$ is the \emph{classical} real line: a complete
ordered Archimedean field with decidable comparison, so that excluded middle is
available for statements $x<y$, $x=y$, $x>y$. This is the ordinary
mathematician's $\mathbb R$, and reasoning classically about it is not a weakness
but the intended semantics.

We need not take this on trust, and the reader need not either. Coq will
\emph{print} the axioms a theorem actually depends on. Running, after
\coq{./run_all.sh},
\begin{lstlisting}
From PNSFormal Require Import AbstractConvergence.
Print Assumptions abstract_convergence.
\end{lstlisting}
returns, for each of the four abstract theorems and for every non-vacuity witness
alike, exactly this list and nothing else:
\begin{lstlisting}
Axioms:
ClassicalDedekindReals.sig_not_dec : forall P : Prop, {~ ~ P} + {~ P}
ClassicalDedekindReals.sig_forall_dec
  : forall P : nat -> Prop,
    (forall n : nat, {P n} + {~ P n}) ->
    {n : nat | ~ P n} + {forall n : nat, P n}
FunctionalExtensionality.functional_extensionality_dep
  : forall (A : Type) (B : A -> Type) (f g : forall x : A, B x),
    (forall x : A, f x = g x) -> f = g
\end{lstlisting}
Three axioms, all from the standard library's \emph{construction} of
$\mathbb R$ (Coq builds the reals rather than postulating them, and these are the
classical principles the construction needs): a decidability of double
negations, a decidability principle for decidable predicates over $\mathbb N$,
and dependent functional extensionality. All three hold in the ordinary
set-theoretic model, so they are consistency-preserving. \textbf{There is no
user-introduced axiom, and no use of propositional excluded middle, proof
irrelevance, or choice.} Everything else a theorem depends on is a section
hypothesis --- discharged by Coq into the statement as an explicit argument,
where it is visible to anyone who prints the type --- and every section
hypothesis is catalogued below.

\emph{Two totality conventions.} In \coq{Coq.Reals}, division and square root
are \emph{total}: $x/0 = 0$ and $\sqrt{y} = 0$ for $y<0$. A development can
accidentally lean on these and prove a statement that is true only because of
the convention. This one does not. Every division in the abstract files has a
denominator that is either a \emph{declared} positive quantity ($h_K$, $c_1$,
$c_2$, $\xi$, $\aK$, $c_J$, the Young parameter $\lambda$) or one \emph{proved}
positive before use ($\tau_{1,\mathrm{NS}}^{-1}$, $\varphi_1$, $\sqrt{c_1}$,
$\Cstab$); every $\sqrt{\cdot}$ is applied to a quantity whose nonnegativity is
likewise declared or proved --- a positive datum, a stabilization parameter
($\tau_1,\tau_2>0$ by \coq{tau1_pos}, \coq{tau2_pos}), an inner product
$\langle x,x\rangle$ (the axiom \coq{ip_pos}), or a sum of such
(\coq{Rsum_nonneg}). The conventions are therefore never load-bearing --- but
note that this is a property established by \emph{reading the development}, not
one the kernel enforces: Coq would happily accept a theorem that silently relied
on $1/0=0$.

% ---------------------------------------------------------------------
\subsection{Stratum 1: the interface --- the implicit assumptions}\label{sec:stratum1}
% ---------------------------------------------------------------------

These are the assumptions that do \emph{not} appear as hypotheses, because they
were made when the mathematics was encoded rather than when it was proved. They
are no less real for that, and they are the ones a reader is likeliest to
mis-read. We give all five.

\subsubsection{(I1) The elemental pairing as a pre-Hilbert space}

\file{InnerSpace.v} models the elemental pairing
$\langle f,g\rangle_K = \int_K f\cdot g\,\mathrm d\Omega$ by a carrier $V$ with
addition, scalar multiplication and a form $\langle\cdot,\cdot\rangle$ subject
to exactly four axioms:
\begin{equation}\label{eq:prehilbert}
  \langle x,y\rangle=\langle y,x\rangle,\qquad
  \langle x+y,z\rangle=\langle x,z\rangle+\langle y,z\rangle,\qquad
  \langle a\,x,y\rangle=a\,\langle x,y\rangle,\qquad
  \langle x,x\rangle\ \ge\ 0 .
\end{equation}

\emph{What is assumed.} Symmetry, linearity in the first slot, and positive
\emph{semi}-definiteness. Nothing else.

\emph{What is deliberately \textbf{not} assumed.} Completeness; definiteness
($\langle x,x\rangle=0\Rightarrow x=0$); finite dimension; the existence of an
orthonormal basis; any topology at all. From \eqref{eq:prehilbert} the file
\emph{derives} full bilinearity, the Cauchy--Schwarz inequality (by the
discriminant argument, with the degenerate direction handled explicitly --- this
is where semi-definiteness rather than definiteness earns its keep), the Gram
expansions of $\lVert x\pm y\rVert^2$, the triangle inequality, the parametrized
Young inequality at vector level, and the difference-of-squares identity
$\langle -a+b,\,a+b\rangle=\lVert b\rVert^2-\lVert a\rVert^2$ that drives the
stabilization cancellation.

\emph{Interpretation.} That the whole a priori theory needs only
\eqref{eq:prehilbert} is a small mathematical result in its own right, made
visible by the formalisation: the stability, continuity, interpolation and
convergence estimates are \emph{algebraic}, not topological. No limit is taken,
no completeness is used, no Riesz representation is invoked. Soundness in the
other direction is immediate: every weighted $L^2(K)^m$ pairing --- in
particular $\int_K \alpha\,\mathbf f\!\cdot\!\mathbf g$ and the Frobenius pairing
on tensors --- satisfies \eqref{eq:prehilbert}, so it is a model and the theorems
transfer to it verbatim. \emph{Tightening: none available.} This is the minimal
interface; removing any of the four axioms breaks Cauchy--Schwarz.

\subsubsection{(I2) The mesh as a finite list, and the absence of geometry}

The triangulation is \coq{Variable Th : list K} over an abstract element type
\coq{K}; the interior faces are \coq{Fl : list F} with two designated
neighbours \coq{e1, e2 : F -> K}; and $\sum_{K\in\Th}$ is the fold
\coq{Rsum Th f}. Three readings follow.

\emph{Finiteness is structural.} A \coq{list} is finite by construction, so
``the mesh is finite'' is not an assumption that could fail --- it is part of
the type. Nothing in the development requires the list to be duplicate-free:
duplicated elements would merely be counted consistently on both sides of every
inequality.

\emph{There is no geometry.} No simplex, no coordinates, no measure, no normal
vector appears anywhere. Every geometric fact the proofs need has been pushed
into (a) the three scalar fields $h_K$, $\aK$, $|\mathbf a|_{\infty,K}$, and
(b) the analytic hypotheses of Stratum~3. This is what makes those hypotheses
the \emph{complete} geometric content of the proofs, in the sense of
\S\ref{sec:complete}.

\emph{Faces carry an assumption about the boundary.} \coq{Fl} is the list of
\emph{interior} faces. The fact that the boundary faces contribute nothing ---
because $\mathbf u_h=\mathbf v_h=\mathbf 0$ on $\partial\Omega$ --- is not
derived; it is folded into the assembly hypotheses \coq{H_assemble_p} and
\coq{H_assemble_c} (\S\ref{sec:stratumG}). The homogeneous Dirichlet condition
therefore enters the continuity proof as part of an assumption, not as a
theorem, and we flag it as such.

\subsubsection{(I3) The fields as independent families --- the crux}

For the continuity lemma the discrete data are declared as
\begin{lstlisting}
Variables (gu gv du dv cxu cxv gpu gpv uu vv pp qq divu divv : K -> V).
\end{lstlisting}
fourteen \emph{mutually independent} functions from elements to vectors, whose
intended readings are recorded only in a comment:
\[
  \mathtt{gu}\,k \sim \bigl(\alpha^{1/2}\mathcal P\nabla\mathbf u\bigr)\big|_K,
  \quad
  \mathtt{du}\,k \sim \bigl(2\nu\,\nabla\!\cdot(\alpha\,\mathcal P\nabla\mathbf u)\bigr)\big|_K,
  \quad
  \mathtt{divu}\,k \sim \bigl(\nabla\!\cdot(\alpha\mathbf u)\bigr)\big|_K,
  \quad \dots
\]
\textbf{Nothing in the Coq text asserts that \coq{gu} is the gradient of
\coq{uu}}, or that \coq{divu} is its divergence. The names are comments; the
kernel sees fourteen unrelated symbols. The \emph{only} facts relating them are
the ones written down as hypotheses --- the integration-by-parts identities and
the inverse/interpolation estimates.

This must be read correctly, in both directions.

\emph{It is not a defect.} The theorem is universally quantified: \emph{for
every} tuple of fourteen families satisfying the listed hypotheses, the estimate
holds. The genuine finite element objects form one such
tuple --- provided each hypothesis is a true statement about them, which is
exactly what Stratum~3 asserts and what the hand audit checks. So the theorem
applies to the real objects, and it applies to more besides. A more general
theorem is not a weaker one.

\emph{It is what makes the audit exhaustive.} Because the kernel cannot use any
unlisted property of $\nabla$, the list of relations between the atoms
\emph{is} the list of analytic facts the proof consumes. That is the content of
\S\ref{sec:complete}, and it is why we can claim the trusted base is complete
rather than merely long.

\emph{What a reader must not do.} Read the intended meanings into the statement.
\coq{abstract_continuity} does not say ``the stabilized form is continuous on
the discrete space''; it says ``any fourteen families satisfying \coq{H_skew},
the \coq{Hw_*} and the rest obey the bound''. Likewise \coq{abstract_stability}
--- whose field list is the shorter one, \emph{seven} families
(\coq{gu uu pp cxu gpu divu du}), with only \coq{H_skew_diag}, \coq{H_ibp_diag},
\coq{S3} and \coq{Heps} to relate them --- does not say ``$B_S$ is coercive'';
it says ``for any seven families satisfying those four hypotheses, the
eighteen-term expression built from them is
$\ge\Cstab|||U_h|||^2$''. The bridge from the second reading to the first is
the instantiation, and the instantiation is exactly the hand-audited claim that
the finite element objects satisfy the hypotheses.

(The stability file's atom list used to be \emph{six} families, with
$\alpha\mathcal X(U_h)$ carried by a single atom \coq{xu}. It is now seven:
\coq{xu} has been split into its convective and pressure-gradient parts
\coq{cxu}, \coq{gpu}, with \coq{xu} \emph{defined} as their sum. This is forced
--- one cannot state ``the convective part is skew on the diagonal'' without
naming the convective part --- and it is faithful, since
$\alpha\mathcal X(U_h)=\alpha\,\mathbf a\!\cdot\!\nabla\mathbf u_h+\alpha\nabla p_h$
is the definition of $\mathcal X$. It costs one further Stratum-1 identification,
of the most trivial kind, and it is what buys the discharge of \coq{HBS}.)

\subsubsection{(I4) Three identifications hidden in \texttt{Definition}s}

Three modelling decisions are encoded as definitions rather than hypotheses, so
they appear in no trusted-base table. All three are sound; all three deserve
naming, and the last is the subtlest thing in this section.

\emph{The eighteen-term form of $B_S$.} In \emph{every} file that mentions it,
$B_S$ is now a \coq{Definition}:
\[
  B_S \;:=\; T_1+T_2+\dots+T_{18},
\]
the eighteen terms of \lbl{eq:Bstab}. \file{AbstractContinuity.v} and
\file{AbstractInterpolation.v} have always defined it this way;
\file{AbstractStability.v} and \file{AbstractConvergence.v} now do too, each
reusing \emph{the same} \coq{AbstractInterpolation.BS} applied to its own atoms
--- the diagonal $(U_h,U_h)$ atoms in the stability file, the $(W,W)$ atoms in
the convergence file --- rather than restating the expression. The claim ``this
expression \emph{is} the paper's $B_S(\mathbf a;U,V)$'' is thus not a hypothesis
but a reading obligation --- a Stratum-1 identification, verified by diffing the
eighteen \coq{Definition}s against the appendix's Step~0 (this is done in
\file{AUDIT.md}). Nothing is lost, and the asymmetry a reader used to have to
explain --- why one file carried an \coq{HBS} and the other did not --- is gone:
the four files now place the same fact in the same logical position, and the
tested forms are \emph{theorems} in both places where they are needed
(\coq{HBS}, and its reuse at the $W$ atoms; \S\ref{sec:stratumG}~G1).

\emph{Linearity of the error decomposition.} In \file{AbstractConvergence.v} the
interpolation error $E=U-\widehat U_h$ is carried by the first-slot atoms and
the discrete error $W=\widehat U_h-U_h$ by the second-slot ones. (The appendix
calls the latter $E_h$; we follow the Coq naming, \coq{BSWW}, \coq{NW}, and
write $W$ throughout this section.) With that, the triple norm of
the total error is \emph{defined} by
\begin{equation}\label{eq:perErr}
  |||U-U_h|||^2 \;:=\;
  \sum_{K\in\Th}\Bigl[
    \nu\,\lVert \mathtt{gu}+\mathtt{gv}\rVert^2
    + \sigt\,\lVert \mathtt{uu}+\mathtt{vv}\rVert^2
    + \varepsilon\,\lVert \mathtt{pp}+\mathtt{qq}\rVert^2
    + \tau_1\lVert \mathtt{xe}+\mathtt{xw}\rVert^2
    + \tau_2\lVert \mathtt{divu}+\mathtt{divv}\rVert^2 \Bigr].
\end{equation}
That is: the atoms of $U-U_h$ are taken to be the \emph{sums} of the atoms of
$E$ and of $W$. Mathematically this is nothing but the linearity of each atom
map ($\nabla$, $\nabla\cdot$, $\mathcal P$, multiplication by $\alpha$,
restriction to $K$) together with $U-U_h=E+W$ --- true, and trivially so. But it
is an identification, made in a \coq{Definition}, and it is the reason the
triangle inequality of \coq{NErr_triangle} can be \emph{proved} from the
pre-Hilbert axioms rather than assumed. We record it here for completeness.

\emph{$B_S$ is never a bilinear form.} This is the subtlest of the
identifications. Nowhere in the development is $B_S$ a function of two
arguments. It appears as \emph{scalars}: in each file, a \coq{Definition} equal
to the eighteen-term expression evaluated at that file's atoms. Bilinearity is
therefore never available to the kernel as a \emph{fact}; where the paper's
proof uses it, it has been absorbed into an assumption --- namely \coq{Horth}
(\S\ref{sec:stratumG}), which assumes not only Galerkin orthogonality but also
the bilinearity step that turns it into $B_S(W,W)=-B_S(E,W)$. With \coq{HBS} now
proved, \coq{Horth} is the \emph{only} place bilinearity is still assumed, and
it is the largest single class-G item left.

One consequence of this used to be a real gap, and has now been closed twice
over. Each of the stability and convergence files needs a \emph{tested} $B_S$ on
a diagonal --- $B_S(U_h,U_h)$ and $B_S(W,W)$ respectively --- and each used to
carry \emph{two} unrelated encodings of it: a free real pinned by a hypothesis
(\coq{HBS}, \coq{HBS_W}), and the eighteen-term expression evaluated at the
corresponding atoms. \emph{Nothing proved they agree.} Each was separately a true
statement about the real $B_S$, so soundness was never at risk --- but a
transcription error in either encoding would have slipped through, and the
development gave the parameters exactly this treatment (\coq{tau1_agree},
\coq{tau2_agree}, \coq{sigt_agree}) while giving $B_S$ none. It does now, in both
files: the diagonal $B_S$ is \emph{defined} as the eighteen-term expression, and
the tested form is \emph{proved} from it (\coq{HBS}, which the convergence file
reuses at its own atoms rather than restating; \S\ref{sec:stratumG}~G1). The number the stability lemma bounds from below and
the number the interpolation and convergence files bound from above are now,
provably, the same closed expression --- reconciled inside the kernel rather than
by a side condition.

\subsubsection{(I5) Which quantities are global and which are elemental}

The scalars $\nu,\sigma,\varepsilon,c_1,c_2,\Cinv,\Cbar,\xi,C_2,c_J,c_J',
C_{\mathrm{face}},N_f,C_I$ are single reals: not one of them varies with the
element. The elementwise objects are the three mesh fields $h_K$, $\aK$,
$|\mathbf a|_{\infty,K}$, the field atoms of (I3), the interpolation-error
functionals $\mathtt{IU},\mathtt{IP}$, and the elemental face terms
$\mathrm{FB}_{p,e},\mathrm{FB}_{c,e}$ --- all of type \coq{K -> R} or
\coq{K -> V}. Two comments.

\emph{On $\sigma$.} That $\sigma$ is a \emph{constant} is not a Coq artefact:
it is the paper's own \lbl{H:data}, which reads ``$\nu>0$ and $\sigma\ge0$ are
constants''. The a priori theory --- in the manuscript and in Coq alike --- is
for the \emph{linearized} problem, in which the Darcy--Brinkman--Forchheimer law
$\sigma=a(\alpha)+b(\alpha)|\mathbf u|$ has been frozen at the previous
iterate. The nonlinearity is the business of the outer iteration and lies
outside the a priori scope of both artefacts. Coq's global \coq{sigma : R} is
therefore faithful; a reader should simply not expect the machine-checked
theorems to say anything about the Forchheimer nonlinearity, because the paper's
do not either.

\emph{On uniformity.} Every constant in Stratum~3 is a single real, quantified
\emph{outside} the ``$\forall k$'' of each estimate. Uniformity of the constants
over the elements of the mesh --- and over the mesh family as $h\to0$ --- is
thus assumed, not derived. This is not a hidden extra: uniformity is precisely
the mathematical content of shape-regularity (\lbl{H:mesh}), and the estimates
would be worthless without it. But it is worth seeing where it lives: it lives
in the quantifier order.

% ---------------------------------------------------------------------
\subsection{Stratum 2 (Class D): data, mesh and parameter conditions}\label{sec:stratumD}
% ---------------------------------------------------------------------

These are hypotheses of the paper's theorems in any formal system. There is
nothing to \emph{prove}; there is something to \emph{check}, namely that the
mesh and the data one actually uses satisfy them --- and, in two places noted
below, something to \emph{disclose}, namely that Coq asks for slightly more than
the manuscript does.

\subsubsection{D1 --- Positivity: Coq now assumes exactly what the paper does}

This is the complete positivity list of the four abstract theorems --- every
sign hypothesis they make, with none omitted (items 1--18 of
Table~\ref{tab:inventory}; the two numerical-parameter conditions, items 19--20,
are D3):
\begin{lstlisting}
(* physical data and mesh *)
Hypothesis nu_pos       : 0 < nu.
Hypothesis sigma_nonneg : 0 <= sigma.
Hypothesis eps_nonneg   : 0 <= eps.
Hypothesis hK_pos       : forall k, 0 < hK k.
Hypothesis aK_pos       : forall k, 0 < aK k.
Hypothesis am_nonneg    : forall k, 0 <= am k.
(* algorithmic and analytic constants *)
Hypothesis c1_pos   : 0 < c1.        Hypothesis c2_pos      : 0 < c2.
Hypothesis Cinv_pos : 0 < Cinv.      Hypothesis Cb_pos      : 0 < Cb.
Hypothesis C2_nonneg: 0 <= C2.       Hypothesis C2_lt_1     : C2 < 1.
Hypothesis cJ_pos   : 0 < cJ.        Hypothesis cJ_le_1     : cJ <= 1.
Hypothesis one_le_cJ' : 1 <= cJ'.    Hypothesis Cface_nonneg: 0 <= Cface.
Hypothesis Nf_nonneg: 0 <= Nf.       Hypothesis CI_pos      : 0 < CI.

(* the nonnegativity of the interpolation-error functionals is NOT
   assumed: it is derived.  See below. *)
Lemma IU_nonneg : forall k, 0 <= IU k.
Proof. intro k. pose proof (nrm_nonneg Hs (uu k)) as Hn.
       pose proof (HI_uu k) as Hi. nra. Qed.
Lemma IP_nonneg : forall k, 0 <= IP k.
Proof. intro k. pose proof (nrm_nonneg Hs (pp k)) as Hn.
       pose proof (HI_pp k) as Hi. nra. Qed.
\end{lstlisting}

\emph{Two rows that used to be here, and the one that pays for them.} An earlier
revision assumed $E_{\mathrm{int},K}(\mathbf u)\ge0$ and
$E_{\mathrm{int},K}(p)\ge0$ outright, alongside $C_I\ge0$. Both are now theorems.
The pre-Hilbert axioms already give $\lVert x\rVert\ge0$
(\coq{nrm_nonneg}, \file{InnerSpace.v}), so the interpolation hypothesis
\coq{HI_uu} reads $0\le\lVert\mathbf e_{\mathbf u}\rVert_K\le
C_I\,E_{\mathrm{int},K}(\mathbf u)$; a nonnegative product with $C_I>0$ forces
the second factor nonnegative. The cost is the strengthening
$C_I\ge0\rightsquigarrow C_I>0$ (item~15), and we flag it as a \emph{trade}
rather than a saving: $\{C_I>0\}$ is logically strictly stronger than
$\{C_I\ge0,\ E_{\mathrm{int}}(\mathbf u)\ge0,\ E_{\mathrm{int}}(p)\ge0\}$. The
degenerate reading $C_I=0$ is no longer an instance --- but it never was a
meaningful one, since it would force $\mathbf e_{\mathbf u}\equiv0$ and $e_p\equiv0$,
i.e.\ an exactly-representable solution. Net: two rows leave the trusted base,
one inequality tightens, and nothing that the manuscript's estimates ever
claimed is given up.

\emph{Role.} Denominators, radicands and the sign of every coefficient
collection. \emph{Reasonable.} $\nu>0$, $h_K>0$, $\aK>0$ are physical;
$|\mathbf a|_{\infty,K}\ge0$ because it is a norm; $c_1,c_2>0$ and
$\Cinv,\Cbar>0$, while $C_{\mathrm{face}},N_f\ge0$ suffice --- all conventions on
constants that are anyway free to be enlarged. $C_I>0$ is the one strictness
here that is not merely conventional: it is what pays for the two derived rows
above, and enlarging a valid $C_I$ keeps it valid, so no interpolation estimate
anyone would cite is excluded.

\emph{Exactly the manuscript's hypotheses --- and this was not always so.}
An earlier revision of the development asked for $\sigma>0$, $\varepsilon>0$ and
$C_2>0$ \emph{strictly}, where \lbl{H:data} and \lbl{eq:epscond} allow
$\sigma\ge0$, $\varepsilon\ge0$ and $C_2\ge0$. The reaction-free case
$\sigma=0$ --- pure Navier--Stokes, no Darcy term --- was therefore not
literally an instance of the machine-checked theorems. That was an artefact, not
mathematics: $\sigma$ occurs only \emph{additively} in the denominators
($\aK\tau_{1,\mathrm{NS}}^{-1}+\sigma$, already positive), no division needs
$\sigma\ne0$, and \file{ContinuityAlgebra.v} proved the same $\sigt$ identities
under $\sigma\ge0$ all along. Only three lemmas of \file{StabilityAlgebra.v}
were genuinely $\sigma$-strict, and each was false at $\sigma=0$ merely by
being stated with $<$ where $\le$ was true:
\begin{center}
\small
\begin{tabular}{@{}lll@{}}
\toprule
was & is & at $\sigma=0$ \\
\midrule
\coq{sigt_pos} : $0<\sigt$ & \coq{sigt_nonneg} : $0\le\sigt$ & $\sigt=0$ \\
\coq{alphaK_tau1_lt_tau1NS} : $\aK\tau_1<\tau_{1,\mathrm{NS}}$
  & \coq{alphaK_tau1_le_tau1NS} : $\aK\tau_1\le\tau_{1,\mathrm{NS}}$
  & equality \\
\coq{sigt_denom_pos} (via $\sigma/\aK>0$) & (via $\sigma/\aK\ge0$) & unchanged \\
\bottomrule
\end{tabular}
\end{center}
With those three weakened, the entire chain goes through unaltered. The
development now assumes \emph{exactly} \lbl{H:data}: $\sigma\ge0$,
$\varepsilon\ge0$, $0\le C_2<1$. The reaction-free limit is a genuine instance
of \coq{abstract_stability} and of \coq{abstract_convergence}, and the last
place where Coq asked for more than the manuscript is gone.

\emph{One redundant hypothesis also went.} \coq{xi_pos} ($\xi>0$) was declared
alongside \coq{xi_large} ($\xi>2$), which implies it; it is now derived.

\noindent
By contrast, one hypothesis the reader might \emph{expect} to find is
gratifyingly \textbf{absent}: $\aK\le1$. It is a standing hypothesis of
\file{ContinuityAlgebra.v} (\coq{alphaK_le_1}, used by the scalar lemmas that
pass from the porosity-weighted estimate to the manuscript's unweighted one),
but it appears nowhere in \file{AbstractContinuity.v},
\file{AbstractInterpolation.v} or \file{AbstractConvergence.v}. The reason is
that those files prove the \emph{sharp}, porosity-weighted forms --- with the
local weight $\aK$ retained: \lbl{eq:sharpcont} for continuity and
interpolation, \lbl{rem:sharperconv} for convergence --- and the step that
discards the weight is exactly the step that needs $\aK\le1$. The abstract
theorems are therefore strictly stronger than the manuscript's displayed
estimates, and they need one hypothesis fewer.

\subsubsection{D2 --- The penalty condition on the mass-stabilization parameter}

\begin{lstlisting}
Hypothesis H_eps : forall k, eps <= C2 * c1 * (aK k)^2 * t1 k / (hK k)^2.
Hypothesis Heps  : forall k, eps <= eps_max nu (hK k) (aK k) sigma (am k) c1 c2 C2.
\end{lstlisting}
Both lines say the same thing --- \coq{eps_max} \emph{is} the right-hand side of
the first --- and both are the elementwise (equivalently, infimum) form of the
manuscript's \lbl{eq:epscond}:
\begin{equation}\label{eq:epscond-here}
  \varepsilon \;\le\; C_2\,c_1 \min_{K\in\Th}
  \frac{\aK^{2}\,\tau_{1,K}}{h_K^{2}},\qquad 0\le C_2<1 .
\end{equation}

\emph{Role.} It is the entire source of pressure coercivity. Through
\lbl{lem:parameters}\,(P5) it gives $\varepsilon\tau_2\le C_2<1$, hence the
pressure coefficient $\varepsilon(1-\varepsilon\tau_2)\ge(1-C_2)\,\varepsilon$,
which is the $1-C_2$ branch of
$\Cstab=\min\{C_{\mathrm{visc}},C_u,1-C_2,1\}$. In the continuity proof it
enters Steps~3, 7 and~9.

\emph{Why it is reasonable.} $\varepsilon$ is not a property of nature: it is
the penalty parameter \emph{we} select. \eqref{eq:epscond-here} is a design rule,
and a mild one --- it is satisfied by construction in the numerical experiments,
where \file{ManufacturedSolution.v} machine-checks the concrete chain at
$C_2=10^{-2}$ and $\varepsilon:=10^{-4}\varepsilon_{\mathrm{ref}}$
(\coq{rhs_closed_form}, \coq{eps_choice_margin}).

\emph{Tightening.} The two ends of the admissible range of $C_2$ pull against
each other and the trade-off is exact: as $C_2\uparrow1$ the coercivity constant
$1-C_2$ collapses, while as $C_2\downarrow0$ the admissible $\varepsilon$
collapses. There is nothing to gain by weakening the condition; there is
something to gain by choosing $C_2$ well.

\subsubsection{D3 --- The numerical-parameter condition}

\begin{lstlisting}
Hypothesis c1_large : c1 > xi * Cb^2.   (* the SHARP form *)
Hypothesis xi_large : xi > 2.
\end{lstlisting}
This is \lbl{eq:conditions_on_num_param}, with $\Cbar$ the \emph{weighted}
inverse constant of \lbl{lem:winv} (not the bare $\Cinv$; this is precisely the
substitution prescribed by \lbl{rem:winvconst}, and Coq performs it --- the
constant passed into $C_u$ and $C_{\mathrm{visc}}$ is \coq{Cb}). The manuscript
states $c_1>2\xi\Cbar^2$; Coq assumes the weaker $c_1>\xi\Cbar^2$, and the
paragraph below explains why both are worth having.

\emph{Role.} Remarkably narrow: in the whole development the \emph{content} of
these two hypotheses is consumed at exactly one place --- the theorem
\coq{stability_constants_positive_sharp} of \file{StabilityAlgebra.v}, reached
from \file{AbstractStability.v}; \file{AbstractConvergence.v} merely forwards
them into \coq{abstract_stability}. Its job is to establish
\[
  C_u \;=\; 1-\frac{\xi\,\Cbar^{2}}{c_1}\;>\;0,
  \qquad
  C_{\mathrm{visc}} \;=\;
  \min\Bigl\{\,2-\frac{4\,\Cbar^{2}}{c_1},\ 2\Bigl(1-\frac{2}{\xi}\Bigr)\Bigr\}\;>\;0,
\]
and hence $\Cstab>0$. Everything else in the stability proof is indifferent to
the size of $c_1$.

\emph{How tight it is --- now machine-checked, both ways.} This is the one
condition in the paper that couples a \emph{numerical} choice ($c_1$) to an
\emph{analytic} constant ($\Cbar$), so it repays a close look. Positivity of the
two constants requires
\[
  C_u>0 \iff c_1>\xi\,\Cbar^{2},
  \qquad
  C_{\mathrm{visc}}>0 \iff c_1>2\,\Cbar^{2}\ \text{ and }\ \xi>2 ,
\]
and since $\xi>2$ the first implies the second. \textbf{The sharp sufficient
condition is therefore $c_1>\xi\Cbar^{2}$ with $\xi>2$ --- a factor of two
weaker than the manuscript's} --- and that is what the development now assumes
(\coq{stability_constants_positive_sharp}). The manuscript's condition is kept
as a corollary (\coq{stability_constants_positive}), since it implies the sharp
one.

The manuscript's extra factor of two is nevertheless not idle: it buys a
quantitative \emph{margin} rather than bare positivity, and this too is now a
theorem (\coq{C_stab_margin}). Under $c_1>2\xi\Cbar^2$ and $\xi>2$,
\[
  \frac{\xi\Cbar^2}{c_1}<\frac12
  \;\Longrightarrow\;
  C_u>\frac12,
  \qquad
  c_1>2\xi\Cbar^2>4\Cbar^2
  \;\Longrightarrow\;
  2-\frac{4\Cbar^2}{c_1}>1,
\]
so that the coercivity constant enjoys a floor that is free of $\Cbar$
altogether:
\[
  \Cstab \;\ge\; \min\Bigl\{\,2\Bigl(1-\tfrac2\xi\Bigr),\ \tfrac12,\ 1-C_2\Bigr\}.
\]
Under the sharp condition alone the constants stay positive but
$C_u\to0$ as $c_1\downarrow\xi\Cbar^2$, and the floor is lost. So the two
conditions do different jobs, and the formalisation now records both: the sharp
one is what the theorem \emph{needs}, the manuscript's is what a practitioner
should \emph{use}.

\emph{In practice.} $\Cinv$ grows like $k^2$ in the polynomial
degree~\cite{codina2018hp}, which is what motivates the manuscript's calibration
$c_1=4k^4$, $c_2=2k^2$; the condition is then a statement about the constant in
front of $k^4$, and the element-aware $c_1$ study in the repository is the
empirical counterpart.

\subsubsection{D4 --- Jump control across interior faces}

\begin{lstlisting}
Hypothesis H_jump :
  forall f, cJ * ph (e1 f) <= ph (e2 f) /\ ph (e2 f) <= cJ' * ph (e1 f).
Hypothesis cJ_pos : 0 < cJ.   Hypothesis cJ_le_1 : cJ <= 1.
Hypothesis one_le_cJ' : 1 <= cJ'.
\end{lstlisting}
This is \lbl{H:jump} (the manuscript's \lbl{eq:jumpcond}): for every interior face
shared by $K_1,K_2$,
\begin{equation}\label{eq:jumpcond-here}
  c_J\,\varphi_{1,K_1}\ \le\ \varphi_{1,K_2}\ \le\ c_J'\,\varphi_{1,K_1},
  \qquad 0<c_J\le1\le c_J' ,
\end{equation}
with $\varphi_{1,K}=\aK\,\tau_{1,\mathrm{NS}}^{-1}$.

\emph{Role.} It is what makes the \emph{jump} of $\tau_1$ across a face small
enough to be absorbed. Via \lbl{lem:jump} it yields
$\sigma\,\bigl|[\![\tau_1]\!]\bigr|\le C\,\sigt|_{K_i}\,\tau_{1,K_i}$, which is
the only reason Steps~6b and~6d of the continuity proof close. It is used
nowhere else.

\emph{Why it is reasonable.} It is the analogue of condition~(39) of
Codina~\cite{codina2001stabilized}. And it is weaker than it looks: by
\lbl{H:porosity} the factors $\alpha_{K_i}$ of neighbouring elements are
automatically comparable, and by \lbl{H:mesh} so are the diameters $h_{K_i}$;
so \eqref{eq:jumpcond-here} reduces to the comparability of
$|\mathbf a|_{\infty,K_1}$ and $|\mathbf a|_{\infty,K_2}$ relative to $\nu/h_K$
--- a mild condition on the advection field, and one that a continuous finite
element velocity from the previous iterate satisfies as a matter of course.

\emph{Tightening.} The manuscript is careful to say that \lbl{H:jump} is
``needed only when $\sigma>0$'': at $\sigma=0$ the jump of $\tau_1$ never
appears. The Coq files assume it unconditionally. This costs nothing (the
theorems are implications, and one may always supply a true hypothesis), but it
does mean the $\sigma=0$ specialisation is not obtained for free from
\coq{abstract_continuity}. A $\sigma$-case split would recover it.

\subsubsection{D5 --- Bounded face multiplicity}

\begin{lstlisting}
Hypothesis H_mult1 :
  forall g : K -> R, (forall k, 0 <= g k) ->
    Rsum Fl (fun f => g (e1 f)) <= Nf * Rsum Th g.
\end{lstlisting}
(and \coq{H_mult2} with \coq{e2}). This is the multiplicity clause of
\lbl{H:mesh}: for every nonnegative elemental quantity $g$,
\begin{equation}\label{eq:mult}
  \sum_{f\in\mathcal F}g\bigl(e_i(f)\bigr)\ \le\ N_f\sum_{K\in\Th}g(K),
  \qquad i=1,2 .
\end{equation}

\emph{Mathematical content.} \eqref{eq:mult} is exactly the statement that each
element occurs at most $N_f$ times in the list $\bigl(e_i(f)\bigr)_{f\in
\mathcal F}$ --- that is, that no element has more than $N_f$ faces. Stating it
as a functional inequality valid for all nonnegative $g$, rather than as a
counting fact, is the clean way to say this in a setting with no measure and no
decidable equality on \coq{K}.

It bundles one further fact, easy to miss and worth naming: applied to a $g$
supported \emph{off} $\Th$, the right-hand side of \eqref{eq:mult} is zero,
so the hypothesis also forces $e_i(f)\in\Th$ for every face --- the neighbours
of a face are elements of the mesh. That is mesh \emph{conformity}, not merely
bounded multiplicity, and it is assumed here rather than stated separately.

\emph{Role.} It is the sole device by which a sum over faces is converted into a
sum over elements, in Steps~6b and~6d. \emph{Reasonable.} It is a standard
consequence of shape-regularity: $N_f$ is bounded by the number of faces of the
reference element. \emph{Tightening.} None needed; \eqref{eq:mult} is already the
exact abstraction of the counting fact.

\subsubsection{D6 --- Porosity, and its resolution by the mesh}

\lbl{H:porosity} asks that $\alpha\in W^{1,\infty}(\Omega)$ with $0<\alpha\le1$,
and that the mesh \emph{resolve} the porosity variation:
\begin{equation}\label{eq:resolved-here}
  h_K\,\lVert\nabla\alpha\rVert_{L^\infty(K)}\ \le\ C_\alpha\,\alpha_{0,K}
  \qquad\forall K\in\Th ,
\end{equation}
whence --- for convex elements, and for shape-regular non-convex ones with
$\delta_\alpha$ enlarged by a factor depending only on the chunkiness parameter
(this qualifier is amendment~F2 of \S\ref{sec:amend}, added because the
formalisation needed it) --- $\alpha_{\infty,K}\le\delta_\alpha\,\alpha_{0,K}$
with $\delta_\alpha=1+C_\alpha$.

\emph{Where it lives in Coq.} Nowhere as a hypothesis of the abstract theorems
--- and that is worth understanding. \eqref{eq:resolved-here} is what the
manuscript uses to \emph{prove} \lbl{lem:winv} from the unweighted inverse
estimate; in the abstract files, \lbl{lem:winv} is assumed (as \coq{S3} and the
\coq{Hw_*}), so \lbl{H:porosity} has already been consumed and survives only
\emph{inside the constants} $\Cbar,\Cinv,C_I,C_{\mathrm{face}}$. Its scalar
core, however, \emph{is} machine-checked: \file{ContinuityAlgebra.v} proves
\coq{delta_alpha_bound} ($\alpha_\infty-\alpha_0\le C_\alpha\alpha_0
\Rightarrow \alpha_\infty\le\delta_\alpha\alpha_0$) and \coq{winv_ratio} (the
step $\alpha_\infty/\alpha_0^{1/2}\le\delta_\alpha^{1/2}\alpha_\infty^{1/2}$
that turns the unweighted estimate into the weighted one). What remains outside
the kernel is only the norm-level gluing.

\emph{Why it is reasonable.} It says: do not use a mesh so coarse that the
porosity swings wildly inside one element. That is not a restriction, it is
good practice, and it is exactly the condition under which the $\alpha$-weighted
inverse estimates carry a constant independent of $\alpha$.

% ---------------------------------------------------------------------
\subsection{Stratum 3 (Class G): the Green-type identities}\label{sec:stratumG}
% ---------------------------------------------------------------------

We come to the analytic base. Class~G collects the identities that come from
integration by parts. Each reduces, ultimately, to the divergence theorem on a
single element for integrands of the form (elementwise polynomial)~$\times$~($W^{1,\infty}$
weight), plus finite bookkeeping across faces --- material that is entirely
standard, and entirely absent from Coq's standard library, which is why it is
assumed. They are hand-audited against the manuscript in \file{AUDIT.md}~\S4 and
targeted by the Lean~4 roadmap.

They are not all of equal weight, and we say so plainly. \textbf{\coq{Horth} is
now the single largest assumption in the development}; \coq{H_face_p}/%
\coq{H_face_c} of \S\ref{sec:stratumI} is the second. The rest are small. An
earlier revision awarded that title to \coq{HBS}, the tested identity; the award
was a mistake, since \coq{HBS} was a free real pinned by its own defining
equation and so constrained no model. It is now a \emph{theorem}, and G1 is the
account of what its removal did and did not buy.

\subsubsection{G1 --- The tested identities: now theorems}

Both diagonals of $B_S$ that the development needs --- $B_S(U_h,U_h)$ for
stability and $B_S(W,W)$ for convergence --- used to be pinned by a monolithic
hypothesis (\coq{HBS}, \coq{HBS_W}). Neither is any longer --- and only one of
them needed a theorem to replace it. The statement at
issue, in the stability file's notation, is

\begin{lstlisting}
Theorem HBS :                      (* was: Hypothesis HBS *)
  BS = 2 * nu * Rsum Th (fun k => << (gu k) , (gu k) >>)
       + sigma * Rsum Th (fun k => << (uu k) , (uu k) >>)
       + eps * Rsum Th (fun k => << (pp k) , (pp k) >>)
       - Rsum Th (fun k => t1 k * << (Lstar_m k) , (L_m k) >>)
       - Rsum Th (fun k => t2 k * << (Lstar_c k) , (L_c k) >>).
\end{lstlisting}
In the manuscript's notation, with
$\mathcal L_m = \alpha\mathcal X(U_h)-G(\mathbf u_h)$,
$\mathcal L_m^{*} = -\alpha\mathcal X(U_h)-G(\mathbf u_h)$,
$\mathcal L_c = \nabla\!\cdot(\alpha\mathbf u_h)+\varepsilon p_h$ and
$\mathcal L_c^{*} = -\nabla\!\cdot(\alpha\mathbf u_h)+\varepsilon p_h$:
\begin{equation}\label{eq:HBS-here}
  B_S(U_h,U_h)
  = 2\nu\bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\bigr\rVert^2
  + \sigma\lVert\mathbf u_h\rVert^2
  + \varepsilon\lVert p_h\rVert^2
  - \sum_{K}\tau_{1,K}\bigl\langle \mathcal L_m^{*},\mathcal L_m\bigr\rangle_K
  - \sum_{K}\tau_{2,K}\bigl\langle \mathcal L_c^{*},\mathcal L_c\bigr\rangle_K .
\end{equation}
This is the manuscript's stability display \lbl{eq:StabilityEstimate} taken at
its \emph{first} line,
$B_S(\mathbf a;U_h,U_h)=B(\mathbf a;U_h,U_h)-\sum_K\langle\mathcal L^{*}U_h,
\boldsymbol\tau\mathcal LU_h\rangle$, composed with the Galerkin identity that
precedes it in the manuscript,
\begin{equation}\label{eq:galerkin-here}
  B(\mathbf a;U_h,U_h)
  = 2\nu\bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\bigr\rVert^2
  + \bigl\lVert\sigma^{1/2}\mathbf u_h\bigr\rVert^2
  + \varepsilon\lVert p_h\rVert^2 ,
\end{equation}
which the manuscript derives ``using the finite element unknown as the test
function and assuming $\nabla\!\cdot(\alpha\mathbf a)=\mathbf 0$ and
$\mathbf u_h=\mathbf 0$ on $\partial\Omega$''.

\emph{What it used to bundle, and where each piece now lives.} A great deal, and
the reader is entitled to know exactly what --- and, now that the hypothesis is
gone, exactly where each piece went. It bundled: (i) the definition of the
Galerkin form $B$ of the linearized problem and of the stabilization term
$-\sum_K\langle\mathcal L^{*}U_h,\boldsymbol\tau\mathcal LU_h\rangle$
(\lbl{eq:OSGSProblem}); (ii) the substitution $V_h:=U_h$; (iii) the
\emph{annihilation} of the Galerkin convective, pressure and mass contributions
at $V_h=U_h$ --- the terms $T_2+T_4$ of \lbl{eq:Bstab} --- which is what
$\nabla\!\cdot(\alpha\mathbf a)=0$ (\lbl{H:advection}) and
$\mathbf u_h\in H^1_0$ buy, and which is the whole content of
\eqref{eq:galerkin-here}; and (iv) the symmetry and idempotence of the pointwise
deviatoric-symmetric viscous projector $\mathcal P$ (i.e.\ $\mathcal
P^{\!\top}\mathcal P=\mathcal P$), which turns the manuscript's
$2(\nabla\mathbf v_h,\alpha\nu\mathcal P\nabla\mathbf u_h)$ --- note the
\emph{full} gradient in the first slot --- into
$2\nu\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\rVert^2$. ($\mathcal P$ here
is the viscous projector of the deviatoric operator, \emph{not} the orthogonal
subgrid projection $\pi_h$ that distinguishes OSGS from ASGS; it is never the
identity.)

\emph{Of those four, exactly one was provable, and it is now proved.} Piece~(iii)
--- the annihilation --- is genuine mathematics with genuine content, and it is
precisely what the two diagonal Green identities \eqref{eq:diag-here} supply. It
is now discharged inside the kernel (\coq{diag_zero}). Pieces (i), (ii) and~(iv)
are of a different character: they are not facts that could be \emph{derived}
from anything the kernel knows, because they are part of what it \emph{means}
for the eighteen-term expression to be the paper's $B_S$. They are the
Stratum-1 reading obligation of \S\ref{sec:stratum1}~(I4) --- the same obligation
that \file{AbstractContinuity.v} and \file{AbstractInterpolation.v} have always
carried, since those files have always \emph{defined} $B_S$ as $T_1+\dots+T_{18}$
rather than assuming a tested form for it.

\emph{So this is a genuine reduction, but not a free one, and we say which.} What
left the trusted base is a row of Table~\ref{tab:inventory}; what replaced it is
two rows (items 28--29) that are strictly smaller --- a reader checks each in a
line. That part is a real shrink: one monolithic row out, two elementary rows in,
and the difference-of-squares algebra that used to ride along inside the
hypothesis is now kernel-checked. But pieces (i), (ii) and~(iv) did not
\emph{vanish}; they \emph{moved}, from an explicit hypothesis in one file into
the definitional identification that the other three files already relied on.
The honest summary is that the stability file has stopped being the odd one out:
its $B_S$ is now identified the same way as everywhere else, and the trust that
identification carries is now visible in exactly one place (I4) instead of two.
The idempotence of $\mathcal P$, in particular, is still assumed --- it is the
reason the atom \coq{gu} may be read as $\alpha^{1/2}\mathcal P\nabla\mathbf u_h$
and $T_1=2\nu\langle\mathtt{gv},\mathtt{gu}\rangle$ read as the manuscript's
$2(\nabla\mathbf v_h,\alpha\nu\mathcal P\nabla\mathbf u_h)$ --- but it is assumed
as a reading of the atoms, not as a hypothesis of a theorem, and it was already
being assumed that way by three of the four files.

\emph{And one precision that cuts against the headline.} In the previous
revision \coq{BS} was a \coq{Variable} and \coq{HBS} an equation pinning it, so
the closed theorem had the shape $\forall b,\ b=t\Rightarrow P(b)$ --- which is
logically just $P(t)$. As a \emph{logical} assumption \coq{HBS} was therefore
eliminable: the row could have been struck at any time, at the price of a lemma
that spoke only about the tested form $t$ and not about the paper's $B_S$.
(Indeed the previous revision's stability file assumed no Green identity at
all.) What that row really bought was the \emph{caller's} obligation to re-prove
the tested identity for whatever $B_S$ it supplied. So the accurate claim is not
that a large assumption was discharged, but that a reader obligation was
internalised --- and items 28--29 are the honest price of making the lemma speak
about the eighteen-term $B_S$ itself. This is why the file's own count goes
\emph{up}, $16\to17$: the arithmetic $53\to50$ is a count of \emph{rows}, and
the reason to care is the reconciliation, not the count.

\emph{Role.} It is the \emph{starting line} of the stability proof. Everything
downstream is proved in Coq \emph{from} \eqref{eq:HBS-here} --- including, it is
worth stressing, the rest of \lbl{eq:StabilityEstimate} itself: the
manuscript's displayed right-hand side, the difference-of-squares form
$\dots+\lVert\tau_1^{1/2}\alpha\mathcal X(U_h)\rVert_h^2
-\lVert\tau_1^{1/2}G(\mathbf u_h)\rVert_h^2
+\lVert\tau_2^{1/2}\nabla\!\cdot(\alpha\mathbf u_h)\rVert_h^2
-\varepsilon^2\lVert\tau_2^{1/2}p_h\rVert_h^2$, is \emph{not} assumed. It is
derived inside the kernel from \eqref{eq:HBS-here} by \coq{momentum_pairing} and
\coq{mass_pairing}. So is everything after it: the Young step, the exact
cancellation of $\Cbar$, the coefficient collection, the assembly. The only
change is that \eqref{eq:HBS-here} is itself now a theorem, so ``everything
downstream'' now reaches all the way up to the starting line.

\emph{The two diagonal identities, which are what is actually assumed.} The
convergence file needs the same identity for the discrete-error pair,
$B_S(W,W)$; the stability file needs it for $(U_h,U_h)$. In an earlier revision
these were \emph{two} monolithic hypotheses, \coq{HBS} and \coq{HBS_W} --- and
both were genuinely independent of \coq{H_skew} and
\coq{H_ibp_vp}/\coq{H_ibp_qu}, which are the \emph{mixed} identities, relating
the $E$-atoms to the $W$-atoms. Obtaining \eqref{eq:HBS-here} on a diagonal
additionally requires their \emph{diagonal} instances,
\begin{equation}\label{eq:diag-here}
  \sum_K\bigl(\mathbf v,\ \alpha\,\mathbf a\!\cdot\!\nabla\mathbf v\bigr)_K = 0,
  \qquad
  \sum_K\bigl(\mathbf v,\ \alpha\nabla q\bigr)_K
  = -\sum_K\bigl(q,\ \nabla\!\cdot(\alpha\mathbf v)\bigr)_K ,
\end{equation}
which \coq{HBS} and \coq{HBS_W} were each silently bundling together with the
whole difference-of-squares expansion.

The development now assumes \eqref{eq:diag-here} --- and \emph{only}
\eqref{eq:diag-here} --- as the two hypotheses \coq{H_skew_diag} and
\coq{H_ibp_diag} (items 28--29 of Table~\ref{tab:inventory}). The stability file
then defines its diagonal $B_S$ to \emph{be} the eighteen-term expression at its
own atoms, and \textbf{proves} the tested identity:
\begin{lstlisting}
(* qualified names elided for readability; see the source for the full form *)
Definition BS : R :=               (* was: Variable BS : R *)
  AbstractInterpolation.BS Hs K Th nu sigma eps c1 c2 hK aK am
    gu gu du du cxu cxu gpu gpu uu uu pp pp divu divu.

Theorem HBS :                      (* was: Hypothesis HBS *)
  BS = 2*nu * Rsum Th (fun k => << gu k , gu k >>) + ... (as displayed above)
\end{lstlisting}
The proof is the stability argument run backwards: the eighteen terms are
expanded, the three pairs that cancel by symmetry of the inner product do so
(\coq{sym_uu_du}, \coq{sym_x_du}, \coq{sym_x_uu}), the momentum and mass pairings
are re-assembled by the difference-of-squares identity (\coq{momentum_group},
\coq{mass_group}, built on \coq{momentum_pairing} and \coq{Bv_expand}), and what
remains --- exactly $T_2+T_4$ --- vanishes by \eqref{eq:diag-here}
(\coq{diag_zero}). No positivity hypothesis is used: the proof consumes items
28--29 and nothing else from the trusted base.

Note the direction of the dependency this creates: \file{AbstractStability.v} now
imports \file{AbstractInterpolation.v}, in order to reuse the \emph{same}
eighteen-term \coq{BS} rather than restate it. That is the point --- the number
the stability lemma bounds from below is then, by construction, the number the
interpolation and convergence files bound from above, with no side condition
reconciling them.

\emph{And \coq{HBS_W} did not even need a theorem of its own.} Once
\coq{abstract_stability} is a closed theorem over
\coq{H_skew_diag}/\coq{H_ibp_diag}, the convergence file gets the $(W,W)$ tested
identity by simply \emph{applying} it at the discrete-error atoms. It defines
\coq{BSWW} as the same eighteen-term expression at those atoms, observes that
this is \emph{definitionally} the stability file's \coq{BS} there, and applies
the theorem:
\begin{lstlisting}
Definition BSWW : R :=
  AbstractInterpolation.BS Hs K Th nu sigma eps c1 c2 hK aK am
    gv gv dv dv cxv cxv gpv gpv vv vv qq qq divv divv.

Lemma BSWW_is_ASBS :               (* one and the same expression *)
  BSWW = AbstractStability.BS Hs K Th nu sigma eps c1 c2 hK aK am
           gv vv qq cxv gpv divv dv.
Proof. reflexivity. Qed.

Lemma stab_W : 0 < Cst /\ BSWW >= Cst * NW ^ 2.
Proof.
  pose proof (abstract_stability ... gv vv qq cxv gpv divv dv
                H_skew_diag H_ibp_diag Hw_dv Heps_W) as [H1 H2].
  split; [exact H1 |]. rewrite NW2_bridge in H2.
  rewrite BSWW_is_ASBS. exact H2.
Qed.
\end{lstlisting}
So there is no \coq{Theorem HBS_W} in the development, and no
\coq{Hypothesis HBS_W} either: the whole item is \emph{gone}, and the
difference-of-squares expansion it used to smuggle in is proved once, in the
stability file, and reused. The convergence file's \coq{H_skew_diag} and
\coq{H_ibp_diag} are the same two identities as the stability file's, stated at
its own atoms --- which is why Table~\ref{tab:inventory} lists them once, marked
\textbf{S},\textbf{V}, rather than twice.

Three things are gained. Both class-G bundles are \emph{gone from the trusted
base}, replaced by two elementary identities that a reader can check in a line
each --- and those same two identities serve \emph{both} files, so the second
tested identity costs nothing at all beyond the first. The two encodings of
$B_S$ --- the eighteen-term expression and the tested form --- are now provably the
same number, which is the internal consistency check the development previously
lacked (\S\ref{sec:stratum1}, I4). And the non-vacuity witness got sharper as a
side effect: $B_S(U_h,U_h)=7/8$ used to be a number \file{NonVacuity.v}
\emph{chose} (a free real, set by fiat, then justified against the assumed
\coq{HBS}); it is now a number that file \emph{computes} from the definition, so
the witness has strictly less freedom and the headline $7/8\ge7/16$ is the
stronger for it (\S\ref{sec:vacuity}).

\subsubsection{G2 --- Galerkin orthogonality}

\begin{lstlisting}
Hypothesis Horth : BSWW = - BSEW.
\end{lstlisting}
i.e., writing $E=U-\widehat U_h$ and $W=\widehat U_h-U_h$,
\begin{equation}\label{eq:horth-here}
  B_S(W,W) \;=\; -\,B_S(E,W).
\end{equation}

\emph{What it bundles.} Consistency --- the identity
$B_S(U-U_h,\,V_h)=0$ for all $V_h\in\mathcal X_h$ (\lbl{eq:consistency}), which
holds because the stabilized form is residual-based and the exact solution has
enough regularity (\lbl{eq:regularity}) for $\mathcal L_{\mathbf a}U\in L^2(K)$
--- together with bilinearity of $B_S$ in its first argument and the algebraic
identity $W=(U-U_h)-E$.

\emph{Role.} It is the single line
$B_S(E_h,E_h)=B_S(\widehat U_h-U,E_h)$ of the manuscript's proof of
\lbl{thm:convergence}, and it is what makes the convergence theorem a statement
about the \emph{actual} discrete solution rather than about an arbitrary pair of
fields. Without it, $W$ would be an unconstrained vector and no bound on it
could hold.

\emph{Why it is reasonable.} Consistency is the design property of a
residual-based stabilization: the exact solution makes every subgrid residual
vanish, so it satisfies the discrete problem exactly. It is the least surprising
assumption in the list, and the easiest to check by hand.
\emph{Tightening.} Like \coq{HBS} before it became a theorem, it becomes
provable the moment a concrete variational instance is formalised; it needs no
primitive beyond the divergence-theorem target T4 of the roadmap. Unlike
\coq{HBS}, it cannot be discharged by the diagonal identities: those kill
$T_2+T_4$ on a diagonal, whereas \coq{Horth} relates two \emph{different}
arguments $E$ and $W$, and so needs the bilinearity that the scalar encoding of
$B_S$ does not give the kernel (\S\ref{sec:stratum1}, I4).

\subsubsection{G3 --- Skew-symmetry of the convective term}

\begin{lstlisting}
Hypothesis H_skew :
  Rsum Th (fun k => << (vv k) , (cxu k) >>)
  = - Rsum Th (fun k => << (uu k) , (cxv k) >>).
\end{lstlisting}
This is the manuscript's \lbl{eq:skew} (identity (i) of Step 6a):
\begin{equation}\label{eq:skew-here}
  \sum_{K}\bigl\langle \mathbf v,\ \alpha\,\mathbf a\!\cdot\!\nabla\mathbf u
  \bigr\rangle_K
  \;=\; -\sum_{K}\bigl\langle \mathbf u,\ \alpha\,\mathbf a\!\cdot\!\nabla\mathbf v
  \bigr\rangle_K .
\end{equation}

\emph{Why it is true, and why it is \emph{global}.} Because
$\alpha\mathbf a\cdot\nabla(\mathbf u\cdot\mathbf v)$ integrates to zero over
$\Omega$ when $\nabla\!\cdot(\alpha\mathbf a)=0$ (\lbl{H:advection}) and
$\mathbf u\cdot\mathbf v\in W^{1,1}_0(\Omega)$. Note carefully that
\eqref{eq:skew-here} is asserted for the \emph{sum over the mesh}, not
element by element --- and that is correct: the elementwise identity carries a
face term, which is exactly the content of \coq{H_elem_conv_ibp} below. The
global version has no face term because the face contributions cancel in pairs.
The formalisation keeps the two apart, as it must.
\emph{Role.} Step~6a of the continuity proof. \emph{Reasonable.} Textbook.

\emph{A boundary condition hides here too.} That
$\mathbf u\cdot\mathbf v\in W^{1,1}_0(\Omega)$ requires \emph{both} arguments to
vanish on $\partial\Omega$. For $\mathbf v_h$ that is the discrete space; for the
first argument in \lbl{lem:continterp} --- the interpolation error
$\mathbf e_{\mathbf u}=\mathbf u-\widehat{\mathbf u}_h$ --- it requires the
\emph{interpolant to preserve the homogeneous Dirichlet condition},
$\widehat{\mathbf u}_h\in\mathcal V_{h0}$. The manuscript says so; Coq assumes
it, silently, inside \coq{H_skew} and the two global identities of G4. It is
true for the Lagrange interpolant of a function vanishing on the boundary, and
it is worth knowing that this is where it is being used.

\subsubsection{G4 --- The global pressure identities}

\begin{lstlisting}
Hypothesis H_ibp_vp :
  Rsum Th (fun k => << (vv k) , (gpu k) >>)
  = - Rsum Th (fun k => << (pp k) , (divv k) >>).
Hypothesis H_ibp_qu :
  Rsum Th (fun k => << (qq k) , (divu k) >>)
  = - Rsum Th (fun k => << (uu k) , (gpv k) >>).
\end{lstlisting}
These are the manuscript's \lbl{eq:globalibp} (identity (ii)):
\begin{equation}\label{eq:globalibp-here}
  \sum_K\bigl\langle\mathbf v,\ \alpha\nabla p\bigr\rangle_K
  = -\sum_K\bigl\langle p,\ \nabla\!\cdot(\alpha\mathbf v)\bigr\rangle_K,
  \qquad
  \sum_K\bigl\langle q,\ \nabla\!\cdot(\alpha\mathbf u)\bigr\rangle_K
  = -\sum_K\bigl\langle\mathbf u,\ \alpha\nabla q\bigr\rangle_K .
\end{equation}
\emph{Content.} The divergence theorem on $\Omega$ with vanishing velocity
traces. \emph{Role.} Step~6a. \emph{Reasonable.} Textbook.

\subsubsection{G5 --- The elemental identities, with their face terms}

\begin{lstlisting}
Lemma H_elem_conv_ibp :          (* was: Hypothesis -- see below *)
  forall k, << (cxu k) , (vv k) >> = - << (cxv k) , (uu k) >> + FBc_e k.
Lemma H_elem_p_ibp :             (* was: Hypothesis -- see below *)
  forall k, << (vv k) , (gpu k) >> = - << (divv k) , (pp k) >> + FBp_e k.
\end{lstlisting}
The first is the manuscript's \lbl{eq:elemibp} (identity (iii)); the second is the
elemental pressure identity of Step~6b:
\begin{equation}\label{eq:elemibp-here}
\begin{aligned}
  \bigl\langle\alpha\,\mathbf a\!\cdot\!\nabla\mathbf u,\ \mathbf v\bigr\rangle_K
  &= -\bigl\langle\alpha\,\mathbf a\!\cdot\!\nabla\mathbf v,\ \mathbf u\bigr\rangle_K
   + \int_{\partial K}\alpha\,(\mathbf n_K\!\cdot\!\mathbf a)\,
     (\mathbf u\!\cdot\!\mathbf v)\,\mathrm d\Gamma,\\
  \bigl\langle\mathbf v,\ \alpha\nabla p\bigr\rangle_K
  &= -\bigl\langle\nabla\!\cdot(\alpha\mathbf v),\ p\bigr\rangle_K
   + \int_{\partial K}\alpha\,(\mathbf n_K\!\cdot\!\mathbf v)\,p\,\mathrm d\Gamma .
\end{aligned}
\end{equation}
These are \eqref{eq:skew-here} and \eqref{eq:globalibp-here} \emph{localised},
with the face terms retained. (The manuscript writes \lbl{eq:elemibp} with the
elementwise-constant factor $\tau_{1,K}$ carried through all three terms; Coq
divides it out here --- legitimately, $\tau_{1,K}$ being a positive constant on
$K$ --- and reintroduces it in the assembly hypotheses of G6.) The first needs $\nabla\!\cdot(\alpha\mathbf a)=0$
to hold \emph{on each element} --- which is not automatic from the distributional
statement, and is exactly why amendment~F1 (\S\ref{sec:amend}) strengthened
\lbl{H:advection} to grant
$\mathbf a|_K\in W^{1,\infty}(K)^d$ elementwise together with the localisation
of $\nabla\!\cdot(\alpha\mathbf a)=0$. This is a good illustration of the
exercise paying for itself: the formalisation \emph{found} the missing clause,
because the Coq hypothesis could not be justified without it.

\emph{These two identities constrained nothing on their own --- so they are now
theorems.} A sharp reader will notice that the face terms
$\mathrm{FB}_{c,e}$ and $\mathrm{FB}_{p,e}$ used to be \emph{free} variables
(\coq{Variable FBc_e : K -> R}). Given \emph{any} atoms whatsoever, one could
simply \emph{define}
$\mathrm{FB}_{c,e}(K):=\langle\alpha\mathbf a\!\cdot\!\nabla\mathbf u,\mathbf
v\rangle_K+\langle\alpha\mathbf a\!\cdot\!\nabla\mathbf v,\mathbf u\rangle_K$
and \eqref{eq:elemibp-here} would hold by construction. Taken in isolation,
then, the two hypotheses were \emph{definitions in disguise}: they named the
face term rather than constraining anything.

That is exactly what the development now does, and openly:
\begin{lstlisting}
Definition FBc_e (k : K) : R := << cxu k , vv k >> + << cxv k , uu k >>.
Definition FBp_e (k : K) : R := << vv k , gpu k >> + << divv k , pp k >>.

Lemma H_elem_conv_ibp :          (* was: Hypothesis *)
  forall k, << cxu k , vv k >> = - << cxv k , uu k >> + FBc_e k.
Proof. intro k. unfold FBc_e. ring. Qed.
\end{lstlisting}
This is faithful, not a dodge: by the divergence theorem on $K$ the boundary
integral \emph{is} the integration-by-parts defect, so defining it as that
defect is what \lbl{eq:elemibp} says. Two hypotheses leave the trusted base of
each of the three files, and nothing is lost --- because the mathematical
content of the boundary terms never lived here. It lives in the two hypotheses
that \emph{do} constrain them: the assembly identities of~G6, which tie
$\sum_K\tau_{1,K}\mathrm{FB}_{\cdot,e}(K)$ to interior-face integrals (this is
where the divergence theorem and the vanishing of the boundary faces actually
enter), and the face estimates of \S\ref{sec:stratumI}, which bound those
integrals by norms. The trio carried the weight all along; now the bookkeeping
says so.

\subsubsection{G6 --- Facewise assembly, Steps 6b and 6d}

\begin{lstlisting}
Hypothesis H_assemble_p :
  Rsum Th (fun k => t1 k * FBp_e k)
  = Rsum Fl (fun f => (t1 (e1 f) - t1 (e2 f)) * FBp f).
\end{lstlisting}
(and \coq{H_assemble_c}, identically, with the convective face integral), i.e.
\begin{equation}\label{eq:assemble-here}
  \sum_{K\in\Th}\tau_{1,K}\int_{\partial K}\!\!\alpha\,(\mathbf n_K\!\cdot\!\mathbf v)\,p
  \;=\;
  \sum_{\Gamma\in\mathcal F}\bigl(\tau_{1,K_1(\Gamma)}-\tau_{1,K_2(\Gamma)}\bigr)
  \int_{\Gamma}\!\alpha\,(\mathbf n\!\cdot\!\mathbf v)\,p .
\end{equation}

\emph{Content.} Two facts, bundled. First, on an interior face the integrand is
single-valued (the fields are continuous across faces) while the two outward
normals are opposite, so the two elemental contributions combine into one
face integral weighted by the \emph{jump} $[\![\tau_1]\!]$ --- and $\tau_1$ may
be pulled out because it is elementwise constant. Second, \textbf{the boundary
faces contribute nothing}, because $\mathbf u_h=\mathbf v_h=\mathbf 0$ on
$\partial\Omega$; this is why the right-hand side sums over interior faces only.

\emph{Honest note.} The second fact --- the homogeneous Dirichlet boundary
condition --- enters the continuity proof \emph{here}, inside an assumption. It
is not derived. A reader auditing \coq{H_assemble_p} is auditing both the
face-pairing combinatorics and the boundary condition at once.
\emph{Role.} Steps~6b and~6d. \emph{Reasonable.} Finite bookkeeping over the
face lattice, once traces of continuous piecewise polynomials are available.

% ---------------------------------------------------------------------
\subsection{Stratum 3 (Class I): inverse and interpolation estimates}\label{sec:stratumI}
% ---------------------------------------------------------------------

Class~I collects the estimates that ``know'' about polynomials and about
shape-regular geometry: the one kind of fact that a formalisation over an
abstract inner-product space cannot possibly derive, and must therefore be
given. Each is assumed with an \emph{explicit constant}, quantified uniformly
over the elements --- and, as noted in (I5), that uniformity is where
shape-regularity does its work.

A pedagogical remark that organises everything below. \textbf{An inverse
estimate is a statement about polynomials, and is false for general Sobolev
functions.} It says that on an element of size $h$, differentiating a
polynomial of bounded degree costs at most a factor $C/h$ --- which is true
because the polynomial space is finite-dimensional and all norms on it are
equivalent, with the $h$-dependence supplied by scaling to a reference element.
Nothing of the sort holds for an arbitrary $H^1$ function. This is why, in
\lbl{lem:continterp}, the estimates applied to the \emph{first} argument must
change when that argument stops being a finite element function and becomes the
interpolation error: they are replaced by \emph{interpolation} estimates. The
Coq files make this substitution visible --- though, as~I2 sets out, not quite
textual.

\subsubsection{I1 --- The weighted inverse estimates}

\emph{One schema, nine instances.} Every weighted inverse estimate in the
development has the \emph{same shape} --- a bounded atom $A_K$ controlled by a
source atom $B_K$, with an $h_K^{-1}$ and a porosity weight $W_K$ in between. The
Coq layer names that shape once, in the infrastructure file
\file{InverseEstimates.v}:
\begin{lstlisting}
Definition winv_est (C : R) (W : K -> R) (A B : K -> carrier Hs) : Prop :=
  forall k, nrm (A k) <= C / hK k * W k * nrm (B k).
\end{lstlisting}
that is, $\lVert A_K\rVert_K\le (C/h_K)\,W_K\,\lVert B_K\rVert_K$ on every
element $K$, with the weight
$W_K\in\{\aK^{1/2},\,\aK,\,\aK|\mathbf a|_{\infty,K}\}$ and $C$ one of
$\Cinv,2\nu\Cbar,\Cbar$. This predicate is \textbf{purely notational}: each of
the nine hypotheses below is stated \emph{as} a \coq{winv_est} instance and
unfolds \emph{definitionally} to exactly the estimate it always was (six of the
nine to nothing more than $\beta$-reduction; the three $\aK|\mathbf a|/\aK$-weight
ones re-associate and match propositionally, which every \coq{pose proof}$\,\ldots\,$\coq{nra}
consumer absorbs). So \coq{winv_est} \emph{replaces no hypothesis} and adds none:
the trusted base is the same nine propositions as before, only written through a
common name --- the row count of Table~\ref{tab:inventory} is unchanged. The nine
\emph{cannot} be collapsed into a single \coq{winv_est} quantified over an
arbitrary vector (\coq{forall x, nrm (D x) <= ...}): that would be
\emph{unsound}. The discrete atoms and the interpolation-error atoms inhabit the
\emph{same} carrier, so a \coq{forall}-$x$ inverse estimate would license an
inverse estimate for a \emph{non-polynomial} interpolation error --- false, and
the very fact I2 turns on. The atoms are therefore kept as nine independent named
hypotheses; the ``true for polynomials / true for interpolation errors, never
both'' firewall is exactly what forbids the merge, and \coq{winv_est} is notation
\emph{over} the nine, never a tenth hypothesis.

The stability lemma needs exactly one of them:
\begin{lstlisting}
Hypothesis S3 :
  forall k, nrm (du k) <= (2 * nu * Cb / hK k) * sqrt (aK k) * nrm (gu k).
\end{lstlisting}
\begin{equation}\label{eq:S3-here}
  \bigl\lVert 2\nu\,\nabla\!\cdot(\alpha\,\mathcal P\nabla\mathbf u_h)\bigr\rVert_K
  \;\le\; \frac{2\nu\,\Cbar}{h_K}\,\aK^{1/2}\,
  \bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\bigr\rVert_K ,
\end{equation}
which is \lbl{eq:winv-divvisc} (the factor $2\nu$ is present because the Coq atom
\coq{du} models the full second-order term $2\nu\nabla\!\cdot(\alpha\mathcal
P\nabla\mathbf u_h)$).

\emph{Role.} It is the \emph{only} analytic input to the stability lemma besides
the two diagonal Green identities (items 28--29), and its role is beautiful: it
is consumed in \coq{du_sq_bound}, whose
bound is then fed through the Young step with parameter
$\lambda=\xi\Cbar^{2}\nu\aK/h_K^{2}$, and the constant $\Cbar$ \emph{cancels
exactly}:
\[
  \frac{4\nu^2\Cbar^2\,\aK/h_K^2}{\lambda} \;=\; \frac{4\nu}{\xi} .
\]
That exact cancellation is the mathematical heart of the stability proof --- it
is why the final constants depend on $\Cbar$ only through the ratio
$\Cbar^2/c_1$ --- and it is machine-checked.

\emph{Why it is reasonable.} It is the standard inverse estimate
(Brenner--Scott~\cite{brenner2008mathematical}, Lemma~4.5.3;
Ern--Guermond~\cite{ernguermond}), reduced from the unweighted case by
\lbl{H:porosity}; the manuscript proves that reduction, obtaining
$\Cbar=\sqrt{d\,\delta_\alpha}\,\Cinv+C_\alpha$. \emph{Tightening.} Note that
$\Cbar>\Cinv$ \emph{strictly}: the porosity is not piecewise polynomial, and
\lbl{rem:winvconst} is the manuscript's acknowledgement of the price. Since $c_1$
must exceed $\xi\Cbar^2$ (the manuscript's rule of thumb being $2\xi\Cbar^2$), a
sharper $\Cbar$ buys a smaller admissible $c_1$ and hence less stabilization ---
a concrete reason to care.

\emph{One relation deliberately not imposed.} In the manuscript
$\Cbar=\sqrt{d\,\delta_\alpha}\,\Cinv+C_\alpha\ \ge\ \Cinv$. Coq declares
\coq{Cinv} and \coq{Cb} as two unrelated positive reals and never assumes
$\Cinv\le\Cbar$. The theorems are correspondingly more general (they hold for
any admissible pair), which is sound and costs nothing --- but it does mean that
the explicit constants the Coq files produce, such as $C_{\mathrm{tot}}$, are
not automatically comparable term-by-term with the manuscript's until the
relation is supplied.

The continuity lemma needs eight instances --- and \emph{exactly} eight, one for
each way the first or second argument enters the eighteen terms. They are the
whole of \lbl{lem:winv}, instantiated. Read through the schema, each is a
\coq{winv_est}$\ C\ W\ A\ B$ line --- bounded atom $A$, source atom $B$, constant
$C$, weight $W$ --- so the family collapses to one display and a data table:

\begin{center}
\small
\begin{tabular}{@{}llllll@{}}
\toprule
Coq & bounded $A_K$ & source $B_K$ & $C$ & $W_K$ & Paper \\
\midrule
\coq{Hw_gu}, \coq{Hw_gv}
  & $\alpha^{1/2}\mathcal P\nabla\mathbf w_h$ & $\mathbf w_h$
  & $\Cinv$ & $\aK^{1/2}$ & \lbl{eq:winv-grad} \\
\coq{Hw_du}, \coq{Hw_dv} ($\equiv$\,\coq{S3})
  & $2\nu\nabla\!\cdot(\alpha\mathcal P\nabla\mathbf w_h)$
  & $\alpha^{1/2}\mathcal P\nabla\mathbf w_h$
  & $2\nu\Cbar$ & $\aK^{1/2}$ & \lbl{eq:winv-divvisc} \\
\coq{Hw_cxu}, \coq{Hw_cxv}
  & $\alpha\,\mathbf a\!\cdot\!\nabla\mathbf w_h$ & $\mathbf w_h$
  & $\Cinv$ & $\aK\,|\mathbf a|_{\infty,K}$ & \lbl{eq:winv-conv} \\
\coq{Hw_gpu}
  & $\alpha\nabla r_h$ & $r_h$
  & $\Cinv$ & $\aK$ & \lbl{eq:winv-gradp} \\
\coq{Hw_divu}
  & $\nabla\!\cdot(\alpha\mathbf w_h)$ & $\mathbf w_h$
  & $\Cbar$ & $\aK$ & \lbl{eq:winv-divu} \\
\bottomrule
\end{tabular}
\end{center}

\noindent
Each row reads back to its full displayed norm by unfolding \coq{winv_est}: the
\coq{Hw_du} row is $\lVert 2\nu\nabla\!\cdot(\alpha\mathcal P\nabla\mathbf
w_h)\rVert_K\le(2\nu\Cbar/h_K)\,\aK^{1/2}\lVert\alpha^{1/2}\mathcal P\nabla\mathbf
w_h\rVert_K$, i.e.\ \lbl{eq:winv-divvisc}, which is \emph{definitionally} the
stability lemma's \coq{S3} \eqref{eq:S3-here} --- the same proposition, so
\file{AbstractConvergence.v} discharges \coq{S3} from the $V$-side \coq{Hw_dv} by
the kernel with no new assumption.

\emph{Composition, once and for all.} Where the proof needs the \emph{double}
inverse estimate --- $\mathtt{du}$ bounded directly by $\mathtt{uu}$ by chaining
\coq{Hw_du} through \coq{Hw_gu}, so that $\aK^{1/2}\!\cdot\!\aK^{1/2}=\aK$ and
$h_K^{-1}\!\cdot\!h_K^{-1}=h_K^{-2}$ --- the generic lemma \coq{winv_compose} of
\file{InverseEstimates.v} derives it from \emph{any} two \coq{winv_est} instances
sharing a middle atom:
\[
  \coq{winv_est}\ C_1\,W_1\,A\,M \ \text{ and } \
  \coq{winv_est}\ C_2\,W_2\,M\,B
  \ \Longrightarrow\
  \lVert A_K\rVert_K \;\le\; \frac{C_1C_2\,(W_1W_2)_K}{h_K^{2}}\,\lVert B_K\rVert_K,
\]
giving $\lVert\mathtt{du}\rVert_K\le 2\nu\Cbar\Cinv\,\aK\,h_K^{-2}
\lVert\mathtt{uu}\rVert_K$ at $C_1=2\nu\Cbar$, $C_2=\Cinv$ --- coefficient
$2\nu\Cbar\Cinv\,\aK/h_K^2$, exactly the appendix's ``double'' estimate. This is
the estimate the continuity and interpolation files formerly re-derived by hand
in a $\approx20$-line block present \emph{byte-for-byte three times}
(\coq{double_inv_u}, \coq{double_inv_v}, and its \file{AbstractInterpolation.v}
twin); \coq{winv_compose} retires that triplication without changing a single
conclusion. It is a \emph{proved lemma}, not a hypothesis --- it adds nothing to
the trusted base, and \coq{Print Assumptions} on it returns only the three
stdlib axioms.

\noindent
The minimality is itself a form of honesty: the trusted base contains the
instances the proof consumes and no others. In \file{AbstractInterpolation.v}
only the three $V$-side instances (\coq{Hw_gv}, \coq{Hw_dv}, \coq{Hw_cxv})
survive, because the first argument is no longer discrete --- see I2. And in
\file{AbstractConvergence.v} the stability lemma's \coq{S3} is discharged
\emph{by the kernel} from \coq{Hw_dv}: the two are literally the same statement,
so the composition introduces no new assumption.

\subsubsection{I2 --- The interpolation estimates}

\begin{lstlisting}
Hypothesis HI_gu   : forall k, nrm (gu k)   <= CI * sqrt (aK k) / hK k * IU k.
Hypothesis HI_du   : forall k, nrm (du k)   <= 2 * nu * CI * aK k / (hK k)^2 * IU k.
Hypothesis HI_cxu  : forall k, nrm (cxu k)  <= CI * aK k * am k / hK k * IU k.
Hypothesis HI_gpu  : forall k, nrm (gpu k)  <= CI * aK k / hK k * IP k.
Hypothesis HI_divu : forall k, nrm (divu k) <= CI * aK k / hK k * IU k.
Hypothesis HI_uu   : forall k, nrm (uu k)   <= CI * IU k.
Hypothesis HI_pp   : forall k, nrm (pp k)   <= CI * IP k.
\end{lstlisting}
With $E=[\mathbf e_{\mathbf u};e_p]=U-\widehat U_h$ and the elemental
interpolation-error sizes of \lbl{eq:Eint},
\[
  E_{\mathrm{int},K}(\mathbf u)=h_K^{k_u+1}\lVert\mathbf u\rVert_{H^{k_u+1}(K)},
  \qquad
  E_{\mathrm{int},K}(p)=h_K^{k_p+1}\lVert p\rVert_{H^{k_p+1}(K)},
\]
these are exactly the manuscript's replacement table
\lbl{eq:interpdivvisc}--\lbl{eq:interpzero}:
\begin{equation}\label{eq:interptable-here}
\begin{aligned}
  \bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf e_{\mathbf u}\bigr\rVert_K
    &\le C\,\aK^{1/2}h_K^{-1}E_{\mathrm{int},K}(\mathbf u), &
  \bigl\lVert 2\nu\nabla\!\cdot(\alpha\mathcal P\nabla\mathbf e_{\mathbf u})\bigr\rVert_K
    &\le C\,\nu\,\aK\,h_K^{-2}E_{\mathrm{int},K}(\mathbf u), \\
  \lVert\alpha\,\mathbf a\!\cdot\!\nabla\mathbf e_{\mathbf u}\rVert_K
    &\le C\,\aK|\mathbf a|_{\infty,K}h_K^{-1}E_{\mathrm{int},K}(\mathbf u), &
  \lVert\alpha\nabla e_p\rVert_K
    &\le C\,\aK\,h_K^{-1}E_{\mathrm{int},K}(p), \\
  \lVert\nabla\!\cdot(\alpha\mathbf e_{\mathbf u})\rVert_K
    &\le C\,\aK\,h_K^{-1}E_{\mathrm{int},K}(\mathbf u), &
  \lVert\mathbf e_{\mathbf u}\rVert_K \le C\,E_{\mathrm{int},K}(\mathbf u),
  \quad & \lVert e_p\rVert_K \le C\,E_{\mathrm{int},K}(p).
\end{aligned}
\end{equation}

\emph{The structural point.} Compare \eqref{eq:interptable-here} with the table
of I1: they carry \emph{the same elemental coefficients}, with
$\lVert\mathbf u_h\rVert_K$ and $\lVert p_h\rVert_K$ replaced by
$E_{\mathrm{int},K}(\mathbf u)$ and $E_{\mathrm{int},K}(p)$. That is the entire
reason the appendix can say ``rerun the continuity proof with first argument
$E$'' and be believed, and the formalisation makes the claim checkable rather
than merely plausible.

Two details of the correspondence are worth stating precisely, because the
substitution is not quite textual. First, the \coq{HI_*} bounds match the
\coq{Hw_*} bounds \emph{as the continuity proof composes them}, not one by one:
\coq{Hw_du} bounds $\mathtt{du}$ by $\mathtt{gu}$ and \coq{Hw_gu} bounds
$\mathtt{gu}$ by $\mathtt{uu}$, and it is their \emph{composition} --- the
appendix's ``double'' inverse estimate, with coefficient
$2\nu\Cbar\Cinv\,\aK/h_K^2$ --- that \coq{HI_du}, with coefficient
$2\nu C_I\,\aK/h_K^2$, replaces. Second, the two inverse constants $\Cinv$ and
$\Cbar$ are \emph{merged into the single} $C_I$, as the source says in so many
words (``one constant \coq{CI} for all of them''). Third, \coq{HI_uu} and
\coq{HI_pp} have no \coq{Hw_} counterpart at all: a \emph{discrete} first
argument needs no zeroth-order bound, because $\lVert\mathbf u_h\rVert_K$ simply
\emph{is} the right-hand side of the other estimates, whereas an interpolation
error must itself be bounded by $E_{\mathrm{int},K}$. Hence seven \coq{HI_*}
against five first-argument \coq{Hw_*}. With those three readings in place, the
file reruns --- and the kernel re-checks --- every one of the eighteen term
bounds.

\emph{Why they are reasonable.} They are Bramble--Hilbert / Deny--Lions
technology on a shape-regular simplex (Brenner--Scott~\cite{brenner2008mathematical},
Thm.~4.4.20; Ern--Guermond~\cite{ernguermond}), combined with
\eqref{eq:resolved-here} for the $\nabla\alpha$ terms and the pointwise
contraction $|\mathcal P\mathbf T|\le|\mathbf T|$. The $L^\infty$ variants
(\lbl{eq:interpinfty}, used inside the face estimates) additionally need
$H^{k+1}(K)\hookrightarrow C^0(\overline K)$, valid for $d\le3$ since
$k_u+1,k_p+1>d/2$ --- and this is where the restriction $d\le3$ in the
manuscript's proof of \lbl{lem:continterp} comes from.

\emph{An important limit of what is proved.} \coq{IU} and \coq{IP} are abstract
nonnegative functionals (\coq{Variables (IU IP : K -> R)} with
\coq{IU_nonneg}/\coq{IP_nonneg}). Coq does \emph{not} know that
$\mathtt{IU}\,k=h_K^{k_u+1}\lVert\mathbf u\rVert_{H^{k_u+1}(K)}$. Consequently
the machine-checked convergence theorem reads
\[
  |||U-U_h||| \;\le\; \Bigl(K_{\mathrm{abs}}^{I}+C_{\mathrm{tot}}^{I}/\Cstab\Bigr)
  \bigl(\Psi_U+\Psi_P\bigr),
\]
with $\Psi_U,\Psi_P$ built from the abstract \coq{IU},~\coq{IP}. \textbf{That
this is an $O(h^{k+1})$ rate is \emph{not} a kernel fact}: it follows only after
instantiating \coq{IU},~\coq{IP} with \lbl{eq:Eint}, which is a Stratum-1
identification. What the kernel certifies is the estimate; what turns the
estimate into a convergence \emph{rate} is the reading of $E_{\mathrm{int},K}$
--- and the $O(h^{k+1})$ plateaus of the manufactured-solution tests are the
empirical confirmation that the reading is the right one.

\subsubsection{I3 --- The face estimates: the most bundled hypothesis}

\begin{lstlisting}
Hypothesis H_face_p :
  forall f, Rabs (FBp f)
  <= Cface * (  aK (e1 f) / hK (e1 f) * (nrm (vv (e1 f)) * nrm (pp (e1 f)))
              + aK (e2 f) / hK (e2 f) * (nrm (vv (e1 f)) * nrm (pp (e2 f)))
              + aK (e1 f) / hK (e1 f) * (nrm (vv (e2 f)) * nrm (pp (e1 f)))
              + aK (e2 f) / hK (e2 f) * (nrm (vv (e2 f)) * nrm (pp (e2 f)))).
\end{lstlisting}
(and \coq{H_face_c}, the same with $\alpha(\mathbf n\!\cdot\!\mathbf a)
(\mathbf u\!\cdot\!\mathbf v)$ and an extra $|\mathbf a|_{\infty,K}$), i.e.\ for
every interior face $\Gamma$ shared by $K_1,K_2$,
\begin{equation}\label{eq:face-here}
  \Bigl|\int_\Gamma \alpha\,(\mathbf n\!\cdot\!\mathbf v)\,p\Bigr|
  \;\le\; C_{\mathrm{face}}\!\!\sum_{i,j\in\{1,2\}}\!\!
  \frac{\alpha_{K_j}}{h_{K_j}}\,
  \lVert\mathbf v\rVert_{K_i}\,\lVert p\rVert_{K_j}.
\end{equation}

\emph{What it bundles --- five ingredients, one constant.} (i) Hölder on the
face; (ii) the face-measure bound $\operatorname{meas}(\Gamma)\le C\,h_K^{d-1}$
(\lbl{H:mesh}); (iii) the $L^\infty$ inverse estimate
$\lVert\psi_h\rVert_{L^\infty(K)}\le \Cinv h_K^{-d/2}\lVert\psi_h\rVert_K$
(\lbl{eq:inverse}) applied on \emph{both} neighbours --- replaced, on the
first-argument side in \file{AbstractInterpolation.v}, by the $L^\infty$
interpolation estimate \lbl{eq:interpinftyE}; (iv) $\alpha\le\alpha_K$ on the
face; and (v) the comparability of neighbouring diameters (\lbl{H:mesh}). All
five collapse into the single constant $C_{\mathrm{face}}$.

\emph{Role.} The jump terms of Steps~6b and~6d --- and, like \lbl{H:jump}, only
when $\sigma>0$. \emph{Reasonable.} Every ingredient is standard and every one
of them appears in the manuscript's Step~6b/6d derivation. A caveat on the
correspondence, though: the manuscript never displays \eqref{eq:face-here}
itself. What it displays is \lbl{eq:jumpest} and the jump-weighted bounds
\lbl{eq:jumpsplit}, in which the factor $\sigma|[\![\tau_1]\!]|$ is already
attached. \eqref{eq:face-here} is that derivation with the jump factor stripped
off --- Coq reattaches it separately, through \coq{H_assemble_*} and
\coq{H_jump} --- so a reader checking this hypothesis against the paper must
perform that regrouping themselves.

\emph{Honest assessment --- and the tightening that is actually available.}
This is the coarsest-grained assumption in the development, and the second
largest after \coq{Horth} --- it moved up the list when \coq{HBS} became a
theorem. A reader auditing it is auditing the manuscript's
Step~6b/6d derivation, not merely checking a textbook citation. It could be
decomposed into its five ingredients --- and \emph{should} be, if the abstract
interface is ever extended with face measures and traces; the roadmap's targets
T2 (the $L^\infty$ inverse estimate and the face-measure bound) and~T5 are
exactly that work. Until then, $C_{\mathrm{face}}$ is a name for a bound whose
derivation lives on paper.

% ---------------------------------------------------------------------
\subsection{The other ten files: no new assumptions --- and four scope boundaries}\label{sec:scope}
% ---------------------------------------------------------------------

Everything so far concerns the four abstract theorems, which carry the whole of
the analytic base. Of the remaining ten files, \emph{eight} are Tier-1 in the
confidence ledger of \S\ref{sec:conclusion}: their hypotheses are pure
positivity --- $\nu,h,\aK>0$ and the like --- so their theorems are
unconditional in the mathematics, and there is nothing to audit in their
assumptions.

The two exceptions are the scalar-algebra modules that feed the abstract
theorems, and they carry no \emph{new} assumptions: \file{StabilityAlgebra.v}
assumes only positivity together with \lbl{H:data}'s $\sigma,\varepsilon,C_2\ge0$
(D1), and takes $\varepsilon\le\varepsilon_{\max}$ (D2) as a premise of
\coq{elemental_coercivity} and $c_1>\xi\Cbar^2$, $\xi>2$ (D3, sharp form) as
premises of \coq{stability_constants_positive_sharp};
\file{ContinuityAlgebra.v} adds $\aK\le1$ (\coq{alphaK_le_1} --- the one
hypothesis the abstract files shed, \S\ref{sec:stratumD}), the
$\varepsilon$-condition (\coq{eps_le}, which \emph{is} \lbl{eq:epscond}),
$C_2<1$, the jump bounds of \lbl{H:jump} (\coq{jump_low}, \coq{jump_high}) and
the porosity-resolution increment of \lbl{H:porosity} (\coq{increment}). Every
one of these has already been audited above as a Class-D item; they reappear
here only because the scalar lemmas are stated in their own sections.

There is, however, something to audit in their \emph{scope}. In four places what
is proved is narrower than what the manuscript claims, and a reader is entitled
to know exactly where.

\emph{(1) The plateau bump is verified through $C^1$, not $C^\infty$.}
\file{PlateauBump.v} has a single hypothesis, \coq{Hypothesis Ha0 : 0 < a0 < 1}.
It proves, by explicit $\varepsilon$--$\delta$, that $\alpha_{\mathrm{mid}}$
meets its neighbours continuously (\coq{am_limit_0}, \coq{am_limit_1}) and with
vanishing first derivative (\coq{am'_limit_0}, \coq{am'_limit_1}). The
manuscript asserts $C^\infty$, a claim that rests on \emph{every} derivative
vanishing at the joins. Only the first-order case is inside the kernel --- and
it is inside no other artefact either, the SymPy script stopping at the same
order. The gap is real, and benign: $C^1$ is what finite element interpolation
of the profile actually consumes.

\emph{(2) The asymptotic limits are partial, with the other number held fixed.}
In \file{Asymptotics.v} the dominant-convection section declares
\coq{Variable Da : R} with \coq{Hypothesis Da_pos : 0 < Da} and \emph{then}
sends $\Reh\to\infty$; the dominant-reaction section declares
\coq{Variable Re : R} with \coq{Re_nonneg} and sends $\Dah\to\infty$. The limits
themselves are honest --- genuine $\varepsilon$--$\delta$ statements
(\coq{tendsto_at_top} of \file{Limits.v}, no filters, no hand-waving) --- but
they are limits in \emph{one} dimensionless number with the other held fixed at
an arbitrary admissible value. This is exactly how the manuscript's \S6 reads
them, so the formalisation is faithful; it simply is not a statement about the
joint limit $(\Reh,\Dah)\to(\infty,\infty)$, and none is claimed.

\emph{(3) The Fourier design of $\tau$ is a design computation, faithfully
formalised --- not a proof of optimality.} \file{TauDesign.v} certifies the
symbol algebra of Appendix~C exactly: the viscous spectrum
$\tfrac43\alpha\nu|k|^2$ and $\alpha\nu|k|^2$, the convective spectral radius
$(\alpha/h)|w\!\cdot\!k|$, the coupling spectrum $\{0,\pm|k|\}$, and the
uniqueness of the positive pair solving $XY=c^2$, $X/Y=\lambda$. What it does
\emph{not} do --- and cannot, and the manuscript does not claim --- is prove
that the resulting $\tau_1,\tau_2$ are in any sense optimal for the PDE. The
Fourier argument is a \emph{frozen-coefficient, plane-wave} heuristic used to
\emph{design} the parameters; its output is then justified after the fact by the
stability and convergence theorems, which hold for these $\tau$'s however they
were arrived at. Coq certifies the computation, not the principle --- and that
is the right division of labour.

\emph{(4) For the manufactured solution, Coq checks the mass equation, not the
forcing.} \file{ManufacturedSolution.v} proves
$\nabla\!\cdot(\alpha\mathbf u)=0$ for the manufactured velocity, and proves it
for an \emph{arbitrary positive} porosity --- not even a differentiable one. The
only hypothesis on the profile is \coq{alpha_pos : forall x y, 0 < alphaf x y}.
The reason is a cancellation the manufactured field is built to exhibit: with
$\mathbf u=(U\alpha_0/\alpha)\,\mathbf S$ one has $\alpha\mathbf u=U\alpha_0
\mathbf S$ identically (\coq{alpha_u1}, \coq{alpha_u2}), so $\alpha$ disappears
from the mass equation before any derivative is taken. The theorem is stated in
the strongest honest form, with the two partial derivatives as premises:
\begin{lstlisting}
Theorem manufactured_mass_conservation_2d :
  forall x y d1 d2 : R,
    derivable_pt_lim (fun t => alphaf t y * u1 t y) x d1 ->
    derivable_pt_lim (fun t => alphaf x t * u2 x t) y d2 ->
    d1 + d2 = 0.
\end{lstlisting}
Read: \emph{whatever} the derivatives of $\alpha u_1$ and $\alpha u_2$ are, they
sum to zero --- the solenoidal structure is exact, and independent of the
profile. (The derivatives are not left dangling: the file computes both, in
\coq{d1_alpha_u1} and \coq{d2_alpha_u2}, as genuine \coq{derivable_pt_lim}
statements.) The momentum \emph{forcing} $\mathbf f$, by far the larger symbolic
computation, is verified by SymPy
(\file{manufactured_solution_verification.py}) and not by Coq. That is a
deliberate division: it is a differentiation exercise, where a proof assistant
adds cost and no assurance.

Beyond these four boundaries, the content of the ten files is unconditional: the
$\tau_1,\tau_2,\sigt$ identities and their asymptotics, the coefficient
expansions and the $\varepsilon$-smallness chain (machine-checked at the
concrete $C_2=10^{-2}$, $\varepsilon=10^{-4}\varepsilon_{\mathrm{ref}}$ of the
numerical experiments), the admissibility of the DBF coefficients on
$0<\alpha\le1$, and the literature porosity profile.

% ---------------------------------------------------------------------
\subsection{The complete hypothesis inventory}\label{sec:inventory}
% ---------------------------------------------------------------------

Everything above, in one place. Table~\ref{tab:inventory} lists \emph{every}
\coq{Hypothesis} of the four abstract theorems --- all $50$ of them, with none
omitted and nothing counted twice --- in the exact form the kernel sees it, its
class, its mathematical content, and the proof steps that consume it. It is the
audit checklist: a reader who satisfies themselves that each of these $50$ lines
is a true statement about the finite element objects has, by
\S\ref{sec:complete}, satisfied themselves of the entire a priori theory.
The nine weighted inverse estimates (items 33--41) are each listed in full but
are all instances of the one \coq{winv_est} schema (\S\ref{sec:stratumI}, I1),
and the ``double'' composites the proof needs are now derived generically by the
\emph{proved} lemma \coq{winv_compose}, retiring three byte-identical hand
derivations; the schema is notational, so those nine remain nine independent
trusted items and the count stays $50$.

\medskip
\noindent\emph{How to read the table.}

\emph{Class} is D (data, mesh and parameter conditions --- hypotheses of the
paper's theorems in any formal system: nothing to prove, something to check),
G (Green-type identities --- integration by parts), or I (inverse and
interpolation estimates). Classes~G and~I are the analytic base; class~D is not.

\emph{In} records which of the four theorems carries the hypothesis:
\textbf{S}~$=$~\coq{abstract_stability} ($17$ hypotheses),
\textbf{C}~$=$~\coq{abstract_continuity} ($36$),
\textbf{I}~$=$~\coq{abstract_continterp} ($39$),
\textbf{V}~$=$~\coq{abstract_convergence} ($44$).
(These are not hand counts: each is the number of anonymous \coq{Prop} binders
in the closed theorem's type as the kernel prints it, and each equals the number
of \coq{Hypothesis} declarations in the corresponding file --- so no declared
hypothesis is silently unused.)
The counts differ because the first slot changes character: \textbf{C} has a
discrete first argument and so carries the five first-slot inverse estimates,
whereas \textbf{I} and \textbf{V} have the interpolation error there and carry
the seven interpolation estimates instead (see~I2, \S\ref{sec:stratumI}).

\emph{Symbols.} $U=[\mathbf u;p]$ and $V=[\mathbf v;q]$ are the first and second
arguments of $B_S$; in \textbf{I} and \textbf{V} the first argument is the
interpolation error $E=[\mathbf e_{\mathbf u};e_p]=U-\widehat U_h$ and the second
is the discrete error $W=\widehat U_h-U_h$. $K$ ranges over the elements $\Th$
and $\Gamma$ over the interior faces, with $K_1=e_1(\Gamma)$, $K_2=e_2(\Gamma)$
its two neighbours; $[\![\tau_1]\!]_\Gamma=\tau_{1,K_1}-\tau_{1,K_2}$;
$\varphi_{1,K}=\aK\tau_{1,\mathrm{NS}}^{-1}$;
$E_{\mathrm{int},K}(\mathbf u)$, $E_{\mathrm{int},K}(p)$ are the elemental
interpolation-error sizes of \lbl{eq:Eint} (\coq{IU}, \coq{IP});
$\mathcal P$ is the deviatoric-symmetric viscous projector; and
$\mathcal L_m,\mathcal L_m^{*},\mathcal L_c,\mathcal L_c^{*}$ are the
residual/adjoint-residual operators of \eqref{eq:HBS-here}.

\emph{Steps} are those of the appendix's proof of \lbl{lem:continuity}
(Steps~0--9), with the Coq lemma that consumes the hypothesis given in
parentheses. These were not assigned by hand: they were extracted mechanically
from the sources by locating, for each hypothesis, every lemma whose proof
mentions it.

\begin{longtable}{@{}r@{\ }L{0.155\textwidth}L{0.455\textwidth}L{0.245\textwidth}@{}}
\caption{Every hypothesis of the four abstract theorems.}\label{tab:inventory}\\
\toprule
\textbf{\#} & \textbf{Coq name (class)} & \textbf{Statement} & \textbf{In; steps that use it} \\
\midrule \endfirsthead
\multicolumn{4}{@{}l}{\emph{Table~\ref{tab:inventory}, continued}}\\
\toprule
\textbf{\#} & \textbf{Coq name (class)} & \textbf{Statement} & \textbf{In; steps that use it} \\
\midrule \endhead
\bottomrule \endfoot

\multicolumn{4}{@{}l}{\emph{\textbf{Class D} --- data, mesh and parameter conditions. Nothing to prove; something to check.}}\\[2pt]

1 & \coq{nu_pos} (D) & $\nu>0$ & S,C,I,V; everywhere \\
2 & \coq{sigma_nonneg} (D) & $\sigma\ge0$ \quad(\lbl{H:data}: $\sigma$ is a \emph{constant}) & S,C,I,V; everywhere \\
3 & \coq{eps_nonneg} (D) & $\varepsilon\ge0$ & S,C,I,V; everywhere \\
4 & \coq{c1_pos} (D) & $c_1>0$ & S,C,I,V; everywhere \\
5 & \coq{c2_pos} (D) & $c_2>0$ & S,C,I,V; everywhere \\
6 & \coq{C2_nonneg} (D) & $C_2\ge0$ & S,C,I,V; (P5), Steps~3,\,7 \\
7 & \coq{C2_lt_1} (D) & $C_2<1$ & S,C,I,V; (P5) (\coq{L_P5}); the $1-C_2$ branch of $\Cstab$ \\
8 & \coq{Cinv_pos} (D) & $\Cinv>0$ \quad(constant of \lbl{eq:inverse}) & C,I,V; carried by the \coq{Hw_*} \\
9 & \coq{Cb_pos} (D) & $\Cbar>0$ \quad(constant of \lbl{lem:winv}) & S,C,I,V; carried by \coq{S3}, \coq{Hw_du/dv} \\
10 & \coq{cJ_pos} (D) & $c_J>0$ & C,I,V; \lbl{lem:jump} (\coq{F_base}) \\
11 & \coq{cJ_le_1} (D) & $c_J\le1$ & C,I,V; \lbl{lem:jump} (\coq{F_base}) \\
12 & \coq{one_le_cJ'} (D) & $1\le c_J'$ & C,I,V; \lbl{lem:jump} (\coq{F_base}) \\
13 & \coq{Cface_nonneg} (D) & $C_{\mathrm{face}}\ge0$ & C,I,V; Steps~6b,\,6d (\coq{face6b}, \coq{face6d}) \\
14 & \coq{Nf_nonneg} (D) & $N_f\ge0$ & C,I,V; facewise C--S (\coq{face_CS_pair}) \\
15 & \coq{CI_pos} (D) & $C_I>0$ \quad(constant of \lbl{eq:interp}; \emph{strict} --- see the note after the table) & I,V; Step~9 (\coq{absorb_elem}); \emph{and} the proofs of \coq{IU_nonneg}, \coq{IP_nonneg} \\
16 & \coq{hK_pos} (D) & $h_K>0$ for every $K\in\Th$ & S,C,I,V; every elemental estimate \\
17 & \coq{aK_pos} (D) & $\aK>0$ for every $K\in\Th$ & S,C,I,V; every elemental estimate \\
18 & \coq{am_nonneg} (D) & $|\mathbf a|_{\infty,K}\ge0$ for every $K\in\Th$ (it is a norm) & S,C,I,V; Steps~5,\,9 \\
19 & \coq{xi_large} (D) & $\xi>2$ \quad(Young parameter; \lbl{eq:conditions_on_num_param}) & S,V; $\Cstab>0$ (\coq{stability_constants_positive_sharp}) \\
20 & \coq{c1_large} (D) & $c_1>\xi\,\Cbar^{2}$ \quad(the \emph{sharp} form; the manuscript's $c_1>2\xi\Cbar^2$ implies it and additionally buys the margin of \coq{C_stab_margin}) & S,V; $\Cstab>0$, \emph{and nowhere else} \\
21 & \coq{H_eps} / \coq{Heps} (D) & $\displaystyle \varepsilon \;\le\; C_2\,c_1\,\frac{\aK^{2}\,\tau_{1,K}}{h_K^{2}}$ \ for every $K\in\Th$ \quad(\lbl{eq:epscond}; the two spellings are definitionally equal) & S,C,I,V; (P5), Steps~3,\,7,\,9 \\
22 & \coq{H_jump} (D) & $c_J\,\varphi_{1,K_1}\le\varphi_{1,K_2}\le c_J'\,\varphi_{1,K_1}$ \ for every interior face $\Gamma$ \quad(\lbl{H:jump}; needed only when $\sigma>0$) & C,I,V; \lbl{lem:jump}, Steps~6b,\,6d \\
23 & \coq{H_mult1} (D) & $\displaystyle \sum_{\Gamma}g\bigl(K_1(\Gamma)\bigr)\;\le\;N_f\sum_{K\in\Th}g(K)$ \ for every $g\ge0$ \quad(bounded face multiplicity, \emph{and} conformity: $K_i(\Gamma)\in\Th$; \lbl{H:mesh}) & C,I,V; Steps~6b,\,6d (\coq{bound_JmpP}, \coq{bound_III}) \\
24 & \coq{H_mult2} (D) & the same with $K_2(\Gamma)$ & C,I,V; Steps~6b,\,6d \\

\midrule
\multicolumn{4}{@{}l}{\emph{\textbf{Class G} --- Green-type identities. The divergence theorem on an element, plus face bookkeeping.}}\\[2pt]

25 & \coq{H_skew} (G) & $\displaystyle \sum_K\bigl(\mathbf v,\ \alpha\,\mathbf a\!\cdot\!\nabla\mathbf u\bigr)_K=-\sum_K\bigl(\mathbf u,\ \alpha\,\mathbf a\!\cdot\!\nabla\mathbf v\bigr)_K$ \quad(\lbl{eq:skew}; uses $\nabla\!\cdot(\alpha\mathbf a)=0$ and $\mathbf u\!\cdot\!\mathbf v\in W^{1,1}_0$) & C,I,V; Step~6a (\coq{step6a}) \\
26 & \coq{H_ibp_vp} (G) & $\displaystyle \sum_K\bigl(\mathbf v,\ \alpha\nabla p\bigr)_K=-\sum_K\bigl(p,\ \nabla\!\cdot(\alpha\mathbf v)\bigr)_K$ \quad(\lbl{eq:globalibp}) & C,I,V; Step~6a \\
27 & \coq{H_ibp_qu} (G) & $\displaystyle \sum_K\bigl(q,\ \nabla\!\cdot(\alpha\mathbf u)\bigr)_K=-\sum_K\bigl(\mathbf u,\ \alpha\nabla q\bigr)_K$ \quad(\lbl{eq:globalibp}) & C,I,V; Step~6a \\
28 & \coq{H_skew_diag} (G) & $\displaystyle \sum_K\bigl(\mathbf v,\ \alpha\,\mathbf a\!\cdot\!\nabla\mathbf v\bigr)_K=0$ \quad(\lbl{eq:skew} on the \emph{diagonal}) & S,V; \coq{diag_zero}, hence the \emph{proof} of \coq{HBS} --- which \textbf{V} then reuses by applying \coq{abstract_stability} at the $W$ atoms, so the former \coq{HBS_W} needs no separate item \\
29 & \coq{H_ibp_diag} (G) & $\displaystyle \sum_K\bigl(\mathbf v,\ \alpha\nabla q\bigr)_K=-\sum_K\bigl(q,\ \nabla\!\cdot(\alpha\mathbf v)\bigr)_K$ \quad(\lbl{eq:globalibp} on the \emph{diagonal}) & S,V; \coq{diag_zero}, hence the \emph{proof} of \coq{HBS}; likewise reused by \textbf{V} \\
30 & \coq{H_assemble_p} (G) & $\displaystyle \sum_{K}\tau_{1,K}\!\!\int_{\partial K}\!\!\!\alpha\,(\mathbf n_K\!\cdot\!\mathbf v)\,p \;=\;\sum_{\Gamma}[\![\tau_1]\!]_\Gamma\!\int_{\Gamma}\!\alpha\,(\mathbf n\!\cdot\!\mathbf v)\,p$ \quad(face pairing $+$ \emph{vanishing of the boundary faces}, $\mathbf u_h=\mathbf v_h=\mathbf 0$ on $\partial\Omega$) & C,I,V; Step~6b (\coq{step6b_identity}) \\
31 & \coq{H_assemble_c} (G) & the same with $\int_\Gamma\alpha(\mathbf n\!\cdot\!\mathbf a)(\mathbf u\!\cdot\!\mathbf v)$ & C,I,V; Step~6d (\coq{bound_III}) \\
32 & \coq{Horth} (G) & $B_S(W,W)=-\,B_S(E,W)$ \quad(Galerkin orthogonality \lbl{eq:consistency} $+$ bilinearity of $B_S$ in its first argument) & V; the convergence chain (\coq{NW_bound}). \textbf{The largest remaining class-G assumption} (\S\ref{sec:stratumG}) \\

\midrule
\multicolumn{4}{@{}l}{\emph{\textbf{Class I} --- inverse and interpolation estimates. True for polynomials (I) and for interpolation errors (I), never for both.}}\\[2pt]

33 & \coq{S3} (I) & $\bigl\lVert 2\nu\nabla\!\cdot(\alpha\mathcal P\nabla\mathbf u_h)\bigr\rVert_K\le \dfrac{2\nu\Cbar}{h_K}\,\aK^{1/2}\bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\bigr\rVert_K$ \quad(\lbl{eq:winv-divvisc}) & S only; the Young step (\coq{du_sq_bound}). In \textbf{V} this slot is filled \emph{by the kernel} from \coq{Hw_dv} \\
34 & \coq{Hw_gu} (I) & $\bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\bigr\rVert_K\le\dfrac{\Cinv}{h_K}\aK^{1/2}\lVert\mathbf u_h\rVert_K$ \quad(\lbl{eq:winv-grad}) & C; Steps~4,\,9 \\
35 & \coq{Hw_gv} (I) & the same for $\mathbf v_h$ & C,I,V; Step~4 (\coq{double_inv_v}) \\
36 & \coq{Hw_du} (I) & $\bigl\lVert 2\nu\nabla\!\cdot(\alpha\mathcal P\nabla\mathbf u_h)\bigr\rVert_K\le\dfrac{2\nu\Cbar}{h_K}\aK^{1/2}\bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf u_h\bigr\rVert_K$ \quad(\lbl{eq:winv-divvisc}) & C; Step~4 (\coq{bound_T6}, \coq{bound_T7}) \\
37 & \coq{Hw_dv} (I) & the same for $\mathbf v_h$ & C,I,V; Step~4; \emph{and} it discharges \coq{S3} in \textbf{V} \\
38 & \coq{Hw_cxu} (I) & $\lVert\alpha\,\mathbf a\!\cdot\!\nabla\mathbf u_h\rVert_K\le\dfrac{\Cinv}{h_K}\,\aK\,|\mathbf a|_{\infty,K}\lVert\mathbf u_h\rVert_K$ \quad(\lbl{eq:winv-conv}) & C; Step~9 (\coq{absorb_elem}) \\
39 & \coq{Hw_cxv} (I) & the same for $\mathbf v_h$ & C,I,V; Step~5 (\coq{bound_T13c}) \\
40 & \coq{Hw_gpu} (I) & $\lVert\alpha\nabla p_h\rVert_K\le\dfrac{\Cinv}{h_K}\,\aK\,\lVert p_h\rVert_K$ \quad(\lbl{eq:winv-gradp}) & C; Step~9 \\
41 & \coq{Hw_divu} (I) & $\lVert\nabla\!\cdot(\alpha\mathbf u_h)\rVert_K\le\dfrac{\Cbar}{h_K}\,\aK\,\lVert\mathbf u_h\rVert_K$ \quad(\lbl{eq:winv-divu}) & C; Step~9 \\
42 & \coq{HI_gu} (I) & $\bigl\lVert\alpha^{1/2}\mathcal P\nabla\mathbf e_{\mathbf u}\bigr\rVert_K\le C_I\,\aK^{1/2}h_K^{-1}E_{\mathrm{int},K}(\mathbf u)$ \quad(\lbl{eq:interpgrad}) & I,V; Step~9 \\
43 & \coq{HI_du} (I) & $\bigl\lVert 2\nu\nabla\!\cdot(\alpha\mathcal P\nabla\mathbf e_{\mathbf u})\bigr\rVert_K\le 2\nu\,C_I\,\aK\,h_K^{-2}E_{\mathrm{int},K}(\mathbf u)$ \quad(\lbl{eq:interpdivvisc}; replaces the \emph{composition} \coq{Hw_du}$\circ$\coq{Hw_gu}) & I,V; Step~4 (\coq{bound_T6}--\coq{bound_T8}) \\
44 & \coq{HI_cxu} (I) & $\lVert\alpha\,\mathbf a\!\cdot\!\nabla\mathbf e_{\mathbf u}\rVert_K\le C_I\,\aK\,|\mathbf a|_{\infty,K}h_K^{-1}E_{\mathrm{int},K}(\mathbf u)$ & I,V; Step~9 \\
45 & \coq{HI_gpu} (I) & $\lVert\alpha\nabla e_p\rVert_K\le C_I\,\aK\,h_K^{-1}E_{\mathrm{int},K}(p)$ & I,V; Step~9 \\
46 & \coq{HI_divu} (I) & $\lVert\nabla\!\cdot(\alpha\mathbf e_{\mathbf u})\rVert_K\le C_I\,\aK\,h_K^{-1}E_{\mathrm{int},K}(\mathbf u)$ & I,V; Step~9 \\
47 & \coq{HI_uu} (I) & $\lVert\mathbf e_{\mathbf u}\rVert_K\le C_I\,E_{\mathrm{int},K}(\mathbf u)$ \quad(\lbl{eq:interpzero}; \emph{no} \coq{Hw_} counterpart --- a discrete field needs no zeroth-order bound) & I,V; Steps~6c,\,9 \\
48 & \coq{HI_pp} (I) & $\lVert e_p\rVert_K\le C_I\,E_{\mathrm{int},K}(p)$ \quad(\lbl{eq:interpzero}; likewise) & I,V; Steps~6b,\,9 \\
49 & \coq{H_face_p} (I,G) & $\displaystyle\Bigl|\int_\Gamma\alpha\,(\mathbf n\!\cdot\!\mathbf v)\,p\Bigr|\le C_{\mathrm{face}}\!\!\sum_{i,j\in\{1,2\}}\!\!\frac{\alpha_{K_j}}{h_{K_j}}\lVert\mathbf v\rVert_{K_i}\,\pi_j$, \ with $\pi_j=\lVert p_h\rVert_{K_j}$ in \textbf{C} and $\pi_j=E_{\mathrm{int},K_j}(p)$ in \textbf{I},\textbf{V}. \emph{Bundles five facts}: H\"older on $\Gamma$, $\mathrm{meas}(\Gamma)\le Ch_K^{d-1}$, the $L^\infty$ inverse estimate \lbl{eq:inverse} (resp.\ the $L^\infty$ interpolation estimate \lbl{eq:interpinftyE}), $\alpha\le\aK$, and comparability of neighbouring diameters & C,I,V; Step~6b (\coq{face6b}). \textbf{The coarsest bundle} (\S\ref{sec:stratumI}) \\
50 & \coq{H_face_c} (I,G) & $\displaystyle\Bigl|\int_\Gamma\alpha\,(\mathbf n\!\cdot\!\mathbf a)(\mathbf u\!\cdot\!\mathbf v)\Bigr|\le C_{\mathrm{face}}\!\!\sum_{i,j\in\{1,2\}}\!\!\frac{\alpha_{K_j}|\mathbf a|_{\infty,K_j}}{h_{K_j}}\,\mu_i\,\lVert\mathbf v\rVert_{K_j}$, \ with $\mu_i=\lVert\mathbf u_h\rVert_{K_i}$ in \textbf{C} and $\mu_i=E_{\mathrm{int},K_i}(\mathbf u)$ in \textbf{I},\textbf{V}; same five ingredients as item~49 & C,I,V; Step~6d (\coq{face6d}) \\
\end{longtable}

\noindent
Four things are worth reading off the table.

\emph{Class~D is not analytic trust.} Items 1--24 are conditions on the data,
the mesh and the numerical parameters. They are hypotheses of the manuscript's
theorems too; a formalisation cannot discharge them, and none of them asks the
reader to believe a piece of mathematics. The analytic trust is items 25--50,
and only those.

\emph{The base is far narrower than the count suggests.} Of the $26$ analytic
items, \emph{sixteen} (items 33--48) are instances of just two textbook facts ---
the inverse estimate and the Bramble--Hilbert interpolation estimate --- applied
to the particular fields the proof touches. The inverse-estimate half of these,
the nine rows 33--41, are now written through the \emph{single} \coq{winv_est}
schema of \file{InverseEstimates.v} (\S\ref{sec:stratumI}, I1), with the
``double'' composites derived generically by \coq{winv_compose}. That
consolidation is \emph{notational}, not a reduction of trust: each row still
instantiates \coq{winv_est} at its own atoms and unfolds to the same proposition,
so the nine remain nine independent hypotheses --- they cannot be fused into one
\coq{forall}-$x$ estimate precisely because of the polynomial/interpolation-error
firewall (an inverse estimate is true for finite-element functions and false for
the interpolation error, which shares the carrier). Discharging the two
primitives in Lean would remove sixteen rows at a stroke. A further \emph{seven}
(items 25--31) are the divergence theorem, its two diagonal instances, and finite
face bookkeeping. What is left is genuinely two things: \coq{Horth}, and the face
bundle \coq{H_face_p}/\coq{H_face_c}.

\emph{One item is strict where the manuscript is not, and it buys two rows.}
Item~15 is \coq{CI_pos} ($C_I>0$), not the $C_I\ge0$ a reader might expect. The
strictness is deliberate and it is a \emph{trade}. Earlier revisions assumed
$C_I\ge0$ together with $E_{\mathrm{int},K}(\mathbf u)\ge0$ and
$E_{\mathrm{int},K}(p)\ge0$ as three separate class-D hypotheses. But the
pre-Hilbert axioms already give $\lVert x\rVert\ge0$ for every $x$
(\coq{nrm_nonneg}), so \coq{HI_uu} reads
$0\le\lVert\mathbf e_{\mathbf u}\rVert_K\le C_I\,E_{\mathrm{int},K}(\mathbf u)$;
with $C_I>0$ the product being nonnegative \emph{forces}
$E_{\mathrm{int},K}(\mathbf u)\ge0$, and \coq{HI_pp} does the same for the
pressure. The two nonnegativity rows are therefore now \emph{theorems}
(\coq{IU_nonneg}, \coq{IP_nonneg}), at the cost of one $\ge$ becoming a $>$.
This is worth stating plainly rather than counting as a free win: $\{C_I>0\}$ is
logically \emph{strictly stronger} than $\{C_I\ge0,\ E_{\mathrm{int}}(\mathbf
u)\ge0,\ E_{\mathrm{int}}(p)\ge0\}$ --- the degenerate reading $C_I=0$ is no
longer an instance. Nothing is lost in practice, since an interpolation constant
is positive in every estimate that has ever been proved, and a $C_I=0$ would
force the interpolation error itself to vanish identically; but the reader is
entitled to see the trade rather than a row count.

\emph{Every item is used.} There are no dead hypotheses: each of the $50$ is
consumed by at least one lemma, as the last column records --- and this is
checked mechanically, not asserted: for each of the four files the number of
\coq{Hypothesis} declarations equals the number of \coq{Prop} arguments the
kernel gives the closed theorem. And the \emph{strict} content of \coq{c1_large}
and \coq{xi_large} --- the inequalities themselves, as opposed to the weaker
$\xi>0$ they imply --- is consumed at exactly \emph{one} place between them
(\S\ref{sec:stratumD}, D3).

% ---------------------------------------------------------------------
\subsection{Are the hypotheses consistent? (Non-vacuity)}\label{sec:vacuity}
% ---------------------------------------------------------------------

A theorem with contradictory hypotheses is true and worthless. Any development
that assumes a dozen quantitative facts owes the reader an answer to the
question: could this bundle be unsatisfiable, making the conclusions vacuous?
There are two answers, and they are different.

\emph{Soundness --- the only one that matters for the paper.} Every hypothesis
of Stratum~3 is a true statement about the finite element objects, either by
textbook citation (Classes~G and~I) or by assumption of the manuscript
(Class~D). The concrete finite element tuple --- $\mathtt{gu}\,K =
(\alpha^{1/2}\mathcal P\nabla\mathbf u_h)|_K$, and so on --- is therefore a
model of the whole bundle, and the abstract theorem, being universally
quantified, applies to it. This is what the crosswalk and the hand audit
establish, and it is the entire content of the claim that the paper's lemmas are
machine-checked.

\emph{Consistency --- the pedant's question, answered.} Independently of the
intended model, the hypothesis bundle is \emph{satisfiable}, so no theorem in
the development is vacuously true. Here is an explicit witness for
\coq{abstract_stability}, small enough to check by hand in a minute.

Take the carrier $V=\mathbb R$ with $\langle x,y\rangle:=xy$ --- a pre-Hilbert
space, since \eqref{eq:prehilbert} is immediate. Let the mesh be a single
element, $\Th=[k_0]$. Fix
\[
  \nu=\sigma=1,\qquad h_{k_0}=\aK=1,\qquad |\mathbf a|_{\infty,k_0}=0,
  \qquad \xi=3,\quad \Cbar=1,\quad c_1=7,\quad c_2=1,\quad C_2=\tfrac12 ,
\]
which satisfies every positivity hypothesis, together with $\xi>2$ and the
sharp numerical-parameter condition $c_1=7>3=\xi\Cbar^{2}$ (\coq{w_c1_large};
the manuscript's stronger $c_1>2\xi\Cbar^2=6$ holds here too). The parameter formulas then give, exactly,
\[
  \tau_{1,\mathrm{NS}}^{-1}=7,\qquad
  \tau_1=\tfrac18,\qquad
  \tau_2=1,\qquad
  \sigt=\tfrac78,\qquad
  \varepsilon_{\max}=C_2c_1\aK^2\tau_1/h^2=\tfrac{7}{16}\;>\;0,
\]
so $\varepsilon:=\tfrac{7}{16}$ is admissible and \coq{Heps} holds. For the
atoms take
\[
  \mathtt{uu}(k_0)=1,\qquad
  \mathtt{gu}=\mathtt{pp}=\mathtt{cxu}=\mathtt{gpu}=\mathtt{divu}=\mathtt{du}=0 .
\]
Then \coq{S3} reads $0\le0$ and holds, and the two diagonal Green identities are
immediate: every atom but $\mathtt{uu}$ vanishes, so \coq{H_skew_diag} reads
$0=0$ and \coq{H_ibp_diag} reads $0=-0$. Note what is \emph{no longer} required
of this witness. $B_S(U_h,U_h)$ used to be a free variable, and the witness
satisfied \coq{HBS} by \emph{defining} $B_S$ to be its right-hand side --- i.e.\
$7/8$ was a number the witness \emph{chose}. It is now no free real: it is the
eighteen-term form at these atoms, so $7/8$ is a number the witness
\emph{computes} (\coq{w_BS_val} merely evaluates it). Computing it:
$G(\mathbf u)=\mathtt{du}-\sigma\,\mathtt{uu}=-1$, so
$\mathcal L_m=\mathcal L_m^{*}=1$ and $\mathcal L_c=\mathcal L_c^{*}=0$, whence
\[
  B_S \;=\; \sigma\,\mathtt{uu}^2-\tau_1\langle\mathcal L_m^{*},\mathcal L_m\rangle
     \;=\; 1-\tfrac18 \;=\; \tfrac78 .
\]
The witness therefore has strictly less freedom than it had, and the headline
$7/8\ge7/16$ is the stronger for it.
On the other side, $C_u=1-\xi\Cbar^2/c_1=\tfrac47$,
$C_{\mathrm{visc}}=\min\{2-\tfrac47,\,2(1-\tfrac23)\}=\tfrac23$, so
$\Cstab=\min\{\tfrac23,\tfrac47,\tfrac12,1\}=\tfrac12$, while
$|||U_h|||^2=\sigt\,\mathtt{uu}^2=\tfrac78$. The conclusion of
\coq{abstract_stability} therefore reads
\[
  \tfrac78 \;=\; B_S \;\ge\; \Cstab\,|||U_h|||^2 \;=\; \tfrac12\cdot\tfrac78
  \;=\; \tfrac{7}{16} ,
\]
which is true, and \emph{non-degenerate}: both sides are strictly positive, so
the theorem is not being satisfied by a $0\ge0$. The hypotheses are jointly
satisfiable with a non-trivial conclusion.

\emph{And it is machine-checked.} The witness above is not a hand argument. It
is the file \file{NonVacuity.v}: it builds the instance ($\mathbb R$ with
$\langle x,y\rangle=xy$ is shown to be a \coq{PreHilbert}; the mesh is
\coq{[tt]}; the data are the rationals above), discharges every hypothesis of
\coq{abstract_stability} --- including \coq{H_skew_diag}, \coq{H_ibp_diag},
\coq{S3} and \coq{Heps} --- and \emph{applies the theorem}:
\begin{lstlisting}
Theorem witness_stability :
  0 < wCstab /\ wBS >= wCstab * wNormSq.
Proof.
  exact (AbstractStability.abstract_stability
           RPH unit wTh wnu wsig weps wc1 wc2 wCb wxi wC2
           w_nu_pos w_sigma_nonneg w_eps_nonneg w_c1_pos w_c2_pos w_Cb_pos
           w_C2_nonneg w_C2_lt_1 w_c1_large w_xi_large
           whK waK wam w_hK_pos w_aK_pos w_am_nonneg
           wgu wuu wpp wcxu wgpu wdivu wdu
           w_skew_diag w_ibp_diag w_S3 w_Heps).
Qed.

Theorem abstract_stability_is_not_vacuous :
  wBS >= wCstab * wNormSq
  /\ 0 < wCstab /\ 0 < wBS /\ 0 < wNormSq
  /\ wCstab = 1/2 /\ wNormSq = 7/8 /\ wBS = 7/8
  /\ wCstab * wNormSq = 7/16.
\end{lstlisting}
The kernel accepts both, and \coq{Print Assumptions
abstract_stability_is_not_vacuous} returns the same three standard-library
axioms as every other theorem here --- nothing more. So the claim ``the hypotheses of
\lbl{lemma:Stability} are consistent, and the coercivity bound it yields is a
non-trivial inequality between positive numbers'' is now itself a
machine-checked theorem, not an assurance.

\emph{A strictly positive advection field.} That first instance takes
$|\mathbf a|_{\infty}=0$ --- the Darcy corner, at which every hypothesis whose
right-hand side carries a factor $|\mathbf a|_\infty$ is discharged at its
boundary, so it does not exhibit that a moving fluid satisfies the bundle.
\file{NonVacuity.v} therefore carries a \emph{second} stability witness,
\coq{abstract_stability_is_not_vacuous_with_advection}, on a two-element mesh
with $|\mathbf a|_{\infty}=1$ and atoms
$\mathtt{gu}=\mathtt{uu}=\mathtt{pp}=\mathtt{divu}=1$,
$\mathtt{gpu}=\mathtt{du}=-1$, $\mathtt{cxu}=(2,-2)$. The parameters shift
($\tau_1=\tfrac19$, $\tau_2=\tfrac87$, $\sigt=\tfrac89$,
$\varepsilon_{\max}=\tfrac7{18}$) and the conclusion is again non-degenerate:
$B_S=\tfrac{5069}{567}\ge\Cstab|||U_h|||^2=\tfrac{167}{42}$ (that is,
$10138\ge4509$ over the common denominator $1134$). With $|\mathbf a|_\infty>0$
strictly, \coq{am_nonneg} is exercised off its boundary; $c_2$ becomes live
(at $|\mathbf a|_\infty=0$ its value does not enter $\tau_1,\tau_2,\sigt$ at
all, which the file records as \coq{w1_c2_is_dead_at_am0}); and the internal
estimate \coq{velocity_slack_identity}, saturated to an \emph{equality} at
$|\mathbf a|_\infty=0$, now holds with strict slack. \coq{H_skew_diag} is met
not as $0=0$ but as a genuine cancellation $1\cdot2+1\cdot(-2)=0$ between two
nonzero contributions --- which on the one-dimensional carrier $\mathbb R$ is
why a \emph{second} element is needed (\coq{w1_cxu_forced_zero_on_singleton}).

\emph{The interpolation and convergence bundles, in the kernel too.}
\file{NonVacuityInterp.v} discharges all $39$ hypotheses of
\coq{abstract_continterp} and \file{NonVacuityConv.v} all $44$ of
\coq{abstract_convergence}, each on a genuine two-element mesh with one interior
face. This is where the one wrinkle worth recording is met head-on:
\coq{H_jump}, \coq{H_face_p} and \coq{H_face_c} are quantified over the face
\emph{type} \coq{F}, not over the face \emph{list} \coq{Fl}, so a witness cannot
discharge them by emptying the list. These do the opposite --- \coq{F}
inhabited, \coq{Fl} non-empty, $e_1\ne e_2$ --- and let the assembly identities
\emph{pin} the face data to a nonzero value (a difference of $\tfrac1{90}$
between the two elements forces $\mathrm{FB}_p=\mathrm{FB}_c=2$). Both
conclusions are non-degenerate with a concrete rational lower bound on the
right-hand side --- $|B_S|=\tfrac{72181}{8400}\le131\le\mathrm{RHS}$ for
interpolation, and $\mathrm{NErr}\le C_{\mathrm{conv}}\,\Psi$ with
$\mathrm{NErr},\Psi>0$ for convergence (its Galerkin identity holding as
$\mathrm{BSWW}=\tfrac{3931}{32}=-\mathrm{BSEW}$, between two nonzero
eighteen-term forms) --- not the vacuous $\mathrm{RHS}>0$ a reader could meet
with $10^{-30}$. Both take $|\mathbf a|_\infty>0$ ($(1,2)$ and $2$), and both
discharge \coq{CI_pos} ($C_I>0$) strictly ($C_I=1$, $C_I=4$) and jointly with
the rest of the bundle --- the first machine check of the one hypothesis this
revision \emph{strengthened} (\S\ref{sec:tighten}).

What these witnesses do \emph{not} show bears stating plainly, and none of it is
hidden. They settle \emph{consistency}, not \emph{soundness}: that the
\emph{finite element} objects satisfy the hypotheses is the other direction, and
it is what the hand audit of Stratum~3 establishes. \coq{abstract_continuity} is
the one theorem with no witness of its own --- its consistency still rests on the
hand argument, not on an instance. And each machine witness has non-sharp
corners, disclosed in its file banner: the advection witness uses a
\emph{uniform} mesh ($h_K=\aK=|\mathbf a|_\infty$ constant across its two
elements, so only the atoms vary) and meets \coq{Heps} with \emph{equality}
($\varepsilon=\varepsilon_{\max}=\tfrac7{18}$), the extremal admissible choice
for an upper bound rather than a strict one; and the interpolation and
convergence witnesses discharge \coq{H_skew}/\coq{H_ibp_vp} and
\coq{H_skew_diag} as $0=\pm0$. That last is not idleness --- on a skew or
antisymmetric diagonal both sides \emph{must} vanish, the witnesses meet them
with nonzero summands that cancel, and it is exactly that cancellation which
pins the face data the rest of the bundle then controls --- but it is the honest
standard for those three rows, not an identity between two nonzero quantities.
Finally, both stability witnesses take $\sigma=1$, so the reaction-free corner
$\sigma=0$ --- itself a genuine \emph{instance} of the theorems
(\S\ref{sec:tighten}) --- is exercised by neither.

% ---------------------------------------------------------------------
\subsection{Summary: what was tightened, and what remains}\label{sec:tighten}

The assessments above, collected. The list is cumulative, so most of its rows
record closures made at earlier revisions: of the fifteen items, ten are closed,
but only \emph{two} of those closed \emph{here} --- \coq{HBS} (G1) and
\coq{IU_nonneg}/\coq{IP_nonneg} (D1), which are precisely the two rows this
revision adds to the ledger. The other eight were already closed at the previous
revision, and are kept so that the ledger reads as a standing account rather than
a changelog. What remains open is exactly the part that cannot be closed without
a concrete variational instance --- the Lean~4 targets of \S\ref{sec:left}. Both
of this revision's closures are \emph{trades} rather than pure reductions, and
are marked as such: \coq{IU_nonneg}/\coq{IP_nonneg} cost the strengthening
$C_I>0$, and \coq{HBS} relocates three of its four bundled pieces into the (I4)
reading obligation rather than discharging them.

\begin{longtable}{@{}L{0.26\textwidth}cL{0.60\textwidth}@{}}
\toprule
\textbf{Item} & & \textbf{Status} \\
\midrule \endhead

$c_1>2\xi\Cbar^{2}$, $\xi>2$ (D3)
  & \textbf{closed}
  & \emph{Sufficient, not necessary} --- and both halves are now theorems.
    Positivity of $C_u,C_{\mathrm{visc}}$ needs only $c_1>\xi\Cbar^2$ (given
    $\xi>2$), and that is what the development now assumes
    (\coq{stability_constants_positive_sharp}); the manuscript's condition is
    kept as a corollary. The extra factor of two is not idle, and
    \coq{C_stab_margin} says exactly what it buys: $C_u>\tfrac12$,
    $2-4\Cbar^2/c_1>1$, hence the $\Cbar$-free floor
    $\Cstab\ge\min\{2(1-2/\xi),\tfrac12,1-C_2\}$. \\

$\sigma>0$, $\varepsilon>0$, $C_2>0$ (D1)
  & \textbf{closed}
  & Was \emph{stronger than the paper}. Now relaxed to \lbl{H:data}'s own
    $\sigma\ge0$, $\varepsilon\ge0$, $0\le C_2<1$: two lemmas of
    \file{StabilityAlgebra.v} were restated with $\le$ in place of $<$
    (\coq{sigt_nonneg}, \coq{alphaK_tau1_le_tau1NS}), a third
    (\coq{sigt_denom_pos}) kept its statement and was reproved from
    $\sigma/\aK\ge0$, and the whole chain went through unchanged. \textbf{The reaction-free case
    $\sigma=0$ is now a genuine instance} of the stability and convergence
    theorems. \\

\coq{BS}, \coq{BSWW} never reconciled with the
eighteen-term $B_S$ (I4)
  & \textbf{closed}
  & Both diagonal $B_S$'s are now \emph{defined} as the eighteen-term
    expression --- the \emph{same} \coq{AbstractInterpolation.BS}, applied to each
    file's own atoms --- and each tested identity is \emph{proved} from it. The
    two encodings are provably the same number in both files --- the check the
    parameters already had (\coq{tau1_agree}\,\dots) and $B_S$ lacked. \\

\coq{HBS_W} (G1)
  & \textbf{closed}
  & \emph{Gone entirely --- not even replaced by a theorem of its own.} Once
    \coq{HBS} became a theorem over \coq{H_skew_diag}/\coq{H_ibp_diag}
    (\eqref{eq:diag-here}), the convergence file obtains the $(W,W)$ tested
    identity by \emph{applying} \coq{abstract_stability} at the discrete-error
    atoms (\coq{BSWW_is_ASBS} by \coq{reflexivity}, then \coq{stab_W}). The
    difference-of-squares algebra it used to smuggle in is proved once and
    reused. A monolithic class-G bundle traded for nothing at all. \\

\coq{HBS} (G1) --- \emph{was ``the single largest assumption''}
  & \textbf{closed}
  & \emph{No longer a hypothesis}, by exactly the route already taken for
    \coq{HBS_W}, and at no extra cost in assumptions: the \emph{same} two
    diagonal identities (items 28--29) discharge both. \file{AbstractStability.v}
    now imports \file{AbstractInterpolation.v}, defines $B_S(U_h,U_h)$ to be its
    eighteen-term expression at the diagonal atoms, and proves the tested
    identity (\coq{diag_zero} $\to$ \coq{HBS}). Two disclosures, since this is a
    trade and not a miracle: the atom $\mathtt{xu}=\alpha\mathcal X(U_h)$ had to
    be split into $\mathtt{cxu},\mathtt{gpu}$ (with $\mathtt{xu}$ now
    \emph{defined} as their sum) --- one cannot state ``the convective part is
    skew'' without naming it; and of the four things \coq{HBS} bundled, only the
    annihilation of $T_2+T_4$ was ever provable. The other three --- the Galerkin
    form, the substitution $V_h:=U_h$, and $\mathcal P^{\!\top}\mathcal
    P=\mathcal P$ --- did not vanish: they \emph{moved} into the eighteen-term
    reading obligation of (I4), which the other three files already carried
    (\S\ref{sec:stratumG}, G1). \\

\coq{IU_nonneg}, \coq{IP_nonneg} (D1)
  & \textbf{closed}
  & \emph{No longer hypotheses.} $E_{\mathrm{int},K}(\mathbf u)\ge0$ and
    $E_{\mathrm{int},K}(p)\ge0$ follow from \coq{HI_uu}/\coq{HI_pp} and
    $\lVert\cdot\rVert\ge0$ (\coq{nrm_nonneg}). The price is $C_I\ge0\rightsquigarrow
    C_I>0$ (item~15): a \emph{strictly stronger} assumption, traded for two rows.
    Flagged rather than banked (\S\ref{sec:stratumD}, D1). \\

\coq{H_elem_conv_ibp}, \coq{H_elem_p_ibp}
alone (G5)
  & \textbf{closed}
  & Were \emph{vacuous in isolation} (their face terms were free variables). The
    face terms are now \coq{Definition}s --- the integration-by-parts defect,
    which is what the divergence theorem says they are --- and the two
    identities are proved. Two hypotheses leave the trusted base of each of the
    three files; no content is lost, because the content lives in
    \coq{H_assemble_*} and \coq{H_face_*}. \\

$\tau_2$: analysed vs.\ implemented
  & \textbf{closed}
  & The bridge is now \emph{wired end to end}. \coq{abstract_convergence_implemented}
    states \lbl{thm:convergence} for the parameter the solver actually forms:
    measured in the triple norm built from the implemented $\tau_2$
    (\lbl{eq:Tau2}) and with $\psi(h)$ built from it too, the estimate holds with
    the same constant inflated by at most $\sqrt{1+C_2}<\sqrt2$. \\

$\xi>0$ (\coq{xi_pos}) alongside $\xi>2$
  & \textbf{closed}
  & Redundant; now derived rather than assumed. \\

\midrule

\coq{Horth} (G2)
  & open
  & \emph{The largest residual trust}, now that both \coq{HBS} and \coq{HBS_W}
    are gone. An identity, verbatim from the manuscript, symbolically
    cross-checked --- but assumed; and it is now the only place where
    bilinearity of $B_S$ in its first argument is assumed. Provable only once a
    concrete variational instance is formalised (roadmap T4/T7). \\

\coq{H_face_p}, \coq{H_face_c} (I3)
  & open
  & \emph{The coarsest bundle}: five standard ingredients behind one constant.
    Decomposable only once face measures and traces enter the interface
    (roadmap T2/T5). \\

\coq{IU}, \coq{IP} abstract (I2)
  & open
  & The kernel proves an \emph{estimate}, not a \emph{rate}. The $O(h^{k+1})$
    reading is a Stratum-1 instantiation, confirmed empirically by the MMS
    plateaus. Closing it means formalising $E_{\mathrm{int},K}$, i.e.\ Sobolev
    spaces --- squarely a Lean task. \\

\midrule

\lbl{H:jump} assumed unconditionally (D4)
  & ---
  & Not a defect. The manuscript needs it only for $\sigma>0$; supplying a true
    hypothesis costs nothing, and by \lbl{H:porosity}+\lbl{H:mesh} it holds
    anyway. A $\sigma$-case split would recover the $\sigma=0$ specialisation
    for free, but buys nothing. \\

$\aK\le1$ \emph{not} assumed
  & ---
  & Not a gap but a bonus: the abstract theorems prove the sharp
    porosity-weighted forms (\lbl{eq:sharpcont}, \lbl{rem:sharperconv}), which
    do not need it. \\

Non-vacuity (\S\ref{sec:vacuity})
  & \textbf{closed}
  & Three witness files (\file{NonVacuity.v}, \file{NonVacuityInterp.v},
    \file{NonVacuityConv.v}) apply \coq{abstract_stability} (twice, the second
    with $|\mathbf a|_\infty>0$), \coq{abstract_continterp} and
    \coq{abstract_convergence} to explicit instances inside the kernel --- each
    bundle consistent, each conclusion non-degenerate, and \coq{CI_pos} now
    witnessed. \coq{abstract_continuity} remains the one theorem without an
    instance of its own. \\
\bottomrule
\end{longtable}

\noindent
The net effect on the analytic base: \coq{HBS} and the two interpolation-error
sign hypotheses are gone --- three items --- and \emph{nothing} is added. The two
elementary diagonal identities that discharge \coq{HBS} were already in the
inventory: \file{AbstractConvergence.v} declared them for its own $B_S(W,W)$
reasoning one revision earlier, and \file{AbstractStability.v} now shares them
rather than adding them (items 28--29). \coq{HBS_W} and the two elemental
integration-by-parts hypotheses left the base at that earlier revision, not this
one. Per theorem: the trusted base of \lbl{thm:convergence} went from $46$ named
hypotheses to $44$ and that of \coq{abstract_continterp} from $41$ to $39$ (both
shedding the two sign rows); \coq{abstract_continuity}'s $36$ is unchanged, its
closed type byte-identical to the previous revision's; and that of
\lbl{lemma:Stability} went from $16$ to $17$ --- the one \emph{increase} in the
development, and the honest price of making the stability lemma speak about the
eighteen-term $B_S$ itself instead of about a free real. Overall the inventory
went from $53$ to $50$.

The count is the weakest part of the claim, and \S\ref{sec:stratumG}~G1 says why:
one of the three departing rows was logically eliminable to begin with. What the
revision really buys is that the difference-of-squares expansion \coq{HBS} used to
smuggle in is now \emph{proved}, and that the number \file{AbstractStability.v}
bounds from below is the same closed expression the other three files bound from
above --- checked by the kernel rather than by a reader comparing two displays.
Against that stand the two disclosures recorded in \S\ref{sec:tighten}:
\coq{CI_pos} is strictly stronger than the \coq{CI_nonneg} it replaces, and three
of \coq{HBS}'s four bundled pieces moved into the (I4) reading obligation rather
than being discharged.

% =====================================================================
\section{The residual trusted base and what is left}\label{sec:left}
% =====================================================================

With the abstract layer in place, the ``what is left'' story is now a short,
explicit list. Every item below is either (i) a named hypothesis of the
abstract theorems --- standard analysis, hand-audited and slated for Lean ---
or (ii) genuinely outside the current scope, and covered by the SymPy suite.

\subsection{The named trusted base of the abstract theorems}

The table below is the \emph{index} of that base. Each entry is stated in full
--- in both its Coq and its mathematical form, with the proof steps that consume
it --- in Table~\ref{tab:inventory}, and assessed for role, plausibility and
tightness in Section~\ref{sec:assumptions}, which also catalogues the
\emph{implicit} assumptions that no table of hypotheses can show. The companion \file{LEAN_ROADMAP.md} classifies each hypothesis into three
classes. Class~D items are assumptions on data/mesh --- hypotheses of the
paper's theorems in any formal system, with nothing to prove. Classes~G and~I
are the textbook analytic facts (Brenner--Scott, Ern--Guermond), hand-audited
against the manuscript in \file{AUDIT.md}~\S4.

\begin{longtable}{@{}L{0.30\textwidth}L{0.34\textwidth}cL{0.16\textwidth}@{}}
\toprule
\textbf{Coq hypothesis} & \textbf{Paper} & \textbf{Class} & \textbf{Status} \\
\midrule \endhead
\coq{S3} & \lbl{eq:winv-divvisc} (\lbl{lem:winv}) & I & hand-audited \\
\coq{Heps}, \coq{H_eps} & \lbl{eq:UpperBoundOnEpsilon} / \lbl{eq:epscond} & D & assumption \\
\coq{H_skew} & \lbl{eq:skew}, identity (i) & G & hand-audited \\
\coq{H_ibp_vp}, \coq{H_ibp_qu} & \lbl{eq:globalibp}, identity (ii) & G & hand-audited \\
\coq{H_skew_diag}, \coq{H_ibp_diag} & \lbl{eq:skew}/\lbl{eq:globalibp} on the \emph{diagonal}; they replace both former tested-identity items: \coq{HBS} is \emph{proved} from them, and \coq{HBS_W} needs no item at all & G & hand-audited \\
\coq{H_assemble_p}, \coq{H_assemble_c} & facewise assembly (Steps~6b,~6d), including the vanishing of the boundary faces & G & hand-audited \\
\coq{H_face_p}, \coq{H_face_c} & \lbl{eq:jumpest}/face estimates: H\"older $+$ $\mathrm{meas}\le Ch^{d-1}$ $+$ $L^\infty$ inverse $+$ $\alpha\le\aK$ & I,G & hand-audited \\
\coq{H_jump} & \lbl{H:jump} & D & assumption \\
\coq{H_mult1}, \coq{H_mult2} & bounded face multiplicity (\lbl{H:mesh}) & D & assumption \\
\coq{Hw_gu} \dots \coq{Hw_divu} (8) & weighted inverse estimates of \lbl{lem:winv} (\lbl{eq:inverse}); \lbl{lem:continterp} keeps only the three V-side ones & I & hand-audited \\
\coq{HI_gu} \dots \coq{HI_pp} (7) & interpolation estimates \lbl{eq:interp}/\lbl{eq:interpinfty} via the replacement table \lbl{eq:interpdivvisc}--\lbl{eq:interpzero} (Bramble--Hilbert) & I & hand-audited \\
\coq{Horth} & Galerkin orthogonality \lbl{eq:consistency} $+$ bilinearity of $B_S$ in its first argument & G & hand-audited \\
\midrule
\multicolumn{4}{@{}l}{\emph{No longer in the base (were hypotheses at the previous revision; now theorems):}}\\
\coq{HBS} & now \emph{proved} from \coq{H_skew_diag}/\coq{H_ibp_diag}: the Galerkin/adjoint identity (\lbl{eq:StabilityEstimate} testing) at the diagonal (\S\ref{sec:stratumG}, G1) & --- & \textbf{discharged} \\
\coq{HBS_W} & \emph{eliminated}: \coq{abstract_stability}, applied at the $W$ atoms, carries the tested identity (\S\ref{sec:stratumG}, G1) & --- & \textbf{discharged} \\
\coq{H_elem_conv_ibp}, \coq{H_elem_p_ibp} & now \emph{proved}: their face terms are \coq{Definition}s, the integration-by-parts defect (\S\ref{sec:stratumG}, G5) & --- & \textbf{discharged} \\
\coq{IU_nonneg}, \coq{IP_nonneg} & now \emph{proved}: with $\lVert\cdot\rVert\ge0$ and $C_I>0$, \coq{HI_uu}/\coq{HI_pp} force them --- a \emph{trade}, since it strengthens item~15 to \coq{CI_pos} (\S\ref{sec:inventory}) & --- & \textbf{discharged} \\
\bottomrule
\end{longtable}

\noindent
Interface integrity is itself machine-checked: stability's
inverse-estimate input is literally the hypothesis \coq{Hw_dv}, its
$\varepsilon$-condition is definitionally \lbl{eq:epscond}
(\coq{Heps_W}), and the $\tau_1,\tau_2,\sigt$ formulas of the two
scalar-algebra modules are proved to agree
(\coq{tau1_agree}/\coq{tau2_agree}/\coq{sigt_agree}), so the composed
statements provably concern the same quantities.

\noindent
The roadmap observes that Classes~G and~I collapse onto two primitives:
\begin{itemize}[nosep]
  \item \textbf{Class~I} reduces to the \emph{unweighted} inverse estimate on a
    shape-regular simplex (an $L^2$ and an $L^\infty$ version) plus a
    weighted-to-unweighted reduction --- whose \emph{scalar} chain
    ($\alpha_\infty/\alpha_0$ comparabilities under \lbl{H:porosity}) is
    \emph{already machine-checked} in \file{ContinuityAlgebra.v}
    (\coq{delta_alpha_bound}, \coq{winv_ratio}); only the norm-level gluing
    remains. With \lbl{lem:continterp} formalised, Class~I now also contains
    the interpolation estimates \coq{HI_*} --- Bramble--Hilbert/Deny--Lions
    technology on a shape-regular simplex, the natural companion Lean
    target.
  \item \textbf{Class~G} reduces to the divergence theorem on a single simplex
    for (polynomial)$\times$($W^{1,\infty}$ weight) data, plus finite
    bookkeeping over elements and faces using continuity of the integrands.
\end{itemize}

\subsection{Hand-audited without a Coq statement}

None remain within the stability/continuity/convergence development: with
\file{AbstractInterpolation.v} and \file{AbstractConvergence.v}, the
interpolation replacements of \lbl{lem:continterp} and the
triangle-inequality assembly of \lbl{thm:convergence} --- previously the
one corner resting on the hand audit alone --- are inside the kernel (the
triangle inequality is \emph{proved}, not assumed). What remains outside
Coq is exactly the SymPy-delegated and out-of-scope material of the next
two subsections, plus the named hypothesis base above.

\subsection{The Lean~4 roadmap (a plan, not a verification)}

\file{LEAN_ROADMAP.md} lays out an assistant-driven plan to discharge
Classes~G and~I in Lean~4 + Mathlib, with a companion statement skeleton
\file{PorousNSToolbox.lean}. \textbf{That skeleton is explicitly a statement
roadmap, not a proof artifact}: every target ends in \coq{sorry} and it is not
expected to elaborate against current Mathlib without API repair. Nothing in
the Lean file is machine-checked; it pins the intended semantics of each
residual target (unweighted inverse estimate on a simplex, its $L^\infty$
sibling and the face-measure bound, the weighted reductions, and the
simplex divergence theorem) for a future formalisation loop. The dependency-minimal
path noted there is the inverse-estimate technology, which alone would
discharge \coq{S3} and all eight \coq{Hw_*}, shrinking
\lbl{lemma:Stability}'s analytic base to the two diagonal Green identities
\coq{H_skew_diag}/\coq{H_ibp_diag} --- everything else it assumes being
conditions on the data and the mesh.

\subsection{Deliberately delegated to the SymPy suite}

The Appendix~A elemental matrices
(\file{elemental_matrices_verification.py},
\file{elemental_bilinear_form_verification.py}) are transcription checks by
symbolic differentiation of shape functions --- a computer-algebra task where a
proof assistant adds cost but no assurance. They are left to SymPy.

\subsection{Out of scope for standard-library Coq}

Two items need machinery absent from stdlib Coq (natural in Lean~4 + Mathlib),
covered only by the SymPy scripts: the subscale-norm equivalence of
\file{subscale_norm_verification.py} (Appendix~B9, Rayleigh-quotient spectral
theory), and the CDR adjoint / Green's-identity computation of
\file{cdr_operator_verification.py} (the divergence theorem in general form).

\subsection{The two standing caveats}

\begin{itemize}[nosep]
  \item \textbf{Statements, not models.} Coq certifies that each identity,
    inequality and limit follows from the real-number axioms (and, for the four
    abstract theorems, the named base). It cannot certify that these are the
    \emph{right} statements about the physics --- that requires reading them,
    which the crosswalk is meant to make easy.
  \item \textbf{Theory, not code.} Coq says nothing about whether the Julia or
    Kratos solvers implement the verified formulas. That correspondence is
    established empirically by the method-of-manufactured-solutions convergence
    tests in the codebase (the $O(h^{k+1})$ plateaus).
\end{itemize}

% =====================================================================
\section{Theory-side amendments applied for compatibility}\label{sec:amend}
% =====================================================================

Formalising the proofs surfaced a few points where the manuscript's hypotheses
or notation were slightly looser than what the proofs consume. The following
amendments (catalogued in \file{proof_verification/AUDIT.md}, findings F1--F8)
have been \textbf{applied} to \file{theory/paper/} so the manuscript is
$100\%$ compatible with the verified statements. None changes a mathematical
result; each closes a precision or hygiene gap. The added text is marked with
the manuscript's \verb|\amend{}| review macro.

\begin{longtable}{@{}cL{0.72\textwidth}c@{}}
\toprule
\textbf{ID} & \textbf{Amendment} & \textbf{Status} \\
\midrule \endhead
F1 & \lbl{H:advection} now grants $\mathbf a|_K\in W^{1,\infty}(K)^d$ elementwise (with the localisation of $\nabla\!\cdot(\alpha\mathbf a)=0$ to each $K$), which identity (iii)/\lbl{eq:elemibp} --- the Coq hypothesis \coq{H_elem_conv_ibp} --- actually uses. & applied \\
F2 & \lbl{H:porosity}'s derived bound $\alpha_{\infty,K}\le\delta_\alpha\alpha_{0,K}$ now states the element-convexity it relies on (matching \coq{delta_alpha_bound}/\coq{winv_ratio}). & applied \\
F3 & the two previously-unlabelled coefficient displays after \lbl{eq:UpperBoundOnEpsilon} now carry stable labels \lbl{eq:ViscousCoefficientBound} and \lbl{eq:VelocityCoefficientBound} (the anchors of \coq{visc_final}/\coq{u_855}); the SymPy headers were updated to match. & applied \\
F4 & the main-text triple norm now writes the elementwise-weighted term $\lVert\sigt_\alpha^{1/2}\mathbf u_h\rVert_h^2$ with the broken-norm subscript, matching \lbl{eq:triplenorm}. & applied \\
F6 & a remark after \lbl{eq:Tau2Final} records that the analysis $\tau_2$ and the implemented \lbl{eq:Tau2} are equivalent up to $1+C_2<2$ (machine-checked: \coq{tau2impl_le_tau2}, \coq{tau2_le_scaled_impl}). & applied \\
F7 & a clause notes that the coefficient bounds are written for $\aK=\alpha_{\infty,K}$; other representatives cost factors of $\delta_\alpha$. & applied \\
F8 & in \lbl{lem:winv} the label \lbl{eq:winv-conv} sat on the \emph{last} line of the display (the pressure-gradient estimate), while the convective line above it had none --- so two of the four references to \lbl{eq:winv-conv} pointed at the wrong estimate. The convective line now carries \lbl{eq:winv-conv} and the pressure-gradient line the new \lbl{eq:winv-gradp}; the four call sites were re-pointed accordingly (the Step-9 ``pressure part'' now cites \lbl{eq:winv-gradp}). Found by the Coq audit, which needed to cite the two estimates separately (\coq{Hw_cxu}/\coq{Hw_cxv} vs \coq{Hw_gpu}). & applied \\
\midrule
F5 & \file{supplement.tex} is untouched SIAM template boilerplate --- a submission-set decision for the authors, left unchanged. & deferred \\
\bottomrule
\end{longtable}

\noindent
After these edits the paper builds clean (\coq{latexmk} exit~0, $0$ undefined
references/citations); the two new equation labels resolve.

% =====================================================================
\section{Conclusion: how much this proves, in plain terms}\label{sec:conclusion}
% =====================================================================

It is worth stating plainly what a reader should take away.

\subsection{What machine-checking actually buys}

An ordinary proof is read by a human referee, once, under time pressure; a
missed sign or an unjustified ``clearly'' can survive into print. A
machine-checked proof is different in kind. Every single inference --- every
rewrite, every inequality, every case split --- is verified by a small, fixed
program (the kernel) that knows only the rules of logic and of real arithmetic
and cannot be hurried, fatigued, or won over by plausibility. Here that check
is then \emph{repeated by a second, independent program} (\coq{coqchk}) which
re-derives everything from the compiled proof terms without trusting the first.

The practical consequence: for every theorem in these eighteen files, the claim
\emph{``the stated conclusion follows from the stated hypotheses''} is no longer
a matter of belief or of referee diligence. The only way such a statement could
still be false is if the two independent kernels shared a soundness bug that
happened to strike exactly here --- for a logic and an implementation hardened
over three decades and exercised across thousands of developments, a
possibility so remote that it is no longer the dominant source of doubt. In the
terms of the request: the machine-checked \emph{deductions} are, to an
overwhelming degree, simply correct --- not ``probably correct'', but correct in
the same sense that $2+2=4$ is correct, with the residual uncertainty pushed
down to whether the checker itself is sound.

\subsection{The three things that still require human trust}

Certainty about a \emph{deduction} is not the same as certainty that a
\emph{theorem is the one you wanted}. A machine-checked result is an
implication --- ``if these hypotheses, then this conclusion'' --- and exactly
three things sit outside the kernel and carry whatever doubt remains.

\begin{enumerate}[leftmargin=1.6em]
  \item \textbf{Reading the statement correctly.} The kernel guarantees the
    proof matches the statement; it cannot guarantee the statement says what the
    author meant. \emph{This document is that check}: every Coq statement is
    mapped to the paper equation it encodes, and the definitions were diffed
    against the manuscript sources.
  \item \textbf{The hypotheses holding in the application.} For all four
    abstract theorems, a named list of standard analytic facts (the trusted base
    of \S\ref{sec:left}, read out in full in \S\ref{sec:assumptions}) is
    \emph{assumed}, not proved here. These are textbook results, hand-audited
    against the manuscript --- but ``hand-audited'' means human-checked, so they
    still carry the ordinary, small-but-nonzero risk of any careful proof. And
    two of them, \coq{Horth} and the face estimates, bundle enough of the paper's
    own derivation that auditing them means auditing the paper, not merely
    checking a citation (\S\ref{sec:stratumG}, \S\ref{sec:stratumI}).
  \item \textbf{The trust base.} The Coq kernel and the standard-library
    axiomatisation of the real numbers. Universally relied upon; named here for
    completeness.
\end{enumerate}

\subsection{The confidence ledger}

Putting it together, the results fall into three honest tiers.

\begin{longtable}{@{}L{0.24\textwidth}L{0.44\textwidth}L{0.26\textwidth}@{}}
\toprule
\textbf{Tier} & \textbf{What is in it} & \textbf{What must still be trusted} \\
\midrule \endhead
\textbf{1. Certain \newline (no analytic trust: positivity, or the paper's own
Class-D conditions)} &
All the scalar / algebraic content. Under positivity alone: the
$\tau_1,\tau_2,\sigt$ identities, the coefficient expansions, the Fourier and
coupling spectra, the asymptotic limits, the manufactured-solution divergence
identity, and the plateau bump's $C^1$ behaviour. Under the Class-D conditions
of \S\ref{sec:stratumD} as \emph{explicit premises} --- $\varepsilon\le
\varepsilon_{\max}$ (\lbl{eq:epscond}), $c_1>\xi\Cbar^2$ with $\xi>2$ (sharp; the
manuscript's $2\xi\Cbar^2$ additionally buys the \coq{C_stab_margin} floor),
$C_2<1$, $\aK\le1$, the jump bounds of \lbl{H:jump} --- the $\varepsilon$-smallness
chain, the coefficient lower bounds, and the parameter and jump inequalities. &
Only the trust base (kernel $+$ real-number axioms). The positivity hypotheses
are physically trivial, and the Class-D conditions are hypotheses of the paper's
theorems in any formal system: they add \emph{no analytic trust}. \\
\midrule
\textbf{2. Certain \emph{as an implication} \newline (conditional)} &
\lbl{lemma:Stability}, \lbl{lemma:Continuity}, \lbl{lem:continterp} and
\lbl{thm:convergence}: the coercivity bound
$B_S\ge\Cstab\,|||U_h|||^2$, the continuity bound
$B_S\le C_{\mathrm{tot}}(\mathrm{Br}_U+\mathrm{Br}_P)|||V|||$, the
interpolation-error continuity
$|B_S(E,V_h)|\le C_{\mathrm{tot}}^{I}(\Psi_U+\Psi_P)|||V_h|||$ and the
convergence estimate
$|||U-U_h|||\le(K_{\mathrm{abs}}^{I}+C_{\mathrm{tot}}^{I}/\Cstab)
(\Psi_U+\Psi_P)$, with explicit constants and every intermediate step
(Cauchy--Schwarz, Young, the eighteen-term bookkeeping, the jump
treatment, the summation over the mesh, the triple-norm triangle
inequality, the composition itself) inside the kernel. &
Trust base, \emph{plus} the named analytic hypotheses of \S\ref{sec:left} ---
some two dozen of them once the eight \coq{Hw_*} and seven \coq{HI_*} are
counted individually (Green-type identities, inverse and interpolation
estimates) --- holding. Standard, hand-audited, Lean-targeted; each is stated
and assessed in \S\ref{sec:assumptions}. Nothing more: the hypotheses are exactly
the manuscript's (\S\ref{sec:stratumD}). \\
\midrule
\textbf{3. Outside the Coq scope \newline (SymPy-verified)} &
The Appendix~A elemental-matrix transcription (left to SymPy by design);
the subscale-norm / CDR-adjoint items outside stdlib-Coq scope. &
The ordinary risk of a careful symbolic computation, mitigated by the
SymPy suite. \\
\bottomrule
\end{longtable}

\subsection{Where the residual gaps are, ranked}

In decreasing order of how much unformalised trust they still carry:
\begin{itemize}[nosep]
  \item \textbf{The analytic trusted base} (the hypotheses of Tier~2): the
    divergence-theorem / integration-by-parts identities and the finite-element
    inverse estimates. Standard and hand-audited, but not yet machine-checked.
    The Lean~4 roadmap targets exactly these, and notes that the
    inverse-estimate technology alone would shrink \lbl{lemma:Stability}'s
    residual analytic base to the two diagonal Green identities
    \coq{H_skew_diag}/\coq{H_ibp_diag}.
  \item \textbf{The remaining Green-type items of the convergence file}
    (\coq{Horth}, and the two diagonal identities \coq{H_skew_diag} /
    \coq{H_ibp_diag} that replaced the former \coq{HBS}/\coq{HBS_W} --- the
    first now \emph{proved} from them, the second eliminated outright):
    hand-audited, of the same character as the other Green identities;
    the interpolation estimates \coq{HI_*} join the Class-I Lean targets.
  \item \textbf{The data / mesh assumptions} (\lbl{H:jump}, \lbl{H:mesh}, the
    $\varepsilon$-condition, positivity): not gaps in any proof but
    \emph{conditions of applicability} --- hypotheses the mesh and data must
    satisfy, present in the paper's theorems in any formal system, with nothing
    to prove --- and, since this revision, with no exception: the development
    assumes exactly \lbl{H:data}'s $\sigma\ge0$, $\varepsilon\ge0$,
    $0\le C_2<1$, so the reaction-free case $\sigma=0$ \emph{is} an instance.
  \item \textbf{Modelling and code} (the two standing caveats of
    \S\ref{sec:left}): Coq checks the mathematics as stated, not that it models
    the physics, nor that the Julia/Kratos solvers implement it. The
    method-of-manufactured-solutions convergence tests are what pin the latter.
\end{itemize}

\subsection{Bottom line}

The parts of this paper that are pure calculation or pure logical deduction ---
the whole of the parameter design, the asymptotics and the manufactured-solution
algebra, and, crucially, the \emph{entire chain of estimates} inside the
stability, continuity, interpolation and convergence proofs --- are now
verified to a standard of certainty that human refereeing cannot reach --- under
exactly the manuscript's own hypotheses. What still rests on human judgement is a
\emph{short, explicit, named} list, and Table~\ref{tab:inventory} is that list in
full: $50$ hypotheses, of which $24$ are conditions on the data and the mesh
(nothing to prove) and $26$ are analytic. Of those $26$, sixteen are instances
of two textbook facts (the inverse estimate and Bramble--Hilbert), seven are the
divergence theorem, its two diagonal instances, and face bookkeeping, and the
genuine residue is two items: \coq{Horth} and the face bundle. Those, and the
physical- and code-modelling questions that no symbolic proof can settle. The
value of the exercise is precisely that it shrinks what must be taken on trust
from ``the whole proof'' to that short list --- and puts the list in plain sight.

% =====================================================================
\section{Reproducing the check}\label{sec:repro}
% =====================================================================

From \file{proof_verification/coq-formal/}:
\begin{lstlisting}[language=bash,keywordstyle=\color{black}]
# one-time: install Coq / Rocq (any of)
brew install coq          # macOS (installs Rocq 9.1.x, provides coqc/coqchk)
apt install coq           # Debian/Ubuntu (Coq 8.x)
opam install coq

# verify
./run_all.sh              # per-file PASS + coqchk kernel re-verification
\end{lstlisting}

\noindent
A passing run prints one \coq{[PASS]} line per file, then the \coq{coqchk} line.
The run recorded for this revision (Coq~8.18.0, Ubuntu~24.04, $\approx75$\,s
wall clock) is:
\begin{lstlisting}[language=bash,keywordstyle=\color{black}]
Coq formal verification suite (PNSFormal)
coqc: The Coq Proof Assistant, version 8.18.0
  [PASS] Limits.v            [PASS] AbstractContinuity.v
  [PASS] DerivKit.v          [PASS] AbstractInterpolation.v
  [PASS] StabilityAlgebra.v  [PASS] AbstractConvergence.v
  [PASS] ContinuityAlgebra.v [PASS] TauDesign.v
  [PASS] AbstractSums.v      [PASS] Asymptotics.v
  [PASS] InnerSpace.v        [PASS] ManufacturedSolution.v
  [PASS] InverseEstimates.v  [PASS] PlateauBump.v
  [PASS] AbstractStability.v [PASS] NonVacuityInterp.v
  [PASS] NonVacuity.v        [PASS] NonVacuityConv.v
RESULT: all 18 files compiled; every theorem is proved.
Running coqchk (independent kernel re-verification) ...
  [PASS] coqchk: modules successfully checked by the trusted kernel.
\end{lstlisting}
(The sources keep the $8.16$--$8.20$-compatible \coq{From Coq} spelling on
purpose; on Rocq~9 it merely emits a deprecation notice in favour of
\coq{From Stdlib}.)

\bigskip
\noindent\emph{Cross-references.} File-level detail and the SymPy$\leftrightarrow$Coq
mapping: \file{README.md} in the \file{coq-formal/} directory. The step-by-step
audit of the analytic steps and the F1--F8 findings: \file{AUDIT.md}. The Lean~4
plan for the residual base: \file{LEAN_ROADMAP.md} and \file{PorousNSToolbox.lean}.
The authoritative theory: \file{theory/paper/article.tex} and its appendices
(\file{continuity_appendix.tex}, \file{fourier_appendix.tex},
\file{elemental_matrices_appendix.tex}).

% =====================================================================
\begin{thebibliography}{9}
% =====================================================================

\bibitem{brenner2008mathematical}
S.~C.~Brenner and L.~R.~Scott,
\emph{The Mathematical Theory of Finite Element Methods}, 3rd~ed.,
Texts in Applied Mathematics~15, Springer, New York, 2008.
\newblock (Inverse estimates: Lemma~4.5.3. Interpolation / Bramble--Hilbert:
Theorem~4.4.20.)

\bibitem{ernguermond}
A.~Ern and J.-L.~Guermond,
\emph{Theory and Practice of Finite Elements},
Applied Mathematical Sciences~159, Springer, New York, 2004.
\newblock (Inverse inequalities on shape-regular meshes and the local
interpolation estimates in the form used here.)

\bibitem{codina2001stabilized}
R.~Codina,
\emph{A stabilized finite element method for generalized stationary
incompressible flows},
Comput.\ Methods Appl.\ Mech.\ Engrg.\ \textbf{190} (2001), 2681--2706.
\newblock (Source of the jump condition of which \lbl{H:jump} is the analogue
--- condition~(39) there --- and of the $L^\infty$ interpolation estimates
(78)--(79).)

\bibitem{codina2018hp}
R.~Codina,
\emph{On $hp$ convergence of stabilized finite element methods for the
convection--diffusion equation},
SeMA Journal \textbf{75} (2018), 591--606.
\newblock (The $k^2$ growth of $\Cinv$ with the polynomial degree, which
motivates the calibration $c_1=4k^4$, $c_2=2k^2$.)

\end{thebibliography}

\end{document}
